### 2020-03-08 data-raw/model_documentation.Rmd
First commit woth a minimal package that builds but does not yet make the new design zone structures.

---
title: "Abalone MSE Model Documentation"
author: "Malcolm Haddon"
date: "`r format(Sys.time(), '%d&period; %B %Y')`"
output: 
    word_document:
      reference_docx: C:\Users\User\Dropbox\A_Book\book\trialreferencestyle.docx
      toc: true
      toc_depth: 3
      fig_caption: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE)

options(knitr.kable.NA = "",
        knitr.table.format = "pandoc")

wkdir <- "C:/Users/User/Dropbox/AbMSE/AbMSE/data-raw/"

options("show.signif.stars"=FALSE,"stringsAsFactors"=FALSE,
        "max.print"=50000,"width"=240)

library(rutilsMH)
library(knitr)
library(captioner)

tab_nums <- captioner(prefix = "Table")
fig_nums <- captioner(prefix = "Figure")
equ_nums <- captioner(prefix = "Equ.")
```

##### l

NOTE THERE ARE ISSUES WITH THE USE OF $L$ TO REFER TO PARTICULAR LENGTHS, AND $i$ TO REFER TO PARTICULAR SIZE-CLASSES. a DECISION IS NEEDED ABOUT SUCH NOMENCLATURE.

## General Model Structure

The focus of this work are Australian abalone stocks, including blacklip abalone (_Haliotis rubra_) and greenlip abalone (_Haliotis laevigata_). The underlying population dynamics model will have an annual time-step (with summer and winter seasons - see later on within year dynamics), and be described using a length-based model using a size-transition matrix to describe the growth in size each time period. The option of moving to a combined length-and-age-based model using a transition matrix will be explored once this new model has been developed to determine whether the advantages of such an extension outweight the disadvantages of extended computation time.

### Model Dynamics

The model dynamics follows the numbers-at-size through time as they are affected by natural mortality, somatic growth, fishing mortality, and recruitment. The model developed in Haddon _et al_. (2013) and Haddon and Helidoniotis (2013) used separate vectors of numbers-at-size to describe the cryptic and emergent components of each population. After exploratory work that found there were no real advantages to using this structure, the model has been simplified through the cryptic and emergent components of the population now being contained in the single vector $\mathbf{N_t}$, where $\mathbf{N_t}$ is assumed to be the numbers-at-size (shell length in mm) at the start of each year $t$. This simplifies the equations and speeds the calculations although with this approach the effect of emergence needs to be included explicitly in some of the equations describing the dynamics.

The model structure adopted to reflect the assumed annual dynamics begins at the start of each year with half of the survivorship from natural mortality occurring followed by individual growth, then survivorship from fishing mortality, followed by the remaining survivorship from natural mortality. Finally any recruitment in that year is added to the first few size classes of of the population vector $\mathbf{N_t}$.  If natural mortality is implemented as the survivorship from half of natural mortality, that is $\mathbf{S_h} = e^{-M/2}$, twice a year, with other dynamic processes in between, then the dynamics for the numbers-at-size can be represented in matrix notation as:

`r equ_nums("m1")`

$${\mathbf{{N}_{t+1}={S}_{h}{AGS}_{h}{N}_{t}+ {R}}}$$

where $\mathbf{S_h}$ is the survivorship following half of the instantaneous natural mortality, $M$, $A$ is the survivorship following the imposition of fishing mortality, $\mathbf{G}$ is the growth transition matrix, and $\mathbf{R}$ is the vector of numbers recruiting into each size-class within $\mathbf{N_{t+1}}$.

The survivorship following fishing mortality is defined as:

`r equ_nums("m1a")`

$$\mathbf{A} = e^{-\mathbf{s}F}$$

where $\mathbf{s}$ is the vector of selectivity-at-length (or size), and $F$ is the fully selected, instantaneous fishing mortality rate. The simplification means that now the transition from cryptic and emergent no longer needs to be included in the annual dynamics.

## Model Initiation

Model initiation will always begin with the population being assumed to be at equilibrium in the absence of fishing. This implies that, using __Equ `r equ_nums("m1", display="num")`__, the survivorship from the annual harvest rate, $\mathbf{A}$ = 1.0, so that, at equilibrium, it has no effect:

`r equ_nums("m2")`

$${\mathbf{{N}^{*}={S}_{h}{AGS}_{h}{N}^{*}+ {R}}}$$

This can be re-arranged to obtain an analytic expression for the equilibrium numbers-at-length $\mathbf{N^*}$, thus:

`r equ_nums("m3")`

$${\mathbf{{N}^{*}-{S}_{h}{AGS}_{h}{N}^{*} = {N}^{*}({I-}{S}_{h}{GS}_{h})= {R}}}$$

which, finally, implies:

`r equ_nums("m4")`

$${\mathbf{{N}^{*}=\left ({I}-{S}_{h}{GS}_{h} \right )^{-1} {R}}}$$

If there is an initial estimated fishing mortality rate, this can be defined as the complement of an annual harvest rate and is distributed down the diagonal of an otherwise zero square matrix $\mathbf{A}$:

`r equ_nums("m5")`

$$A_L = \left (1 - s_{L,t}H_t \right )$$

where $A_L$ is the survivorship of length-class $L$, $s_{L,t}$ is the selectivity of length-class $L$ in year $t$ (which will relate to the Legal Minimum Length in that year = _LML_), and $H_t$ is the fully selected harvest rate in year $t$. With an initial constant fishing mortality rate the equilibrium numbers for the emergent population would then become:

`r equ_nums("m6")`

$${\mathbf{{N}^{*}=\left ({I}-{S}_{h}{AGS}_{h} \right )^{-1} {R}}}$$

One important aspect of __Equ `r equ_nums("m6", display="num")`__ is that it can be used to search for a constant harvest rate that will return a specified initial depletion level. This might be required should such an initial state be wanted for testing how a given harvest strategy might operate when it begins to be applied to a stock when that stock is in different initial states of depletion.

## Biology and Stock Related Statistics

### Emergence

A logistic curve (Haddon, 2011) can be used to describe the transition from the cryptic to the emergent component of the population, when this is required. Given the presence of a Legal Minimum Length (LML), crypsis and emergence only become influential on the dynamics if natural mortality differs between the two components or if the emergence logistic overlaps with the selectivity curve, and or the LML.

`r equ_nums("m7")`

$$E_L=\frac{1}{1+exp \left [{-log(19)(L-L_{E50})/(L_{E95}-L_{E50})} \right ]}=\frac{1}{1+exp \left [{-log(19)(L-L_{E50})/\delta_E} \right ]}$$

where $E_L$ is the proportion of size-class $L$ that are emergent, and $L_{E50}$ and $L_{E95}$ are the usual logistic parameters defining the lengths at which 50% and 95% are emergent. The term $\delta_E$ is the constant $(L_{E95}-L_{E50})$. Emergence from crypsis only becomes an issue for the dynamics of the model if they are considered to have different natural mortality rates and/or when the emergence curve overlaps with the selectivity curve (which it can do when the LML is low, e.g. 127mm early on in Tasmania, especially on the west coast). Where the selectivity and emergence curves overlap then the proportion remaining in crypsis would act as a refuge, effectively reducing the fishing mortality on those size classes.

### Selectivity

Selectivity, $S_{L,t}$ for length $L$ (= size-class $i$) in year $t$ needs to be defined by year to permit changes in the LML to be reflected in the selectivity by divers. This implies that rather than a single vector of values, a matrix of selectivity values will be required in the simulations one column per year. Each year's selectivity is defined as:

`r equ_nums("m13")`

$$S_{L,t}=\frac{1}{1+{exp \left [ {-log(19)( L-{L_{s50}})/{\delta_S}} \right ] }}$$

where $\delta_S=L_{s95}-L_{s50}$.

### Growth

The growth from size-class $j$ to size-class $i$ is described by the elements of a growth transition matrix defined by:

`r equ_nums("m8")`

$$\begin{align}
  & \begin{matrix}
   {G_{i,j}}=\int\limits_{-\infty }^{{L_i}+{LW/2}}{\frac{1}{\sqrt{2\pi }{\sigma_{j}}}\exp \left( -\left[ \frac{{L_i}-{{\bar{L}}_{j}}}{2{({\sigma_{j}})^{2}}} \right] \right)}dL & {L_i}={L_{Min}}  \\
\end{matrix} \\ 
 & \begin{matrix}
   {G_{i,j}}=\int\limits_{{L_i}-{LW/2}}^{{{L}_{i}}+{LW/2}}{\frac{1}{\sqrt{2\pi }{\sigma_{j}}}\exp \left( -\left[ \frac{{L_i}-{{{\bar{L}}}_{j}}}{2{({\sigma_{j}})^{2}}} \right] \right)}dL & {{L}_{Min}}<{L_{i}}\le {L_{Max}}  \\
\end{matrix} \\ 
\end{align}$$

where $G_{i,j}$ is the probability of growing from size class $j$ into size class $i$, $LW$ is the size-class width, $\sigma_j$ is the standard deviation of the normal curve describing the growth increments of animals starting in size class $j$, $L_i$ is the length of size class $i$, and $\bar{L}_{j}$ is the mean growth increment of animals starting from the mean of size-class $j$. $L_{Min}$ and $L_{Max}$ are the minimum and maximum size-classes, with the maximum being treated as a plus group. To make $L_{Max}$ a plus group, and at the same time ensure that all columns sum to 1.0 (to prevent growth implying losses of its own), the final row of the matrix is modified for each column $j$ as:

`r equ_nums("m9")`

$${{G}_{{L}_{Max,j}}}={{G}_{{L}_{Max,j}}}+\left( 1-\sum\limits_{i={{L}_{1}}}^{{{L}_{Max}}}{{{G}_{i,j}}} \right)$$

The expected mean growth increment for each size-class $i$ is defined using an inverse logistic growth curve that has been found to describe blacklip abalone growth well (Haddon et al. 2008; Helidoniotis et al., 2011):

`r equ_nums("m10")`

$${\bar{L}_{i,j}={L_j}+\frac{Max\Delta L}{1+{exp \left[ {log\left( 19 \right)\left( {{L}_{j}}-{L_{g50}} \right)/\delta_g} \right ]}}+{{\varepsilon }_{{L_j}}}}$$

$Max \Delta L$ is the maxiumum growth increment for the population, $L_g50$ and $L_g95$ are the usual logistic parameters defining the lengths at which 50% and 5% of the maximum growth increment are expressed. The ${\varepsilon}_{{L}_{i}}$ is the 
Variation around the mean expected growth increment. It is assumed to be normally distributed with a standard deviation that varies with the growth increment (Haddon et al. 2008):

`r equ_nums("m11")`

$${{\sigma }_{L_j}}=\frac{Max{{\sigma }_{L}}}{1+{exp \left[ {log\left( 19 \right)\left( {L_j}-{L_{g95}} \right)/\delta_g} \right ]}}$$

The $\left( L_{Max}-{L_{g95}} \right)$ remains a constant and can be parametereized as such ($\delta_{g\sigma}$).

### Weight-at-Length

The weight-at-length, $W_L$, relationship invovles two constants:

`r equ_nums("m12")`

$$W_L=aL^b$$

### Maturity-at-Length

Maturity at size, $m_L$,  uses an alternative logistic curve, again with two parameters, only this time $L_{50}\% = -{\alpha}/{\beta}$ and the inter-quartile distance is $2{\text{Ln}}3 \beta$.

`r equ_nums("m12")`

$$m_L=\frac{exp(\alpha + \beta L)}{1+exp(\alpha + \beta L)}=\frac{1}{1+\left(exp(\alpha + \beta L)  \right)^{-1}}$$

### Spawning and Exploitable Biomass

Mature or spawning biomass needs to include numbers-at-size by maturity-at-size and weight-at-size:

`r equ_nums("m14")`

$$B_{t}^{S}=\sum\limits_{L={L_{Min}}}^{L_{Max}}{\left( {m_L}{W_L}{N_{L,t}} \right)}$$

Spawning biomass is, like exploitable biomass, calculated in the same units as the $W_L$ equation. If that is in grams then it requires division by 1000,000 to estimate tonnes, if in kg then division by 1000 is required. Exploitable biomass, here, is estimated after half of natural mortality and growth have occurred and before any fishing mortality occurs in any single year. Only emergent biomass is considered as no fishing mortality is imposed on the cryptic component, but this is only important is the emergence curve overlaps the selectivity curve:

`r equ_nums("m15")`

$$B_{t}^{E}=\sum\limits_{L={{L}_{Min}}}^{{{L}_{Max}}}{{{s}_{L,t}}{{E}_{L}}{{W}_{L}}G_{L,j}{e^{-M/2}}N_{L,t}}$$

where the exploitable numbers-at-size $L$ in year $t$, ${{N}_{L,t}^{E}}$, is obtained from

`r equ_nums("m16")`

$${{N}_{L,t}^{E}}=G_{L,j}{e^{-M/2}}{{s}_{L,t}}{{E}_{L}}N_{L,t}$$

where $N_{L,t}$ is the numbers-at-size $L$ at the start of year $t$.

### Catchability

Catchability in a stock assessment model can be estimated analytically as:

`r equ_nums("m17")`

$$q=\exp \left[ \sum\limits_{t=1}^{n}{Ln\left( {{I}_{t}}/B_{t}^{E} \right)}/n \right]$$

where $n$ is the number of years across which the observed catch rates, $I_t$, and predicted exploitable biomass are considered. In the simulation/operating model a maximum catch rate, $CE_{Max}$ was used to scale the unfished exploitable biomass to generate a catchability value for each population. The maximum catch rates were randomly selected from a pre-specified distribution, and then the following equation used:

`r equ_nums("m18")`

$$q_p=\left( CE_{Max}/{B_{0,p}^{E}} \right)$$

where $p$ is the index for each population and ${B_{0,p}^{E}}$ is the unfished exploitable biomass for population $p$.

### Annual Model Dynamics

Once each population is initiated its dynamics can be projected forwards a year at a time depending on how much catch is expected to be taken or how much effort is expected to be focussed into each population. The population initiation sets up the equilibrium numbers for the properties defined for each population. Then, given a specific harvest rate for each population they can be projected forward in yearly steps. This projection is based around how the numbers-at-size change through fishing, growth, natural mortality, and recruitment. As before, the fishing mortality rate over a year is defined as the complement of an annual harvest rate and is distributed down the diagonal of an otherwise zero matrix __A__:

`r equ_nums("m19")`

$$A_L=\left(1-{s_{L,t}H_t} \right)$$

where $A_L$ is the survivorship of length class $L$, $s_{L,t}$ is the selectivity of length class $L$ in year $t$, and $H_t$ is the fully selected harvest rate in year $t$ (the harvest rate being the proportion of exploitable biomass taken as catch). We can define the survivorship from applying half of natural mortality as follows:

`r equ_nums("m20")`

$$O_S=e^{-M/2}$$

$O_S$ does not need to be a vector as multiplying a matrix or vector by a constant is simpler. We apply this survivorship twice in a year with the other dynamics occurring between:

$${\mathbf{N_{t+1}}}={O_S}\left[{\mathbf{GA_t}}{O_S}{\mathbf{N_t}}\right] + {\mathbf{R}}  $$

##### l
## References

Haddon, M. (2011) _Modelling and Quantitative Methods in Fisheries_. 2nd Ed. CRC/Chapman & Hall. 449p.

Haddon, M., Mundy, C., and D. Tarbath (2008) Using an inverse-logistic model to describe growth increments of blacklip abalone (_Haliotis rubra_) in Tasmania. _Fishery Bulletin_ __106__:58-71

Haddon, M. and F. Helidoniotis (2013) Legal minimum lengths and the management of abalone fisheries. _Journal of Shellfish Research_ __32__:197-208

Haddon, M., Mayfield, S., Helidoniotis, F., Chick, R. and C. Mundy (2013) _Identification and Evaluation of Performance Indicators for Abalone Fisheries_. FRDC Final Report 2007/020. CSIRO Oceans and Atmosphere and Fisheries Research Development Corporation. 295 p.

Helidoniotis, F., Haddon, M., Tuck, G., and D. Tarbath (2011) The relative suitability of the von Bertalanffy, Gompertz and inverse logistic models for describing growth in blacklip abalone populations (_Haliotis rubra_) in Tasmania, Australia. _Fisheries Research_ __112__: 13-21

### 2020-03-18 data-raw/development.Rmd
Beginning the transition to region/SMU and the breaking up of the data files and zone structure.

--- 
title: "Issues Developing the new Abalone MSE"
author: "Malcolm Haddon"
date: "`r Sys.Date()`"
font: Times
fontsize: 12pt
output: html_document

---

```{r setup, include=FALSE}
# output: bookdown::gitbook
# bookdown::epub_book: default
# before_body: latex/before_body.tex from output.yml
knitr::opts_chunk$set(size="scriptsize")
knitr::opts_chunk$set(
  cache = FALSE,
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	comment="#",
	dpi = 300,
	fig.align = 'center')

options(knitr.kable.NA = "",
        knitr.table.format = "pandoc")

#wkdir <- "./../../A_Book/book/"

options(show.signif.stars=FALSE,
        stringsAsFactors=FALSE,
        comment="#",
        str = strOptions(strict.width = "cut"))

library(rutilsMH)
library(aMSE)
suppressPackageStartupMessages(library(knitr))
#library(captioner)
library(diagrams)

setpalette("R4")

```

## Main Issues

* How best to define each population in a sequential manner reflective of their relative productivity? Need to keep the option of random allocation of properties. See section.

* Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular region, there is a need to include the option of having larval dispersal.

* Rather than use one data-file and one ctrl file, perhaps break up the data required into a file for PDF definitions by population, a file for the variables common across the region, and a control file.

* There needs to be a simpler implementation of larval dispersal this time through. This will have implications for any initial equilibrium conditions. See section.

* In an attempt to speed the model runs it will be worthwhile exploring the advantages and disadvantages of splitting the region into two, the constant part, regionC, and the dynamic part, regionD. Is the constant part always separate from the global part?

* Need to develop a new twoblock.csv file containing definitions for 6 populations rather then generalize by block. This would imply that the variance for eahc population would be smaller than that used for each block.

### function: definepops

When generating the properties of each population, currently, the population properties, as defined by the PDfs describing each property, are allocated randomly. With the advent of the spatial data it should be possible to condition the sequence of populations along the coastline in a manner more reflective of the relative productivity of each small area selected to represent a model population.

Random is relatively easy, but both random and specific will be driven by the conditioning. Perhaps best done in relative terms. A detailed examination of hte distribution of ctaches among and between the eastern Block 13 may be illuminating.

Currently the data-file used to condition the definepops function simply lists the mean values of the parameters for all populations per block. It will become necessary to define each populatin spearately to implement sequential definitions. Rather than conditioning entailing continual running until a combination arises that generates about the correct totals, it will be necessary to develop routines to assist searching for suitable levels of productivity for each population. Perhaps there are some simplifiying relationships between Me, G, and AvRec that can aid in this search conditioning.

### larval dispersal

To enable larval dispersal it will be necessary to alter the algorithm whereby each population is dealt with completely before movingon to the next population. Thus, it will be necessary to complete the annual dynamics, except for recruitment, for all populations and only when all are available undertake the recruitment with (or without dispersal). One detailed part will be achieving whatever initial equilibrium conditions are desired (unfished, some given depletion level, etc.).

There is no analytical solution to the initial equilibrium if there is to be larval dispersal.

The dispersal matrix could be initially defined as a matrix and that can be tested to determine whether only the diagonal is filled.

## Highest Level Outline of Abalone MSE

1) Initiate simulated region - use a ctrl file to define details of data-file, HCR function, and other initial conditions. May involve setting an initial depletion level.

2) conduct assessment of region status at start of each year (option for partial year?)

3) Using selected HCR set catches for subsequent year(s), and allocate among populations using model fleet dynamics.

4) run the model dynamics for one year

5) If simulation duration not complete return to item 2)

6) If complete then summarize and store results

### Requirements

1) A data-file that contains details for each population

* SMUname    - eg block1, block2
* popnum - population index   1, 2, 3, etc
* MaxDL      - maximum growth increment  eg 32
* sMaxDL     - variation of maxDL  eg  2.5
* L50        - Length at 50% MaxDL  eg 123
* sL50       - variation in L50  eg  5
* L50inc - L95 - L50 =  delta  eg 45
* sL50inc - variation of L50inc  eg 1.25
* SigMax - maximum variation around growth increment eg 4.6
* sSigMax - variation in SigMax  eg 0.1
* Wtb - weight-at-length exponent eg 3.161963
* sWtb - variation of Wtb  eg 0.148
* Wtbtoa - intercept power relationship betweeb Wtb and Wta 962.8098
* sWtbtoa - exponent of power relationship betweeb Wtb and Wta -14.3526
* Me - Natural mortality eg 0.2 maxage = 23, 0.1 maxage = 46 
* sMe - variation in Me  eg 0.003
* AvRec - log of unfished recruitment LnR0  eg 13
* sAvRec -   eg 1.0
* defsteep - Beverton-Holt Steepness  eg 0.6
* sdefsteep         eg 0.02
* L50C - length at 50% emergent  126.422
* sL50C                           1.5
* deltaC - length at increment 95% emergent 95 = 50 + deltaC eg  10
* sdeltaC                                      0.5
* MaxCEpar - maximum cpue tonnes/hour for defining catchability  0.37
* sMaxCEpar                              0.02
* selL50p - L50 of selectivity           0   (deviation from LML)
* selL95p - L95 of selectivity           1.5
* SaMa - maturity logistic a parameter  eg -16
* L50Mat - L50 for maturity b = -a/L50  eg  123.384
* sL50Mat                                   4.0
* larvalD - larval drift                    0.02  = 2%

2) A data-file containing details relating to the whole region

* numpop - number of populations
* PDFS   - The number of parameters with probability density functions
* minc   - minimum size class  2mm
* cw     - class width         2mm 
* Nclass - number of size classes 105
* nSMU   - number of Spatial Management Units  eg  2
* LML    - current Legal Minimum Length (LML, MLL, MLS, etc) e.g. 140
* randomseed - for repeatability of results - use a different one for each simulation
* Nyrs   - number of years for each simulation   eg 40
* firstYear - first year of simulation eg 1 (hypothetical) or 2014 (conditioned on fishery)
* fixYear - ?  eg 3  used in plotsimulation3
* projLML - the LML to be used in the simulated dynamics

3) A control-file containing details of each run

* label for run
* region file name
* datafile name
* HCR details file
* replicates
* HCRlabel
* output directory

### 2020-03-18 R/inputfiles.R

@@ -102,9 +102,9 @@ diagnostics <- function(inzone,plot=TRUE) {   # inzone <- testzone
 }  # end of diagnostics
 
 
-#' @title datafileTemplate generates a template input datafile
+#' @title datafiletemplate generates a template input datafile
 #'
-#' @description datafileTemplate generates a standard input datafile to use
+#' @description datafiletemplate generates a standard input datafile to use
 #'     as a template. It is possible to define the number of blocks and
 #'     then, once the data file is created, go in an edit it appropriately
 #'     to suit exactly your own needs.
@@ -122,16 +122,16 @@ diagnostics <- function(inzone,plot=TRUE) {   # inzone <- testzone
 #'     included in the data file. These would be used in any stock reduction
 #'     that becomes part of the conditioning; defaults to FALSE
 #'
-#' @return a standard definition data file ready to be read by readdataFile
+#' @return a standard definition data file ready to be read by readdatafile
 #' @export
 #'
 #' @examples
 #' \dontrun{
 #'  filename <- dataTemplate()
-#'  condDat <- readdataFile(filename)
+#'  condDat <- readdatafile(filename)
 #'  str(condDat,max.level=1)
 #' }
-datafileTemplate <- function(numblock=3,Nyrs=NA,filename="tmpdat.csv",
+datafiletemplate <- function(numblock=3,Nyrs=NA,filename="tmpdat.csv",
                              FisheryData=FALSE) {
    genconst <- function(invect) {
       nlab <- length(invect)
@@ -210,107 +210,6 @@ datafileTemplate <- function(numblock=3,Nyrs=NA,filename="tmpdat.csv",
 
 # Utility functions used within parseFile, not exported
 
-#' @title getConst extracts 'nb' numbers from a line of text
-#'
-#' @description getConst parses a line of text and extracts 'nb' pieces of
-#'     text as numbers
-#'
-#' @param inline text line to be parsed, usually obtained using readLines
-#' @param nb the number of numbers to extract
-#' @param index which non-empty object to begin extracting from?
-#'
-#' @return a vector of length 'nb'
-#' @export
-#'
-#' @examples
-#' \dontrun{
-#'   # Not exported, prefix with AbMSE:::
[…]

### 2020-03-18 data-raw/MakeDatafor_AbMSE.r

@@ -1,66 +1,0 @@
-#rm(list=ls())   # Cleanup the R console if required
-# Set up the run ----------
-
-
-# library(r4cpue)
-library(mhutils)
-library(aMSE)
-library(microbenchmark)
-
-
-options("show.signif.stars"=FALSE,"stringsAsFactors"=FALSE,
-        "max.print"=50000,"width"=240)
-#     listFunctions("C:/A_Mal/A_Book/rcode/agestruct/age_utils.r")
-
-# two block abalone zone with 6 populations -----------
-
-
-# Setup data for aMSE -----------------------------------------------------
-
-datadir <- "./../../rcode2/aMSE/data-raw/"
-
-infile <- paste0(datadir,"twoblock.csv")
-
-#datafile <- datafileTemplate(numblock=3,filename=outfile)
-condDat <- readdataFile(infile)
-str(condDat,max.level = 2)
-
-save(condDat,file=paste0(datadir,"condDat.RData"))
-
-
-# check and transfer -------------------------------------------------------
-
-
-
-tools::checkRdaFiles(paths=datadir)
-tools::resaveRdaFiles(paths=datadir,compress="auto")
-tools::checkRdaFiles(paths=datadir)
-
-devtools::use_data(condDat,
-                   pkg="C:/A_mal/Rcode/Abalone/AbMSE",
-                   internal=FALSE, overwrite=TRUE)
-
-devtools::use_data(condDat3,
-                   pkg="C:/A_mal/Rcode/Abalone/AbMSE",
-                   internal=FALSE, overwrite=TRUE)
-
-devtools::use_data(abdat,
-                   pkg="C:/A_mal/Rcode/Abalone/abspatial",
-                   internal=FALSE, overwrite=TRUE)
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-

### 2020-04-13 README.Rmd
Moved to a readme.Rmd file to generate the GitHub readme.md file.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aMSE

<!-- badges: start -->
<!-- badges: end -->

This packages up a new Abalone Management Strategy Evaluation framework. It has a novel structure that is based around the spatial ideas of populations within spatial management units (SMUs), within a region:

* region - highest geogrpahical level. Is simply the totality of the spatial management units.

* SMU - spatial management units. In Tasmania these would currently be the classical statistical blocks.

* population - literally a population. These are the active components within the simulation framework. The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighboiuring populations on each other.

This conceptual structure has implications for how the simulations are conditioned during the management strategy testing. While it is still the case tha tthe general properties of each SMU are used to set the scene for each SMU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population. Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SMU.

An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant. This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.

Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examin its implications.

## Installation

You can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:

``` r
# install.packages("devtools")
devtools::install_github("haddonm/aMSE")
devtools::install_github(haddonm/rutilsMH) # also useful for plotting, etc
```
## Example

This is a basic example which illustrates the generation of an initial equilibrium. It uses built in data-sets but usually you would generate regionC (constants) and regionD (the dynamic components) from the conditioning data:

```{r example}
# basic example code
library(aMSE)
library(rutilsMH)

  data(region1)
  glb <- region1$globals
  data(testregC)
  data(testregD)
  regDe <- testequil(regC=testregC,regD=testregD,glob=glb)
  str(regDe)

```

### 2020-04-14 data-raw/development.Rmd
added findF1, cleaned the ctrl file, the makehtml files, plotprod and setupdirs, removed file.path in favour of filenametopath as more general.

@@
 
+* The region is simply to container for all the SMU (there may only be one).
+The notion is that the variation of population properties within an SMU can differ, but conditioning the populations is a whole subject for discussion in itself.
+
 * Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular region, there is a need to include the option of having larval dispersal.
 
-* Rather than use one data-file and one ctrl file, perhaps break up the data required into a file for PDF definitions by population, a file for the variables common across the region, and a control file.
+* Currently using a control.csv file to identify conditioning files, HCR files, and related.
+There is a region file containnig the strucutre of the simulated region, and a datafile containing the conditioning data.
 
-* There needs to be a simpler implementation of larval dispersal this time through. This will have implications for any initial equilibrium conditions. See section.
+* When constructing a region the constant parts are kept in regionC and the dynamic parts in regionD.
+Is the constant part always separate from the global part?
+Certainly not having to move such large objects around all the time should speed things along.
 
-* In an attempt to speed the model runs it will be worthwhile exploring the advantages and disadvantages of splitting the region into two, the constant part, regionC, and the dynamic part, regionD.
-Is the constant part always separate from the global part?
+## Design Principles
 
-* Need to develop a new twoblock.csv file containing definitions for 6 populations rather then generalize by block.
-This would imply that the variance for eahc population would be smaller than that used for each block.
+* Each run will be conducted in a specified directory, which will always contain at least subdirectories named 'data', 'plots', and 'tables'.
+
+* Each run will be able to generate an html review of all output (see makehtml_utils.R)
+
+* Currently using a spatial hierarchy region/SMU/population instead of zone/block/population, so this is now more general than only Tasmanian terminology.
+Stick to this hierarchy throughout the code.
+Consistency.
+
+* Naming things is always a complex issue, the LML term is still being used, the other States each appear to use different terms: eg. MLL and other three letter acronyms, so stick with LML.
+
+* Once a particular run is constructed, but before the run proceeds, generate a series of plots and tables to characterize the simulated region.
+
+* All plots should be generated as .png files and stored in the 'plots' directory of each runs directory.
+In addition, any tables should be saved in the same place as csv files (for matrices) or RData files (for more complex objects) ready to be read in for printing in the html output.
+Each plot and table will have a unique name that will include the run's name.
+This will nbe necessary to automate the html output and later for any .Rmd output.
 

### 2020-04-14 R/aMSE_utils.R

@@ -12,7 +12,11 @@
 #'     whose existence is to be checked before it is created if it
 #'     does not already exist.
 #' @param make if the directory does NOT exist should it be created.
-#'     default = TRUE
+#'     default = TRUE; if make=FALSE and a directory does not exist
+#'     a warning will be given to the console.
+#' @param verbose default=TRUE, prints directory status to the console,
+#'     If make is set to FALSE and a directory does not exist a
+#'     warning will always be given.
 #'
 #' @return a message to the screen if the directory exists or is
 #'     created; if make is TRUE then it also creates the directory as
@@ -22,15 +26,15 @@
 #' @examples
 #' indirect <- getwd()
 #' dirExists(indirect)
-dirExists <- function(indir,make=TRUE) {
+dirExists <- function(indir,make=TRUE,verbose=TRUE) {
   if (dir.exists(indir)) {
-    cat(indir,":  exists  \n")
+    if (verbose) cat(indir,":  exists  \n")
   } else {
     if (make) {
-      dir.create(file.path(indir))
-      cat(indir,":  created  \n")
+      dir.create(indir)
+      if (verbose) cat(indir,":  created  \n")
     } else {
-      cat(indir,":  does not exist \n")
+      warning(cat(indir,":  does not exist \n"))
     }
   }
 }  # end of dirExists
@@ -51,7 +55,7 @@ dirExists <- function(indir,make=TRUE) {
 #' @export
 #'
 #' @examples
-#' indir <- "C:/Users/Malcolm/Dropbox/rcode2/aMSE/data-raw"
+#' indir <- tempdir()
 #' infile <- "control.csv"
 #' filenametopath(indir,infile)
 filenametopath <- function(inpath,infile) {
@@ -178,8 +182,6 @@ getStr <- function(inline,nb) {
   return(outconst)
 } # end of getStr
 
-
-
 #' @title pathend determines what character is at the end of a path
 #'
 #' @description pathend determines what character is at the end of a
@@ -225,5 +227,33 @@ pathtype <- function(inpath) {
   return(typepath)
 } # end of pathtype
 
+#' @title setupdirs sets up and checks the directories used in a run
+#'
+#' @description setupdirs sets up and checks the directories used in
+#'     a run, if they do not exist to start with they will be created.
+#'     Note the use of the <<- funciton, which assigns the directory
+#'     names in teh global environment rather than only within the
[…]

### 2020-04-14 R/defineZone.R

@@ -351,6 +351,47 @@ fillzoneDef <- function(regC,regD,prod) {  # inzone=zone; prod=production
    return(zDef)
 }  # end of fillzoneDef
 
+#' @title findF1 approximates the F0.1 harvest rate for each population
+#'
+#' @description findF1 generates estimates of the annual harvest rate
+#'     that would be equivalent to the F0.1 for each population. It
+#'     does this by calculating the gradient of the production curve
+#'     for the selected pop number and selecting the best approximation
+#'     by searching for the gradient that is cloest to 0.1.
+#'
+#' @param produyct the 3D array from doproduction
+#' @param location return the index (default=TRUE) or the gradient
+#'     vector if location=FALSE
+#'
+#' @return either the index within the production array or the vector
+#'     of gradients
+#' @export
+#'
+#' @examples
+#' data(product) #for pop=1, a 28percent drop from Hmsy leads to a
+#' findF1(product=product) # loss of 3 tonnes, 4 percent of MSY
+#' findmsy(product)  # compare the AnnH, Deplet, and RelCE levels.
+findF1 <- function(product) {
+  npop <- dim(product)[3]
+  label <- c(colnames(product),"index")
+  xval <- matrix(0,nrow=npop,ncol=length(label),
+                 dimnames=list(1:npop,label))
+  for (pop in 1:npop) {
+    harv <- product[,"AnnH",pop]
+    nH <- length(harv)
+    catch <- product[,"Catch",pop]
+    grad <- numeric(nH-1)
+    for (i in 1:(nH-1)) {
+      divisor <- harv[i+1] - harv[i]
+      numerator <- catch[i+1] - catch[i]
+      grad[i] <- numerator/divisor
+    }
+    pickF1 <-  which.closest(0.1,grad/grad[1])
+    xval[pop,] <- c(product[pickF1,,pop],pickF1)
+  }
+  return(xval)
+} # end of findF1
+
 #' @title findmsy identifies the closest productivity value to MSY
 #'
 #' @description findmsy for each population in the region, identifies

### 2020-04-14 R/makehtml_funs.R

@@ -1,0 +1,203 @@
+
+
+add_css <- function(htmlfile) {
+    cat('<html><head><title>', 'aMSEout', '</title>\n',
+      '    <!-- source for text below is http://unraveled.com/publications/css_tabs/ -->\n',
+      '    <!-- CSS Tabs is licensed under Creative Commons Attribution 3.0 - http://creativecommons.org/licenses/by/3.0/ -->\n',
+      '    <!-- When visiting unraveled.com/publications/css_tabs it appeared to be a toxic website - BE CAREFUL -->\n',
+      '    \n',
+      '    <style type="text/css">\n',
+      '    \n',
+      '    body {\n',
+      '      font: 100% verdana, arial, sans-serif;\n',
+      '      background-color: #fff;\n',
+      '      margin: 50px;\n',
+      '    }\n',
+      '    \n',
+
+      '    /* begin css tabs */\n',
+      '    \n',
+      '    ul#tabnav { /* general settings */\n',
+      '      text-align: left; /* set to left, right or center */\n',
+      '      margin: 1em 0 1em 0; /* set margins as desired */\n',
+      '      font: bold 11px verdana, arial, sans-serif; /* set font as desired */\n',
+      '      border-bottom: 1px solid #6c6; /* set border COLOR as desired */\n',
+      '      list-style-type: none;\n',
+      '      padding: 3px 10px 2px 10px; /* THIRD number must change with respect to padding-top (X) below */\n',
+      '    }\n',
+      '    \n',
+      '    ul#tabnav li { /* do not change */\n',
+      '      display: inline;\n',
+      '    }\n',
+      '    \n',
+      '    body#tab1 li.tab1, body#tab2 li.tab2, body#tab3 li.tab3, body#tab4 li.tab4 { /* settings for selected tab */\n',
+      '      border-bottom: 1px solid #fff; /* set border color to page background color */\n',
+      '      background-color: #fff; /* set background color to match above border color */\n',
+      '    }\n',
+      '    \n',
+      '    body#tab1 li.tab1 a, body#tab2 li.tab2 a, body#tab3 li.tab3 a, body#tab4 li.tab4 a { /* settings for selected tab link */\n',
+      '      background-color: #fff; /* set selected tab background color as desired */\n',
+      '      color: #000; /* set selected tab link color as desired */\n',
+      '      position: relative;\n',
+      '      top: 1px;\n',
[…]

### 2020-04-14 data-raw/sourcer.R

@@ -2,36 +2,23 @@
 
 
 
-
 #' Title
 #'
-#' @param pop
-#' @param res
-#' @param location
+#' @param plotdir
+#' @param tabledir
+#' @param runname
 #'
 #' @return
 #' @export
 #'
 #' @examples
-findF1 <- function(pop,res,location=TRUE) {
-  harv <- results[,"AnnH",pop]
-  catch <- results[,"Catch",pop]
-  grad <- numeric(nH-1)
-  for (i in 1:(nH-1)) {
-    divisor <- harv[i+1] - harv[i]
-    numerator <- catch[i+1] - catch[i]
-    grad[i] <- numerator/divisor
-  }
-  pickF1 <-  which.closest(0.1,grad/grad[1])
-  if (location) {
-    return(pickF1)
-  } else {
-    return(grad)
-  }
-}
-
-
-
+setuphtml <- function(pldir, tabdir,rname) {
+  plottabfile <<- filenametopath(pldir,paste0("plotFileTable_",rname,".csv"))
+  tabletabfile <<- filenametopath(tabdir,paste0("TableFileTable_",rname,".csv"))
+  label <- c("file","caption","category","TimeMade")
+  cat(label,"\n",file = plottabfile,sep=",")
+  cat(label,"\n",file = tabletabfile,sep=",")
+} # end of setuphtml
 
 
 

### 2020-04-14 data-raw/test_aMSE.R

@@ -1,37 +1,0 @@
-
-
-library(rutilsMH)
-library(aMSE)
-library(microbenchmark)
-
-# read data files ----------------------------------------------------
-rundir <- "./../../rcode2/aMSEuse/run2"
-datadir <- file.path(rundir,"data")
-ctrlfile <- "control.csv"
-ctrl <- readctrlfile(datadir,ctrlfile)
-reg1 <- readregionfile(datadir,ctrl$regionfile)
-glb <- reg1$globals
-constants <- readdatafile(datadir,ctrl$datafile,glb)
-
-# Define the Zone ----------------------------------------------------
-ans <- makeregionC(reg1,constants)
-regionC <- ans$regionC
-popdefs <- ans$popdefs
-ans <- makeregion(glb,regionC)
-
-regionC <- ans$regionC  # region constants
-regionD <- ans$regionD  # region dynamics
-
-ans2 <- findunfished(regionC,regionD,glb)
-str(ans2,max.level=2)
-
-regionC <- ans2$regionC  # region constants
-regionD <- ans2$regionD  # region dynamics
-product <- ans2$product
-
-sapply(regionC,"[[","MSY")
-sapply(regionC,"[[","SaM")
-str(regionD)
-
-regionD$exploitB[1,]
-round(regionD$Nt[,1,])

### 2020-04-14 README.Rmd
Started adjusting the make_html function and generating some figures to start experiments with output.

@@
 ```
+
+Alternatively, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio.
+Open the New Project option in the project dialog at the top right of the RStudio screen and selection Version Control, then use 'https://github.com/haddonm/aMSE' in the top box, identify where you want the new directory put, and press return.
+
+It would be a good idea to read Hadley Wickham's draft chpater on Git and GitHub at https://r-pkgs.org/index.html.
+
 ## Example

### 2020-04-26 data-raw/devel.R
Began Running_aMSE.Rmd, which led to changes to template and data read files.

@@ -56,6 +56,11 @@ source(filenametopath(sourcedir,"plotproductivity_source.R"))
 source(filenametopath(sourcedir,"plotbiology_source.R"))
 # end characterize biology -------------------------------------------
 
+if (ctrl$initdepl < 1.0) {
+  regionD <- dodepletion(regionC, regionD, glb, product)
+}
+
+
  endtime <- as.character(Sys.time())
 
 

### 2020-04-26 data-raw/plotbiology_source.R

@@ -1,10 +1,10 @@
-
+# These plots use both regionC and regionD
 # characterize productivity ------------------------------------------
 # plot of Yield vs Spawning biomass
 mids <- glb$midpts
 numpop <- glb$numpop
 
-# maturation ---------------------------------------------------------
+# maturation uses regionC---------------------------------------------
 maturity <- getlistvar(regionC,"Maturity")
 rownames(maturity) <- mids
 
@@ -22,7 +22,7 @@ if (nchar(filename) > 0) dev.off()
 caption <- "The maturity vs length for each population."
 addfilename(filename,tabfile=plottabfile,"Biology","plot",caption)
 
-# weight-at-length ---------------------------------------------------
+# weight-at-length uses regionC---------------------------------------
 WtL <- getlistvar(regionC,"WtL")
 rownames(WtL) <- mids
 
@@ -40,7 +40,7 @@ if (nchar(filename) > 0) dev.off()
 caption <- "The weight-at-length for each population. The x-axis is constrained to encompass legal sizes."
 addfilename(filename,tabfile=plottabfile,"Biology","plot",caption)
 
-# emergence  ---------------------------------------------------------
+# emergence uses regionC----------------------------------------------
 emerg <- getlistvar(regionC,"Emergent")
 rownames(emerg) <- mids
 
@@ -58,7 +58,7 @@ if (nchar(filename) > 0) dev.off()
 caption <- "The emergence-at-length for each population. The x-axis is constrained to emphasize differences."
 addfilename(filename,tabfile=plottabfile,"Biology","plot",caption)
 
-# initial numbers-at-size --------------------------------------------
+# initial numbers-at-size uses regionD--------------------------------
 
 mids <- glb$midpts
 numpop <- glb$numpop
@@ -86,7 +86,7 @@ caption <- "The numbers-at-size for the whole region and for each population sep
 addfilename(filename,tabfile=plottabfile,"Biology","plot",caption)
 
 
-# Tabulate biological properties -------------------------------------
+# Tabulate biological properties uses regionC-------------------------
 
 getvar <- function(indexvar,regC=regionC) {
         invar <- getlistvar(regC,indexvar)

### 2020-04-26 data-raw/plotproductivity_source.R

@@ -1,4 +1,4 @@
-
+# All these plots only use the product array
 # characterize productivity ------------------------------------------
 # plot of Yield vs Spawning biomass
 xval <- findmsy(product)

### 2020-04-28 R/getfunctions.R
Now can do an initial depletion and report on population and regional properties as needed. No Errors, Warnings, or Notes.

@@ -158,46 +158,69 @@ getunFished <- function(inzone,glb) {  # inzone=zone
   return(nofished)
 }  # End of getUnfished
 
-#' @title getzoneProps extracts the depletion level for a given year
+#' @title getregionprops extracts the depletion level for a given year
 #'
-#' @description getzoneProps extracts the depletion level along with an
-#'     array of other statistics for a given year.
+#' @description getregionprops extracts a set of properties for each
+#'     population and summarizes them for the region as well. These
+#'     properties include effB0, matureB, legalmatB, propprot, MSY,
+#'     effexB0, exploitB, SpBDepl, ExBDepl, legalDepl, MSYDepl, LML,
+#'     and bLML. See the model documentation for the meaning of each
+#'     of these.
 #'
 #' @param regC the constants for the region being simulated
+#' @param regD the dynamic part of the region being simulated
 #' @param glb the global constants
-#' @param year which years information is wanted
+#' @param year which years information is wanted, default=1
 #'
-#' @return a list of size objects
+#' @return a matrix of population properties with a column of totals
 #' @export
 #'
 #' @examples
-#' \dontrun{
-#'  print("wait on internal data sets")
-#' }
-getzoneProps <- function(regC,glb,year=1) { # inzone=zone; glb=glb; year=1
+#' data(testregC)
+#' data(testregD)
+#' data(region1)
+#' glb <- region1$globals
+#' round(getregionprops(testregC,testregD,glb),4)
+getregionprops <- function(regC,regD,glb,year=1) { # regC=regionC; regD=regionD; glb=glb; year=1
   numpop <- glb$numpop
   Nclass <- glb$Nclass
-  totB0 <- sum(sapply(regC,"[[","B0"))
-  ExB0 <- sum(sapply(regC,"[[","ExB0"))
-  matB <- 0.0
-  ExB <- 0.0
-  legalMatB <- 0.0
-  ce <- 0.0
-  for (pop in 1:numpop) {
-    ExB <- ExB + regC[[pop]]$ExploitB[year]
-    matB <- matB + regC[[pop]]$MatureB[year]
[…]

### 2020-04-28 data-raw/sourcer.R

@@ -1,1 +1,17 @@
 
+
+
+
+
+#
+#
+# round(getregionprops(regionC,regionD,glb),4)
+#
+# round(getregionprops(regionC,regionDD,glb),4)
+#
+#
+#
+
+
+
+

### 2020-04-29 README.Rmd
Changed the idea of a plots to a results subdirectory within the rundir. Also simplified the plotting of the production graphs by altering plotprod. No errors, no warnings, no notes.

@@
 These are the active components within the simulation framework.
-The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighboiuring populations on each other.
+The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.
 
 This conceptual structure has implications for how the simulations are conditioned during the management strategy testing.
-While it is still the case tha tthe general properties of each SMU are used to set the scene for each SMU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population.
+While it is still the case that the general properties of each SMU are used to set the scene for each SMU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population.
 Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SMU.
@@
 ```
+
+### 29/04/2020
+
+Now developing a vignette entitled _Running_aMSE_, which contains some worked examples to get a user moving in the uise of __aMSE__. There remains much development still to do.

### 2020-04-29 R/inputfiles.R

@@ -336,7 +336,7 @@ readctrlfile <- function(indir,infile="control.csv") {
 #' popdefs <- readdatafile(reg1$globals,datadir,ctrl$datafile)
 #' print(popdefs)
 #' }
-readdatafile <- function(numpop,indir,infile) {  # indir=yourdir;infile="constdat.csv";numpop=6
+readdatafile <- function(numpop,indir,infile) {  # indir=datadir;infile="reg1smu2pop6.csv";numpop=6
    filename <- filenametopath(indir,infile)
    indat <- readLines(filename)   # reads the whole file as character strings
    begin <- grep("PDFs",indat)

### 2020-04-29 R/makehtml_funs.R

@@ -8,12 +8,12 @@
 #'     results. The origin came from Ian Taylor although I have
 #'     modified the styles and ensure correct Version 3 CSS
 #'
-#' @param plotdir the directory within the run directory that contains
+#' @param resdir the directory within the run directory that contains
 #'     all the plot results
 #'
-#' @return nothing but it does generate a .css file in the plotdir
-write_css <- function(plotdir) {
-  filename <- filenametopath(plotdir,"aMSEout.css")
+#' @return nothing but it does generate a .css file in the resdir
+write_css <- function(resdir) {
+  filename <- filenametopath(resdir,"aMSEout.css")
   cat('    \n',
       '    body {\n',
       '      font-size:  18px; \n',
@@ -86,7 +86,6 @@ write_css <- function(plotdir) {
       '      font-size: 15px; \n',
       '      border: 1px solid black;\n',
       '      border-collapse: collapse; \n',
-      '      width: 100%; \n',
       '    } \n',
       '    th, td {\n',
       '      padding: 5px; \n',
@@ -116,6 +115,7 @@ write_head <- function(htmlfile) {
       '<html> \n',
       '  <head>',
       '    <meta charset="utf-8"> \n',
+      '    <meta name="format-detection" content="telephone=no"/> \n',
       '    <title>', 'aMSEout', '</title>\n',
       '    <!-- source for text below is http://unraveled.com/publications/css_tabs/ -->\n',
       '    <!-- CSS Tabs is licensed under Creative Commons Attribution 3.0 - http://creativecommons.org/licenses/by/3.0/ -->\n',
@@ -152,21 +152,20 @@ htmltable <- function(inmat,filename,caption) {
   cat('<br><br> \n',file=filename,append=TRUE)
   cat('<table> \n',file=filename,append=TRUE)
   cat('<caption> ',caption,'</caption> \n',file=filename,append=TRUE)
+  cat('<br> \n',file=filename,append=TRUE)
   cat('<tr> \n',file=filename,append=TRUE)
-  cat('<th>Var</th> \n',file=filename,append=TRUE)
-  tmp <- NULL
-  for (cl in 1:numcol) tmp <- paste(tmp,paste('<th>',columns[cl],'</th>',collapse=""),collapse="")
-  cat(tmp,file=filename,append=TRUE)
+  cat(' <th>Var</th> \n',file=filename,append=TRUE)
+  for (cl in 1:numcol)
+     cat(' <th>',columns[cl],'</th> \n',file=filename,append=TRUE)
[…]

### 2020-05-01 data-raw/plotproductivity_source.R
Continued development of Running_aMSE.Rnd, which emtailed improving the logfilename and making a getextension funciton. No Errors, Warnings, or Notes.

@@ -11,7 +11,7 @@ plotprod(product,xname="MatB",xlab="Spawning Biomass t",
          ylab="Production t",filename = filename)
 
 caption <- "The production curve relative to each population's spawning biomass. The vertical lines identify the Bmsy values."
-addfilename(filename,resfile=resfile,"Production","plot",caption)
+logfilename(filename,resfile=resfile,"Production",caption)
 
 # plot of Yield vs Annual Harvest Rate
 file <- paste0("production_AnnH_",ctrl$runlabel,".png")
@@ -19,7 +19,7 @@ filename <- filenametopath(resdir,file)  #  filename=""
 plotprod(product,xname="AnnH",xlab="Annual Harvest Rate",filename = filename)
 
 caption <- "The production curve relative to the Annual Harvest Rate applied to each population. The vertical lines identify the Hmsy values."
-addfilename(filename,resfile=resfile,"Production","plot",caption)
+logfilename(filename,resfile=resfile,"Production",caption)
 
 # plot of Yield vs population depletion
 file <- paste0("production_Deplet_",ctrl$runlabel,".png")
@@ -33,7 +33,7 @@ for (pop in 1:numpop) abline(v=xval[pop,"Deplet"],lwd=2,col=pop)
 if (nchar(filename) > 0) dev.off()
 
 caption <- "The production curve relative to the depletion level of each population. The vertical lines identify the Depletion level giving rise to the MSY."
-addfilename(filename,resfile=resfile,"Production","plot",caption)
+logfilename(filename,resfile=resfile,"Production",caption)
 
 
 # plot of Yield vs population depletion but constrained to within
@@ -50,7 +50,7 @@ for (pop in 1:numpop) abline(v=xval[pop,"Deplet"],lwd=2,col=pop)
 if (nchar(filename) > 0) dev.off()
 
 caption <- "The production curve relative to the depletion level of each population. Here the x-axis is shortened to clarify the flatness of the production curve about the MSY points."
-addfilename(filename,resfile=resfile,"Production","plot",caption)
+logfilename(filename,resfile=resfile,"Production",caption)
 
 # Now do total production --------------------------------------------
 yield <- rowSums(product[,"Catch",])
@@ -99,7 +99,7 @@ abline(v=Ht[pickmsy],col=2,lwd=2)
 if (nchar(filename) > 0) dev.off()
 
 caption <- "The production curves for the region. Also the relationships between spawning biomass depletion and harvest rate."
[…]

### 2020-05-11 README.Rmd
Cleaned up makehtml and extended running_aMSE.Rmd. No Errors, Warnings, or Notes
Higher level functions, such as setupregion now simplify the running. No Errors, Warnings, or Notes

@@
 ```{r example}
-# basic example code
+# a minimal example
+starttime <- as.character(Sys.time())
 library(aMSE)
-library(rutilsMH)
+# Obviously you should modify the rundir to suit your own computer
+rundir <- "./../../rcode2/aMSEUse/run2"
+outdir <- setupdirs(rundir)
+datadir <- outdir$datadir # this usually has the data files 
+resdir <- outdir$resdir   # this holds all results
 
+data(ctrl)
+data(constants)
 data(region1)
 glb <- region1$globals
-  data(testregC)
-  data(testregD)
-  regDe <- testequil(regC=testregC,regD=testregD,glob=glb)
-  str(regDe)
+runname <- ctrl$runlabel
+
+out <- setupregion(constants, glb, region1)
+regionC <- out$regionC
+regionD <- out$regionD
+product <- out$product  # important bits usually saved in resdir
+# did the larval dispersal level disturb the equilibrium?
+regDe <- testequil(regionC,regionD,glb)
+
+resfile <- setuphtml(resdir,runname)  # prepare to log results
+plotproductivity(resdir,runname,product,glb)
+biology_plots(resdir, runname, glb, regionC, regionD, product)
+numbersatsize(resdir, runname, glb, regionD)
+
+endtime <- as.character(Sys.time())
+
+reportlist <- list(runname=runname,
+  starttime=starttime,endtime=endtime,
+  regionC=regionC,
+  regionD=regionD,
+  product=product,
+  glb=glb,constants=constants
+)
+
+runnotes <- "This is a bare-bones example."
+# If you unhash this last function call it will generate and open a 
+# local website inside resdir so you can see the results so far.
+
+# make_html(replist=reportlist,rundir=rundir,width=500,
+#           openfile=TRUE,runnotes=runnotes,verbose=FALSE)
 
@@
 
-### 29/04/2020
+See the vignette Running_aMSE.Rmd for a more detailed example.
 
-Now developing a vignette entitled _Running_aMSE_, which contains some worked examples to get a user moving in the uise of __aMSE__. There remains much development still to do.
+See the New.Rd for recent developments

### 2020-05-11 data-raw/plotbiology_source.R

@@ -1,165 +1,0 @@
-# These plots use both regionC and regionD
-# characterize productivity ------------------------------------------
-# plot of Yield vs Spawning biomass
-mids <- glb$midpts
-numpop <- glb$numpop
-
-# maturation uses regionC---------------------------------------------
-maturity <- getlistvar(regionC,"Maturity")
-rownames(maturity) <- mids
-
-file <- paste0("maturity_v_Length_",ctrl$runlabel,".png")
-filename <- filenametopath(resdir,file)  #  filename=""
-plotprep(width=7,height=4,newdev=FALSE,filename=filename,cex=0.9,
-         verbose=FALSE)
-plot(mids,maturity[,1],type="l",lwd=2,xlab="Shell Length mm",
-     ylab="Proportion Mature",panel.first=grid(),xlim=c(50,210))
-for (pop in 2:numpop) lines(mids,maturity[,pop],lwd=2,col=pop)
-legend("topright",paste0("P",1:numpop),lwd=3,col=c(1:numpop),bty="n",
-       cex=1.2)
-if (nchar(filename) > 0) dev.off()
-
-caption <- "The maturity vs length for each population."
-logfilename(filename,resfile=resfile,"Biology",caption)
-
-# weight-at-length uses regionC---------------------------------------
-WtL <- getlistvar(regionC,"WtL")
-rownames(WtL) <- mids
-
-file <- paste0("Weight_at_Length_",ctrl$runlabel,".png")
-filename <- filenametopath(resdir,file)  #  filename=""
-plotprep(width=7,height=4,newdev=FALSE,filename=filename,cex=0.9,
-         verbose=FALSE)
-plot(mids,WtL[,1],type="l",lwd=2,xlab="Shell Length mm",
-     ylab="Weight Kg",panel.first=grid(),xlim=c(110,210))
-for (pop in 2:numpop) lines(mids,WtL[,pop],lwd=2,col=pop)
-legend("topleft",paste0("P",1:numpop),lwd=3,col=c(1:numpop),bty="n",
-       cex=1.2)
-if (nchar(filename) > 0) dev.off()
-
-caption <- "The weight-at-length for each population. The x-axis is constrained to encompass legal sizes."
-logfilename(filename,resfile=resfile,"Biology",caption)
-
-# emergence uses regionC----------------------------------------------
[…]

### 2020-05-12 README.Rmd
Simplified the output directory structure and cleaned up all related files and vignettes. No Errors, Warnings, Notes (No EWN)

--- 2020-04-29 README.Rmd
@@
 
-Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examin its implications.
+Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.
 
@@
 
-``` r
-# install.packages("devtools")
-devtools::install_github("haddonm/aMSE")
-devtools::install_github(haddonm/rutilsMH) # also useful for plotting, etc
+```{r echo=TRUE, eval=FALSE}
+if (!require(devtools)){install.packages("devtools")}
+
+devtools::install_github("https://github.com/haddonm/aMSE",build_vignettes = TRUE)
 ```
@@
 ```{r example}
-# basic example code
+# a minimal example
+starttime <- as.character(Sys.time())
 library(aMSE)
-library(rutilsMH)
+# Obviously you should modify the resdir to suit your own computer
+resdir <- "./../../rcode2/aMSEUse/out/testrun"
+dirExists(resdir,make=TRUE,verbose=TRUE)
+# You now need to ensure that there is a control.csv, reg1smu2pop6.csv
+# and region1.csv file in the data directory
+ctrlfiletemplate(resdir)  # puts a template ctrlfile into resdir
+regionfiletemplate(resdir)  # puts a template region file into resdir
+datafiletemplate(6,resdir,filename="reg1smu2pop6.csv") # etc
+ctrl <- checkresdir(resdir) # checks the data files are present
+runname <- ctrl$runlabel    # and reads the control file, if present
+region1 <- readregionfile(resdir,ctrl$regionfile)
+glb <- region1$globals
+constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
 
-  data(region1)
-  glb <- region1$globals
-  data(testregC)
-  data(testregD)
-  regDe <- testequil(regC=testregC,regD=testregD,glob=glb)
-  str(regDe)
+out <- setupregion(constants, glb, region1) # make operating model 
+regionC <- out$regionC     # constant bits
+regionD <- out$regionD     # dynamics bits
+product <- out$product     # all usually saved in resdir, not here
+  # did the larval dispersal level disturb the equilibrium?
+regDe <- testequil(regionC,regionD,glb)
+
+resfile <- setuphtml(resdir,runname)# prepare to save and log results
+
+plotproductivity(resdir,runname,product,glb)
+biology_plots(resdir, runname, glb, regionC)
+numbersatsize(resdir, runname, glb, regionD)
+
+endtime <- as.character(Sys.time())
+
+reportlist <- list(runname=runname,
+                   starttime=starttime,endtime=endtime,
+                   regionC=regionC, regionD=regionD, product=product,
+                   glb=glb,constants=constants
+)
+
+runnotes <- "This is a bare-bones example."
+# If you unhash the make_html component it will generate a local 
+# website inside resdir and open it so you can see the results so far.
+#
+# make_html(replist=reportlist,resdir=resdir,width=500,
+#          openfile=TRUE,runnotes=runnotes,verbose=FALSE)
 

### 2020-05-12 data-raw/bare-bones.r

@@ -3,45 +3,44 @@
 # a minimal example
 starttime <- as.character(Sys.time())
 library(aMSE)
-# Obviously you should modify the rundir to suit your own computer
-rundir <- "./../../rcode2/aMSEUse/run2"
-outdir <- setupdirs(rundir)
-datadir <- outdir$datadir
-resdir <- outdir$resdir
-
-data(ctrl)
-data(constants)
-data(region1)
-glb <- region1$globals
+# Obviously you should modify the resdir to suit your own computer
+resdir <- "./../../rcode2/aMSEUse/out/testrun"
+dirExists(resdir,make=TRUE,verbose=TRUE)
+# You now need to ensure that there is a control.csv, reg1smu2pop6.csv
+# and region1.csv file in the data directory
+ctrlfiletemplate(resdir)
+regionfiletemplate(resdir)
+datafiletemplate(6,resdir,filename="reg1smu2pop6.csv")
+ctrl <- checkresdir(resdir)
 runname <- ctrl$runlabel
+region1 <- readregionfile(resdir,ctrl$regionfile)
+glb <- region1$globals
+constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
 
-out <- setupregion(constants, glb, region1)
+out <- setupregion(constants, glb, region1) # make operating model
 regionC <- out$regionC
 regionD <- out$regionD
-product <- out$product
-
+product <- out$product     # important bits usually saved in resdir
+          # did the larval dispersal level disturb the equilibrium?
 regDe <- testequil(regionC,regionD,glb)
 
-resfile <- setuphtml(resdir,runname)
+resfile <- setuphtml(resdir,runname)# prepare to save and log results
 
 plotproductivity(resdir,runname,product,glb)
-biology_plots(resdir, runname, glb, regionC, regionD, product)
+biology_plots(resdir, runname, glb, regionC)
 numbersatsize(resdir, runname, glb, regionD)
 
 endtime <- as.character(Sys.time())
 
-
 reportlist <- list(runname=runname,
-  starttime=starttime,endtime=endtime,
-  regionC=regionC,
-  regionD=regionD,
-  product=product,
-  glb=glb,constants=constants
+                   starttime=starttime,endtime=endtime,
+                   regionC=regionC, regionD=regionD, product=product,
+                   glb=glb,constants=constants
 )
 
 runnotes <- "This is a bare-bones example."
 # If you unhash this component it will generate a local website inside
[…]

### 2020-05-12 data-raw/regiontable.R

@@ -1,28 +1,0 @@
-
-
-filename <- filenametopath(resdir,"regiontable.html")
-
-
-
-capt <- "Table Regional Population Biological Properties."
-filename <- filenametopath(resdir,"regiontable.html")
-htmltable(results,filename,capt)
-
-
-
-cat('<caption> </caption> \n',file=filename,append=TRUE)
-cat('<tr> \n',file=filename,append=TRUE)
-for (cl in 1:numcol) {
-  cat('<th>',columns[cl],'</th> \n',file=filename,append=TRUE)
-}
-for (rw in 1:numrow) {
-  cat('<tr> \n',file=filename,append=TRUE)
-  for (cl in 1:numcol) { cat('<th>',res[rw,cl],'</th> \n',file=filename,append=TRUE) }
-  cat('</tr> \n',file=filename,append=TRUE)
-}
-cat('</table>',file=filename,append=TRUE)
-
-
-tmp <- NULL
-for (cl in 1:numcol) tmp <- paste(tmp,paste('<th>',columns[cl],'</th>',collapse=""),collapse="")
-tmp

### 2020-05-15 data-raw/bare-bones.r
Began developing code relating to SMUs and fleet dynamics. Replaced all data files. NoEWN.

@@ -41,6 +41,6 @@ reportlist <- list(runname=runname,
 runnotes <- "This is a bare-bones example."
 # If you unhash this component it will generate a local website inside
 # resdir and open it so you can see the results so far.
-# make_html(replist=reportlist,resdir=resdir,width=500,
-#          openfile=TRUE,runnotes=runnotes,verbose=FALSE)
+ make_html(replist=reportlist,resdir=resdir,width=500,
+          openfile=TRUE,runnotes=runnotes,verbose=FALSE)
 

### 2020-05-17 data-raw/development.Rmd
Big changes involving the larval dispersal dynamics, which are now integrated into the dynamics more intuitively. Other functions and examples all changed to match. No Errors, or Warnings, but a note about example files it creates being found?

@@
 * output directory
+
+## The Dynamics Described
+
+### The Initial Equilibrium
+
+The unfished equilibrium is a classical starting point for any model, even if a decision is made to begin any simulations at some lower depletion level.
+We are using a one-year time-step but the dynamics occurs within each year.
+The dynamic part of the region has the following structure:
+
+regionD - the dynamic part of hte regional structure
+
+* SMU     : num [1:6] 1 1 2 2 2 2
+* matureB : num [1:40, 1:6] 782 0 0 0 0 ...
+* exploitB: num [1:40, 1:6] 892 0 0 0 0 ...
+* catch   : num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
+* harvestR: num [1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
+* cpue    : num [1:40, 1:6] 399 0 0 0 0 ...
+* recruit : num [1:40, 1:6] 630563 0 0 0 0 ...
+* deplsB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
+* depleB  : num [1:40, 1:6] 1 0 0 0 0 0 0 0 0 0 ...
+* catchN  : num [1:105, 1:40, 1:6] 0 0 0 0 0 0 0 0 0 0 ...
+* Nt      : num [1:105, 1:40, 1:6] 6.31e+05 2.34e-06 4.31e-05 ...
+
+This represents the dynamics before any simulations have occurred, the unfished equilibrium state.
+Similarly, the constant part of the region, made up of a list for each population, has the following structure:
+
+* Me      : num 0.198
+* R0      : num 636147
+* B0      : num 782
+* effB0   : num 782
+* ExB0    : num 892
+* effExB0 : num 892
+* MSY     : num 54.3
+* MSYDepl : num 0.304
+* bLML    : num 153
+* popq    : num 0.000447
+* SaM     : num 139
+* popdef  : Named num [1:19] 34.45 122.75 164.92 4.61 -16 ....
+* LML     : num [1:40] 132 132 132 132 132 132 132 132 132 132 ...
+* G       : 'STM' num [1:105, 1:105] 2.12e-13 4.52e-12 8.33e-11 1.27e-09 1.62e-08 ...
+* Maturity: num [1:105] 1.42e-07 1.78e-07 2.24e-07 2.83e-07 3.56e-07 ...
+* WtL     : num [1:105] 0.000351 0.003438 0.013072 0.033719 0.070319 ...
+* Emergent: num [1:105] 1.92e-16 3.44e-16 6.18e-16 1.11e-15 1.99e-15 ...
+* Select  : num [1:105, 1:40] 2.87e-127 2.61e-125 2.37e-123 2.16e-121 1.96e-119 ...
+* SelWt   : num [1:105, 1:40] 1.01e-130 8.97e-128 3.10e-125 7.28e-123 1.38e-120 ...
+* MatWt   : num [1:105] 4.97e-11 6.13e-10 2.93e-09 9.53e-09 2.50e-08 ...
+* SMU     : num 1

### 2020-05-17 R/datasets.r

@@ -111,12 +111,12 @@ NULL
 #' @description product is the productivity curve matrix from
 #'     doproduction when the example region is generated using the
 #'     inbuilt datasets ctrl, region1, and constants. The slowest
-#'     part of building the whole is to use the findunfished function
-#'     to adjust the regionC and regionD values to account for any
-#'     larval dispersal. To save that time in the examples (to avoid
-#'     time limits of examples should this package ewver go to CRAN),
-#'     then this dataset can be used instead. This is a three
-#'     dimensional array of productivity variables. Catch is the explected
+#'     part of building the whole is to use the modregC function
+#'     to adjust the regionC and generate the production array. To
+#'     save that time in the examples (to avoid time limits on
+#'     examples should this package go to CRAN), then this dataset can
+#'     be used instead. This is a three dimensional array of
+#'     productivity variables.
 #'
 #' @name product
 #'
@@ -233,7 +233,11 @@ NULL
 #'  data(testregD)
 #'  data(region1)
 #'  glb <- region1$globals
+#'  r0 <- getvar(testregC,"R0")
+#'  move <- makemove(glb$numpop,r0,glb$larvdisp)
+#'  glb$move <- move
 #'  ans <- testequil(testregC, testregD, glb)
+#'  str(testregC[[1]])
 NULL
 
 #' @title testregD is a list of 8 matrices and 2 arrays defining the dynamics of a region
@@ -263,5 +267,9 @@ NULL
 #'  data(testregD)
 #'  data(region1)
 #'  glb <- region1$globals
+#'  r0 <- getvar(testregC,"R0")
+#'  move <- makemove(glb$numpop,r0,glb$larvdisp)
+#'  glb$move <- move
 #'  ans <- testequil(testregC, testregD, glb)
+#'  str(testregD)
 NULL

### 2020-05-18 README.Rmd
Fixed up the application of larval dispersal and other details in the dynamics when applying catches rather than harvest rates. No Errors or Warning, same Note about unknown directories in the check directory.

@@
 
-It would be a good idea to read Hadley Wickham's draft chpater on Git and GitHub at https://r-pkgs.org/index.html.
+It would be a good idea to read Hadley Wickham's draft chapter on Git and GitHub at https://r-pkgs.org/index.html.
 
@@
 
-out <- setupregion(constants, glb, region1) # make operating model 
+out <- setupregion(constants, region1) # make operating model 
 regionC <- out$regionC     # constant bits
@@
 product <- out$product     # all usually saved in resdir, not here
-  # did the larval dispersal level disturb the equilibrium?
+glb <- out$glb             # now contains the larval movement matrix
+  # Are we at equilibrium?
 regDe <- testequil(regionC,regionD,glb)

### 2020-05-18 data-raw/development.Rmd
Fixed up the application of larval dispersal and other details in the dynamics when applying catches rather than harvest rates. No Errors or Warning, same Note about unknown directories in the check directory.

@@
 * SMU     : num 1
+
+In the code (_setupregion_) the sequence is:
+
+1.
+first generate regionC from the constants and region1.
+At that point each population structure contains everything except those things relating to productivity: B0, ExB0, MSY, MSYDepl and popq.
+makeregionC also updates glb by including the move matrix.
+2.
+running the _makeregion_ updates B0, ExB0, and popq in regionC and generates the initial equilibrium for regionD.
+As this is the unfished equilibrium the catch, harvestR, and catchN matrices are empty.
+3. The final run of _moderegC_ internally runs _doproductivity_ which allows regionC's MSY and MSYDepl entries to be updated for each population, as well as producing the product matrix.
+
+In the regionD matrices, the numbers-at-size are the numbers at the start of the year, similarly, the matureB is the mature biomass at the start of the year.
+The ExB0 is the exploitable biomass implied by the initial numbers-at-size, which means to calcualte it we need to apply half of natural mortality and growth before multiplyiung by selectivity and weight-at-length.

### 2020-05-25 data-raw/Tas_HS_PMs.Rmd
Added getgrad1, getgrad4, and getscore1. Started Tas_HS_PMs.Rmd.

---
title: "Abalone Performance Measures"
author: "Malcolm Haddon"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  html_document:
    df_print: paged
    fig_caption: yes
    fig_height: 5.5
    fig_witdh: 6.5
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document:
    fig_caption: yes
    fig_height: 5.75
    fig_width: 7.5
    includes:
      in_header: C:\A_Mal\Rcode\header.tex
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
  word_document:
    reference_docx: C:\A_Mal\Rcode\wordreferencestyle.docx
    toc: true
    fig_caption: true
---

<style type="text/css">
  body, td {
     font-size: 15px;
  }
  code.r{
    font-size: 30px;
  }
  pre {
    font-size: 15px
  }
</style>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return '3.'+n}
      } 
  }
});
</script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE)
  options(knitr.kable.NA = "",
          knitr.table.format = "pandoc")

options(tinytex.verbose = TRUE)  # change this to suit

options("show.signif.stars"=FALSE,
        "stringsAsFactors"=FALSE,
        "max.print"=50000,
        "width"=240)

library(aMSE)
library(rutilsMH)
suppressPackageStartupMessages(library(knitr))
library(captioner)

tab_nums <- captioner(prefix = "Table")
fig_nums <- captioner(prefix = "Figure")
pgwid=75     
```

# Introduction

The Tasmanian abalone harvest strategy uses a Multi-Criterion Decision Analysis (MCDA) on fishery data from each statistical block. The MCDA is based upon the weighted scores of three performance measures, each of which relate to a measure of abalone catch-per-unit-effort. This analysis is used as the basis for decisions concerning the aspirational catch levels intended for each statistical block, and the sum of these intended catches across the blocks within each quota zone is used as the recommended biological catch (RBC). This is then translated by the managers into a total allowable catch (TAC) by the managers. This document aims to introduce the three performance measures and illustrate how they may be used in the management strategy evaluation framework (aMSE).

## The Three Performance Measures

The three PMs use different aspects of the commercial fishery cpue to each produce a score, which is combined in a weighted sum which is directly translated into a recommended catch for each statistical block. In all cases below, where reference is made to using cpue, this is referring to standardized commercial cpue. Currently the docket book based cpue is still what is used, and that is limited to fishery data from 1992 to the present (prior to 1992 the cpue data was reported in a different format).

1. Target CPUE - involves a comparison of the current standardized cpue with a previously agreed target cpue, with each statistical block having their own target.

2. Gradient1 - the gradient or slope of the change in cpue from the current year to the previous year, which aims to represent short-term changes in the fishery.

3. Gradient4 - the gradient change in cpue over the last four years (including the current year $t$, $y_t, y_{t-1}, y_{t-2}, y_{t-3}$)

We will use a standardized time-series, 1992 - 2019 from the eastern zone block 13 (data-set _blockE13_ within __aMSE__) to illustrate the use and calculation of each of the performance measures. The column of data of most interest is labelled _cpue_ and is the standardized cpue scaled to the bias-corrected geometric mean cpue across the whole time series.

```{r echo=TRUE}
data("blockE13")
ab <- blockE13
```

`r tab_nums("T1", caption="The blockE13 data-set from aMSE")`

```{r echo=FALSE } 
 #tabulate the data-set being used 
kable(ab,digits=c(0,4,4,4,4,3,3,1)) 
```

### Gradient1

Gradient1 is simply the proportional change in cpue from one year to the next. This implies that with a time-series of 28 years (as in the blockE13 data-set) we can obtain 27 comparisons. Formally, where $CE_{b,t}$ is the cpue in block/SMU $b$ in year $t$, then the Gradient1 for block/SMU $b$ is:

$$Gradient1_b=\left (\frac{CE_{b,t}}{CE_{b,t-1}} - 1 \right )$$

The central idea for the Gradient1 (and Gradient4) performance measure is that the gradient obtained in each year can be converted into a score between 0 and 10, cenrted on a score of 5 for a gradient of 0.0.

In __aMSE__ we have the function _getgrad1_ to calculate the gradient. The function requires the input of a vector of cpue values (a minimum of two values) and it will output a vector of length one less than the input vector. If we input the vector _ab$cpue_ from teh blockE13 data-set we obtain Gradient1 values for 1993 - 2019, which have a range from -0.208 to 0.286, which is obviously asymmetric around the zero value.

```{r echo=TRUE}
grad1 <- getgrad1(ab$cpue)
range(grad1,na.rm=TRUE)
```

Earlier versions of the harvest strategy used a single range of Gradient1 values for all blocks, but this made it ineffective for some blocks, so in assessments since 2017, the observed range of values within each block was extended by 10% either side and each block treated separately. The reference period for setting this range was one year less than the year of the assessment. Thus for 2020 we would use the full time-series from 1992 - 2019. This change, to customize the Gradient1 (and Gradient4) performance measures to each block, was a significant improvement to the harvest strategy as the sensitivity to the Gradient1 and Gradient4 PMs is now the same in each block. However, it does mean that the relationship between the slope value (gradient) and the resulting score is no longer necessarily linear. We can see this with our example data-set and use the function _getscore1_ to get the score for each gradient1 value.

```{r echo=TRUE, plot.width=5.75, plot.height=2.75}
# round the range to get workable numbers
bounds <- round((range(grad1,na.rm=TRUE) * 1.1),2)
low <- seq(bounds[1],0.0,length=6) 
high <- seq(0.0,bounds[2],length=6)
xax <- c(low[1:5],high)  # combine above with below zero
xax2 <- seq(-0.32,0.32,length=11)
wrong <- lm(0:10 ~ xax2)
par(cex=0.9,font.axis=7,font=7,font.lab=7)
plot(xax,0:10,type="n",xaxt="n",xlab="Gradient1 Value",yaxt="n",
     ylab="Gradient1 Score")
axis(side=1,at=xax,label=xax)     # need to define the axes independently
axis(side=2,at=0:10,label=0:10)   
abline(v=xax,lwd=1,col="grey",lty=3)
abline(h=0:10,lwd=1,col="grey",lty=3)
abline(wrong,lty=2)
lines(c(-0.23,0.0),c(0,5),lwd=2)  # add the linear relationships
lines(c(0.0,0.32),c(5,10),lwd=2)
score1 <- getscore1(grad1)  # examine the code using getscore1 without brackets
points(grad1,score1,pch=16,col=2,cex=1.5)
```

`r fig_nums("F1", caption="The Gradient1 scores for the Gradient1 values obtained for the cpue column in the blockE13 data-set from aMSE. The asymmetric response with a break at zero is apparent in comparison to extending the value axis from -0.32 to 0.32, as in the dashed line.")`

There were 16 negative gradients and 11 positive gradients. As these are proportions it should not be unexpected that the negative terms will have a smaller range than the positive. If we were to force a linear relationship on to this data by making the range of Gradient1 values extend from -0.32 to +0.32 this would automatically increase the scores for the negative Gradient1 values and hence reduce any constraining influence the negative values would be having.

### Gradient4

This algorithm needs clarification. Currently, _getgrad4_, as below, works by applying it to the grad1 values, but that implies we are using five years each time.

```{r  echo=TRUE}
grad4 <- getgrad4(grad1,wid=4) # uses five years of data; red in the figure 
grad3 <- getgrad4(grad1,wid=3) # uses four years of data; blue in the figure
range(grad4)
range(grad3)
```

```{r  echo=TRUE, plot.width=5.75, plot.height=2.75}
#plotprep(width=7,height=5,newdev=FALSE)
parset(plots=c(2,1))
plot(ab$year,ab$cpue,type="l",lwd=2,ylim=c(0,100))
plot(0:27,c(NA,grad1),type="l",lwd=2)        # black
lines(0:27,c(NA,NA,NA,NA,grad4),col=2,lwd=2) # red
lines(0:27,c(NA,NA,NA,grad3),col=4,lwd=2)    # blue
abline(h=0.0,col=3)

```

### 2020-05-25 R/getfunctions.R

@@ -92,7 +92,6 @@ getextension <- function(filename) {
 #'    which provides the annual proportional change in CPUE.
 #'
 #' @param vectce vector of cpue for a given spatial scale
-#' @param maxGradient the scaling factor used in a particular zone
 #'
 #' @return a vector of gradient1 for each year, starting from year 2
 #' @export
@@ -109,6 +108,36 @@ getgrad1 <- function(vectce) {
   return(grad1)
 } # end of getgrad1
 
+#' @title getgrad4 applies a linear regression in steps of wid to input
+#'
+#' @description getgrad4 takes an input vector and applies a linear
+#'     regression along the vector in steps of the parameter wid
+#'
+#' @param vectce the inout vector of cpue or proportional changes in
+#'     cpue
+#' @param wid the number of years of the input data to use in each
+#'     regression
+#'
+#' @return a vector of length (wid-1) shorter than the input vector
+#' @export
+#'
+#' @examples
+#' x <- c(0.0169,0.1953,0.1102,0.1511,-0.0403,-0.0247,-0.0255,-0.1089,
+#'        -0.1458,-0.2082,0.0289,-0.0267)
+#' grad4 <- getgrad4(x,wid=4)
+#' grad3 <- getgrad4(x,wid=3)
+#' cbind(c(NA,grad4),grad3)
+getgrad4 <- function(vectce,wid=4) { # vectce=ab$cpue; wid=4
+  nyr <- length(vectce)
+  inc <- wid-1
+  num <- nyr-wid+1
+  x <- 1:wid
+  grad4 <- numeric(num)
+  for (i in 1:num) {
+    grad4[i] <- coef(lm(vectce[i:(i+wid-1)] ~ x))[2]
+  }
+  return(grad4)
+} # end of getgrad4
 
 #' @title getlistvar extracts a vector or matrix from regionC
 #'
@@ -281,7 +310,6 @@ getscore1 <- function(grad1) {
   score1 <- grad1
   pickl0 <- which(grad1 <= 0)
   score1[pickl0] <- grad1[pickl0]*vars[2] + vars[1]
-  points(grad1[pickl0],score1[pickl0],pch=16,cex=1.25,col=2)
   model2 <- lm(5:10 ~ xax[6:11])
   vars2 <- coef(model2)
   pickg0 <- which(grad1 > 0)

### 2020-05-25 data-raw/devel.R

@@ -253,7 +253,9 @@ fish
 class(fish)
 
 
-grad1 <- grad1PM(ab$cpue)
+data("blockE13")
+ab <- blockE13
+grad1 <- getgrad1(ab$cpue)
 range(grad1,na.rm=TRUE)
 bounds <- round((range(grad1,na.rm=TRUE) * 1.1),2)
 low <- seq(bounds[1],0.0,length=6)
@@ -289,9 +291,15 @@ cbind(1993:2019,grad1,score1)
 
 
 
+grad4 <- getgrad4(grad1)
+range(grad4)
 
-
-
+plotprep(width=7,height=5,newdev=FALSE)
+parset(plots=c(2,1))
+plot(ab$year,ab$cpue,type="l",lwd=2,ylim=c(0,100))
+plot(0:27,c(NA,grad1),type="l",lwd=2)
+lines(0:27,c(NA,NA,NA,NA,grad4),col=2,lwd=2)
+abline(h=0.0,col=3)
 
 
 

### 2020-05-26 data-raw/Tas_HS_PMs.Rmd
Correccted getgrad4 and modified getscore1 into getscore, which now works for both grad PMs.

@@
 lines(c(0.0,0.32),c(5,10),lwd=2)
-score1 <- getscore1(grad1)  # examine the code using getscore1 without brackets
+score1 <- getscore(grad1)  # examine the code using getscore1 without brackets
 points(grad1,score1,pch=16,col=2,cex=1.5)
@@
 ```{r  echo=TRUE}
-grad4 <- getgrad4(grad1,wid=4) # uses five years of data; red in the figure 
-grad3 <- getgrad4(grad1,wid=3) # uses four years of data; blue in the figure
+grad4 <- getgrad4(ab$cpue,wid=4) # uses five years of data; red in the figure 
+grad3 <- getgrad4(ab$cpue,wid=3) # uses four years of data; blue in the figure
+range(grad1)
 range(grad4)
@@
 plot(ab$year,ab$cpue,type="l",lwd=2,ylim=c(0,100))
-plot(0:27,c(NA,grad1),type="l",lwd=2)        # black
-lines(0:27,c(NA,NA,NA,NA,grad4),col=2,lwd=2) # red
-lines(0:27,c(NA,NA,NA,grad3),col=4,lwd=2)    # blue
+plot(0:27,c(NA,grad1),type="l",lwd=2,ylim=c(-0.21,0.29))        # black
+lines(0:27,c(NA,NA,NA,grad4),col=2,lwd=2) # red
+lines(0:27,c(NA,NA,grad3),col=4,lwd=2)    # blue
 abline(h=0.0,col=3)
@@
 ```
+
+```{r  echo=TRUE, plot.width=5.75, plot.height=2.75}
+# round the range to get workable numbers
+bounds <- round((range(grad4,na.rm=TRUE) * 1.1),2)
+low <- seq(bounds[1],0.0,length=6) 
+high <- seq(0.0,bounds[2],length=6)
+xax <- c(low[1:5],high)  # combine above with below zero
+xax2 <- seq(-0.31,0.31,length=11)
+wrong <- lm(0:10 ~ xax2)
+par(cex=0.9,font.axis=7,font=7,font.lab=7)
+plot(xax,0:10,type="n",xaxt="n",xlab="Gradient4 Value",yaxt="n",
+     ylab="Gradient4 Score")
+axis(side=1,at=xax,label=xax)     # need to define the axes independently
+axis(side=2,at=0:10,label=0:10)   
+abline(v=xax,lwd=1,col="grey",lty=3)
+abline(h=0:10,lwd=1,col="grey",lty=3)
+abline(wrong,lty=2)
+lines(c(-0.15,0.0),c(0,5),lwd=2)  # add the linear relationships
+lines(c(0.0,0.31),c(5,10),lwd=2)
+score4 <- getscore(grad4)  # examine the code using getscore1 without brackets
+points(grad4,score4,pch=16,col=2,cex=1.5)
+
+```
+
+```{r}
+cbind(score1,c(NA,NA,score4))
+```
+
+```{r  echo=TRUE, plot.width=5.75, plot.height=2.75}
+years <- 1993:2019
+score42 <- c(NA,NA,score4)
+parset()
+plot(years,score1,type="p",pch=16,cex=1.25,panel.first=grid(),
+     ylim=c(0,10),ylab="Score 1 and 4")
+lines(years,score1,lwd=1)
+points(years,score42,type="p",pch=16,cex=1.25,col=2)
+lines(years,score42,lwd=1,col=2)
+abline(h=5.0,col=3)
+legend("topright",c("score1","score4"),col=c(1,2),lwd=3,bty="n",cex=1.5)
+```
+
+```{r}
+
+```

### 2020-05-27 data-raw/Tas_HS_PMs.Rmd
Updated Tas_HS_PMs.Rmd to clarify the Gradient4 analysis and the potential biases in both Grad1 and Grad4

--- 2020-05-25 data-raw/Tas_HS_PMs.Rmd
@@
 
-$$Gradient1_b=\left (\frac{CE_{b,t}}{CE_{b,t-1}} - 1 \right )$$
+$$Grad1_b=\left (\frac{CE_{b,t}}{CE_{b,t-1}} - 1 \right )$$
 
@@
 lines(c(0.0,0.32),c(5,10),lwd=2)
-score1 <- getscore1(grad1)  # examine the code using getscore1 without brackets
+score1 <- getscore(grad1)  # examine the code using getscore1 without brackets
 points(grad1,score1,pch=16,col=2,cex=1.5)
+lines(c(-0.23,0.32),c(0,10),col=1,lty=3) # scores all biased low
 ```
@@
 
-This algorithm needs clarification. Currently, _getgrad4_, as below, works by applying it to the grad1 values, but that implies we are using five years each time.
+Gradient4 calculates the gradient of the proportional changes over four years that occurs in each Spatial Management Unit (though the four year time-frame can/may be varied).
+This can be applied to a time-series of cpue just as with Gradient1, but here the minimum input requirement would be a vector of four cpue values.
+The objective with a longer time-series owuld be to calculate how the gradient changes for each set of four values moving along the series one year at a time.
+Each value divides through the cpue values by the value in the first year of each group of four to give the proportional change through time relative to the first year in each group:
+
+$$p{\Delta}_t =CE_t/CE_1$$
+
+A linear regression is fitted to each group of four years where the independent variable is either the specific years in which the cpue occurred or, more simply, a series of 1, 2, 3, and 4.
+The independent variable only affects the intercept and for the purposes of the performance measure we are only interested in the gradient.
+
+$$Grad4=\frac{\sum{t.p{\Delta}_t}-\sum{t}\sum{p{\Delta}_t}/n}{\sum{t^2}-\left(\sum{t} \right)^2/n}$$
+
+where $t$ is the year (1 to 4), $p{\Delta}_t$ is the proportional change of cpue in year $t$ relative to the first year of the group, and $n$ is, in this case, four.
+This is merely the standard analytic equation for the gradient of a linear regression.
+More simply, in R, we can use `grad4 = coef(lm(pCE ~ 1:4))[2]`, which calculates the complete linear regression and extracts the gradient.
+We can compare the range of values obtained for Gradient1, Gradient4, and even try Gradient3.
 
 ```{r  echo=TRUE}
-grad4 <- getgrad4(grad1,wid=4) # uses five years of data; red in the figure 
-grad3 <- getgrad4(grad1,wid=3) # uses four years of data; blue in the figure
-range(grad4)
-range(grad3)
+grad4 <- getgrad4(ab$cpue,wid=4) # uses five years of data; red in the figure 
+grad3 <- getgrad4(ab$cpue,wid=3) # uses four years of data; blue in the figure
+cat("Grad1 ",range(grad1),"\n")
+cat("Grad4 ",range(grad4),"\n")
+cat("Grad3 ",range(grad3),"\n")
 ```
+
+These different performance measures may take on more meaning by plotting them against the respective cpue values that gave rise to them.
+Of course, Gradient1 will begin one year after the start of the time series while Gradient 4 will start in the fourth year of the series.
 
@@
 parset(plots=c(2,1))
-plot(ab$year,ab$cpue,type="l",lwd=2,ylim=c(0,100))
-plot(0:27,c(NA,grad1),type="l",lwd=2)        # black
-lines(0:27,c(NA,NA,NA,NA,grad4),col=2,lwd=2) # red
-lines(0:27,c(NA,NA,NA,grad3),col=4,lwd=2)    # blue
+plot(ab$year,ab$cpue,type="l",lwd=2,ylim=c(0,100),panel.first=grid(),
+     xlab="",ylab="CPUE kg/hr")
+plot(ab$year,c(NA,grad1),type="l",lwd=2,ylim=c(-0.21,0.29),xlab="",
+     panel.first=grid(),ylab="Gradient Value")                  # black
+lines(ab$year,c(NA,NA,NA,grad4),col=2,lwd=2) # red
+lines(ab$year,c(NA,NA,grad3),col=4,lwd=2)    # blue
 abline(h=0.0,col=3)
+legend("topright",c("Grad1","Grad3","Grad4"),col=c(1,4,2),lwd=3,bty="n")
+```
+
+`r fig_nums("F2", caption="The cpue from block13 in the eastern zone compared with the time-series of values for the Gradient1, Gradient3, and Gradient4 values obtained for the cpue column in the blockE13 data-set from aMSE.")`
+
+Of course, before these performance measure values can be used the values need to be translated into scores between 0 and 10.
+Because the observed range below zero is smaller than the range of observations above zero we need to use two linear relationships to apply the same range of scores to those below as those above zero.
+There are two possible biased outcomes from using only one relationship.
+If we were to extend the lower range out to match the absolute range above zero this would bias the scores of any negative gradients high.
+Alternatively if we were to place a line between the lower limit and the upper limit this would bias the scores of all Gradient4 values low.
+
+```{r  echo=TRUE, plot.width=5.75, plot.height=2.75}
+# round the range after extending by 10 percent to get workable numbers
+bounds <- round((range(grad4,na.rm=TRUE) * 1.1),2) # 1.1  = 10 perc
+low <- seq(bounds[1],0.0,length=6) 
+high <- seq(0.0,bounds[2],length=6)
+xax <- c(low[1:5],high)  # combine above with below zero
+xax2 <- seq(-0.31,0.31,length=11)
+wrong <- lm(0:10 ~ xax2)
+par(cex=0.9,font.axis=7,font=7,font.lab=7)
+plot(xax,0:10,type="n",xaxt="n",xlab="Gradient4 Value",yaxt="n",
+     ylab="Gradient4 Score")
+axis(side=1,at=xax,label=xax)     # need to define the axes independently
+axis(side=2,at=0:10,label=0:10)   
+abline(v=xax,lwd=1,col="grey",lty=3)
+abline(h=0:10,lwd=1,col="grey",lty=3)
+abline(wrong,lty=2)               # add the biased line
+lines(c(-0.15,0.0),c(0,5),lwd=2)  # add the linear relationships
+lines(c(0.0,0.31),c(5,10),lwd=2)
+score4 <- getscore(grad4)  # examine the code using getscore1 without brackets
+points(grad4,score4,pch=16,col=2,cex=1.5)
+lines(c(-0.15,0.31),c(0,10),col=1,lty=3) # scores all biased low
 
 ```
+
+`r fig_nums("F3", caption="The Gradient4 values translated into scores in the same manner as used with the Gradinet1 values.
+The broken stick approach accounts for the differences in the observed range of values below 0.0 and those above.
+The dashed line illustrates the effect of extending the lower range to match the extent of the upper range, which would lead to scores being biased high for negative gradients.
+The dotted lines illustrates how the scores would all be biased low if we merely joined the lower and upper limits of Gradient4 values.")`
+
+```{r}
+cbind(score1,c(NA,NA,score4))
+```
+
+```{r  echo=TRUE, plot.width=5.75, plot.height=2.75}
+years <- 1993:2019
+score42 <- c(NA,NA,score4)
+parset()
+plot(years,score1,type="p",pch=16,cex=1.25,panel.first=grid(),
+     ylim=c(0,10),ylab="Score 1 and 4")
+lines(years,score1,lwd=1)
+points(years,score42,type="p",pch=16,cex=1.25,col=2)
+lines(years,score42,lwd=1,col=2)
+abline(h=5.0,col=3)
+legend("topright",c("score1","score4"),col=c(1,2),lwd=3,bty="n",cex=1.5)
+```
+
+```{r}
+
+```

### 2020-06-10 README.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

@@
 This packages up a new Abalone Management Strategy Evaluation framework.
-It has a novel structure that is based around the spatial ideas of populations within spatial management units (SMUs), within a region:
+It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:
 
-* region - highest geogrpahical level. Is simply the totality of the spatial management units.
+* zone - highest geogrpahical level. Is simply the totality of the spatial management units.
 
-* SMU - spatial management units. In Tasmania these would currently be the classical statistical blocks.
+* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.
 
@@
 This conceptual structure has implications for how the simulations are conditioned during the management strategy testing.
-While it is still the case that the general properties of each SMU are used to set the scene for each SMU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population.
-Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SMU.
+While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population.
+Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU.
 
@@
 This is a basic example which illustrates the generation of an initial equilibrium.
-It uses built in data-sets but usually you would generate regionC (constants) and regionD (the dynamic components) from the conditioning data:
+It uses built in data-sets but usually you would generate zoneC (constants) and zoneD (the dynamic components) from the conditioning data:
 
@@
 library(aMSE)
+library(makehtml)
 # Obviously you should modify the resdir to suit your own computer
@@
 dirExists(resdir,make=TRUE,verbose=TRUE)
-# You now need to ensure that there is a control.csv, reg1smu2pop6.csv
-# and region1.csv file in the data directory
+# You now need to ensure that there is a control.csv, reg1sau2pop6.csv
+# and zone1.csv file in the data directory
 ctrlfiletemplate(resdir)  # puts a template ctrlfile into resdir
-regionfiletemplate(resdir)  # puts a template region file into resdir
-datafiletemplate(6,resdir,filename="reg1smu2pop6.csv") # etc
+zonefiletemplate(resdir)  # puts a template zone file into resdir
+datafiletemplate(6,resdir,filename="zone1sau2pop6.csv") # etc
 ctrl <- checkresdir(resdir) # checks the data files are present
 runname <- ctrl$runlabel    # and reads the control file, if present
-region1 <- readregionfile(resdir,ctrl$regionfile)
-glb <- region1$globals
+zone1 <- readzonefile(resdir,ctrl$zonefile)
+glb <- zone1$globals
 constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
 
-out <- setupregion(constants, region1) # make operating model 
-regionC <- out$regionC     # constant bits
-regionD <- out$regionD     # dynamics bits
+out <- setupzone(constants, zone1) # make operating model 
+zoneC <- out$zoneC     # constant bits
+zoneD <- out$zoneD     # dynamics bits
 product <- out$product     # all usually saved in resdir, not here
@@
   # Are we at equilibrium?
-regDe <- testequil(regionC,regionD,glb)
+regDe <- testequil(zoneC,zoneD,glb)
 
@@
 plotproductivity(resdir,runname,product,glb)
-biology_plots(resdir, runname, glb, regionC)
-numbersatsize(resdir, runname, glb, regionD)
+biology_plots(resdir, runname, glb, zoneC)
+numbersatsize(resdir, runname, glb, zoneD)
 
@@
                    starttime=starttime,endtime=endtime,
-                   regionC=regionC, regionD=regionD, product=product,
+                   zoneC=zoneC, zoneD=zoneD, product=product,
                    glb=glb,constants=constants

### 2020-06-10 data-raw/Tas_HS_PMs.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

@@
 This implies that with a time-series of 28 years (as in the blockE13 data-set) we can obtain 27 comparisons.
-Formally, where $CE_{b,t}$ is the cpue in block/SMU $b$ in year $t$, then the Gradient1 for block/SMU $b$ is:
+Formally, where $CE_{b,t}$ is the cpue in block/SAU $b$ in year $t$, then the Gradient1 for block/SAU $b$ is:
 
@@
 The function requires the input of a vector of cpue values (a minimum of two values) and it will output a vector of length one less than the input vector.
-If we input the vector _ab$cpue_ from teh blockE13 data-set we obtain Gradient1 values for 1993 - 2019, which have a range from -0.208 to 0.286, which is obviously asymmetric around the zero value.
+If we input the vector _ab$cpue_ from the blockE13 data-set we obtain Gradient1 values for 1993 - 2019, which have a range from -0.208 to 0.286, which is obviously asymmetric around the zero value.
 

### 2020-06-10 data-raw/development.Rmd
A large change through most of the package because I have changed the term region to zone, SMU to SAU, and smu to sau. Most of the internal data files have also changed. I also removed the makehtml_funs set of files and imported makehtml instead. No errors or warnings, but that note about extraneous directorys is still there.

--- 2020-05-17 data-raw/development.Rmd
@@
 
-* The region is simply to container for all the SMU (there may only be one).
-The notion is that the variation of population properties within an SMU can differ, but conditioning the populations is a whole subject for discussion in itself.
+* The zone is simply to container for all the SAU (there may only be one).
+The notion is that the variation of population properties within an SAU can differ, but conditioning the populations is a whole subject for discussion in itself.
 
-* Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular region, there is a need to include the option of having larval dispersal.
+* Running the MSE once conditioned should be simple, the difficult part is conditioning the model to represent a particular zone, there is a need to include the option of having larval dispersal.
 
 * Currently using a control.csv file to identify conditioning files, HCR files, and related.
-There is a region file containnig the strucutre of the simulated region, and a datafile containing the conditioning data.
+There is a zone file containnig the strucutre of the simulated zone, and a datafile containing the conditioning data.
 
-* When constructing a region the constant parts are kept in regionC and the dynamic parts in regionD.
+* When constructing a zone the constant parts are kept in zoneC and the dynamic parts in zoneD.
 Is the constant part always separate from the global part?
@@
 
-* Currently using a spatial hierarchy region/SMU/population instead of zone/block/population, so this is now more general than only Tasmanian terminology.
+* Currently using a spatial hierarchy zone/SAU/population instead of zone/block/population, so this is now more general than only Tasmanian terminology.
 Stick to this hierarchy throughout the code.
@@
 
-* Once a particular run is constructed, but before the run proceeds, generate a series of plots and tables to characterize the simulated region.
+* Once a particular run is constructed, but before the run proceeds, generate a series of plots and tables to characterize the simulated zone.
 
@@
 
-1) Initiate simulated region - use a ctrl file to define details of data-file, HCR function, and other initial conditions. May involve setting an initial depletion level.
+1) Initiate simulated zone - use a ctrl file to define details of data-file, HCR function, and other initial conditions. May involve setting an initial depletion level.
 
-2) conduct assessment of region status at start of each year (option for partial year?)
+2) conduct assessment of zone status at start of each year (option for partial year?)
 
@@
 
-* SMUname    - eg block1, block2
+* SAUname    - eg block1, block2
 * popnum - population index   1, 2, 3, etc
@@
 
-2) A data-file containing details relating to the whole region
+2) A data-file containing details relating to the whole zone
 
@@
 * Nclass - number of size classes 105
-* nSMU   - number of Spatial Management Units  eg  2
+* nSAU   - number of Spatial Management Units  eg  2
 * LML    - current Legal Minimum Length (LML, MLL, MLS, etc) e.g. 140
@@
 * label for run
-* region file name
+* zone file name
 * datafile name
@@
 We are using a one-year time-step but the dynamics occurs within each year.
-The dynamic part of the region has the following structure:
+The dynamic part of the zone has the following structure:
 
-regionD - the dynamic part of hte regional structure
+zoneD - the dynamic part of the zonal structure
 
-* SMU     : num [1:6] 1 1 2 2 2 2
+* SAU     : num [1:6] 1 1 2 2 2 2
 * matureB : num [1:40, 1:6] 782 0 0 0 0 ...
@@
 This represents the dynamics before any simulations have occurred, the unfished equilibrium state.
-Similarly, the constant part of the region, made up of a list for each population, has the following structure:
+Similarly, the constant part of the zone, made up of a list for each population, has the following structure:
 
@@
 * MatWt   : num [1:105] 4.97e-11 6.13e-10 2.93e-09 9.53e-09 2.50e-08 ...
-* SMU     : num 1
+* SAU     : num 1
+
+In the code (_setupzone_) the sequence is:
+
+1.
+first generate zoneC from the constants and zone1.
+At that point each population structure contains everything except those things relating to productivity: B0, ExB0, MSY, MSYDepl and popq.
+makezoneC also updates glb by including the move matrix.
+2. running the _makezone_ updates B0, ExB0, and popq in zoneC and generates the initial equilibrium for zoneD. As this is the unfished equilibrium the catch, harvestR, and catchN matrices are empty.
+3. The final run of _moderegC_ internally runs _doproductivity_ which allows zoneC's MSY and MSYDepl entries to be updated for each population, as well as producing the product matrix.
+
+In the zoneD matrices, the numbers-at-size are the numbers at the start of the year, similarly, the matureB is the mature biomass at the start of the year.
+The ExB0 is the exploitable biomass implied by the initial numbers-at-size, which means to calcualte it we need to apply half of natural mortality and growth before multiplyiung by selectivity and weight-at-length.

### 2020-06-10 R/makehtml_funs.R

@@ -1,482 +1,0 @@
-
-#' @title dirExists: Checks for the existence of a directory
-#'
-#' @description dirExists: does a directory exist? It uses dir.exists
-#'     and reports existence if already present and uses dir.create
-#'     it it does not exist, but avoids the warning message is one
-#'     already exists. The option of not creating a new directory is
-#'     also present.
-#'
-#' @param indir a character string containing the name of the directory
-#'     whose existence is to be checked before it is created if it
-#'     does not already exist.
-#' @param make if the directory does NOT exist should it be created.
-#'     default = TRUE; if make=FALSE and a directory does not exist
-#'     a warning will be given to the console.
-#' @param verbose default=TRUE, prints directory status to the console,
-#'     If make is set to FALSE and a directory does not exist a
-#'     warning will always be given.
-#'
-#' @return a message to the screen if the directory exists or is
-#'     created; if make is TRUE then it also creates the directory as
-#'     listed in 'indir'.
-#' @export
-#'
-#' @examples
-#' \dontrun{
-#' indirect <- getwd()
-#' dirExists(indirect,make=FALSE)
-#' }
-dirExists <- function(indir,make=TRUE,verbose=TRUE) {
-  if (dir.exists(indir)) {
-    if (verbose) cat(indir,":  exists  \n")
-  } else {
-    if (make) {
-      dir.create(indir)
-      if (verbose) cat(indir,":  created  \n")
-    } else {
-      warning(cat(indir,":  does not exist \n"))
-    }
-  }
-}  # end of dirExists
-
-#' @title filenametopath safely add a filename to a path
-#'
-#' @description filenametopath add a filename to a path safely, using
-#'     pathtype to get the seperator and then checks the end character.
-#'     If the separator is nothing or a '/' or a '//' then it adds to
-#'     the path appropriately. Without this one can unwittingly include
-#'     extra separators or none at all.
-#'
-#' @param inpath the path to be analysed
-#' @param infile the filename to be added to the path 'inpath'
-#'
-#' @return the completed filename or extended path
-#' @export
-#'
-#' @examples
-#' indir <- tempdir()
-#' infile <- "control.csv"
-#' filenametopath(indir,infile)
-filenametopath <- function(inpath,infile) {
-  typepath <- pathtype(inpath)
[…]

### 2020-09-02 README.Rmd
Modified those files that needed changes because makehtml had changed how it dealt with addtable and addplot. Hence bare-bones.r and README.Rmd have changed. No Errors, No Warnings, but notes about finding directories in teh check directory (which gets deleted), and the future time stamp issue.

@@
 library(aMSE)
+library(rutilsMH)
 library(makehtml)
 # Obviously you should modify the resdir to suit your own computer
-resdir <- "./../../rcode2/aMSEUse/out/testrun"
+resdir <- "./../../A_code/aMSEUse/out/testrun"
 dirExists(resdir,make=TRUE,verbose=TRUE)
-# You now need to ensure that there is a control.csv, reg1sau2pop6.csv
-# and zone1.csv file in the data directory
-ctrlfiletemplate(resdir)  # puts a template ctrlfile into resdir
-zonefiletemplate(resdir)  # puts a template zone file into resdir
-datafiletemplate(6,resdir,filename="zone1sau2pop6.csv") # etc
-ctrl <- checkresdir(resdir) # checks the data files are present
-runname <- ctrl$runlabel    # and reads the control file, if present
+# The following function calls will ensure that there is a control.csv,
+# a zone1sm\au2pop6.csv
+# and region1.csv file in the data directory
+ctrlfiletemplate(resdir)
+zonefiletemplate(resdir)
+datafiletemplate(6,resdir,filename="zone1sau2pop6.csv")
+
+ctrl <- checkresdir(resdir)
+runname <- ctrl$runlabel
 zone1 <- readzonefile(resdir,ctrl$zonefile)
-glb <- zone1$globals
+glb <- zone1$globals     # glb without the movement matrix
 constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
 
-out <- setupzone(constants, zone1) # make operating model 
-zoneC <- out$zoneC     # constant bits
-zoneD <- out$zoneD     # dynamics bits
-product <- out$product     # all usually saved in resdir, not here
-glb <- out$glb             # now contains the larval movement matrix
-  # Are we at equilibrium?
+out <- setupzone(constants,zone1) # make operating model
+zoneC <- out$zoneC
+zoneD <- out$zoneD
+glb <- out$glb        # glb now has the movement matrix
+product <- out$product     # important bits usually saved in resdir
+          # did the larval dispersal level disturb the equilibrium?
 regDe <- testequil(zoneC,zoneD,glb)
@@
                    zoneC=zoneC, zoneD=zoneD, product=product,
-                   glb=glb,constants=constants
-)
+                   glb=glb,constants=constants)
 
 runnotes <- "This is a bare-bones example."
-# If you unhash the make_html component it will generate a local 
-# website inside resdir and open it so you can see the results so far.
-#
+# If you unhash the make_html call it will generate a local website 
+# inside resdir and open it so you can see the results so far.
 # make_html(replist=reportlist,resdir=resdir,width=500,
-#          openfile=TRUE,runnotes=runnotes,verbose=FALSE)
-
+#          openfile=TRUE,runnotes=runnotes,verbose=FALSE,
+#          packagename="aMSE",analysis="aMSE")
 ```

### 2020-09-04 README.Rmd
Simplified the characterization of the properties of initial conditions. No Errors, Warnings, and 2 empty notes.

@@
 # Obviously you should modify the resdir to suit your own computer
-resdir <- "./../../A_code/aMSEUse/out/testrun"
+resdir <- "./../../A_code/aMSEUse/out/run1"
 dirExists(resdir,make=TRUE,verbose=TRUE)
@@
 
-resfile <- setuphtml(resdir,runname)# prepare to save and log results
+resfile <- setuphtml(resdir)# prepare to save and log results
 
-plotproductivity(resdir,runname,product,glb)
-biology_plots(resdir, runname, glb, zoneC)
-numbersatsize(resdir, runname, glb, zoneD)
+plotproductivity(resdir,product,glb)
+biology_plots(resdir, glb, zoneC)
+numbersatsize(resdir, glb, zoneD)
 
@@
 #          openfile=TRUE,runnotes=runnotes,verbose=FALSE,
-#          packagename="aMSE",analysis="aMSE")
+#          packagename="aMSE",htmlname="testrun")
 ```

### 2020-09-04 data-raw/Running_aMSE.Rmd
Simplified the characterization of the properties of initial conditions. No Errors, Warnings, and 2 empty notes.

---
title: "Running the MSE"
author: "Malcolm Haddon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Running the MSE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type="text/css">
  body, td {
     font-size: 15px;
  }
  code.r{
    font-size: 15px;
  }
  pre {
    font-size: 8px
  }
</style>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return '3.'+n}
      } 
  }
});
</script>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

options(knitr.kable.NA = "",
        knitr.table.format = "pandoc")

options("show.signif.stars"=FALSE,
        "stringsAsFactors"=FALSE,
        "max.print"=50000,
        "width"=240)

library(aMSE)
library(rutilsMH)
suppressPackageStartupMessages(library(knitr))
library(captioner)

tab_nums <- captioner(prefix = "Table")
fig_nums <- captioner(prefix = "Figure")
pgwid=75     
```

# Introduction

The final intent to develop the MSE code base to the point where it should be possible to produce the data files, the control file, and a harvest control file and set multiple batch jobs running. It is assumed that even with maximum code optimization each MSE run would take too long to be run in an interactive manner. Hence, there is a requirement to be able to set off multiple jobs at once, taking advantage of modern computers having multiple CPU cores. Batch jobs are an ideal way to handle this.

However, during the development of the MSE framework it is expected that this will proceed in a sequential, although multi-faceted, fashion and many of the details of each run will need to be apparent to the development team. This document is to assist with that process. Once the MSE project is nearing completion a different 'How to Run an MSE' document will be needed. But for now, a document with many more details is required, not just for the developers but for those who may work to maintain and possibly further develop the code after this project finishes.

Despite this being more of a development document than a final document it remains important to install from the beginning the underlying requirements of the final batch running environment. The details of these requirements may change through the project development but the fundamental structures should remain the same.

## The Underlying Spatial Structure and Population Dynamics

The spatial complexity of abalone fisheries is the main reason that structurally simple stock assessment models fail to capture the dynamics of abalone fisheries. Hence, an explicit spatial structure is required for the operating model. Here, the upper geographical scale being used is that of a zone, which is the scale at which quota or TACs are set. In Tasmania this is equivalent to the quota zone. Within the single zone, there will be (can be) multiple spatial management units (SAUs) for which aspirational catches are set (which sum to the TAC; in Tasmania these are equivalent to the statistical blocks), and within each SAU there can be multiple populations (sub-areas) of abalone, each with their own set of biological properties, which reflect, to some extent, the average properties for the SAU.

Within each population the population dynamics are described using difference equations operating at an annual time-step with the biological processes of growth, mortality, and recruitment occurring in a modelled sequence. The dynamics in this size-based model are described by following the fate of the numbers-at-size. In each year the dynamics begin with the numbers-at-size at the start of each year. The first process undergone is that growth occurs, then half natural mortality is applied. At that point the exploitable biomass can be calculated from the numbers-at-size before fishing occurs. The remaining dynamics involve the removal of the catch, the application of the last half of natural mortality and the addition of recruits, at that point larval dispersal occurs (if it is > 0.0), which produces the numbers-at-size at the start of the next year.

## The Input File Requirements

To implement a run using the __aMSE__ R package one currently needs to install both __aMSE__ and __rutilsMH__, both of which can be downloaded or installed from the GitHub sites:

https://github.com/haddonm/aMSE and https://github.com/haddonm/rutilsMH

Instructions for how to make those downloads are included in the _ReadMe.md_ files for each package, but as an example one could use:

```{r echo=TRUE, eval=FALSE}
if (!require(devtools)){install.packages("devtools")}

devtools::install_github("https://github.com/haddonm/aMSE",build_vignettes=TRUE)

```

### Package Dependency

Apart from __rutilsMH__, the __aMSE__ package now uses the new package __makehtml__ that contains functions that allow for the rapid local website generation of results from a particular run. It is thus necessary to also install that package.

Apart from installing the R packages __aMSE__, __rutilsMH__, and __makehtml__, for each run of the MSE you will need to identify and name a directory somewhere on your computer, here we will refer to it as the _resdir_, which will contain all the results. You will need such a _resdir_ for each separate MSE run that is made. Each such result directory will be the focus for all the stored results from each particular run. The idea is that given the _resdir_ it will be possible to access all the plots, tables, and other results generated from the given run (indeed it should be possible to repeat the entire run if required). Later functions will be written that will take a set of such run directories and summarize and compare the results found within.

In the 'resdir' directory there needs to be at least four .csv files:

1. A control file that identifies ideally, a unique run name, the other input file names, the number of replicates, and other details of a particular MSE run, such as which harvest control rule (HCR) to use.

2. A zonefile, which contains a set of constants that relate to the whole zone. this includes the definition of the spatial structure, constants relating to the size structure, and a set of global constants used throughout the modelling.

3. A file of biological constants for each population making up the zone, including details of growth, maturity, weight-at-length, natural mortality, recruitment, emergence, selectivity and expected initial unfished catch rates.

4. A harvest control-rule file containing any constants or values used within the HCR.

Functions are provided within __aMSE__ for writing out template versions of the first three files. These mimic the worked example we will use below and have been used to generate some of the built-in data-sets within __aMSE__, which are used to illustrate the operation of the various R functions that have been developed for conducting the simulations. In practice, one would run the _ctrlfiletemplate_, the _datafiletemplate_, and the _zonefiletemplate_ R functions to produce such files and then use them after editing the contents to suit your own fishery and particular run details (such editing would constitute much of the conditioning of the operating model.

You should not be surprised to find that just as there are _xxxxfiletemplate_ functions there are _readctrlfile_, _readdatafile_, and _readzonefile_ R functions (though eventually these may be hidden within more all encompassing functions). The intent (hope) is that most users should eventually not need to know such details.

The 'resdir' directory (referred to in the R code as _resdir_) is the repository for all the plots and tables that are produced by __aMSE__ for a particular run. When the directories for a particular run are setup and checked (using _setuphtml_), a 'resultTable_runname.csv' file is also created. After that, if desired, each plot or table that are generated can be stored as a '.png' or '.csv' file in the _resdir_, and the filename of each result logged by adding it to the 'resultTable_runname.csv' (the function _logfilename_ is used to add the filename to the list). Actually, now it is best to use the functions _addtable_ and _addplot_ to do the work for you. At the same time one should add a category (productivity, catches, etc.) and caption (see the help page for _logfilename_, _addplot_, and _addtable_). At any stage in the run, it is then possible to call the function _make_html_ and this will produce a local webpage inside the resdir and one can open it in a local browser (or it can open automatically). This provides for a quick way to visualize the results from a run and allow for the planning of later comparisons.

## A Possible Workflow

There are thus a number of steps involved when running __aMSE__:

1. Initial decisions on the run directory, the runname, filenames for the zone, data, and HCR files, the spatial structure to be simulated, the zone wide properties (such as larval dispersal rate), the number of simulations to run, the number of years to project the simulation, and other aspects relating to the details of a particular run (see the details of the _ctrl_ and _zone_ objects and data files).

2. Condition the operating model on a fishery using biological characteristics taken from the fishery, or, alternatively, use a hypothetical set of characters as a testbed (see the details of the _constants_ R object).

3. The initiation of the operating model sets up the dynamics in an unfished state at equilibrium. Once that is achieved the properties of the unfished 'fishery' are characterized. This characterization includes a summary of the basic biological properties for each population including the selectivity, maturation, growth, and productivity at all spatial scales (population, SAU, zone). This characterization results in a series of plots and tables all of which are stored in _resdir_ and logged in _resfile_.

4. If the initial depletion for the run is < 1.0, then one uses _dodepletion_ to achieve the desired depletion for the zone. This zone-wide depletion level is found by searching for the harvest rate that, when applied to all populations, depletes the zone to close to the desired initial level. This means that the depletion within each population may differ (the example code within the _getzoneprops_ function illustrates this).

5. Once again one can characterize the properties of the zone and populations to contrast the fished state with the unfished.

6. Once all this has been achieved then one can move to conducting projections under the control to the selected harvest control rule.

## Initial Decisions

When conducting a management strategy evaluation to compare a set of harvest strategies, running such simulations requires a number of decisions about how to condition the model to reflect a particular fishery. The following questions require an answer:

1. How many Spatial Assessment Units are going to be included in the simulated zone?

2. How many separate populations will make up each SAU?

3. How will the biological properties of each population be allocated? Will they be characterized as if they are linearly arranged along the simulated coast, or will their properties be allocated randomly? Without detailed spatial information about productivity and yield such allocation when mimicking a real fishery can take a lot of trial and error.

4. Will larval dispersal among the populations be set at zero, or will a particular proportion of larval production be allocated to dispersal to adjacent populations.

These questions are important because they determine the spatial structure imposed on the operating model of the fisheries dynamics (see the vignette on MSE for explanations of the components that make up the MSE model).

The first question about the number of SAUs will depend upon how management is arranged. For example, in Tasmania, a total allowable catch is allocated to each quota zone. But spatial management is taken further by allocating intended catch levels to each statistical block to avoid excessive fishing in any single block should abalone become available in greater numbers than usual (__`r fig_nums("f1",display="cite")`__). Elsewhere, however, it may be the case that management decisions are made at the level of each SAU and the spread of effort or catch among its component populations is decided each year by the industry members. Such possible variation can be accommodated by __aMSE__ by changing the number of SAU and associated populations.

<br>

```{r pressure, echo=FALSE, fig.cap="A caption", out.width = '65%'}
knitr::include_graphics("C:/Users/User/Dropbox/A_code/aMSEUse/figures/Tas_Abalone_Zones.png")
```

`r fig_nums("f1", caption=" A schematic map of the Tasmanian quota zones and associated legal minimum lengths. A trimmed version of the figure on page 189 of Mundy and McAllister (2019).")`

<br>

The number of populations making up each SAU is also an important decision, which is fundamental to the level of detail to which it is possible to take the model's conditioning. The expectation is that, within Tasmania, using the GPS-logger data it will be possible to characterize smaller more homogeneous areas within each statistical block (SAU). These may be characterized using their average relative productivity through the available time-series, or by any other form of data available for the selected  area. In answer to the third question, it should also be possible, using the GPS-logger data, to characterize such populations in terms of relative productivity in sequence around the coastline, which can then be mimicked within the simulation model. If such detailed data are not available then conditioning the operating model might involve randomly allocating biological properties across populations repeatedly until a schema arises that generates data with strong similarities to the observed history of the fishery.

The fourth question about larval dispersal will be influenced by the area and the species being considered. With blacklip abalone in Tasmania mostly larval settlement occurs locally with only minor levels of larval dispersal to populations no large distance away (Miller _et al_., 2008). But even low levels of larval dispersal (the example we will use below has a dispersal rate of 3% - 0.03) have an influence on equilibrium conditions and potential yields, especially when large productive populations lie next to less productive smaller populations.

## Conditioning on Biological Characteristics

An array of biological characteristics are required for each population (see the contents of _data(constants)_). These relate to maturity and growth, and other aspects of productivity such as unfished recruitment levels. Translating the biological data that has been collected from the various fisheries through time into these parameters is a large part of what is entailed in conditioning the operating model. The full requirements will become clear when we discuss the format and contents of the various files required to conduct a model run.

## Conditioning Options

The abalone MSE package (__aMSE__) is designed to permit three different approaches to simulation testing. Each of these would requires a somewhat different initiation. The value of having these alternatives is that the amount of time taken to condition the model differs considerably between them.

### Unfished equilibrium - Purely hypothetical testing.

This could be used when testing the sensitivity to variation or different conditions of a given harvest strategy

* begin from an unfished equilibrium
* apply a given time-series of catches or harvest rates using a given harvest strategy throughout

### An Initial Hypothetical Depletion Level

Similar to version 1, but would include the sensitivity to initial conditions in any testing. It could answer the question of how well a harvest strategy operates at different depletion levels.

* begin from an unfished equilibrium
* apply a constant series of catches or a constant harvest rate to attain a given depletion level before applying the chosen harvest strategy.

### Condition the Operating Model on a Real Fishery

* begin from an unfished equilibrium
* apply an historically accurate level of fishing at the SAU scale, at least, and ideally at the population level (would require identifying populations using the GPS data logger data and using those area identifications to sub-divide the historical catches appropriately).
* first apply with respect to size, any known biological characteristics of weight, growth, maturity.and emergence to each population within each SAU.
* select plausible values for those constants, such as natural mortality, unfished recruitment, and unfished mature biomass
* then adjust productivity terms R0, B0, and others if required, so that the predicted dynamics of the individual SAU, under the historical catches reflect the observed dynamics. 
* if there is any knowledge of recruitment variation then include that as well.
* Use surplus production models or even size-structured models (if size-composition data is available) to gain insight into the relative productivity of each SAU.

It should be clear that option three should be most valuable to a particular fishery but also the hardest option to condition the operating model in an appropriate manner.

## A Worked Example

We have already gone through quite a few details, much of which will easily have been forgotten, so now we will examine setting up a model in practice and getting it operational, which will include displaying the initial results. In this example, we will be creating directories on your hard-drive. At that point in the code you may wish to change the names of the directory structure so that the example does not upset your, no doubt, neat and tidy 'C:' drive directory layout. Throughout the worked example, we will be calling numerous R functions.

I recommend that you examine the help page for each of these (for example using _?setupdirs_). To see the code for each function just type the name of the function into the console but without brackets. Finally, to see what arguments each function have you can use _args(function_name)_ or _formals(function_name)_.

```{r echo=TRUE}
# much of these details should be hidden within high level functions
# once the code development is complete. 
starttime <- as.character(Sys.time()) # When did the run start?
library(aMSE)  # you might check out the help pages for each function
library(makehtml)
 # Obviously you should change the resdir to something that suits you!
 resdir <- "C:/Users/User/Dropbox/A_code/aMSEUse/out/run1"
 dirExists(resdir,make=TRUE,verbose=TRUE)
 # You now need to ensure that there is a control.csv, zone1sau2pop6.csv
 # and zone1.csv file in the data directory
ctrlfiletemplate(resdir)
zonefiletemplate(resdir)
datafiletemplate(6,resdir,filename="zone1sau2pop6.csv")

ctrl <- checkresdir(resdir)
zone1 <- readzonefile(resdir,ctrl$zonefile)
glb <- zone1$globals
constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)

out <- setupzone(constants, zone1)  # can take a few seconds
zoneC <- out$zoneC  # the constant part
zoneD <- out$zoneD  # the dynamic part
product <- out$product  # the productivity defined
glb <- out$glb

resfile <- setuphtml(resdir) #needed to formalize the outputs
 # store the initial objects
save(ctrl,file=filenametopath(resdir,"ctrl.RData"))
save(glb,file=filenametopath(resdir,"glb.RData"))
save(product,file=filenametopath(resdir,"product.RData"))
save(zoneC,file=filenametopath(resdir,"zoneC.RData"))
save(zoneD,file=filenametopath(resdir,"zoneD.RData"))
numpop <- glb$numpop
```

<br>

Assuming you got this far then you have successfully generated template files, and we will assume that lots of changes have been made to them to customize them to a particular fishery (see the vignette on conditioning the model). Then you will have read them back into the R environment (in sequence because each depends on its predecessor). Of course, once edited the _ctrl..._, _zone..._, and _datafiletemplate_ function calls will need to be deleted or hashed into comments, otherwise if you run the task again you will over-write your edits.

### The zone Structure

With all the inputs in the system the next thing to do is to generate the simulated zone. This largest of spatial scales is made up of two components. The first contains the constants relating to the zone and this large object is a list of _numpop_ populations, each of which is a list of all the properties that will not change through the run (see its structure just below). Then there is the dynamic part of the zone, which is a list of matrices and arrays that contain the dynamics, the things that change through time. We will describe the structure of each in some detail once they have been generated. The generation is a two step process, initially an analytical solution to unfished equilibrium conditions is used and then we use the function _findunfished_ to numerically find the equilibrium when there is larval dispersal present in the zone.

```{r}
# Is there an equilibrium even with larval dispersal?
zoneDe <- testequil(zoneC,zoneD,glb)
```

<br>

Generating the zone entails the production of what we have called _zoneC_, _zoneD_, and _product_. Now we can examine their individual structure and components. If we first consider the structure of _product_ this is the simplest of the R objects being a 3-D array of harvest rates vs population statistic vs population.

```{r}
str(product)
```

<br>

Essentially it is a matrix of of _numpop_ matrices with the exploitable biomass, the mature biomass, the effective harvest rate, the catch or yield, the depletion level, and the relative cpue. Each row of values is obtained by running the dynamics with each constant harvest rate (listed in the rownames) until an equilibrium is reached. Thus, by searching for the maximum equilibrium or sustainable yield one can characterize what is required to obtain the $MSY$, the $B_{MSY}$, the $H_{MSY}$ and what level of stock depletion will be expected to provide the circumstances that can lead to the maximum sustainable yields.

The productivity characteristics can be illustrated by a simple plot (__`r fig_nums("f2",display="cite")`__). It is important to note that the productivity of each population differs as does the harvest rate that would produce the optimum yield from each population. In this case the depletion level that gives rise to the maximum production appears to be approximately the same, but this is not always the case. The implication of this is that attempting to manage the stock at the zone level would not permit the identification of a single optimum harvest rate for all blocks. The _MSY_ in each case is the peak of each curve in the plots with yield on the y-axis.

```{r fig.width=6, fig.height=5, echo=FALSE}
parset(plots=c(2,2))
maxy <- getmax(product[,"MatB",])
plot(product[,"AnnH",1],product[,"MatB",1],type="l",lwd=1,col=1,ylim=c(0,maxy),
     ylab="Mature Biomass",panel.first=grid(),xlab="Harvest Rate")
for (i in 2:numpop)lines(product[,"AnnH",i],product[,"MatB",i],col=i)

maxy <- getmax(product[,"Catch",])
plot(product[,"MatB",1],product[,"Catch",1],type="l",lwd=1,col=1,ylim=c(0,maxy),
     ylab="Yield",panel.first=grid(),xlab="Mature Biomass",xlim=c(0,1750))
for (i in 2:numpop)lines(product[,"MatB",i],product[,"Catch",i],col=i)

maxy <- getmax(product[,"Catch",])
plot(product[,"AnnH",1],product[,"Catch",1],type="l",lwd=1,col=1,ylim=c(0,maxy),
     ylab="Yield",panel.first=grid(),xlab="Harvest Rate")
for (i in 2:numpop)lines(product[,"AnnH",i],product[,"Catch",i],col=i)

maxy <- getmax(product[,"Catch",])
plot(product[,"Deplet",1],product[,"Catch",1],type="l",lwd=1,col=1,ylim=c(0,maxy),
     ylab="Yield",panel.first=grid(),xlab="Depletion")
for (i in 2:numpop)lines(product[,"Deplet",i],product[,"Catch",i],col=i)

```

`r fig_nums("f2", caption=" Four plots illustrating different aspects of the production for each population.")`

<br>

The values for the $MSY$, the $B_{MSY}$, the $H_{MSY}$, and related values of depletion,  are searched for within _product_ and then included into each population list within the  _zoneC_ object, which is why generating that object takes two steps.

```{r }
 # the structure of only one population is shown; the black lines in Figure 2
str(zoneC[[1]],width=pgwid,strict.width="cut")
```

<br>

The structure of each of the lists within the _zoneC_ object has the constant scalars come first followed by the vectors and matrices. The _Select_ matrix has as many columns as there will be years in the forward projection, which allows for any changes that might occur in the LML during the MSE run. The vector _MatWt_ is simply a shortcut to avoid having to multiple the maturity-at-length vector by the weight-at-length each time the mature or spawning biomass is calculated. Similarly, _SelWt_ is a matrix of selectivity-at-length by weight-at-length to avoid having to repeatedly calculate this every time the exploitable biomass is calculated. Once setup, none of these constants should change through the period of the particular run being made.

The dynamic part of the zone, _zoneD_ is derived within the _makezone_ function using the constant part of the zone, ie _zoneC_. It is a list of one vector, eight matrices, and two arrays. Thus, the mature biomass, exploitable biomass, catch, and other _Nyrs_ x _numpop_ matrices provide repositories for the projection of each variable. The two arrays, _catchN_ and _Nt_ have dimensions of length-class x year x population, so the complete dynamics is captured within these. Both of these objects _zoneC_ and _zoneD_ will be saved in _resdir_.

```{r}
str(zoneD,width=pgwid,strict.width="cut")
```

* for each population, _matureB_ = mature biomass, _exploitB_ = exploitable biomass, _catch_ = catch in tonnes, _harvestR_ = annual harvest rate, _cpue_ is the expected cpue, _recruit_ = recruitment in each year, _deplsB_ = depletion of mature biomass, _depleB_ = depletion of exploitable biomass, _catchN_ = numbers-at-length in the catch each year, and finally, _Nt_ is the population numbers-at-length each year.

<br>

The properties of each population and the zone can be tabulated using the function _getzoneprops_.

```{r}
unfishprops <- getzoneprops(zoneC=zoneC,zoneD=zoneD,glb=glb,year=1)
# check out the help for getzoneprops
```

As well as printing out them out we would want to save the unfished zone properties for later reference. We can do that using the usual _write.table_ function but also by logging the filename used into _resfile_ defined earlier.

```{r }
# Use a unique file name and put it into the resdir
filename <- "unfishprops.csv"
write.table(unfishprops,file = filename,sep=",")
# make up a caption for use when the table is printed
caption <- paste("The unfished equilibrium properties of the populations ",
                 "and zone, before any initial depletion.",collapse=" ")
# The category names the tab in local website where the table would be found 
addtable(unfishprops,filen=filename,resdir=resdir,category="Tables",caption)
```

Here we print out the table to see the typical contents

`r tab_nums("t1", caption=" The summary of a zone's properties for the unfished newly generated zone.")`

```{r echo=FALSE}
suppressMessages(kable(unfishprops,digits=c(4,4,4,4,4,4,4)))

```

## Starting at a Depletion < 1.0

During the comparisons of different harvest strategies it will be necessary to determine how well each HS performs starting from different initial depletion levels. This would be answering questions such as 'How well does the HS allow for stock recovery from a depleted stock?'. The intended initial depletion is included as an entry in the _ctrl_ file. And a function, _dodepletion_ is provided to impose fishing mortality onto the stock so that the zone is depleted until it is close to the intended level. The zonal depletion is determined from a weighted mean depletion across all populations, where the weighting factor is the relative unfished mature biomass predicted for each population (see the equations describing the operating model). If, for example, we wanted to deplete our initial zone to a level close to 20% unfished mature biomass we would implement the following code (see the function's help page for a description of the six arguments.

```{r}
zoneDD <- dodepletion(zoneC=zoneC,zoneD=zoneD,glob=glb,depl=0.20,
                        product=product)
str(zoneDD,width=pgwid,strict.width="cut")
```

<br>

If we now run the _getzoneprops_ function, of course we would expect a different set of answers for the depletion of mature and exploitable biomass.

```{r}
initprops <- getzoneprops(zoneC=zoneC,zoneD=zoneDD,glb=glb,year=1)
# check out the help for getzoneprops
filename <- "initprops.csv"
write.table(initprops,file = filename,sep=",")
#  or use tmp <- read.csv(file=filename,header=TRUE,row.names=1)
caption <- paste("The properties of the populations and zone after ",
                 "the zone has been depleted to its starting point.",collapse=" ")
logfilename(filename,resfile=resfile,"Tables",caption)
```

`r tab_nums("t2", caption=" The summary of a zone's properties for the zone once it has been depleted to approximately 0.20B0.")`

```{r echo=FALSE}
kable(initprops,digits=c(4,4,4,4,4,4,4))
```

<br>

Notice that while the spawning biomass (=mature biomass = SpbDepl) is 0.1967 (almost 20%B0), the values for each of the populations varies between 0.14 and 0.27, which reflects the effects of the same fishing mortality on the different productivity levels of each population.

## Fleet Dynamics

The application of any harvest control rule in this complex spatial model will entail the translation of SAU level aspirational catches (which should sum to the zone's TAC), into catches applied to each individual population within each SAU. Of course, the aspirational catches per SAU derived from the harvest control rules (HCR) within each harvest strategy will not be taken exactly, The overall TAC will eventually constrain total catches, but some SAU will experience larger catches than expected, and others will experience smaller catches. SO the application of catches to populations requires two steps.

The first step is to include variation into the aspirational catches from the HCR and then scale them until their sum equals the original TAC. The second step requires that each SAU's actual catch is distributed across their component populations in some  manner that reflects plausible fleet dynamics.

If $C_{u,HCR}$ is the aspiration catch for each SAU $u$ from the HCR, then the first step on the way to setting the actual SAU catch is to include Normal variation:

$$C_{u}^*=N(C_{u,HCR},\sigma_u)$$

where $C_{u}^*$ is the SAU catch before re-scaling to the TAC, and $\sigma_u$ is the standard deviation of the variation that occurs between the aspiration and the actual catches within SAU's. We need to find a scaling factor $\lambda$ to re-scale the total SAU catches to sum to the agreed TAC:

$$\lambda=\frac{\sum\limits_{u=1}^{nSAU}C_{u}^*}{TAC}$$

Thus, the actual SAU catches, $C_u$, after including variation and re-scaling are:

$$C_u=\frac{C_{u}^*}{\lambda}$$

Each SAU will have various numbers of component populations, so now the problem becomes how to allocate a proportion of the actual SAU catch, $C_u$, to each of those populations.

At least in the Tasmanian blacklip abalone fishery the relationship between catch-rates and catches is usually observed to be a linear, hence the relation between catches and effort is also linear. Because of this catch-rates are assumed to have at least some influence over the distribution of catches among areas. Observed catch-rates (cpue)
would naturally be expected to be variable through time and across areas and so are modelled as:

$$I_{t,p}={q_p}\left(B_{t,p}^E\right)^{\lambda}{e^{N(0,\sigma_q)}}$$

where $I_{t,p}$ is the cpue in year $t$ for area $p$, $q_p$ is the catchability coefficient within area $p$, ${B_{t,p}^E}$ is the exploitable biomass in year $t$ and area $p$, with a non-linearity coefficient of $\lambda$, and ${e^{N(0,\sigma_q)}}$ is a Log-Normal random deviate, with $\sigma_p$ being the standard deviation of the catchability coefficient $q_p$. If $\lambda = 1.0$ then the relationship between cpue and exploitable biomass is linear, values other than $\lambda = 1.0$ lead to non-linear relationships. Given how important the use of cpue is in all Australian abalone harvest strategies, this is one assumption whose influence is in need of testing. We are using $p$ as a sub-script for area, because in the conditioning each population within each Spatial Assessment Unit corresponds to a particular area within that SAU. Effectively, area and population are the same thing in the model.

In the operating model, given a total allowable catch (TAC) or quota, if the harvest strategy selected does not allocated the desired level of catch among the different populations (sub-areas) in any given year then some fleet dynamics process will need to be adopted to make that distribution. Given the previous reasoning about cpue, then previous catch rates across the different populations may be considered able to serve as a guide to where to fish in subsequent years. This is a reasonable assumption if the fishery regularly leaves behind a significant proportion of the legal sized animals. However, catch-rates in one year do not give any indication of the availability of undersized abalone that are expected to grow into the fishery. In fisheries that are being fully exploited the advent of new recruits will be an important component of each year’s fishery. Fortunately, the abalone fishery depends on divers literally handling their catch and this automatically provides them an opportunity to identify visually those areas that would be expected to be productive in the next year and also those areas that would be expected to become less productive. Their own observations
would be made with some error and discussions among divers would also rarely be precise. Such diver expectations provide an indication of exploitable biomass and this can thus be used in an algorithm for distributing catches by area:

$$C_{t,p}=TAC \frac{B_{t,p}^E{e^{N(0,\sigma_d)}}}{\sum\limits_{p=1}^{n}{B_{t,p}^{E}{{e}^{N(0,\sigma_d)}}}}$$

where $C_{t,p}$ is the expected catch in area $p$ in year $t$, $TAC$ is the total allowable catch, and $\sigma_d$ is the standard deviation of the catchability interpreted as the diver’s observations on available biomass in the $n$ areas assessed within the year in question.

This approach reflects a system adopted in Dichmont et al (1999) and by Dichmont and Brown (2010) for distributing a TAC among areas. Their approach was related directly to catch-rates rather than exploitable biomass (despite their equation implying a catchability coefficient of 1.0). However, here the exploitable biomass is directly related to catch-rates (even if not always linearly) and so, especially with the random noise added to the biomass values, this can adequately drive the distribution of catches.

As these proxies are for the diver perception of relative abundance they automatically include their knowledge of catches and catch rates from previous years. It would be expected that as $\sigma_d$ increased the ability of divers to appropriately distribute catches between areas within SAUs would decline, which would, in turn, be expected to lead to poor outcomes for the fishery in terms of depletion levels within SAUs.

In practice, in Tasmania, each SAU is considered separately and an aspirational catch is allocated. After all SAUs within a zone have been considered, the individual predicted catches are summed and the whole is then proposed as a TAC for the following year. Alternatively, a formal harvest strategy developed in Tasmania proposes aspirational catches for each SAU (statistical block) based on a multi-criteria decision analysis founded on different but related fishery performance measures.

# References

Miller, K.J., Maynard, B.T. and C.N. Mundy (2008) Genetic diversity and gene flow in
collapsed and healthy abalone fisheries. _Molecular Ecology_ __18__:200-211.

Mundy, C. and J. McAllister (2019) _Tasmanian abalone fishery assessment 2018_, Institute of Marine and Antarctic Studies, University of Tasmania. 190p.

### 2020-09-04 data-raw/bare-bones.r

@@ -15,7 +15,6 @@ zonefiletemplate(resdir)
 datafiletemplate(6,resdir,filename="zone1sau2pop6.csv")
 
 ctrl <- checkresdir(resdir)
-runname <- ctrl$runlabel
 zone1 <- readzonefile(resdir,ctrl$zonefile)
 glb <- zone1$globals     # glb without the movement matrix
 constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
@@ -28,16 +27,15 @@ product <- out$product     # important bits usually saved in resdir
           # did the larval dispersal level disturb the equilibrium?
 regDe <- testequil(zoneC,zoneD,glb)
 
-resfile <- setuphtml(resdir,runname)# prepare to save and log results
+resfile <- setuphtml(resdir)# prepare to save and log results
 
-plotproductivity(resdir,runname,product,glb)
-biology_plots(resdir, runname, glb, zoneC)
-numbersatsize(resdir, runname, glb, zoneD)
+plotproductivity(resdir,product,glb)
+biology_plots(resdir, glb, zoneC)
+numbersatsize(resdir, glb, zoneD)
 
 endtime <- as.character(Sys.time())
 
-reportlist <- list(runname=runname,
-                   starttime=starttime,endtime=endtime,
+reportlist <- list(starttime=starttime,endtime=endtime,
                    zoneC=zoneC, zoneD=zoneD, product=product,
                    glb=glb,constants=constants)
 
@@ -45,5 +43,6 @@ runnotes <- "This is a bare-bones example."
 # If you unhash this component it will generate a local website inside
 # resdir and open it so you can see the results so far.
  make_html(replist=reportlist,resdir=resdir,width=500,
-          openfile=TRUE,runnotes=runnotes,verbose=FALSE)
+          openfile=TRUE,runnotes=runnotes,verbose=FALSE,
+          packagename = "aMSE",htmlname="testrun")
 

### 2020-09-29 README.Rmd
added dynamics.r, added depleteSAU, and improved the conditioning code, though still more to do.

@@
 
-ctrl <- checkresdir(resdir)
+ctrl <- checkctrldat(resdir)
 runname <- ctrl$runlabel

### 2020-09-29 data-raw/Running_aMSE.Rmd
added dynamics.r, added depleteSAU, and improved the conditioning code, though still more to do.

@@
 
-ctrl <- checkresdir(resdir)
+ctrl <- checkctrldat(resdir)
 zone1 <- readzonefile(resdir,ctrl$zonefile)

### 2020-09-29 data-raw/conditionWest.r

@@ -72,10 +72,65 @@ runnotes <- "This is the first conditioning on the west coast."
  glb <- zone1$globals     # glb without the movement matrix
  constants <- readdatafile(glb$numpop,datdir,ctrl$datafile)
 
- out <- setupzone(constants,zone1) # make operating model
+ out <- setupzone(constants,zone1,inc=0.0025) # make operating model
  zoneC <- out$zoneC
  zoneD <- out$zoneD
  glb <- out$glb        # glb now has the movement matrix
  product <- out$product     # important bits usually saved in resdir
  # did the larval dispersal level disturb the equilibrium?
  regDe <- testequil(zoneC,zoneD,glb)
+
+ # do the depletions
+ origzoneD <- zoneD
+ zoneD <- origzoneD
+ origdepl <- c(0.6,0.6,0.7,0.6,0.6,0.6,0.6,0.6)
+ zoneDD <- depleteSAU(zoneC,zoneD,glb,depl=origdepl,product,len=15)
+ zprop <- getzoneprops(zoneC,zoneDD,glb,year=1)
+ round(zprop,4)
+ zoneD <- zoneDD
+
+ nSAU <- glb$nSAU
+ str(zone1,max.level = 1)
+ if (ctrl$hcrname == "historicalC") {
+   histC <- zone1$histCatch
+   histCE <- zone1$histCE
+   histyr <- zone1$histyr
+   yrs <- histyr[,"year"]
+   histLML <- histyr[,"histLML"]
+   nyrs <- length(yrs)
+   for (yr in 2:nyrs) {  # yr=2
+     year <- yrs[yr]
+     lml <- histLML[yr]
+     catchpop <- histC[yr,]
+     zoneD <- oneyearC(zoneC=zoneC,zoneD=zoneD,Ncl=glb$Nclass,
+                       catchp=catchpop,year=yr,sigmar=1e-08,
+                       npop=glb$nSAU,movem=glb$move)
+   }
+
+ }
+ zpropF <- getzoneprops(zoneC,zoneD,glb,year=nyrs)
+ round(zpropF,4)
+
+ outB <- zoneD$deplsB
+ plotprep(width=7,height=6,newdev = FALSE)
+ parset(plots=c(3,1))
+ ymax <- getmax(outB[,c(1,2,3,8)])
[…]

### 2020-10-09 R/generate_results.R
Added numbersatsizeSAU, began documenting the requirements for a generalized MSE.

@@ -147,7 +147,8 @@ compzoneN <- function(unfN,curN,glb,yr,depl,LML=0,resdir="") {
 #'     at-size distribution, omitting the first four size classes to
 #'     avoid the recruitment numbers dominating the plot.
 #'
-#' @param resdir the results directory
+#' @param resdir the results directory, if set to "" then plot is sent to
+#'     the console instead
 #' @param glb the globals list
 #' @param zoneD the dynamic part of the zone, zoneD
 #'
@@ -160,10 +161,14 @@ numbersatsize <- function(resdir, glb, zoneD) {
   # some globals
   mids <- glb$midpts
   numpop <- glb$numpop
-  # initial numbers-at-size uses zoneD--------------------------------
+  # initial numbers-at-size uses zoneD--
   Nt <- zoneD$Nt[,1,]/1000.0
   Ntt <- rowSums(zoneD$Nt[,1,])/1000.0  # totals
-  filen <- filenametopath(resdir,"Initial_N-at-Size.png")
+  if (nchar(resdir) > 0) {
+    filen <- file.path(resdir,"Initial_N-at-Size.png")
+  } else {
+    filen <- ""
+  }
   plotprep(width=7,height=6,newdev=FALSE,filename=filen,cex=0.9,verbose=FALSE)
   parset(plots=c(2,1))
   plot(mids[5:105],Ntt[5:105],type="l",lwd=2,xlab="Shell Length mm (5 - 210mm)",
@@ -175,10 +180,77 @@ numbersatsize <- function(resdir, glb, zoneD) {
   abline(h=0.0,col="darkgrey")
   legend("topright",paste0("P",1:numpop),lwd=3,col=c(1:numpop),bty="n",
          cex=1.2)
-    caption <- "The numbers-at-size for the whole zone and for each population separately. The recruitment numbers are omitted for clarity."
-  addplot(filen,resdir=resdir,category="NumSize",caption)
-
+  if (nchar(resdir) > 0) {
+    caption <- paste0("The numbers-at-size for the whole zone and for each ",
+                      "population separately. The recruitment numbers are",
+                      " omitted for clarity.")
+    addplot(filen,resdir=resdir,category="NumSize",caption)
+  }
 } # end of numbersatsize
+
+#' @title numbersatsizeSAU plots the numbers-at-size for a given SAU
+#'
+#' @description numbersatsizeSAU plots the initial unfished numbers-
[…]

### 2020-10-09 R/getfunctions.R

@@ -118,7 +118,7 @@ getLogical <- function(inline,nb) {  #inline <- txtline; nb=2
 #' @param glob the global variables object
 #' @param zone the zone constants
 #'
-#' @return a matrix of numpop + nSAU + 1 columns of numbers-at-size
+#' @return a matrix of numpop + 1 columns of numbers-at-size
 #' @export
 #'
 #' @examples
@@ -127,18 +127,18 @@ getLogical <- function(inline,nb) {  #inline <- txtline; nb=2
 #' data(testzoneD)
 #' nas <- getnas(testzoneD,yr=1,glob=glb,zone=zone1)
 #' round(nas[1:30,],1)
-getnas <- function(zoneD,yr,glob,zone) {
+getnas <- function(zoneD,yr,glob,zone) {# zoneD=zoneD;yr=1;glob=glb;zone=zone1
   numpop <- glob$numpop
   nSAU <- glob$nSAU
-  columns <- c(paste0("p",1:numpop),zone$SAUnames,"zone")
+  columns <- c(paste0("p",1:numpop),"zone") # ,zone$SAUnames
   nas <- matrix(0,nrow=glob$Nclass,ncol=length(columns),
                 dimnames=list(glob$midpts,columns))
   for (pop in 1:numpop) nas[,pop] <- zoneD$Nt[,yr,pop]
   nas[,"zone"] <- rowSums(nas[,1:numpop],na.rm=TRUE)
-  for (mu in 1:nSAU) {
-    pickcol <- which(zoneD$SAU == mu)
-    nas[,(numpop+mu)] <- rowSums(nas[,pickcol],na.rm=TRUE)
-  }
+  # for (mu in 1:nSAU) {
+  #   pickcol <- which(zoneD$SAU == mu)
+  #   nas[,(numpop+mu)] <- rowSums(nas[,pickcol],na.rm=TRUE)
+  # }
   return(nas)
 } #end of getnas
 
@@ -299,65 +299,66 @@ getsum <- function(inmat,index) {
 #'    unfished. Of course, this will only work as long as movezoneYear
 #'    hasn't been called, which would disrupt the contents of the first
 #'    year. But if applied to zone rather than zone1 etc, this should be
[…]

### 2020-10-14 README.Rmd
Internal consistncy changes with large modifications to readzonefile
Modified the datafile formats from the template making functions and the read daat file functions. This will have broken a number of examples but will get to those.

@@
 # The following function calls will ensure that there is a control.csv,
-# a zone1sm\au2pop6.csv
-# and region1.csv file in the data directory
+# a zone1.csv, and a zone1sau2pop6.csv or population data file 
+# in the data directory, which in this case is the same as the resdir
 ctrlfiletemplate(resdir)
 zonefiletemplate(resdir)
-datafiletemplate(6,resdir,filename="zone1sau2pop6.csv")
+datafiletemplate(numpop=6,resdir,filename="zone1sau2pop6.csv")
 

### 2020-11-08 README.Rmd
Implemented a constCatch HS and illustrated it in the readme.

--- 2020-09-04 README.Rmd
@@
 
-```{r example}
-# a minimal example
-starttime <- as.character(Sys.time())
+```{r example, echo=TRUE}
+# a constant TAC example
+starttime <- (Sys.time())
 library(aMSE)
@@
 library(makehtml)
+library(knitr)
 # Obviously you should modify the resdir to suit your own computer
-resdir <- "./../../A_code/aMSEUse/out/run1"
+if (dir.exists("c:/Users/User/DropBox")) {
+  ddir <- "c:/Users/User/DropBox/A_code/"
+} else {
+  ddir <- "c:/Users/Malcolm/DropBox/A_code/"
+}
+resdir <- paste0(ddir,"aMSEUse/conddata/generic2")  # data and results directory
 dirExists(resdir,make=TRUE,verbose=TRUE)
-# The following function calls will ensure that there is a control.csv,
-# a zone1sm\au2pop6.csv
-# and region1.csv file in the data directory
-ctrlfiletemplate(resdir)
-zonefiletemplate(resdir)
-datafiletemplate(6,resdir,filename="zone1sau2pop6.csv")
-
-ctrl <- checkresdir(resdir)
-runname <- ctrl$runlabel
-zone1 <- readzonefile(resdir,ctrl$zonefile)
+# equilibrium zone -------------------------------------------------------------
+# You now need to ensure that there is, at least, a control.csv, zone1.csv
+# and region1.csv file in the data directory plus some other data .csv files
+# depending on how conditioned you want the model to be. Templates for the
+# correct format can be produced using ctrlfiletemplate(), datafiletemplate(),
+# and zonefiletemplate.
+zone1 <- readctrlzone(resdir,infile="control.csv")
+ctrl <- zone1$ctrl
 glb <- zone1$globals     # glb without the movement matrix
 constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
-
+#zone1$initLML <- 140
 out <- setupzone(constants,zone1) # make operating model
@@
 # did the larval dispersal level disturb the equilibrium?
-regDe <- testequil(zoneC,zoneD,glb)
+zoneD <- testequil(zoneC,zoneD,glb)
+zoneC <- resetexB0(zoneC,zoneD) # rescale exploitB to avexplB after dynamics
+initialC <- zoneC
+unfishedD <- zoneD         # keep a copy of the unfished zone
 
-resfile <- setuphtml(resdir)# prepare to save and log results
+equiltime <- (Sys.time())
+print(equiltime - starttime)
+# deplete generic zone ---------------------------------------------------------
+zoneC <- initialC
+zoneD <- unfishedD
+# set the initial depletion levels if not defined in the data file
+# the values below manage to achieve c(0.7,0.65,0.7,0.6,0.5,0.45,0.475,0.45)
+zone1$condC$initdepl <- c(0.53,0.53,0.48,0.48,0.49,0.49,0.52,0.52,
+                          0.51,0.51,0.45,0.45,0.484,0.484,0.467,0.467)
+origdepl <-  zone1$condC$initdepl
+zoneDD <- depletepop(zoneC,zoneD,glb,depl=origdepl,product,len=12)
+propD <- getzoneprops(zoneC,zoneDD,glb,year=1)
+#round(propD,2)
+# setup for projection ---------------------------------------------------------
+inityrs <- 10
+projyrs <- zone1$projC$projyrs + inityrs
+reps <- ctrl$reps
+projC <- modprojC(zoneC,glb,zone1)
+zoneC <- modzoneCSel(zoneC,projC$Sel,projC$SelWt,glb,projyrs)
+zoneDR <- makezoneDR(projyrs,reps,glb,zoneDD) # zoneDReplicates
+zoneDRp <- addrepvar(zoneC,zoneDR,zoneDR$harvestR,glb,ctrl)
+midtime <- (Sys.time())
+print(midtime - equiltime)
+# prepare the HS --------------------------------------------------------------
+if (projC$HS == "constantCatch") {
+  hsFunc <- constCatch
+  inTAC <- projC$HSdetail
+}
+if (projC$HS == "MCDA") {  # not complete yet, this will do nothing
+  hsFunc <- doMCDA
+  mcdafile <- projC$HSdetail
+  optCE <- MCDAdata(resdir,mcdafile,zone1$SAUnames)
+}
+# Do the replicates ------------------------------------------------------------
+inityr <- 10
+saunames <- zone1$SAUnames
+sauindex <- glb$sauindex
+pyrs <- projC$projyrs + inityr
+B0 <- tapply(sapply(zoneC,"[[","B0"),sauindex,sum)
+exB0 <- tapply(sapply(zoneC,"[[","ExB0"),sauindex,sum)
+# a TAC = 870t = zoneal MSY
+zoneDP <- constCatch(870,zoneDRp,glb,ctrl,projC$projyrs,inityrs=10)
+sauzoneDP <- asSAU(zoneDP,sauindex,saunames,B0,exB0)
 
-plotproductivity(resdir,product,glb)
-biology_plots(resdir, glb, zoneC)
-numbersatsize(resdir, glb, zoneD)
+endtime <- (Sys.time())
+print(endtime - midtime)
+#calculate the relative MSY weighted MSY-depletion level for each SAU
+pmsydepl <- sapply(zoneC,"[[","MSYDepl")
+pmsy <- sapply(zoneC,"[[","MSY")
+smsy <- tapply(pmsy,sauindex,sum)
+smsydepl <- tapply((pmsydepl * pmsy) / smsy[sauindex],sauindex,sum)
 
-endtime <- as.character(Sys.time())
+# plot the projected depletion levels of mature biomass
+ans <- sauzoneDP$saudeplsB
+label <- "Mature Depletion"
+uplim <- 0.7
+plotprep(width=7,height=8,newdev=FALSE)
+parset(plots=c(4,2),byrow=FALSE,margin=c(0.25,0.4,0.1,0.05),
+       outmargin=c(1,1,0,0))
+yrs <- 1:pyrs
+for (sau in 1:8) {
+  maxy <- uplim
+  if (is.null(uplim)) maxy <- getmax(ans[,sau,])
+  plot(yrs,ans[,sau,1],lwd=1,col="grey",panel.first=grid(),ylim=c(0,maxy),
+       xlab="",ylab=saunames[sau])
+  for (iter in 1:reps) lines(yrs,ans[,sau,iter],lwd=1,col="grey")
+  abline(h=smsydepl[sau],lwd=2,col=4)
+  abline(v=inityr,lwd=1,col=2)
+}
+mtext("Years",side=1,line=-0.1,outer=TRUE,cex=1.0,font=7)
+mtext(label,side=2,line=-0.1,outer=TRUE,cex=1.0,font=7)
 
-reportlist <- list(runname=runname,
-                   starttime=starttime,endtime=endtime,
-                   zoneC=zoneC, zoneD=zoneD, product=product,
-                   glb=glb,constants=constants)
-
-runnotes <- "This is a bare-bones example."
-# If you unhash the make_html call it will generate a local website 
-# inside resdir and open it so you can see the results so far.
-# make_html(replist=reportlist,resdir=resdir,width=500,
-#          openfile=TRUE,runnotes=runnotes,verbose=FALSE,
-#          packagename="aMSE",htmlname="testrun")
 ```

### 2020-11-08 R/defineZone.R

@@ -633,6 +633,7 @@ makezoneC <- function(zone,const) { # zone=zone1; const=constants
   midpts <- glb$midpts
   blkdef <- zone$SAUpop
   SAUindex <- defineBlock(nSAU,blkdef,numpop)
+  glb$sauindex <- SAUindex
   SAU <- as.numeric(const["SAU",])
   if (zone$randomseed > 0) set.seed(zone$randomseed)
   if (zone$condition > 0) {

### 2020-11-08 R/inputfiles.R

@@ -394,7 +394,7 @@ read_conddata <- function(filename) {  # filename=filen
 #' ctrl <- readctrlzone(resdir)
 #' ctrl
 readctrlzone <- function(datadir,infile="control.csv") {
-   # datadir=datdir; infile="newcontrol.csv"
+   # datadir=resdir; infile="control.csv"
    filenames <- dir(datadir)
    if (length(grep(infile,filenames)) != 1)
       stop(cat(infile," not found in datadir \n"))
@@ -419,7 +419,7 @@ readctrlzone <- function(datadir,infile="control.csv") {
    SAUpop <-  getConst(indat[begin],nSAU) # how many populations per SAU
    numpop <- sum(SAUpop)
    SAUnames <- getStr(indat[begin+1],nSAU)
-   initdepl <- getConst(indat[begin+2],nSAU)
+   initdepl <- getConst(indat[begin+2],numpop)
    minc <-  getsingleNum("minc",indat) # minimum size class
    cw    <- getsingleNum("cw",indat) # class width
    Nclass <- getsingleNum("Nclass",indat) # number of classes
@@ -442,6 +442,10 @@ readctrlzone <- function(datadir,infile="control.csv") {
       }
       start <- grep("HARVESTS",indat)
       HS <- getStr(indat[start],1)
+      if (HS %in% c("MCDA","Other")) # get filename
+         HSdetail <- removeEmpty(unlist(strsplit(indat[start+1],",")))
+      if (HS == "constantCatch")  # get constant inTAC
+         HSdetail <- as.numeric(removeEmpty(unlist(strsplit(indat[start+1],","))))
    }
    condition <- getsingleNum("CONDITION",indat)
    if (condition > 0) {
@@ -494,13 +498,14 @@ readctrlzone <- function(datadir,infile="control.csv") {
    condC <- list(histCatch=histCatch,histyr=histyr,
                  histCE=histCE,yearCE=yearCE,initdepl=initdepl,
                  compdat=compdat,Sel=NULL,SelWt=NULL)
-   projC <- list(projLML=projLML,HS=HS,projyrs=projyrs,Sel=NULL,SelWt=NULL)
+   projC <- list(projLML=projLML,HS=HS,HSdetail=HSdetail,projyrs=projyrs,
+                 Sel=NULL,SelWt=NULL)
    outctrl <- list(runlabel,datafile,batch,reps,withsigR,withsigB,
                    withsigCE,condition,projyrs)
[…]

### 2020-11-08 data-raw/bare-bones.r

@@ -1,48 +1,160 @@
 # Used in readme.Rmd to illustrate the running of aMSE
 
-# a minimal example
-starttime <- as.character(Sys.time())
+# setup run + resdir  --------------------------------------------------------
+starttime <- (Sys.time())
 library(aMSE)
 library(rutilsMH)
 library(makehtml)
+library(knitr)
 # Obviously you should modify the resdir to suit your own computer
-resdir <- "./../../A_code/aMSEUse/conddata/generic2"
+if (dir.exists("c:/Users/User/DropBox")) {
+  ddir <- "c:/Users/User/DropBox/A_code/"
+} else {
+  ddir <- "c:/Users/Malcolm/DropBox/A_code/"
+}
+resdir <- paste0(ddir,"aMSEUse/conddata/generic2")
 dirExists(resdir,make=TRUE,verbose=TRUE)
-# You now need to ensure that there is a control.csv, zone1sm\au2pop6.csv
-# and region1.csv file in the data directory
+
+# equilibrium zone -------------------------------------------------------------
+# You now need to ensure that there is, at least, a control.csv, zone1.csv
+# and region1.csv file in the data directory plus some other data .csv files
+# depending on how conditioned you want the model to be. Templates for the
+# correct format can be produced using ctrlfiletemplate(), datafiletemplate(),
+# and zonefiletemplate.
 zone1 <- readctrlzone(resdir,infile="control.csv")
 ctrl <- zone1$ctrl
 glb <- zone1$globals     # glb without the movement matrix
 constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
-
 #zone1$initLML <- 140
 out <- setupzone(constants,zone1) # make operating model
 zoneC <- out$zoneC
 zoneD <- out$zoneD
-glb <- out$glb        # glb now has the movement matrix
+glb <- out$glb             # glb now has the movement matrix
 product <- out$product     # important bits usually saved in resdir
 # did the larval dispersal level disturb the equilibrium?
+zoneD <- testequil(zoneC,zoneD,glb)
+zoneC <- resetexB0(zoneC,zoneD) # rescale exploitB to avexplB after dynamics
[…]

### 2020-11-08 data-raw/candidatefunctions.R

@@ -96,6 +96,76 @@ plotSAUdepl <- function(outB) {
   legend("topright",legend=c(11),col=c(6),lwd=3,bty="n",cex=1.0)
 } # end of plotSAUdepl
 
+cpue <- zoneDRp$cpue[1:10,1,1]
+years <- 1:10
+getgrad1(cpue)
+getgrad4(cpue)
+
+plotprep()
+parset()
+plot1(years,cpue,lwd=2)
+
+
+#' @title MCDAdata returns some conditioning cpue data from a file
+#'
+#' @description MCDAdata uses the projC$HS to identify the file to be
+#'     used to input the MCDA cpue data to be used to condition the MCDA
+#'
+#' @param resdir the results directory containing all input csv filers
+#' @param filen the projC$HS filename
+#' @param saunames the names of each SAU (from zone1$SAUnames)
+#'
+#' @return a matrix of cpue data used to condition the MCDA
+#' @export
+#'
+#' @examples
+#' print("wait on example data being available")
+MCDAdata = function(resdir,filen,saunames) {
+  # filen = projC$HS; resdir=resdir; saunames=zone1$SAUnames
+  nSAU <- length(saunames)
+  filename = filenametopath(resdir,filen)
+  if (file.exists(filename)) {
+      indat = readLines(filename)
+  } else {
+    label <- paste0(filename," does not exist in ",resdir," \n")
+    stop(label)
+  }
+  yrce <- getsingleNum("OPTCPUE",indat)
+  begin <- grep("OPTCPUE",indat)+1
+  optCE <- matrix(NA,nrow=yrce,ncol=nSAU)
+  yearCE <- numeric(yrce)
+  colnames(optCE) <- saunames
+  cenum <- rep(NA,(length(saunames) + 1))
+  for (i in 1:yrce) { # i=1
+    begin <- begin + 1
+    tmp <- unlist(strsplit(indat[begin],","))
+    pick <- grep("NA",tmp)
+    if (length(pick) > 0) {
+       cenum[-pick] <- as.numeric(tmp[-pick])
+    } else {
+       cenum <- as.numeric(tmp)
+    }
+    yearCE[i] <- cenum[1]
+    optCE[i,] <- cenum[2:(nSAU+1)]
+  }
+  rownames(optCE) <- yearCE
+  return(optCE)
+} # end of MCDAdata
+
+condcpue <- MCDAdata(resdir,projC$HS,zone1$SAUnames)
+
[…]

### 2020-11-08 README.Rmd
included new data files, ctrl, zone1, and constants with which to test the framework.

@@
 
-You can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:
+Once the MSE framework is made public you can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:
 
@@
 
-Alternatively, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio.
+Alternatively, while the development version remains private, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio.
 Open the New Project option in the project dialog at the top right of the RStudio screen and selection Version Control, then use 'https://github.com/haddonm/aMSE' in the top box, identify where you want the new directory put, and press return.
+ALternatively, you could download the zip file from inside the 'code' button and establish an R project from that.
 
@@
 # correct format can be produced using ctrlfiletemplate(), datafiletemplate(),
-# and zonefiletemplate.
-zone1 <- readctrlzone(resdir,infile="control.csv")
-ctrl <- zone1$ctrl
-glb <- zone1$globals     # glb without the movement matrix
-constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
+# and zonefiletemplate.  In the meantime it is easier to use the included 
+# data files. The original csv files are included in the data-raw directory
+data(ctrl)
+data(zone1)
+data(constants)
+ctrl$reps <- 250  # the original = 1000, we use 250 to use only 25% of the time
 #zone1$initLML <- 140

### 2020-11-22 README.Rmd
Working MCDA but now need data files to reflect current structure

--- 2020-10-14 README.Rmd
@@
 library(makehtml)
+library(knitr)
 # Obviously you should modify the resdir to suit your own computer
-resdir <- "./../../A_code/aMSEUse/out/run1"
+if (dir.exists("c:/Users/User/DropBox")) {
+  ddir <- "c:/Users/User/DropBox/A_code/"
+} else {
+  ddir <- "c:/Users/Malcolm/DropBox/A_code/"
+}
+resdir <- paste0(ddir,"aMSEUse/conddata/generic2")  # data and results directory
 dirExists(resdir,make=TRUE,verbose=TRUE)
-# The following function calls will ensure that there is a control.csv,
-# a zone1.csv, and a zone1sau2pop6.csv or population data file 
-# in the data directory, which in this case is the same as the resdir
-ctrlfiletemplate(resdir)
-zonefiletemplate(resdir)
-datafiletemplate(numpop=6,resdir,filename="zone1sau2pop6.csv")
+# equilibrium zone -------------------------------------------------------------
+# You now need to ensure that there is, at least, a control.csv, zone1.csv
+# and region1.csv file in the data directory plus some other data .csv files
+# depending on how conditioned you want the model to be. Templates for the
+# correct format can be produced using ctrlfiletemplate(), datafiletemplate(),
+# and zonefiletemplate.  In the meantime it is easier to use the included 
+# data files. The original csv files are included in the data-raw directory
 
-ctrl <- checkctrldat(resdir)
-runname <- ctrl$runlabel
-zone1 <- readzonefile(resdir,ctrl$zonefile)
-glb <- zone1$globals     # glb without the movement matrix
-constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
+# Many changes now require the data files to be updated and tested, but that will 
+# happen soon.
 
-out <- setupzone(constants,zone1) # make operating model
-zoneC <- out$zoneC
-zoneD <- out$zoneD
-glb <- out$glb        # glb now has the movement matrix
-product <- out$product     # important bits usually saved in resdir
-          # did the larval dispersal level disturb the equilibrium?
-regDe <- testequil(zoneC,zoneD,glb)
-
-resfile <- setuphtml(resdir)# prepare to save and log results
-
-plotproductivity(resdir,product,glb)
-biology_plots(resdir, glb, zoneC)
-numbersatsize(resdir, glb, zoneD)
-
-endtime <- as.character(Sys.time())
-
-reportlist <- list(runname=runname,
-                   starttime=starttime,endtime=endtime,
-                   zoneC=zoneC, zoneD=zoneD, product=product,
-                   glb=glb,constants=constants)
-
-runnotes <- "This is a bare-bones example."
-# If you unhash the make_html call it will generate a local website 
-# inside resdir and open it so you can see the results so far.
-# make_html(replist=reportlist,resdir=resdir,width=500,
-#          openfile=TRUE,runnotes=runnotes,verbose=FALSE,
-#          packagename="aMSE",htmlname="testrun")
 ```

### 2020-11-23 README.Rmd
New data files, and have added a working example in the Readme.

@@
 }
-resdir <- paste0(ddir,"aMSEUse/conddata/generic2")  # data and results directory
+resdir <- paste0(ddir,"aMSEUse/conddata/generic")  # data and results directory
 dirExists(resdir,make=TRUE,verbose=TRUE)
 # equilibrium zone -------------------------------------------------------------
-# You now need to ensure that there is, at least, a control.csv, zone1.csv
-# and region1.csv file in the data directory plus some other data .csv files
+# You now need to ensure that there is, at least, a control.csv, and a 
+# constantsdata.csv file in the data directory plus some other data .csv files
 # depending on how conditioned you want the model to be. Templates for the
-# correct format can be produced using ctrlfiletemplate(), datafiletemplate(),
-# and zonefiletemplate.  In the meantime it is easier to use the included 
-# data files. The original csv files are included in the data-raw directory
+# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
+# In the meantime it is easier to use the included data file "zone".
+data(zone)
+# Of course, usually one would use data files, control.csv and a zone.csv, which is 
+# listed as the datafile within the control.csv. 
+# zone <- makeequilzone(resdir,"control2.csv") # normally would read in a file
+    equiltime <- (Sys.time())   # let's change the initial depletion
+    origdepl <-  c(0.30,0.31,0.29,0.32,0.30,0.31,0.29,0.32) 
+zoneDD <- depleteSAU(zone$zoneC,zone$zoneD,zone$glb,origdepl,zone$product,len=12)
+out <- prepareprojection(zone$zone1,zone$zoneC,zone$glb,zoneDD,zone$ctrl)
+  zoneDR <- out$zoneDP
+  projC <- out$projC
+  zoneCP <- out$zoneC
+  midtime <- (Sys.time())
 
-# Many changes now require the data files to be updated and tested, but that will 
-# happen soon.
+  glb <- zone$glb
+  ctrl <- zone$ctrl
+  print(equiltime - starttime)
+  print(midtime - equiltime)
+  propD <- getzoneprops(zone$zoneC,zoneDD,glb,year=1)
+  # round(propD,3)
+# Do the replicates ------------------------------------------------------------
+  inityr <- zone$zone1$projC$inityrs
+  saunames <- zone$zone1$SAUnames
+  sauindex <- zone$glb$sauindex
+  pyrs <- projC$projyrs + projC$inityr
+  B0 <- tapply(sapply(zone$zoneC,"[[","B0"),sauindex,sum)
+  exB0 <- tapply(sapply(zone$zoneC,"[[","ExB0"),sauindex,sum)
+
+  midtime <- (Sys.time())
+mseproj <- applymcda(zoneCP,zoneDR,glb,ctrl,projC$projyrs,projC$inityrs)
+sauzoneDP <- asSAU(mseproj,sauindex,saunames,B0,exB0)
+  endtime <- (Sys.time())
+  print(endtime - midtime)
+# Now plot some results
+  plotC <- function(nsau,saunames,reps,projyrs,plts=c(4,2)) {
+    return(list(nsau=nsau,saunames=saunames,reps=reps,projyrs=projyrs,plts=plts))
+  }
+
+  saunames <- zone$zone1$SAUnames
+  pyrs <- projC$projyrs + projC$inityr
+  reps <- ctrl$reps
+  label <- "Catches t"
+  yrs <- 1:pyrs
+
+plotconst <- plotC(nsau=length(saunames),saunames,reps,pyrs,plts=c(4,2))
+
+plotprep(width=7,height=8,newdev=FALSE)
+plotproj(sauzoneDP$cpue,"CPUE",plotconst,vline=projC$inityrs,addqnts=TRUE)
+
+plotproj(sauzoneDP$saudeplsB,"Spawning Biomass Depletion",plotconst,vline=projC$inityrs,addqnts=TRUE)
+
+plotproj(sauzoneDP$recS/1000,"Recruitment '000s",plotconst,vline=projC$inityrs,addqnts=TRUE)
+
+plotproj(sauzoneDP$catS,"Catches t",plotconst,vline=projC$inityrs,addqnts=TRUE)
 

### 2020-11-23 data-raw/MakeDatafor_aMSE.r

@@ -67,7 +67,7 @@ save(product,file=paste0(datadir,"product.RData"))
 
 # some cpue data -----------------------------------------------------
 
-datadir <- "./../../rcode2/aMSE/data-raw/"
+datadir <- "./../../A_code/aMSE/data-raw/"
 
 blockE13 <- read.csv(paste0(datadir,"block13e.csv"),header=TRUE)
 

### 2020-11-24 README.Rmd
Started the_data_files.Rmd vignette

@@
 
-This is a basic example which illustrates the generation of an initial equilibrium.
-It uses built in data-sets but usually you would generate zoneC (constants) and zoneD (the dynamic components) from the conditioning data:
+This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 100 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000).
+It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.
 

### 2020-12-03 R/inputfiles.R
Worked on vignettes and a few functions to simplify implementation.

@@ -76,7 +76,9 @@ ctrlzonetemplate <- function(indir,filename="control.csv") { # indir=resdir; fil
        file=filename,append=TRUE)
    cat("\n",file=filename,append=TRUE)
    cat("RANDOM \n",file=filename,append=TRUE)
-   cat("randomseed, 0, for repeatability of results, alter for each separate run \n",
+   cat("randomseed, 4024136, for repeatability of population definitions set to 0 otherwise \n",
+       file=filename,append=TRUE)
+   cat("randomseedP, 7999647, for repeatability of projections set to 0 otherwise \n",
        file=filename,append=TRUE)
    cat("\n",file=filename,append=TRUE)
    cat("initLML, 140, the initial LML for generating the unfished zone \n",
@@ -431,6 +433,7 @@ readctrlzone <- function(datadir,infile="control.csv") {
    midpts <- seq(minc,minc+((Nclass-1)*cw),2)
    larvdisp <- getsingleNum("larvdisp",indat)
    randomseed <- getsingleNum("randomseed",indat)
+   randomseedP <- getsingleNum("randomseedP",indat)
    initLML <- getsingleNum("initLML",indat)
    projyrs <- getsingleNum("PROJECT",indat)
    firstyear <- getsingleNum("firstyear",indat)
@@ -505,10 +508,11 @@ readctrlzone <- function(datadir,infile="control.csv") {
                  compdat=compdat,Sel=NULL,SelWt=NULL)
    projC <- list(projLML=projLML,HS=HS,HSdetail=HSdetail,projyrs=projyrs,
                  inityrs=inityrs,Sel=NULL,SelWt=NULL)
-   outctrl <- list(runlabel,datafile,batch,reps,randomseed,withsigR,withsigB,
-                   withsigCE,condition,projyrs)
-   names(outctrl) <- c("runlabel","datafile","batch","reps","randseed","withsigR",
-                       "withsigB","withsigCE","condition","projection")
+   outctrl <- list(runlabel,datafile,batch,reps,randomseed,randomseedP,
+                   withsigR,withsigB,withsigCE,condition,projyrs)
+   names(outctrl) <- c("runlabel","datafile","batch","reps","randseed",
+                       "randseedP","withsigR","withsigB","withsigCE",
+                       "condition","projection")
    globals <- list(numpop=numpop, nSAU=nSAU, midpts=midpts,
[…]

### 2020-12-10 README.Rmd
working on adding alternativeHS.

@@
 # Of course, usually one would use data files, control.csv and a zone.csv, which is 
-# listed as the datafile within the control.csv. 
+# listed as the datafile within the control.csv. These must be stored in resdir
 # zone <- makeequilzone(resdir,"control2.csv") # normally would read in a file
@@
   midtime <- (Sys.time())
-mseproj <- applymcda(zoneCP,zoneDR,glb,ctrl,projC$projyrs,projC$inityrs)
+  applyHS <- mcdahcr   # name the HS/HCR to be used
+mseproj <- doprojection(zoneCP,zoneDR,glb,ctrl,projC$projyrs,projC$inityrs)
 sauzoneDP <- asSAU(mseproj,sauindex,saunames,B0,exB0)

### 2020-12-10 R/HS-HCR.R

@@ -1,106 +1,4 @@
 
-#' @title applymcda applies the Tasmanian MCDA to a simulated zone
-#'
-#' @description applymcda applies the Tasmanian MCDA to a simulated zone. It
-#'     is still in need of optimization as 1000 iterations would currently take
-#'     about 20 minutes.
-#'
-#' @param zoneCP the modified zoneC object ready for projections
-#' @param zoneDP the modified zoneD object ready for projections
-#' @param glob the global variables object
-#' @param ctrl the ctrl object
-#' @param projyrs the number of years of projection
-#' @param inityrs the number of years kept from the period of conditioning,
-#'     defult=10
-#' @param wid the number of years for the gradient PM, default=4
-#' @param targqnt the target quantile for the cpue target, default=0.55
-#' @param pmwts the relative weights given to the PMs when generating the score,
-#'     default = c(0.65, 0.25, 0.1) targce, grad4, grad1
-#' @param hcr the TAC multiplier for each score default=c(0.25,0.75,0.8,0.85,
-#'     0.9,1,1.05,1.1,1.15,1.2)
-#'
-#' @return the filled in zoneDP list
-#' @export
-#'
-#' @examples
-#' print("wait on data files")
-#' #  zoneCP=zoneCP;zoneDP=zoneDR;glob=glb;ctrl=ctrl;projyrs=projC$projyrs;inityrs=projC$inityrs
-#' # wid=4;targqnt=0.55;pmwts=c(0.65, 0.25,0.1);hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2)
-applymcda <- function(zoneCP,zoneDP,glob,ctrl,projyrs,inityrs=10,wid = 4,
-                      targqnt = 0.55, pmwts = c(0.65, 0.25,0.1),
-                      hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2)) {
-  sigmar <- ctrl$withsigR # needed to add recruitment variation
-  npop <- glob$numpop
-  nsau <- glob$nSAU
-  Ncl <- glob$Nclass
-  nyrs <- projyrs + inityrs
-  movem <- glob$move
-  reps <- ctrl$reps
[…]

### 2020-12-10 R/inputfiles.R

@@ -24,7 +24,7 @@
 #' # changes needed now I have simplified to two files and changed their formats
 #'  yourdir <- tempdir()
 #'  ctrlzonetemplate(yourdir,filename="testctrl.csv")   #
-#'  control <- readctrlzone(yourdir,"testctrl.csv")
+#'  control <- readctrlfile(yourdir,"testctrl.csv")
 #'  str(control,max.level=1)
 #' }
 ctrlzonetemplate <- function(indir,filename="control.csv") { # indir=resdir; filename="control2.csv"
@@ -375,9 +375,9 @@ read_conddata <- function(filename) {  # filename=filen
 } # end of read_condata
 
 
-#' @title readctrlzone checks datadir contains the required csv files
+#' @title readctrlfile checks datadir contains the required csv files
 #'
-#' @description readctrlzone checks datadir contains the required csv
+#' @description readctrlfile checks datadir contains the required csv
 #'     files including the named control file, which then contains
 #'     the names of the region data file, and the population data
 #'     file. The run stops if any are not present or are misnamed.
@@ -397,10 +397,10 @@ read_conddata <- function(filename) {  # filename=filen
 #' resdir <- tempdir()
 #' ctrlzonetemplate(resdir)
 #' datafiletemplate(6,resdir,filename="zone1sau2pop6.csv")
-#' ctrl <- readctrlzone(resdir)
+#' ctrl <- readctrlfile(resdir)
 #' ctrl
 #' }
-readctrlzone <- function(datadir,infile="control.csv") {
+readctrlfile <- function(datadir,infile="control.csv") {
    # datadir=resdir; infile="control2.csv"
    filenames <- dir(datadir)
    if (length(grep(infile,filenames)) != 1)

### 2020-12-24 README.Rmd
added explicit mentio nof applyHS to doprojection  No EWN

@@
   midtime <- (Sys.time())
-  applyHS <- mcdahcr   # name the HS/HCR to be used
-mseproj <- doprojection(zoneCP,zoneDR,glb,ctrl,projC$projyrs,projC$inityrs)
+mseproj <- doprojection(zoneCP,zoneDR,glb,ctrl,projC$projyrs,applyHS=mcdahcr,projC$inityrs)
 sauzoneDP <- asSAU(mseproj,sauindex,saunames,B0,exB0)

### 2020-12-24 data-raw/candidatefunctions.R

@@ -352,10 +352,7 @@ readctrlfile2 <- function(datadir,infile="control.csv") {
                      "initLML","condC","projC","globals","ctrl",
                      "catchyrs","projyrs")
   return(totans)
-} # end of readctrlzone
-
-
-
+} # end of readctrlfile2
 
 
 

### 2021-03-11 README.Rmd
Modified resdir to rundir and included option to vary final years of recruitment in the historicasl conditioning.

--- 2020-11-24 README.Rmd
@@
 library(knitr)
-# Obviously you should modify the resdir to suit your own computer
+# Obviously you should modify the rundir to suit your own computer
 if (dir.exists("c:/Users/User/DropBox")) {
@@
 }
-resdir <- paste0(ddir,"aMSEUse/conddata/generic")  # data and results directory
-dirExists(resdir,make=TRUE,verbose=TRUE)
+rundir <- paste0(ddir,"aMSEUse/conddata/generic")  # data and results directory
+dirExists(rundir,make=TRUE,verbose=TRUE)
 # equilibrium zone -------------------------------------------------------------
@@
 # correct format can be produced using ctrlfiletemplate(), datafiletemplate().
-# In the meantime it is easier to use the included data file "zone".
-data(zone)
-# Of course, usually one would use data files, control.csv and a zone.csv, which is 
-# listed as the datafile within the control.csv. 
-# zone <- makeequilzone(resdir,"control2.csv") # normally would read in a file
-    equiltime <- (Sys.time())   # let's change the initial depletion
-    origdepl <-  c(0.30,0.31,0.29,0.32,0.30,0.31,0.29,0.32) 
-zoneDD <- depleteSAU(zone$zoneC,zone$zoneD,zone$glb,origdepl,zone$product,len=12)
-out <- prepareprojection(zone$zone1,zone$zoneC,zone$glb,zoneDD,zone$ctrl)
-  zoneDR <- out$zoneDP
-  projC <- out$projC
-  zoneCP <- out$zoneC
-  midtime <- (Sys.time())
+# 
+# Of course, usually one would use data files, control.csv and a zone.csv, which
+# is listed as the datafile within the control.csv. These must be stored in 
+# rundir. There are example control and data files in the DropBox folder:
+# C:\Users\User\Dropbox\National abalone MSE\aMSE_files. Copy the control2.csv
+# and zonewest.csv into you rundir. Then, assuming yo have the very latest
+# version of aMSE = 5300, the following code should work. Recently both aMSE and
+# makehtml have been altered so that all references to resdir have been changed
+# to rundir, so both packages will need updating. See their respective GitHub
+# readme pages for details.
 
+# TasmanianHS.R should be in rundir, but during development is in data-raw
+source(paste0(ddir,"aMSE/data-raw/","TasmanianHS.R"))
+# generate equilibrium zone ----------------------------------------------------
+starttime2 <- (Sys.time())
+zone <- makeequilzone(rundir,"control2.csv",cleanslate = TRUE) # normally would read in a file
+equiltime <- (Sys.time()); print(equiltime - starttime2)
+# declare main objects ---------------------------------------------------------
 glb <- zone$glb
 ctrl <- zone$ctrl
-  print(equiltime - starttime)
-  print(midtime - equiltime)
-  propD <- getzoneprops(zone$zoneC,zoneDD,glb,year=1)
-  # round(propD,3)
-# Do the replicates ------------------------------------------------------------
-  inityr <- zone$zone1$projC$inityrs
-  saunames <- zone$zone1$SAUnames
-  sauindex <- zone$glb$sauindex
-  pyrs <- projC$projyrs + projC$inityr
-  B0 <- tapply(sapply(zone$zoneC,"[[","B0"),sauindex,sum)
-  exB0 <- tapply(sapply(zone$zoneC,"[[","ExB0"),sauindex,sum)
+zone1 <- zone$zone1
+projC <- zone1$projC
+condC <- zone1$condC
+zoneC <- zone$zoneC
+zoneD <- zone$zoneD
+product <- zone$product
+# add some tables and plots into rundir
+biology_plots(rundir, glb, zoneC)
+plotproductivity(rundir,product,glb)
+numbersatsize(rundir, glb, zoneD)
+# Condition on Fishery ---------------------------------------------------------
+zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,sigR=1e-08,sigB=1e-08)
+# save conditioned results -----------------------------------------------------
+save(zoneDD,file=filenametopath(rundir,"zoneDD.RData"))
+# Illustrate productivity
+propD <- getzoneprops(zoneC,zoneDD,glb,year=47)
+addtable(round(t(propD),4),"propertyDD.csv",rundir,category="zoneDD",caption=
+           "Properties of zoneD after conditioning on historical catches.")
+addtable(round(t(zoneDD$harvestR[45:47,]),4),"final_harvestR.csv",rundir,
+         category="zoneDD",caption="Last three years of harvest rate.")
+popdefs <- getlistvar(zone$zoneC,"popdef")
+addtable(round(t(popdefs),3),"popdefs.csv",rundir,category="zoneDD",caption=
+           "Population specific definitions")
+# Prepare projections ----------------------------------------------------------
+cmcda <- mcdahcr(arrce=condC$histCE,hsargs=hsargs,
+                 yearnames=rownames(condC$histCE),saunames=glb$saunames)
+pms <- cmcda$pms
+multTAC <- cmcda$multTAC
+#out <- prepareprojection(projC,zoneC,glb,zoneDD,ctrl,multTAC)
+out <- prepareprojectionnew(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
+                            zoneDD=zoneDD,ctrl=ctrl,varyrs=6,
+                            multTAC=multTAC,lastsigR = 0.25)
 
-  midtime <- (Sys.time())
-mseproj <- applymcda(zoneCP,zoneDR,glb,ctrl,projC$projyrs,projC$inityrs)
-sauzoneDP <- asSAU(mseproj,sauindex,saunames,B0,exB0)
-  endtime <- (Sys.time())
-  print(endtime - midtime)
-# Now plot some results
-  plotC <- function(nsau,saunames,reps,projyrs,plts=c(4,2)) {
-    return(list(nsau=nsau,saunames=saunames,reps=reps,projyrs=projyrs,plts=plts))
-  }
+zoneDP <- out$zoneDP
+projC <- out$projC
+zoneCP <- out$zoneCP
 
-  saunames <- zone$zone1$SAUnames
-  pyrs <- projC$projyrs + projC$inityr
-  reps <- ctrl$reps
-  label <- "Catches t"
-  yrs <- 1:pyrs
+zoneDP <- doTASprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,mcdahcr,hsargs)
+# save more plots to rundir ----------------------------------------------------
+out <- plotbysau(zoneDP,glb,rundir)
 
-plotconst <- plotC(nsau=length(saunames),saunames,reps,pyrs,plts=c(4,2))
+projtime <- Sys.time()
+print(projtime - starttime)
 
-plotprep(width=7,height=8,newdev=FALSE)
-plotproj(sauzoneDP$cpue,"CPUE",plotconst,vline=projC$inityrs,addqnts=TRUE)
+# make results webpage ---------------------------------------------------------
+replist <- list(starttime=as.character(starttime),endtime=as.character(projtime))
 
-plotproj(sauzoneDP$saudeplsB,"Spawning Biomass Depletion",plotconst,vline=projC$inityrs,addqnts=TRUE)
-
-plotproj(sauzoneDP$recS/1000,"Recruitment '000s",plotconst,vline=projC$inityrs,addqnts=TRUE)
-
-plotproj(sauzoneDP$catS,"Catches t",plotconst,vline=projC$inityrs,addqnts=TRUE)
+# Unhash the make_html command to generate the results webspage
+# make_html(
+#   replist = replist,
+#   rundir = rundir,
+#   width = 500,
+#   openfile = TRUE,
+#   runnotes = NULL,
+#   verbose = FALSE,
+#   packagename = "aMSE",
+#   htmlname = "aMSE"
+# )
 

### 2021-03-11 R/dynamics.R

@@ -193,7 +193,7 @@ initiateHS <- function(zoneDP,zoneCP,exb,inN,acatch,sigmar,sigmab,glb) {
                         Ncl=glb$Nclass,sauindex=glb$sauindex,movem=glb$move,
                         sigmar=sigmar,sigmab=sigmab)
     dyn <- outy$dyn
-    saudyn <- popcetosauce(dyn["catch",],dyn["cpue",],glb$sauindex)
+    saudyn <- poptosauCE(dyn["catch",],dyn["cpue",],glb$sauindex)
     zoneDP$exploitB[1,,iter] <- dyn["exploitb",]
     zoneDP$matureB[1,,iter] <- dyn["matureb",]
     zoneDP$catch[1,,iter] <- dyn["catch",]

### 2021-03-11 R/generate_results.R

@@ -9,7 +9,7 @@
 #'     addition, it also tabulates the biological properties of each
 #'     population and SAU and the total zone
 #'
-#' @param resdir the results directory
+#' @param rundir the results directory
 #' @param glb the globals list
 #' @param zoneC the zonal constants by population, zoneC
 #'
@@ -18,7 +18,7 @@
 #'
 #' @examples
 #' print("this will be quite long when I get to it")
-biology_plots <- function(resdir, glb, zoneC) {
+biology_plots <- function(rundir, glb, zoneC) {
   mids <- glb$midpts
   numpop <- glb$numpop
   popdef <- getlistvar(zoneC,"popdef")
@@ -28,7 +28,7 @@ biology_plots <- function(resdir, glb, zoneC) {
   # maturation uses zoneC
   matur <- getlistvar(zoneC,"Maturity")
   rownames(matur) <- mids
-  filen <- filenametopath(resdir,"maturity_v_Length_pop.png")
+  filen <- filenametopath(rundir,"maturity_v_Length_pop.png")
   plotprep(width=7,height=4,newdev=FALSE,filename=filen,cex=0.9,
            verbose=FALSE)
   plot(mids,matur[,1],type="l",lwd=2,xlab="Shell Length mm",
@@ -37,12 +37,12 @@ biology_plots <- function(resdir, glb, zoneC) {
   legend("topright",paste0("P",1:numpop),lwd=3,col=c(1:numpop),bty="n",
          cex=1.2)
   caption <- "The maturity vs length for each population."
-  addplot(filen,resdir=resdir,category="Biology",caption)
+  addplot(filen,rundir=rundir,category="Biology",caption)
 
   # weight-at-length using zoneC--------------------------------------
   WtL <- getlistvar(zoneC,"WtL")
   rownames(WtL) <- mids
-  filen <- filenametopath(resdir,"Weight_at_Length_pop.png")
+  filen <- filenametopath(rundir,"Weight_at_Length_pop.png")
   plotprep(width=7,height=4,newdev=FALSE,filename=filen,cex=0.9,
            verbose=FALSE)
   plot(mids,WtL[,1],type="l",lwd=2,xlab="Shell Length mm",
@@ -51,12 +51,12 @@ biology_plots <- function(resdir, glb, zoneC) {
   legend("topleft",paste0("P",1:numpop),lwd=3,col=c(1:numpop),bty="n",cex=1.2)
   caption <- paste0("The weight-at-length for each population. ",
[…]

### 2021-03-11 data-raw/MakeDatafor_aMSE.r

@@ -14,8 +14,8 @@ if (dir.exists("c:/Users/User/DropBox")) {
 } else {
   ddir <- "c:/Users/Malcolm/DropBox/A_code/"
 }
-resdir <- paste0(ddir,"aMSEUse/conddata/generic2")
-dirExists(resdir,make=TRUE,verbose=TRUE)
+rundir <- paste0(ddir,"aMSEUse/conddata/generic2")
+dirExists(rundir,make=TRUE,verbose=TRUE)
 
 # equilibrium zone -------------------------------------------------------------
 # You now need to ensure that there is, at least, a control.csv, zone1.csv
@@ -23,9 +23,9 @@ dirExists(resdir,make=TRUE,verbose=TRUE)
 # depending on how conditioned you want the model to be. Templates for the
 # correct format can be produced using ctrlfiletemplate(), datafiletemplate(),
 # and zonefiletemplate.
-zone1 <- readctrlfile(resdir,infile="control.csv")
+zone1 <- readctrlfile(rundir,infile="control.csv")
 ctrl <- zone1$ctrl
-constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
+constants <- readdatafile(glb$numpop,rundir,ctrl$datafile)
 
 datadir <- "C:/Users/User/Dropbox/A_Code/aMSE/data/"
 
@@ -42,15 +42,15 @@ library(microbenchmark)
 
 datadir <- "./../../rcode2/aMSE/data-raw/"
 # read data files ----------------------------------------------------
-resdir <- "./../../rcode2/aMSEUse/out/run1"
-dirExists(resdir,make=TRUE,verbose=TRUE)
+rundir <- "./../../rcode2/aMSEUse/out/run1"
+dirExists(rundir,make=TRUE,verbose=TRUE)
 # You now need to ensure that there is a control.csv, zone1sau2pop6.csv
 # and zone1.csv file in the data directory
-ctrl <- checkctrldat(resdir)
+ctrl <- checkctrldat(rundir)
 runname <- ctrl$runlabel
-zone1 <- readzonefile(resdir,ctrl$zonefile)
+zone1 <- readzonefile(rundir,ctrl$zonefile)
 glb <- zone1$globals
-constants <- readdatafile(glb$numpop,resdir,ctrl$datafile)
+constants <- readdatafile(glb$numpop,rundir,ctrl$datafile)
 
 out <- setupzone(constants, zone1)
 zoneC <- out$zoneC

### 2021-03-16 README.Rmd
Added zonetosau for summary plots. MSE functioning correctly.

@@
 out <- prepareprojectionnew(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
-                            zoneDD=zoneDD,ctrl=ctrl,varyrs=6,
+                            zoneDD=zoneDD,ctrl=ctrl,varyrs=8,
                             multTAC=multTAC,lastsigR = 0.25)
@@
 zoneCP <- out$zoneCP
+zoneDDR <- out$zoneDDR
 
@@
 
+After running the whole, even if you do not generate the results webpage, you could also try:
+
+```{r combined_plot, fig.height=7, fig.width=6 }
+B0 <- getvar(zoneC,"B0")
+ExB0 <- getvar(zoneC,"ExB0")
+
+zoneDsau <- zonetosau(zoneDDR,glb,B0,ExB0)
+zonePsau <- zonetosau(zoneDP,glb,B0,ExB0)
+
+nsau <- glb$nSAU
+label <- glb$saunames
+CIprobs <- c(0.05,0.5,0.95)
+sauCI <- vector("list",nsau)
+postyrs <- ctrl$projection
+startyr <- 40
+
+invar <- "cpue"
+prerep <- zoneDsau[[invar]]
+allrep <- zonePsau[[invar]]
+preyrs <- dim(prerep)[1]
+allyrs <- preyrs + postyrs
+reps <- dim(prerep)[3]
+
+plotprep(width=8, height=8,newdev=FALSE)
+parset(plots=c(4,2))
+for (sau in 1:nsau) {
+  ymax <- getmax(rbind(prerep[startyr:preyrs,sau,],allrep[,sau,]))
+  traj <- c(prerep[startyr:preyrs,sau,1],allrep[,sau,1])
+  plot(startyr:allyrs,traj,type="l",lwd=1,col="grey",panel.first=grid(),
+       ylab=paste0(invar,"  ",label[sau]),ylim=c(0,ymax))
+  for (i in 2:reps)
+    lines(startyr:allyrs,c(prerep[startyr:preyrs,sau,i],allrep[,sau,i]),lwd=1,col="grey")
+  CI <- apply(allrep[,sau,],1,quantile,probs=CIprobs)
+  lines((preyrs+1):allyrs,CI[2,],lwd=2,col=4)
+  lines((preyrs+1):allyrs,CI[1,],lwd=1,col=2)
+  lines((preyrs+1):allyrs,CI[3,],lwd=1,col=2)
+  sauCI[[sau]] <- CI
+  abline(v=(preyrs+0.5),col=2)
+ if (invar == "cpue") abline(h=150,col=2)
+}
+
+```
+
 See the vignette Running_aMSE.Rmd for a more detailed example.

### 2021-03-16 data-raw/oldnews.rmd
Added zonetosau for summary plots. MSE functioning correctly.

---
title: "Untitled"
author: "Malcolm Haddon"
date: "17/03/2021"
output: html_document
---

5

* Currently can generate a full region, characterize its productivity, and if run with no catch or variability maintains an equilibrium, even in the presence of larval drift.
6
5
7

* 16-04-2020 worked on _setuphtml_ and _addfilename_ to simplify the use of make-html when adding a plot or table to the results for a simulation.
8
6
9

* 17-04-2020 0.0.0.8920: Cleaned up the make_html code to produce valid HTML5 and CSS3, separated the CSS code into a separate file that is linked into the different html files. Cleaned up the imports and fixed a bunch of the examples, so that there are now no errors, no warnings, and no notes. Although the examples for the read[ctrl][data][region1] functions all currently have a 'dontrun' status. An issue to fix.

7
* 17/03/2021 0.0.0.5000 Well behind in this documentation. The MSE now operates successfully, with the README page on GitHub providing a worked example. The latest versions of aMSE, makehtml, and rutilsMH, as tar.gz source files along with a control2.csv and westzone.csv data files, and a first-run.R file are to be found in National abalone MSE/aMSE_files. Currently, the development of additional functions to generate further summary plots and tables are still being worked on. And work towards correctly conditioning the Operating Model onto the Tasmanian west coast is underway.
10
8
11

* 01-05-2020 0.0.0.8600: Continued with the development of Running_aMSE.Rmd. This has the great advantage of forcing me to clarify the text and the relationships between the R functions and the tasks they do. It even suggests ways to improve on the names of the functions. Naming things continues to be one of the harder things (if I want to keep these things meaningful, nothing wrong with a1, a2, x1, x2, etc). Have discovered how to include some CSS code to customize the vignette format. But, more importantly, I have improved the text, changed _addfilename_ to _logfilename_ and automated recording the file type (which is used by make_html to process the file correctly on the local webpage).
12

13

* 10-05-2020: 0.0.0.8500 Continued the development of Running_aMSE.Rmd. Clarified the intent of this document. Cleaned up the makehtml code and added the dodepletion section. Now need to implement at least one HCR function and associated file.
14

15

* 11-05-2020 0.0.0.8450 Put the results generation into internal functions so their production can be automated as new ones are developed. Ensured all internal functions were operating consistently with regard the various name changes I have made. e.g. fixed the emergence calculations, which was really messing with the results! Set up a better worked example in the readme in the github version so it now produces plots and tables and can generate the local website should the user wish.
16

17

* 12-05-2020 0.0.0.8400 Simplified the intended directory structure to a single results directory _resdir_. Cleaned out old, now unused code, and tidied the help files.
18

19

* 15-05-2020 0.0.0.8350 Started including estimates relating to SMUs as well as populations and whole of region. This entailed making changes to the design of regionD, to ensure internal consistency and ease of use. Began adding details of the fleet dynamics to Running-aMSE.Rmd.
20

21

* 17-05-2020 0.0.0.8300 Some big changes. I have now put a larval dispersal movement matrix into the global glb object. This is now applied to the predicted recruitments before they settle so that the equilibrium and dynamic conditions now take this potential movement into account within the usual dynamics. All other examples and functions now have been modified to account for this. In addition, I have included a src directory to contain the Rcpp and RcppArmadillo code, which I hoped to use to speed things up. However, potential speedups with inverting matrices and multiplying vectors by matrices appear to provide no benefit or make things worse! Further explorations are needed. However, I have introduced two small functions to replace the use of sapply when accessing components of regionC, and this has sped thing up by about 10%. A strange error where the example code from the vignettes and their output are put into files in the aMSE.check directory and then that is all deleted but 'check' finds these files and complains, even though it put them there? Again, further investigation required. 
22

23

* 18-05-2020 0.0.0.8250 Now implemented the movement matrix into oneyearC and oneyearcat, which use catches rather than harvest rates to impose a fishery onto the simulation. The larval dispersal matrix is now applied after the calculation of the expected recruitment levels using _oneyearrec_. That way if no fishing is applied everything remains static (when there is no variation on anything). But now, we can apply catches rather than harvest rates so all harvest control rule outputs (in catches) will be valid, even when they fail miserably. The check note about unknown directories is continuing and I think it is due to my explicit use of tempdir in a few examples, especially in makehtml.R. I will keep looking into it, even though it does not affect functionality in any way.
24

25

* 22-05-2020 0.0.0.8200 Added a fishery dataset from East coast block 13 that includes cpue, catch, and effort. This is for use with exploring the operation of the estimators of the performance measures. I have also now included two new get functions (among others) one to calculate the grad1 values and then one to calculate the Tasmanian harvest strategy gradient1 scores (as yet unweighted).
26

27

* 25-05-2020 0.0.0.8150 Added getgrad4 and modified getgrad1 and getscore1 on the path to having an experimental harvest control rule that approximates the Tasmanian HS. Started a Exploring_Tas_HS.Rmd file.
28

29

* 26-05-2020 0.0.0.8100 Modified getgrad4 to match current Tasmanian practice (divide each cpue chunk by the cpue in the first year of the chunk; basically I forgot but will document it thoroughly), changed getscore1 to getscore as it now works for both grad1 and grad4, and will work for gradX. Now need to add a targCE and a way to combine them in a weighted fashion and the HS and HCR are ready to go.
30

31

* 10-06-2020 0.0.0.8000 Altered all cases of region and reg into zone. Also altered all cases of SMU and smu to SAU and sau. So the data files have been altered accordingly.
32

33

* 10/08/2020 0.0.0.7950 Added to some of the vignettes and made minor improvements to some functions.
34

35

* 01/09/2020 0.0.0.7900 Added more to the Running the MSE vignette, which included the addition of a getzoneprod to summarize the productivity of the whole zone.
36

37

* 02/09/2020 0.0.0.7850 Modified generate_results.r bare-bones.r, devel.r, and README.Rmd to reflect changes I have made both in makehtml and in aMSE.
38

39

* 04/09/2020 0.0.0.7800 Have now stabilized makehtml and the places in aMSE where it is used (generate_results.R). Now it behaves as it should and it will be simple to add further plots and tables to the characterization.
40

41

* 13/09/2020 0.0.0.7750 Have removed 'diagrams' from the requirements of running the MSE and started adding functions and data-sets to faciliate the full blown operating model conditioning.
42

43

* 22/09/2020 0.0.0.7700 Have started adding files to begin conditioning the operating model on teh west coast data. This is designed to bring out the issues and problems raised when including such things insode the package. So far this has entailed changes to the readfiles for the control, zone, and data files. This probably means I will need now to fix the bare-bones example in teh readme.md file on github.
44

45

* 29/09/2020 0.0.0.7650 Have added functions to assist with the conditioning of the OM. Need to update the vignette. All examples should work, but some of them too need modification. 
46

47

* 01/10/2020 0.0.0.7600 Am working on conditioning the model and am now using the observed CPUE in an attempt to match it to the predicted CPUE from teh OM after initial depletion levels are set and the historical catches area applied. I have added a new candidatefunctions.R file to the data-raw folder and will expect to make changes pretty much every day, some small some large.
48

49

* 09/10/2020 0.0.0.7550 Have begun making changes to the characterization of zones, both unfished and depleted. Making functions internally consistent with the requirements of getting an MSZE up and running. Added more characterization functions.
50

51

* 11/10/2020 0.0.0.7500 Have transferred work to my travel laptop. Today have started to implement setting up the harvest strategy to be used in the projections after having generated a generic MSE operating model. This entailed modifying or adding to the zone definition .csv file as well as the readzonefile function and others to ensure internal consistency. I may eventually scrap the separation of the control file and the zone files and have it all in the control file, which currently has little remaining.
52

53

* 14/10/2020 0.0.0.7450 Have started to iron out inconsistencies between different ways of running the MSE. It needs an initial LML in case there is no conditioning and no projection planned. The initial LML will influecen estimates of exlpoitable biomass, MSY, and other aspects of production. So have modified the template generating files and the read data file functions.
54

55

* 26/10/2020 0.0.0.7400 Have altered the input control file to contain the zone wide information that was previously read in seperately. This entailed changing the ctrlfiletemplate and the readctrlzone functions. I have also been modifying the functions that assist with the automatic documentation of functions and the aMSE functions they reference. Not surprisingly, the makezone and oneyear related functions are referencing the most aMSE functions, while many others reference none.
56

57

* 01/11/2020 0.0.0.7350 Now have the route through the generation of the pre-projection zone clarified and consistent. Using the generic MSE, once depleted to a pre-determined level the use of the function addrepvar literally adds variation to the beginning of each individual replicate prior to application of the chosen HCR.
58

59

* 08/11/2020 0.0.0.7300 Implemented a constant Catch (constant TAC) harvest strategy so as to test the run time of a typical run. 1000 replicates in about 73 seconds.
60

61

* 09/11/2020 0.0.0.7250 Included new test data zone1, ctrl, and constants, with which to begin testing the MSE framework
62

63

* 22/11/2020 0.0.0.7000 Now includes an operational mcdahcr and applymcda functions. Currently 1000 replicates should take about 2.5 minutes as I have been optimizing again. There is much still to wrap up inside functions and plot convenient outputs, but it is already interesting.

9
* 26/02/2021 0.0.0.5600 Produced SAU summary using alltosau. 
64
10
65

* 24/11/2020 0.0.0.6950 Started a the_data_files.Rmds vignette, that attempts to describe the structure and contents of the data files (being the control.csv and populatiom.csv). Sone other minor adjustments, such as implementing the use of the randomseed.

11
* 19/02/2021 0.0.0.5750 Starting to improve the HCR to more closely match current usage
66
12
67

* 25/11/2020 0.0.0.6900 Added first zone wide summary of outputs, still need to include length-composition of catch

13
* 16/02/2021 0.0.0.6000 Now all components are currently operational.
68
14
69

* 03/12/2020 0.0.0.6800 Added separate randomseeds for population definition and projections. Added wtedmean and worked on vignettes. Modified zone.RData.

15
* 12/01/2021 0.0.0.6200 Increasing generality now have calibrateMCDA and dohistoricC, plus extra HS related functions. More to come.
70
16
71

* 05/12/2020 0.0.0.6700 Exploring the initiation of projections.

17
* 05/01/21 0.0.0.6300 Added the preferred option of using a time-series of historical catches to deplete the unfished zone. Also added 'qmult' so as to adjust the cpue output, but may revert to just using the 'MaxCEpar'
72
18
73

* 06/12/2020 0.0.0.6600 Started developing a method to calibrate the MCDA using fishery data, without needing to condition the model completely. Fixed up examples so they now work with available data-sets.

19
* 24/12/2020 0.0.0.6400 Strangely readctrlfile seemed to have reverted to readctrlzone!!? Fixed that and included applyHS as an explicit pointer to the harvest strategy function in doprojection.
74
20
75
21
* 10/12/2020 0.0.0.6500 Generalized the selection of the applicable HS but now need to define applyHS before running doprojection.
76
22
77

* 24/12/2020 0.0.0.6400 Strangely readctrlfile seemed to have reverted to readctrlzone!!? Fixed that and included applyHS as an explicit pointer to the harvest strategy function in doprojection.
78

79

* 05/01/21 0.0.0.6300 Added the preferred option of using a time-series of historical catches to deplete the unfished zone. Also added 'qmult' so as to adjust the cpue output, but may revert to just using the 'MaxCEpar'

23
* 06/12/2020 0.0.0.6600 Started developing a method to calibrate the MCDA using fishery data, without needing to condition the model completely. Fixed up examples so they now work with available data-sets.
80
24
81

* 12/01/2021 0.0.0.6200 Increasing generality now have calibrateMCDA and dohistoricC, plus extra HS related functions. More to come.

25
* 05/12/2020 0.0.0.6700 Exploring the initiation of projections.
82
26
83

* 16/02/2021 0.0.0.6000 Now all components are currently operational.

27
* 03/12/2020 0.0.0.6800 Added separate randomseeds for population definition and projections. Added wtedmean and worked on vignettes. Modified zone.RData.
84
28
85

* 19/02/2021 0.0.0.5750 Starting to improve the HCR to more closely match current usage

29
* 25/11/2020 0.0.0.6900 Added first zone wide summary of outputs, still need to include length-composition of catch
86
30
87

* 26/02/2021 0.0.0.5600 Produced SAU summary using alltosau.

### 2021-04-14 README.Rmd
added NumNe to zoneDP and plotZone and condition to the webpage.

--- 2021-03-11 README.Rmd
@@
 ```
+
+# LATEST UPDATE
+
+* 2021-04-14 0.0.0.3200 Added NumNe to zoneDP output ready for its use in estimating FIS results by population and other numbers-at-size related indices.
+NumNe are the numbers-at-size following growth and the application of half of natural mortality but before any fishing mortality has occurred.
+(see News.ms for history of development)
 
@@
 library(knitr)
-# Obviously you should modify the rundir to suit your own computer
+# OBVIOUSLY you should modify the rundir and datadir to suit your own computer
 if (dir.exists("c:/Users/User/DropBox")) {
@@
 }
-rundir <- paste0(ddir,"aMSEUse/conddata/generic")  # data and results directory
-dirExists(rundir,make=TRUE,verbose=TRUE)
+rundir <- paste0(ddir,"aMSEUse/scenarios/HS652510")
+datadir <- paste0(ddir,"aMSEUse/scenarios/tasdata")
+alldirExists(rundir,datadir)
 # equilibrium zone -------------------------------------------------------------
@@
 # TasmanianHS.R should be in rundir, but during development is in data-raw
-source(paste0(ddir,"aMSE/data-raw/","TasmanianHS.R"))
+source(paste0(rundir,"/TasmanianHS.R"))
 # generate equilibrium zone ----------------------------------------------------
 starttime2 <- (Sys.time())
-zone <- makeequilzone(rundir,"control2.csv",cleanslate = TRUE) # normally would read in a file
+zone <- makeequilzone(rundir,"controlsau.csv",datadir=datadir,cleanslate = TRUE)
 equiltime <- (Sys.time()); print(equiltime - starttime2)
@@
 zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,sigR=1e-08,sigB=1e-08)
-# save conditioned results -----------------------------------------------------
-save(zoneDD,file=filenametopath(rundir,"zoneDD.RData"))
-# Illustrate productivity
-propD <- getzoneprops(zoneC,zoneDD,glb,year=47)
-addtable(round(t(propD),4),"propertyDD.csv",rundir,category="zoneDD",caption=
+
+propD <- t(getzoneprops(zoneC,zoneDD,glb,year=47))
+addtable(round(propD,4),"propertyDD.csv",rundir,category="zoneDD",caption=
            "Properties of zoneD after conditioning on historical catches.")
-addtable(round(t(zoneDD$harvestR[45:47,]),4),"final_harvestR.csv",rundir,
-         category="zoneDD",caption="Last three years of harvest rate.")
+addtable(round(t(zoneDD$harvestR[40:47,]),4),"final_harvestR.csv",rundir,
+         category="zoneDD",caption="Last eight years of harvest rate.")
 popdefs <- getlistvar(zone$zoneC,"popdef")
@@
 # Prepare projections ----------------------------------------------------------
+plotconditioning(zoneDD,glb,zoneC,condC$histCE,1973:2019,rundir)
 cmcda <- mcdahcr(arrce=condC$histCE,hsargs=hsargs,
-                 yearnames=rownames(condC$histCE),saunames=glb$saunames)
+                 yearnames=rownames(condC$histCE),saunames=glb$saunames,
+                 acatches=tail(condC$histCatch,1))
 pms <- cmcda$pms
 multTAC <- cmcda$multTAC
-#out <- prepareprojection(projC,zoneC,glb,zoneDD,ctrl,multTAC)
 out <- prepareprojectionnew(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
-                            zoneDD=zoneDD,ctrl=ctrl,varyrs=6,
-                            multTAC=multTAC,lastsigR = 0.25)
-
+                            multTAC=multTAC,zoneDD=zoneDD,ctrl=ctrl,varyrs=7,
+                            lastsigR = ctrl$withsigR)
 zoneDP <- out$zoneDP
@@
 # save more plots to rundir ----------------------------------------------------
-out <- plotbysau(zoneDP,glb,rundir)
-
+histCE <- condC$histCE
+B0 <- getvar(zoneC,"B0")
+ExB0 <- getvar(zoneC,"ExB0")
+sauCI <- sauplots(zoneDP,zoneDDR,glb,rundir,B0,ExB0,startyr=25,addCI=TRUE,
+                  histCE=histCE)
+outzone <- poptozone(zoneDP,glb,
+                     B0=sum(getvar(zoneC,"B0")),
+                     ExB0=sum(getvar(zoneC,"ExB0")))
+plotZone(outzone,rundir,CIprobs=c(0.05,0.5,0.95),addfile=TRUE)
 projtime <- Sys.time()
 print(projtime - starttime)
-
 # make results webpage ---------------------------------------------------------
@@
 
+After running the whole, even if you do not generate the results webpage, you could also try:
+
+```{r }
+str(zoneDP,max.level=1)
+
+```
+
+To see the structure of the dynamic object generated by the projections.
+
 See the vignette Running_aMSE.Rmd for a more detailed example.

### 2021-04-14 data-raw/listallfiles.R

@@ -1,114 +1,13 @@
 
  # infile="C:/Users/User/Dropbox/A_Code/aMSE/R/inputfiles.R"; allfuns=allfilesort[,"function"]
 
-findfuns <- function(infile,allfuns) {
-  numfun <- length(allfuns)
-  content <- readLines(con=infile)
-  rfun <- tail(unlist(strsplit(infile,"/")),1)
-  rfile <- substr(rfun,1,nchar(rfun)-2)
-  funLines <- grep("function",content)
-  titles <- grep("@title",content)
-  testhash <- substr(content[funLines],1,4)
-  omit <- grep("#",testhash)
-  if (length(omit) > 0) {
-    funLines <- funLines[-omit]
-    testhash <- testhash[-omit]
-  }
-  omit2 <- grep("  ",testhash) # remove functions internal to other functions
-  if (length(omit2) > 0) funLines <- funLines[-omit2]
-  nfun <- length(funLines)
-    outf <- as.data.frame(matrix("",nrow=nfun,ncol=1))
-    bounds <- matrix(0,nrow=nfun,ncol=2,
-                     dimnames=list(paste0(rfile,1:nfun),c("start","end")))
-    bounds[,1] <- funLines + 1
-    if (nfun > 1) {
-      bounds[,2] <- c((titles[2:nfun] - 2),length(content))
-  } else {
-    bounds[,2] <- length(content)
-  }
-  for (i in 1:nfun) { # i=1
-    funname <- removeEmpty(unlist(strsplit(content[funLines[i]],"<-"))[1])
-    funcont <- content[bounds[i,1]:bounds[i,2]]
-    testhash <- substr(funcont,1,5)
-    omit <- grep("#",testhash)
-    if (length(omit) > 0) funcont <- funcont[-omit]
-    whichfun <- ", "
-    for (j in 1:numfun) {  #  j = 65
-      if (allfuns[j] != funname)
-        if (length(grep(allfuns[j],funcont)) > 0)
-          whichfun <- paste0(whichfun,allfuns[j],", ")
-    }
-    outf[i,] <- whichfun
-  }
-  return(outf)
-} # end of findfuns
 
 
-#' @title listfuns produces a listing of all functions in an input R file
-#'
-#' @description listfuns reads in a given R file and then identifies each
-#'     function header within it and pulls out the function name, its syntax,
-#'     the line-number in the file, and associates that with the filename.
[…]

### 2021-04-19 README.Rmd
No EWN. Began generalizing the structure of the dynamics to allow for alternative functions for sampling. More to be done

@@
 
-* 2021-04-14 0.0.0.3200 Added NumNe to zoneDP output ready for its use in estimating FIS results by population and other numbers-at-size related indices.
-NumNe are the numbers-at-size following growth and the application of half of natural mortality but before any fishing mortality has occurred.
-(see News.ms for history of development)
+* 2021-04-19 0.0.0.3000 Started development required to generalize the use of the doprojection function (it was previously called doTASprojection).
+This generalization requires the generation of separate functions to process the data for cpue, any FIS, and the numbers-at-size.
+Numerous diganostic plots have been added, also providing a template for adding more  (see News.md for history of development).
 
@@
 plotconditioning(zoneDD,glb,zoneC,condC$histCE,1973:2019,rundir)
-cmcda <- mcdahcr(arrce=condC$histCE,hsargs=hsargs,
-                 yearnames=rownames(condC$histCE),saunames=glb$saunames,
-                 acatches=tail(condC$histCatch,1))
+
+indata <- list(arrce=condC$histCE,yearnames=rownames(condC$histCE),
+               acatches=tail(condC$histCatch,1),fis=NA,nas=NA)
+cmcda <- mcdahcr(indata,hsargs=hsargs,saunames=glb$saunames)
 pms <- cmcda$pms
@@
 
-zoneDP <- doTASprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,mcdahcr,hsargs)
+zoneDP <- doprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,mcdahcr,hsargs,
+                        tasCPUE,tasFIS,tasNaS)
 # save more plots to rundir ----------------------------------------------------
@@
 ExB0 <- getvar(zoneC,"ExB0")
-sauCI <- sauplots(zoneDP,zoneDDR,glb,rundir,B0,ExB0,startyr=25,addCI=TRUE,
+sauout <- sauplots(zoneDP,zoneDDR,glb,rundir,B0,ExB0,startyr=25,addCI=TRUE,
                   histCE=histCE)
+diagnosticsproj(sauout$zonePsau,glb,rundir,nrep=3)
 outzone <- poptozone(zoneDP,glb,

### 2021-04-20 README.Rmd
Further development needed to implement the inclusion of NumNe and midyexpB into zoneD and its derivatives. These are the pre-fishing numbers and expoitable biomass.

@@
 
-* 2021-04-19 0.0.0.3000 Started development required to generalize the use of the doprojection function (it was previously called doTASprojection).
-This generalization requires the generation of separate functions to process the data for cpue, any FIS, and the numbers-at-size.
-Numerous diganostic plots have been added, also providing a template for adding more  (see News.md for history of development).
+* 2021-04-20 0.0.0.2900 Continuing process of modifying the core functions to allow for the numbers-at-size prior to fishing and the midyear-exploitB (midyexpB) to be included in teh zoneD object.
+doprojection still requires fursther modification to account for applying to chosen HCR and HS to the completed conditioned zoneDD to provide the first year of expected catches in the projections (which are to be defined by teh conditoned data prior to projections beginning).
 
@@
 
+#explicit definition of indata rather than using the tasdata function
 indata <- list(arrce=condC$histCE,yearnames=rownames(condC$histCE),
-               acatches=tail(condC$histCatch,1),fis=NA,nas=NA)
+               acatches=tail(condC$histCatch,1),fis=NULL,nas=NULL)
 cmcda <- mcdahcr(indata,hsargs=hsargs,saunames=glb$saunames)
@@
 multTAC <- cmcda$multTAC
-out <- prepareprojectionnew(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
+out <- prepareprojection(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
                             multTAC=multTAC,zoneDD=zoneDD,ctrl=ctrl,varyrs=7,
                             lastsigR = ctrl$withsigR)
+# prepareprojection needs to be modified so that is uses expected catches rather
+# than calculating them internally (which will differ by jurisdiction)
 zoneDP <- out$zoneDP

### 2021-04-20 R/plotfuns.r

@@ -13,24 +13,27 @@
 #' @param rundir the rundir for the scenario
 #' @param nrep the number of replicate trajectories to plot; default=3
 #'
-#' @return the residuals between actual catches and aspirational SAU catches
+#' @return Nothing but it does add some plots to rundir
 #' @export
 #'
 #' @examples
 #' print("wait on suitable data sets")
 diagnosticsproj <- function(zonePsau,glb,rundir,nrep=3) {
-  plotvar <- function(invar,nsau,saunames,nrep,filen,caption,label) {
-    nyrs <- dim(invar)[1]
-    reps <- dim(invar)[3]
+  plotvar <- function(var1,nsau,saunames,nrep,filen,caption,label,var2=NULL) {
+    # a helper function that plots nrep trajectories from invar
+    nyrs <- dim(var1)[1]
+    reps <- dim(var1)[3]
     plotprep(width=7,height=7,filename=filen,cex=1.0,verbose=FALSE)
     parset(plots=c(4,2))
     for (sau in 1:nsau) {
       pickrep <- sample(1:reps,nrep,replace=FALSE)
-      ymax <- getmax(invar[,sau,pickrep])
+      ymax <- getmax(var1[,sau,pickrep])
       ylabel <- paste0(paste0("SAU_",saunames[sau],label))
-      plot(1:nyrs,invar[,sau,pickrep[1]],type="l",lwd=2,ylim=c(0,ymax),
+      plot(1:nyrs,var1[,sau,pickrep[1]],type="l",lwd=2,ylim=c(0,ymax),
            ylab=ylabel,xlab="Years",panel.first=grid())
-      for (i in 2:nrep) lines(1:nyrs,invar[,sau,pickrep[i]],lwd=2,col=i)
+      for (i in 2:nrep) lines(1:nyrs,var1[,sau,pickrep[i]],lwd=2,col=i)
+      if (is.numeric(var2))
+        for (i in 1:nrep) lines(1:nyrs,var2[,sau,pickrep[i]],lwd=2,col=i,lty=3)
     } # end of actual catches
     addplot(filen,rundir=rundir,category="DiagProj",caption)
   } # end of internal function plotvar
@@ -55,9 +58,11 @@ diagnosticsproj <- function(zonePsau,glb,rundir,nrep=3) {
   addplot(filen,rundir=rundir,category="DiagProj",caption)
   # plot the limited trajectory plots # filen=""
   catch <- zonePsau$catch
+  acatch <- zonePsau$acatch
[…]

### 2021-05-11 README.Rmd
Some modifications preparing for further generalizations.

@@
 # LATEST UPDATE
+
+* 2021-05-07 0.0.0.2800
 
@@
 # Condition on Fishery ---------------------------------------------------------
-zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,sigR=1e-08,sigB=1e-08)
+zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,calcexpectpopC,sigR=1e-08,sigB=1e-08)
 
@@
 # Prepare projections ----------------------------------------------------------
-plotconditioning(zoneDD,glb,zoneC,condC$histCE,1973:2019,rundir)
+plotconditioning(zoneDD,glb,zoneC,condC$histCE,rundir)
 
@@
 out <- prepareprojection(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
-                            multTAC=multTAC,zoneDD=zoneDD,ctrl=ctrl,varyrs=7,
-                            lastsigR = ctrl$withsigR)
+                         multTAC=multTAC,calcexpectpopC,zoneDD=zoneDD,
+                         ctrl=ctrl,varyrs=7,lastsigR = ctrl$withsigR)
 # prepareprojection needs to be modified so that is uses expected catches rather
@@
 
-zoneDP <- doprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,mcdahcr,hsargs,
-                        tasCPUE,tasFIS,tasNaS)
+zoneDP <- doprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,hcrfun=mcdahcr,hsargs,
+                        sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
+                        getdata=tasdata,calcpopC=calcexpectpopC)
 # save more plots to rundir ----------------------------------------------------
@@
                      ExB0=sum(getvar(zoneC,"ExB0")))
-plotZone(outzone,rundir,CIprobs=c(0.05,0.5,0.95),addfile=TRUE)
+plotZone(outzone,rundir,glb,CIprobs=c(0.05,0.5,0.95),addfile=TRUE)
 projtime <- Sys.time()

### 2021-05-11 src/usearma.cpp

@@ -48,3 +48,17 @@ double svvC(const arma::vec lhs, const arma::vec rhs){
   double ans = arma::as_scalar(lhs.t() * rhs);
   return ans;
 }
+
+
+//' @title mult multiplies two matrices using RcppArmadillo
+//'
+//' @description mult multiplies two matrices using RcppArmadillo. There are no
+//'    tests for conformity so use at your own risk
+//'
+//' @param A the first matrix
+//' @param B the second matrix
+//' @export
+// [[Rcpp::export("mult")]]
+arma::mat mult(arma::mat A, arma::mat B) {
+  return A*B;
+}

### 2021-05-26 README.Rmd
Major changes. Simplified conducting only conditioning, removed the src directory (no Rcpp currently), now each SAU can have only one population if desired.

--- 2021-04-20 README.Rmd
@@
 # LATEST UPDATE
+
+* 2021-05-07 0.0.0.2500 Some relatively large changes this time.
+I have pulled most of the code used to run the MSE into a separate source file 'MSE_source.R'.
+You can find all required files in  \Dropbox\National abalone MSE\aMSE_files\scenarios.
+I have moved TasmanianHS.R into tasdata as it is currently common to different uses.
+I will endeavour to generalize the contents of MSE_source.R to turn it into a function that can be included in the package.
+Then any changes made there will flow naturally to other instances of each scenario file.
+Also, now there is a minimu requirement of 2 SAU but each SAU can have a minimum of 1 population.
 
@@
 if (dir.exists("c:/Users/User/DropBox")) {
-  ddir <- "c:/Users/User/DropBox/A_code/"
+  ddir <- "c:/Users/User/DropBox/A_codeUse/"
 } else {
-  ddir <- "c:/Users/Malcolm/DropBox/A_code/"
+  ddir <- "c:/Users/Malcolm/DropBox/A_codeUse/"
 }
+doproject <- TRUE  # change to FALSE if only conditioning is required
+verbose <- TRUE
 rundir <- paste0(ddir,"aMSEUse/scenarios/HS652510")
@@
 # TasmanianHS.R should be in rundir, but during development is in data-raw
-source(paste0(rundir,"/TasmanianHS.R"))
-# generate equilibrium zone ----------------------------------------------------
-starttime2 <- (Sys.time())
-zone <- makeequilzone(rundir,"controlsau.csv",datadir=datadir,cleanslate = TRUE)
-equiltime <- (Sys.time()); print(equiltime - starttime2)
-# declare main objects ---------------------------------------------------------
-glb <- zone$glb
-ctrl <- zone$ctrl
-zone1 <- zone$zone1
-projC <- zone1$projC
-condC <- zone1$condC
-zoneC <- zone$zoneC
-zoneD <- zone$zoneD
-product <- zone$product
-# add some tables and plots into rundir
-biology_plots(rundir, glb, zoneC)
-plotproductivity(rundir,product,glb)
-numbersatsize(rundir, glb, zoneD)
-# Condition on Fishery ---------------------------------------------------------
-zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,sigR=1e-08,sigB=1e-08)
+source(paste0(datadir,"/TasmanianHS.R"))
+controlfile <- "controlsau.csv"
+# run the scenario ----------------------------------------------------
+source(paste0(ddir,"aMSEUse/scenarios/MSE_Source.R"))
 
-propD <- t(getzoneprops(zoneC,zoneDD,glb,year=47))
-addtable(round(propD,4),"propertyDD.csv",rundir,category="zoneDD",caption=
-           "Properties of zoneD after conditioning on historical catches.")
-addtable(round(t(zoneDD$harvestR[40:47,]),4),"final_harvestR.csv",rundir,
-         category="zoneDD",caption="Last eight years of harvest rate.")
-popdefs <- getlistvar(zone$zoneC,"popdef")
-addtable(round(t(popdefs),3),"popdefs.csv",rundir,category="zoneDD",caption=
-           "Population specific definitions")
-# Prepare projections ----------------------------------------------------------
-plotconditioning(zoneDD,glb,zoneC,condC$histCE,1973:2019,rundir)
-
-#explicit definition of indata rather than using the tasdata function
-indata <- list(arrce=condC$histCE,yearnames=rownames(condC$histCE),
-               acatches=tail(condC$histCatch,1),fis=NULL,nas=NULL)
-cmcda <- mcdahcr(indata,hsargs=hsargs,saunames=glb$saunames)
-pms <- cmcda$pms
-multTAC <- cmcda$multTAC
-out <- prepareprojection(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
-                            multTAC=multTAC,zoneDD=zoneDD,ctrl=ctrl,varyrs=7,
-                            lastsigR = ctrl$withsigR)
-# prepareprojection needs to be modified so that is uses expected catches rather
-# than calculating them internally (which will differ by jurisdiction)
-zoneDP <- out$zoneDP
-projC <- out$projC
-zoneCP <- out$zoneCP
-zoneDDR <- out$zoneDDR
-
-zoneDP <- doprojections(ctrl,zoneDP,zoneCP,condC$histCE,glb,mcdahcr,hsargs,
-                        tasCPUE,tasFIS,tasNaS)
-# save more plots to rundir ----------------------------------------------------
-histCE <- condC$histCE
-B0 <- getvar(zoneC,"B0")
-ExB0 <- getvar(zoneC,"ExB0")
-sauout <- sauplots(zoneDP,zoneDDR,glb,rundir,B0,ExB0,startyr=25,addCI=TRUE,
-                  histCE=histCE)
-diagnosticsproj(sauout$zonePsau,glb,rundir,nrep=3)
-outzone <- poptozone(zoneDP,glb,
-                     B0=sum(getvar(zoneC,"B0")),
-                     ExB0=sum(getvar(zoneC,"ExB0")))
-plotZone(outzone,rundir,CIprobs=c(0.05,0.5,0.95),addfile=TRUE)
-projtime <- Sys.time()
-print(projtime - starttime)
 # make results webpage ---------------------------------------------------------

### 2021-05-26 R/aMSE_utils.R

@@ -86,8 +86,7 @@ alldirExists <- function(indir1,indir2=indir1,make=FALSE,verbose=TRUE) {
 #'     completely different path and will create the 'todir' if it
 #'     does not already exist.
 #'
-#' @param sourcedir the full path to the directory that contains rundir
-#' @param fromdir the name of the current rundir
+#' @param fromdir the full path of the current rundir
 #' @param todir the name of the new destination rundir
 #' @param filename the filename of the file to be copied
 #' @param makenew if the 'todir' does not exist should it be created using
@@ -104,25 +103,70 @@ alldirExists <- function(indir1,indir2=indir1,make=FALSE,verbose=TRUE) {
 #' # any new requirements.
 #' copyto(fromdir=rundir,todir=destdir,filename="control.csv")
 #' }
-copyto <- function (sourcedir,fromdir, todir, filename="control.csv",
-                    makenew = TRUE,verbose=TRUE) {
-  fdir <- paste0(sourcedir,fromdir)
-  if (!dir.exists(fdir)) stop(cat(fdir, " does not exist!   \n\n"))
-  filen <- filenametopath(fdir,filename)
+copyto <- function (fromdir, todir, filename="control.csv",makenew = TRUE,
+                    verbose=TRUE) {
+  if (!dir.exists(fromdir)) stop(cat(fromdir, " does not exist!   \n\n"))
+  filen <- filenametopath(fromdir,filename)
   if (!file.exists(filen)) stop(cat(filename, " does not exist \n"))
-  tdir <- paste0(sourcedir,todir)
-  if (!dir.exists(tdir)) {
-    if (verbose) cat(tdir," did not exist  \n")
+  if (!dir.exists(todir)) {
+    if (verbose) cat(todir," did not exist  \n")
     if (makenew) {
       dir.create(todir, recursive = TRUE)
       if (verbose) cat(todir," has been created  \n")
     }
   }
-  fileout <- filenametopath(tdir, filename)
+  fileout <- filenametopath(todir, filename)
   file.copy(filen, fileout, overwrite = TRUE, copy.date = TRUE)
   if (verbose) cat(filename, " has been copied to ",todir,"\n")
-}
-# end of copyto
+} # end of copyto
+
+#' @title makewidedat converts long data to a wide data format
+#'
[…]

### 2021-05-26 R/dynamics.R

@@ -411,7 +411,12 @@ oneyearD <- function(zoneC,zoneD,inHt,year,sigmar,Ncl,npop,movem) {
 #'
 #' @description oneyearrec calculates the Beverton-Holt recruitment for the
 #'    input populations in a single year; parameterized with steepness,
-#'    R0, and B0. To drop variation to insignificant levels set sigmar-1-08
+#'    R0, and B0. To drop variation to insignificant levels set sigmar-1-08. To
+#'    allow for inclusion of recruitment deviates, which can be required to
+#'    explain high catches when stock biomass would otherwise be low, we have
+#'    introduced the devR argument. As the deviates are all multiplicative and
+#'    range in value > 0, then the default = -1 so this will be ignored unless
+#'    a deviate is input.
 #'
 #' @param steep the steepness for the population; scalar
 #' @param R0 the unfished recruitment levels for the population; scalar
@@ -419,6 +424,9 @@ oneyearD <- function(zoneC,zoneD,inHt,year,sigmar,Ncl,npop,movem) {
 #' @param Bsp the current spawning biomass; scalar
 #' @param sigR standard deviation of the recruitment residuals;
 #'     scalar. set this to 1e-08 to avoid recruitment variability
+#' @param devR if recruitment deviates are available then they should be input
+#'     here. If negative values then a random epsilon is used, otherwise
+#'     epsilon is given the value of devR
 #' @return an absolute number of recruits from a given spawning biomass
 #' @export
 #'
@@ -431,8 +439,12 @@ oneyearD <- function(zoneC,zoneD,inHt,year,sigmar,Ncl,npop,movem) {
 #' insigmar <- 0.3
 #' oneyearrec(steep,R0,B0,Bsp,insigmar)
 #' }
-oneyearrec <- function(steep,R0,B0,Bsp,sigR) {
-  epsilon <- exp(rnorm(length(Bsp),mean=0,sd=sigR) - (sigR * sigR)/2)
+oneyearrec <- function(steep,R0,B0,Bsp,sigR,devR=-1) {
+  if (devR > 0) {
+    epsilon <- devR
+  } else {
+    epsilon <- exp(rnorm(length(Bsp),mean=0,sd=sigR) - (sigR * sigR)/2)
+  }
[…]

### 2021-05-27 README.Rmd
encapsulated the main MSE within do_MSE, No EWN

@@
 
-* 2021-05-07 0.0.0.2500 Some relatively large changes this time.
-I have pulled most of the code used to run the MSE into a separate source file 'MSE_source.R'.
-You can find all required files in  \Dropbox\National abalone MSE\aMSE_files\scenarios.
-I have moved TasmanianHS.R into tasdata as it is currently common to different uses.
-I will endeavour to generalize the contents of MSE_source.R to turn it into a function that can be included in the package.
-Then any changes made there will flow naturally to other instances of each scenario file.
-Also, now there is a minimu requirement of 2 SAU but each SAU can have a minimum of 1 population.
-
-* 2021-04-20 0.0.0.2900 Continuing process of modifying the core functions to allow for the numbers-at-size prior to fishing and the midyear-exploitB (midyexpB) to be included in teh zoneD object.
-doprojection still requires fursther modification to account for applying to chosen HCR and HS to the completed conditioned zoneDD to provide the first year of expected catches in the projections (which are to be defined by teh conditoned data prior to projections beginning).
+* 2021-05-27 aMSE 0.0.0.2400 Have encapsulated the bulk of running the MSE for a particular scenario into the function 'do_MSE'.
+Check out its help page but also see it in action in the example within this readme file or in the new version of first_use_saudata.R in ~Dropbox\National abalone MSE\aMSE_files\scenarios\HS652510.
+I have also updated the aMSE_0.0.0.2400.tar.gz file.
+Any issues get in touch.
 
@@
 
-This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 100 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000).
+This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 50 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000).
 It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.
@@
 datadir <- paste0(ddir,"aMSEUse/scenarios/tasdata")
-alldirExists(rundir,datadir)
+alldirExists(rundir,datadir,verbose=TRUE)
 # equilibrium zone -------------------------------------------------------------
@@
 # run the scenario ----------------------------------------------------
-source(paste0(ddir,"aMSEUse/scenarios/MSE_Source.R"))
-
+#source(paste0(ddir,"aMSEUse/scenarios/MSE_Source.R"))
+out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,
+              hcrfun=mcdahcr,sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
+              getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=32,
+              cleanslate=TRUE,verbose=TRUE,doproject=TRUE,ndiagprojs=3)
 # make results webpage ---------------------------------------------------------
-replist <- list(starttime=as.character(starttime),endtime=as.character(projtime))
+replist <- list(starttime=as.character(out$starttime),
+                endtime=as.character(out$projtime))
+glb <- out$glb
+runnotes <- paste0(out$ctrl$runlabel,":  RunTime = ",out$tottime,
+                   "  replicates = ",glb$reps,",   years projected = ",glb$pyrs,
+                   "  Populations = ",glb$numpop," and SAU = ",glb$nSAU,
+                   "  Randomseed for conditioning = ",out$ctrl$randseed)
 
@@
 #   packagename = "aMSE",
-#   htmlname = "aMSE"
+#   htmlname = "hs652510"
 # )
@@
 ```{r }
-str(zoneDP,max.level=1)
+str(out$zoneDP,max.level=1)
 

### 2021-05-30 R/aMSE.R
Updates vignettes and expansion of output from do_MSE.

@@ -7,6 +7,7 @@
 #' @importFrom utils write.table head tail
 #' @importFrom rutilsMH getmax getmin setpalette plotprep removeEmpty
 #' @importFrom rutilsMH splitDate which.closest parset getparplots
+#' @importFrom rutilsMH digitsbyrow
 #' @importFrom makehtml addplot addtable dirExists filenametopath
 #' @importFrom makehtml getextension htmltable logfilename make_html pathend
 #' @importFrom makehtml pathtype setuphtml write_css write_head

### 2021-06-24 README.Rmd
Removed deprecated functions to data-raw. Improved documentation.

@@
 
-* 2021-05-27 aMSE 0.0.0.2400 Have encapsulated the bulk of running the MSE for a particular scenario into the function 'do_MSE'.
-Check out its help page but also see it in action in the example within this readme file or in the new version of first_use_saudata.R in ~Dropbox\National abalone MSE\aMSE_files\scenarios\HS652510.
-I have also updated the aMSE_0.0.0.2400.tar.gz file.
-Any issues get in touch.
+* 2021-06-24 aMSE 0.0.0.2100 Big jump in number as some larger changes.
+I have removed a bunch of deprecated functions (which are now in 'deprecated.R' in the data-raw directory).
+I have tidied many other functions, and have modified listall files to aid in the auto-documentation of the package.
 

### 2021-06-28 README.Rmd
added addNA and changevar, and ability to alter starting year of projection at will. No EWN

@@
 
-* 2021-06-24 aMSE 0.0.0.2100 Big jump in number as some larger changes.
-I have removed a bunch of deprecated functions (which are now in 'deprecated.R' in the data-raw directory).
-I have tidied many other functions, and have modified listall files to aid in the auto-documentation of the package.
+* 2021-06-28 aMSE 0.0.0.2000 The AMSE can now conduct what are effectively retrospective analyses.
+This is affected by changing the 'CATCHES, 47, if>1 this is the number of historical catch years' line in the control.csv file, which can now be done in a loop using the new 'changevar' function.
+This has necessitated also changing the 'CEYRS, 20, The first year of catches for which there are CPUE records, ie 1992 is 20th year of catches', which used to take a direct value of how any years of CPUE data there are.
+We could have just changed two numbers but along that path errors lay.
 
@@
 controlfile <- "controlsau.csv"
-# run the scenario ----------------------------------------------------
-#source(paste0(ddir,"aMSEUse/scenarios/MSE_Source.R"))
-out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,
-              hcrfun=mcdahcr,sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
-              getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=32,
-              cleanslate=TRUE,verbose=TRUE,doproject=TRUE,ndiagprojs=3)
+# run the scenario --------------------- Obviously unhash this to make it work
+#out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,
+#              hcrfun=mcdahcr,sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
+#              getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=32,
+#              cleanslate=TRUE,verbose=TRUE,doproject=TRUE,ndiagprojs=3)
 # make results webpage ---------------------------------------------------------
-replist <- list(starttime=as.character(out$starttime),
-                endtime=as.character(out$projtime))
-glb <- out$glb
-runnotes <- paste0(out$ctrl$runlabel,":  RunTime = ",out$tottime,
-                   "  replicates = ",glb$reps,",   years projected = ",glb$pyrs,
-                   "  Populations = ",glb$numpop," and SAU = ",glb$nSAU,
-                   "  Randomseed for conditioning = ",out$ctrl$randseed)
-
+# replist <- list(starttime=as.character(out$starttime),
+#                 endtime=as.character(out$projtime))
+# glb <- out$glb
+# runnotes <- paste0(out$ctrl$runlabel,":  RunTime = ",out$tottime,
+#                    "  replicates = ",glb$reps,",   years projected = ",glb$pyrs,
+#                    "  Populations = ",glb$numpop," and SAU = ",glb$nSAU,
+#                    "  Randomseed for conditioning = ",out$ctrl$randseed)
 # Unhash the make_html command to generate the results webspage
@@
 ```{r }
-str(out$zoneDP,max.level=1)
+#str(out$zoneDP,max.level=1)
 

### 2021-06-28 R/inputfiles.R

@@ -368,7 +368,7 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
    withsigR <- getsingleNum("withsigR",indat)
    withsigB <- getsingleNum("withsigB",indat)
    withsigCE <- getsingleNum("withsigCE",indat)
-   hyrs=40 # minimum to set up equilibrium unfished population
+   hyrs=40 # minimum to set up equilibrium; should this be altered?
    filenames2 <- dir(datadir)
    if (length(grep(datafile,filenames2)) != 1)
       stop("population data file not found \n")
@@ -389,9 +389,6 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
    randomseedP <- getsingleNum("randomseedP",indat)
    initLML <- getsingleNum("initLML",indat)
    projyrs <- getsingleNum("PROJECT",indat)
-   firstyear <- getsingleNum("firstyear",indat)
-   lastyear <- getsingleNum("lastyear",indat)
-   pyrnames <- firstyear:lastyear # projection year names
    projLML <- NULL
    HS <- NULL
    histCatch <- NULL
@@ -415,7 +412,7 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
       colnames(histCatch) <- SAUnames
       histyr <- matrix(0,nrow=hyrs,ncol=2)
       colnames(histyr) <- c("year","histLML")
-      for (i in 1:catches) {
+      for (i in 1:hyrs) {
          begin <- begin + 1
          asnum <- as.numeric(unlist(strsplit(indat[begin],",")))
          histyr[i,] <- asnum[1:2]
@@ -424,7 +421,8 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
       rownames(histCatch) <- histyr[,1]
       rownames(histyr) <- histyr[,1]
    } # end of catches loop
-   yrce <- getsingleNum("CEYRS",indat)
+   startce <- getsingleNum("CEYRS",indat)
+   yrce <- hyrs - startce + 1
    if (yrce == 0) {
       warning("CPUE calibration has no data")
    } else {
@@ -441,6 +439,8 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
       rownames(histCE) <- yearCE
    } # end of if(yrce == 0)
[…]

### 2021-07-04 README.Rmd
fixed recruitment deviates, tidied output objects, extra plots available, some deprecated functions removed.

@@
 # LATEST UPDATE
+
+* 2021-07-04 aMSE 0.0.0.1900 More big changes.
+I have removed the numbers-at-size large matrices from the zoneDP object into a NAS object.
+zoneDP is now only 12Mb instead of 450Mb and so can now be stored conveniently for each run.
+In addition, fixed recruitment deviates have now been implemented (although the default is to not use them - all are set to -1).
+But, expreimentally, in Tasmania, setting the deviate in 1991 = 2.0 meaning that the base level of recruitment off the recruitment curve is doubled, improves the relationship between predicted cpue and observed cpue so this holds great promise for improved conditioning.
 

### 2021-07-04 R/inputfiles.R

@@ -1,5 +1,5 @@
 
-# rutilsMH::listFunctions("C:/Users/User/Dropbox/rcode2/makehtml/R/makehtml_funs.R")
+# rutilsMH::listFunctions("C:/Users/User/Dropbox/A_code/aMSE/R/inputfiles.R")
 
 #' @title ctrlzonetemplate generates a template input control file
 #'
@@ -453,10 +453,22 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
          compdat[[i]] <- read.csv(file=filename,header=TRUE)
       }
    }  # end of sizecomp loop
+   recdevs <- matrix(-1,nrow=hyrs,ncol=nSAU,dimnames=list(hyrnames,SAUnames))
+   rdevs <- getsingleNum("RECDEV",indat)
+   if (rdevs > 0) {
+      if (rdevs != hyrs)
+         stop("rows of recdevs not equal to conditioning years \n")
+      begin <- grep("RECDEV",indat) + 1
+      for (i in 1:rdevs) {
+         begin <- begin + 1
+         devs <- as.numeric(unlist(strsplit(indat[begin],",")))
+         recdevs[i,] <- devs[2:(nSAU+1)]
+      }
+   } # end of recdev loop
    # make output objects
    condC <- list(histCatch=histCatch,histyr=histyr,
                  histCE=histCE,yearCE=yearCE,initdepl=initdepl,
-                 compdat=compdat,Sel=NULL,SelWt=NULL)
+                 compdat=compdat,recdevs=recdevs,Sel=NULL,SelWt=NULL)
    projC <- list(projLML=projLML,projyrs=projyrs,
                  Sel=NULL,SelWt=NULL,histCE=histCE)
    outctrl <- list(runlabel,datafile,batch,reps,randomseed,randomseedP,

### 2021-07-08 README.Rmd
Functions added to aid conditioning. No EWN

@@
 
+* 2021_07-04 aMSE 0.0.0.1500 Another big change. Functions added to aid conditioning, and aspects of the projections streamlined.  No errors, warnings, or notes
+
 * 2021-07-04 aMSE 0.0.0.1900 More big changes.
@@
 But, expreimentally, in Tasmania, setting the deviate in 1991 = 2.0 meaning that the base level of recruitment off the recruitment curve is doubled, improves the relationship between predicted cpue and observed cpue so this holds great promise for improved conditioning.
-
-* 2021-06-28 aMSE 0.0.0.2000 The AMSE can now conduct what are effectively retrospective analyses.
-This is affected by changing the 'CATCHES, 47, if>1 this is the number of historical catch years' line in the control.csv file, which can now be done in a loop using the new 'changevar' function.
-This has necessitated also changing the 'CEYRS, 20, The first year of catches for which there are CPUE records, ie 1992 is 20th year of catches', which used to take a direct value of how any years of CPUE data there are.
-We could have just changed two numbers but along that path errors lay.
 

### 2021-07-08 R/aMSE_utils.R

@@ -134,6 +134,41 @@ catchweightCE <- function(cedat,cdat,nsau) {
   return(totCE)
 } # end of catchweightCE
 
+#' @title changeline replaces a given line in a given file with new text
+#'
+#' @description changeline enables a text file to be changed line by line.
+#'     One identifies a given line, perhaps by using 'findlinenumber', and
+#'     this can then be replaced by an input character string. Obviously this
+#'     is a fine way to mess up a data file so use with care.
+#'
+#' @param rundir the directory path in which to find the text file.
+#' @param filename the full name of the text file in quotations.
+#' @param linenumber the line number of the text file to be changed
+#' @param newline the character string with which to replace the line
+#' @param verbose should confirmation be output to the console. default=FALSE
+#'
+#' @seealso{
+#'  \link{findlinenumber}, \link{changevar}
+#' }
+#'
+#' @return nothing but it does alter a line in a text file. Optionally it may
+#'     confirm the action to the console
+#' @export
+#'
+#' @examples
+#' print("wait on an example")
+changeline <- function(rundir, filename, linenumber, newline,verbose=FALSE) {
+  # rundir=rundir; filename="controlsau.csv"; linenumber=210; newline="1991,-1,-1,-1,-1,-1,-1,-1,-1,"
+  filen <- filenametopath(rundir,filename)
+  dat <- readLines(filen)
+  origtext <- dat[linenumber]
+  dat[linenumber] <- newline
+  writeLines(dat,filen)
+  if (verbose) {
+    cat(origtext," \n")
+    cat("replaced with ",newline)
+  }
+} # end of changeline
 
 #' @title changevar can alter the value in a single line in, eg, the control file
 #'
@@ -157,13 +192,17 @@ catchweightCE <- function(cedat,cdat,nsau) {
 #' @param prompt should allowable names first be listed with current values?
 #' @param verbose should concluding remarks to console be made. default=TRUE
 #'
+#' @seealso{
+#'  \link{findlinenumber}, \link{changeline}
+#' }
+#'
 #' @return nothing but a file is changed - be careful
 #' @export
 #'
 #' @examples
 #' print("Wait on some time passing")
 changevar <- function(filename,rundir,varname,newvalue,prompt=FALSE,verbose=TRUE) {
-  #  varname = "CATCHES"; newvalue=47; filename="controlsau.csv";rundir=rundir
[…]

### 2021-07-08 R/getfunctions.R

@@ -63,9 +63,12 @@ getConst <- function(inline,nb,index=2) { # parses lines containing numbers
 #'     the SAU. See the 'data-file-description.docx', and the example data files
 #'     to see examples of the required format.
 #'
-#' @param rundir the run directory containing the control and data files for a
-#'     particular run
+#' @param datadir the directory containing the LF data files
 #' @param filename the complete filename for the length-composition data file
+#'
+#' @seealso{
+#'  \link{getzoneLF}, \link{getnas}
+#' }
 #'
 #' @return a list containing the lfs array (lengths x years x sau) and palfs
 #'     matrix of the count of observations in each year for each sau
@@ -73,8 +76,8 @@ getConst <- function(inline,nb,index=2) { # parses lines containing numbers
 #'
 #' @examples
 #' print("wait on suitable internal data files")
-getLFdata <- function(rundir,filename) {
-  filen <- filenametopath(rundir,filename)
+getLFdata <- function(datadir,filename) {
+  filen <- filenametopath(datadir,filename)
   lfdat <- read.csv(filen,header=TRUE)
   colnames(lfdat) <- tolower(colnames(lfdat))
   sau <- sort(unique(lfdat[,"sau"]))
@@ -170,26 +173,28 @@ getLogical <- function(inline,nb) {  #inline <- txtline; nb=2
   return(outtmp)
 }
 
-#' @title getnas gets the numbers-at-size for all spatial scales
+#' @title getnas gets the numbers-at-size for all populations and the zone
 #'
 #' @description getnas extracts numbers-at-size for all populations,
-#'     SAUs, and the zone. There will therefore be numpop + nSAU + 1
+#'     SAUs, and the zone. There will therefore be numpop + 1
 #'     columns of numbers-at-size fr the particular year given.
 #'
 #' @param zoneD the dynamic portion of the zone
 #' @param yr which yr from the range available should be summarized
 #' @param glob the global variables object
-#' @param zone the zone constants
+#'
+#' @seealso{
+#'  \link{getLFdata}, \link{getzoneLF}, \link{prepareprojection},
+#'  \link{doprojections}
+#' }
 #'
 #' @return a matrix of numpop + 1 columns of numbers-at-size for each population
 #'     and for the zone
 #' @export
 #'
[…]

### 2021-07-08 R/inputfiles.R

@@ -445,12 +445,15 @@ readctrlfile <- function(rundir,infile="control.csv",datadir=rundir,verbose=TRUE
    if (sizecomp > 0) {
       lffiles <- NULL
       locsizecomp <- grep("SIZECOMP",indat)
-      for (i in 1:sizecomp)
-         lffiles <- c(lffiles,getStr(indat[locsizecomp+i],1))
-      compdat <- vector("list",sizecomp)
-      for (i in 1:sizecomp) {
-         filename <- filenametopath(datadir,lffiles[i])
-         compdat[[i]] <- read.csv(file=filename,header=TRUE)
+      if (sizecomp > 1) {
+         compdat <- vector("list",sizecomp)
+         for (i in 1:sizecomp) {
+            lffilename <- removeEmpty(unlist(strsplit(indat[locsizecomp+i],",")))
+            compdat[[i]] <- getLFdata(datadir,lffilename)
+         }
+       } else {
+         lffilename <- removeEmpty(unlist(strsplit(indat[locsizecomp+1],",")))
+         compdat <- getLFdata(datadir,lffilename)
       }
    }  # end of sizecomp loop
    recdevs <- matrix(-1,nrow=hyrs,ncol=nSAU,dimnames=list(hyrnames,SAUnames))

### 2021-07-12 README.Rmd
Combined historical and projected dynamics and made the rest of the code base consistent with that. No EWN.

--- 2021-06-28 README.Rmd
@@
 
-* 2021-06-28 aMSE 0.0.0.2000 The AMSE can now conduct what are effectively retrospective analyses.
-This is affected by changing the 'CATCHES, 47, if>1 this is the number of historical catch years' line in the control.csv file, which can now be done in a loop using the new 'changevar' function.
-This has necessitated also changing the 'CEYRS, 20, The first year of catches for which there are CPUE records, ie 1992 is 20th year of catches', which used to take a direct value of how any years of CPUE data there are.
-We could have just changed two numbers but along that path errors lay.
+* 2021_07-12 aMSE 0.0.0.1000 Announcing some rather large changes.
+Early on in development I made a strategic decision which has proved to make the documentation of the code and implementation of extraction of results overly complex.
+It is now clear that it was a strategic mistake to separate the historical conditioning from the future projections.
+Combining them, for purposes of applying a Harvest control rule, or tabulating or plotting results, was adding complexity to every scenario, new plot and table.
+For example, previously, plots where there were traces of the historical data added to the projections were greatly flawed by having the final year of historical data repeated in the projection data.
+Obviously, that plotting error could have been solved by selecting which years to plot, instead I decided to simplify everything (for both me and other maintainers) by combining the conditioned dynamic component of the zone and the projection years of the dynamics.
+Now, in the 'zoneDP' object (see 'out$zoneDP' after a run), for each replicate, we have a continuous set of years from the first year of historical catches out to the last year of the projections.
+This simplifies everything from the documentation of the code base, the application of the harvest strategy, the plotting of results, basically everything I can think of.
+The next steps are to expand the results sections and re-write the documentation.
+The current Tasmania example has now been expanded to 56 populations among the 8 SAU.
+The fit of the operating model the historical CPUE is now rather improved because now it is possible to include recruitment deviates in the historical conditioning.
+Currently in the included example it is very ad hoc.
+One can use ad hoc recruitment deviates or ones derived from some form of localized model fitting.
+The ad hoc ones currently in use were mainly used to test the new utility functions provided to assist with such things.
 
@@
 
-* zone - highest geogrpahical level. Is simply the totality of the spatial management units.
+* zone - highest geographical level. Is simply the totality of the spatial management units.
 
@@
 if (dir.exists("c:/Users/User/DropBox")) {
-  ddir <- "c:/Users/User/DropBox/A_codeUse/"
+  ddir <- "c:/Users/User/DropBox/A_codeUse/aMSEUse/scenarios/"
 } else {
-  ddir <- "c:/Users/Malcolm/DropBox/A_codeUse/"
+  ddir <- "c:/Users/Malcolm/DropBox/A_codeUse/aMSEUse/scenarios/"
 }
@@
 verbose <- TRUE
-rundir <- paste0(ddir,"aMSEUse/scenarios/HS652510")
-datadir <- paste0(ddir,"aMSEUse/scenarios/tasdata")
-alldirExists(rundir,datadir,verbose=TRUE)
+postdir <- "testnew"  # this is used to label the output HTML file 
+rundir <- paste0(ddir,postdir)
+datadir <- paste0(ddir,"tasdata")
+alldirExists(rundir,datadir,verbose=verbose)
+source(paste0(datadir,"/TasmanianHS.R"))
+controlfile <- "controlsau.csv"
 # equilibrium zone -------------------------------------------------------------
@@
 
-# TasmanianHS.R should be in rundir, but during development is in data-raw
+# HarvestStrategy.R should be in datadir, if it is to be shared by scenarios
 source(paste0(datadir,"/TasmanianHS.R"))
-controlfile <- "controlsau.csv"
 # run the scenario --------------------- Obviously unhash this to make it work
-#out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,
-#              hcrfun=mcdahcr,sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
-#              getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=32,
-#              cleanslate=TRUE,verbose=TRUE,doproject=TRUE,ndiagprojs=3)
+# out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,hcrfun=mcdahcr,
+#                  sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
+#                  getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=50,
+#                  cleanslate=TRUE,verbose=TRUE,doproject=doproject,ndiagprojs=4)
 # make results webpage ---------------------------------------------------------
@@
 # glb <- out$glb
+# projy <- ifelse(doproject,glb$pyrs,0)
 # runnotes <- paste0(out$ctrl$runlabel,":  RunTime = ",out$tottime,
-#                    "  replicates = ",glb$reps,",   years projected = ",glb$pyrs,
+#                    "  replicates = ",glb$reps,",   years projected = ",projy,
 #                    "  Populations = ",glb$numpop," and SAU = ",glb$nSAU,
 #                    "  Randomseed for conditioning = ",out$ctrl$randseed)
-# Unhash the make_html command to generate the results webspage
-# make_html(
-#   replist = replist,
-#   rundir = rundir,
-#   width = 500,
-#   openfile = TRUE,
-#   runnotes = NULL,
-#   verbose = FALSE,
-#   packagename = "aMSE",
-#   htmlname = "hs652510"
-# )
+# 
+# make_html(replist = replist,  rundir = rundir,  width = 500,  openfile = TRUE,
+#           runnotes = runnotes,   verbose = FALSE,  packagename = "aMSE",
+#           htmlname = postdir)
 
@@
 #str(out$zoneDP,max.level=1)
+#str1(out$hcrout)
 

### 2021-07-12 R/aMSE_utils.R

@@ -406,6 +406,7 @@ poptosau <- function(invar,glb) {  # invar=zoneDP$matureB; glb=glb
 #'
 #' @examples
 #' print("wait on appropriate built-in data files")
+#' # catvect=zoneDD$catch[1:finalyr,]; cpuevect=zoneDD$cpue[1:finalyr,]
 poptosauCE <- function(catvect,cpuevect,sauindex) {
   saucatch <- tapply(catvect,sauindex,sum,na.rm=TRUE)
   wts <- catvect/saucatch[sauindex]
@@ -450,6 +451,7 @@ poptozone <- function(inzone,NAS=NULL,glb, B0, ExB0) {
   depleB <- exploitB/ExB0
   recruit <- addpops(inzone$recruit,nyrs,reps)
   catch <- addpops(inzone$catch,nyrs,reps)
+  TAC <- addpops(inzone$acatch,nyrs,reps)
   catchN <- array(data=0,dim=c(N,nyrs,reps), # define some arrays
                   dimnames=list(glb$midpts,1:nyrs,1:reps))
   Nt <- array(data=0,dim=c(N,nyrs,reps),
@@ -473,7 +475,7 @@ poptozone <- function(inzone,NAS=NULL,glb, B0, ExB0) {
     }
   }
   outzone <- list(matureB=matureB,exploitB=exploitB,catch=catch,
-                  TAC=inzone$TAC,harvestR=harvestR,cpue=cpue,
+                  acatch=inzone$acatch,TAC=TAC,harvestR=harvestR,cpue=cpue,
                   recruit=recruit,deplsB=deplsB,depleB=depleB,
                   catchN=catchN,Nt=Nt)
   return(outzone)
@@ -670,7 +672,7 @@ sumpop2sau <- function(invect,sauindex) {
 #'
 #' @examples
 #' print("wait on suitable internal data sets ")
-zonetosau <- function(inzone,NAS=NULL,glb, B0, ExB0) { # inzone=zoneDDR; NAS=NULL; glb=glb; B0=B0; ExB0=ExB0
+zonetosau <- function(inzone,NAS=NULL,glb, B0, ExB0) { # inzone=zoneDP; NAS=NAS; glb=glb; B0=B0; ExB0=ExB0
   nsau <- glb$nSAU
   sauindex <- glb$sauindex
   saunames <- glb$saunames
[…]

### 2021-07-20 README.Rmd
added do_condition and compareCPUE. Some incomplete developments underway. No EWN

@@
 # LATEST UPDATE
+
+* 2021-07-20 aMSE 0.0.0.900 Added do_condition and compareCPUE.
+Both used to speed the conditioning of the operating model, though a number of developments are still under development.
+Currently can automatically search for the AvRec value that will optimize the sum-of-squared residual fit between the
+observed CPUE in the historical period, and those predicted by the conditioned model. The undeveloped bit related to the ad hoc recruitment deviates.
 

### 2021-07-20 R/defineZone.R

@@ -583,6 +583,8 @@ makeabpop <- function(popparam,midpts,projLML) {
 #'     numpop populations
 #' @param datadir the directory containing the data csv files. This defaults to
 #'     equal the rundir, but can be different if desired. default=rundir
+#' @param doproduct boolean, should the productivity calculations be made
+#'     during the conditioning. Set to FALSE conditionOM
 #' @param cleanslate a boolean determining whether any old results are deleted
 #'     from rundir before starting. Default=FALSE
 #' @param verbose Should progress comments be printed to console, default=TRUE
@@ -593,7 +595,7 @@ makeabpop <- function(popparam,midpts,projLML) {
 #' @examples
 #' print("wait on datafiles")
 makeequilzone <- function(rundir,ctrlfile="control.csv",datadir=rundir,
-                          cleanslate=FALSE,verbose=TRUE) {
+                          doproduct=TRUE,cleanslate=FALSE,verbose=TRUE) {
  #  rundir=rundir;ctrlfile="controlsau.csv";cleanslate = TRUE;datadir=datadir;verbose=verbose
   zone1 <- readctrlfile(rundir,infile=ctrlfile,datadir=datadir,verbose=verbose)
   ctrl <- zone1$ctrl
@@ -606,7 +608,7 @@ makeequilzone <- function(rundir,ctrlfile="control.csv",datadir=rundir,
     constants <- readdatafile(glb$numpop,datadir,ctrl$datafile)
   }
   if (verbose) cat("Files read, now making zone \n")
-  out <- setupzone(constants,zone1,verbose=verbose) # make operating model
+  out <- setupzone(constants,zone1,doproduct,verbose=verbose) # make operating model
   zoneC <- out$zoneC
   zoneD <- out$zoneD
   glb <- out$glb             # glb now has the movement matrix
@@ -1013,6 +1015,8 @@ resetexB0 <- function(zoneC,zoneD) {
 #'
 #' @param constants the population constants derived from readdatafile
 #' @param zone1 the zonal object driving the construction
+#' @param doproduct boolean, should the productivity calculations be made
+#'     during the conditioning. defined in do_MSE and makeequilzone
 #' @param uplim the upper limit of harvest rate applied, default=0.4
[…]

### 2021-07-20 R/docondition.R

@@ -1,0 +1,105 @@
+
+#
+# rundir=rundir
+# controlfile="controlsau.csv"
+# datadir=datadir
+# hsargs=hsargs
+# hcrfun=mcdahcr
+# sampleCE=tasCPUE
+# sampleFIS=tasFIS
+# sampleNaS=tasNaS
+# getdata=tasdata
+# calcpopC=calcexpectpopC
+# varyrs=7
+# startyr=42
+# cleanslate=FALSE
+# verbose=TRUE
+# doproject=doproject
+# ndiagprojs=4
+
+#' @title do_condition is a function to assist with conditioning the model
+#'
+#' @description do_condition is a utility function that can be used to speed the
+#'     conditioning of the operating model on available fishery data. During the
+#'     generation of the simulated zone the operation that takes the longest is
+#'     the estimation of the productivity of each population. The productivity
+#'     is estimated at equilibrium and getting there is what tales the time.
+#'     When conditioning the model the objective is to match the predicted
+#'     fishery response to the historical catches to the observed responses.
+#'     'do_condition' generates the simulated zone much more rapidly, which
+#'     simplifies any searches for optimal values of AvRec (average unfished
+#'     recruitment), and especially for suitable values of recruitment deviates.
+#'
+#' @param rundir the full path to the directory in which all files relating to a
+#'     particular run are to be held. If datadir != rundir then the main data
+#'     files are kept in datadir. The juridictionHS.R can be held in either.
+#' @param controlfile the filename of the control file present in rundir
+#'     containing information regarding the particular run.
+#' @param datadir the directory in which the SAU data file is to be found. This
+#'     will usually be the rundir for the scenario run but if the data file is
+#'     to be shared among a set of scenarios it will be more efficient to define
+#'     a separate datadir
+#' @param calcpopC a function that takes the output from hcrfun (either the
+#'     aspirational catch x SAU or TAC x zone) and generates the actual catch
+#'     per population in each SAU expected in the current year.
+#' @param cleanslate should the directory be emptied of all files first? All
+#'     html, png, RData, and css files in the directory will be deleted.
+#'     default = FALSE. This is obviously a very powerful and potentially
[…]

### 2021-07-20 R/generate_results.R

@@ -177,9 +177,11 @@ fishery_plots <- function(rundir,glb,select,histyr,projLML, rge=50:90) {
   plot(mids[rge],selplot[rge,1],type="l",lwd=2,col=1,panel.first=grid(),
        xlab="Shell Length (mm)",ylab="Selectivity")
   for (i in 2:nsel) lines(mids[rge],selplot[rge,i],lwd=2,col=i)
-  legend("right",legend=paste0("lml ",selyrs),col=c(1:nsel),lwd=3,
+  legend("right",legend=paste0("lml ",values," ",selyrs),col=c(1:nsel),lwd=3,
          bty="n",cex=1.2)
-  caption <- "The selectivity vs length through the years of the Dynamics."
+  caption <- paste0("The selectivity vs length through the years of the Dynamics.",
+                      "The years after the LML value are the first year in the data",
+                      " in which that LML is expressed.")
   addplot(filen,rundir=rundir,category="Fishery",caption)
   return(invisible(selplot))
 } # end of fishery_plots
@@ -353,6 +355,44 @@ plotproductivity <- function(rundir,product,glb) {
                     "level of each population. Here the x-axis is ",
                     "shortened to clarify the flatness of the production ",
                     "curve about the MSY points.")
+  addplot(filen,rundir=rundir,category="Production",caption)
+
+  # what CPUE at Bmsy would be exhibited at MSY
+  filen <- filenametopath(rundir,"production_CPUE_at_Bmsy.png")
+  nsau <- glb$nSAU
+  npop <- glb$numpop
+  nh <- dim(product)[1]
+  label <- dimnames(product)
+  sauindex <- glb$sauindex
+  wts <- matrix(0,nrow=nh,ncol=npop,dimnames=list(label[[1]],label[[3]]))
+  sauyield <- matrix(0,nrow=nh,ncol=nsau,dimnames=list(label[[1]],glb$saunames))
+  saucpue <- matrix(0,nrow=nh,ncol=nsau,dimnames=list(label[[1]],glb$saunames))
+  # Now do sau production
+  for (i in 1:nh) {
+    sauyield[i,] <- tapply(product[i,"Catch",],sauindex,sum,na.rm=TRUE)
+    wts[i,] <- product[i,"Catch",]/sauyield[i,sauindex]
[…]

### 2021-07-28 README.Rmd
Added functions to assist in auto-conditioning.

--- 2021-07-08 README.Rmd
@@
 
-* 2021_07-04 aMSE 0.0.0.1500 Another big change. Functions added to aid conditioning, and aspects of the projections streamlined.  No errors, warnings, or notes
+* 2021_07-28 aMSE 0.0.0.800 Added functions (changecolumn, getrecdevcolumn, gettasdevssq) to assist in automated operating model conditioning on the fishery.
+Also modified the biology_plots and oneyear and oneyearcat to assist with speeding the processes.
+Lots of additional minor changes but some important ones to do_MSE (read the help, ?do_MSE).
 
-* 2021-07-04 aMSE 0.0.0.1900 More big changes.
-I have removed the numbers-at-size large matrices from the zoneDP object into a NAS object.
-zoneDP is now only 12Mb instead of 450Mb and so can now be stored conveniently for each run.
-In addition, fixed recruitment deviates have now been implemented (although the default is to not use them - all are set to -1).
-But, expreimentally, in Tasmania, setting the deviate in 1991 = 2.0 meaning that the base level of recruitment off the recruitment curve is doubled, improves the relationship between predicted cpue and observed cpue so this holds great promise for improved conditioning.
+* 2021-07-20 aMSE 0.0.0.900 Added do_condition and compareCPUE.
+Both used to speed the conditioning of the operating model, though a number of developments are still under development.
+Currently can automatically search for the AvRec value that will optimize the sum-of-squared residual fit between the
+observed CPUE in the historical period, and those predicted by the conditioned model. The undeveloped bit related to the ad hoc recruitment deviates.
 
@@
 verbose <- TRUE
-rundir <- paste0(ddir,"aMSEUse/scenarios/HS652510")
-datadir <- paste0(ddir,"aMSEUse/scenarios/tasdata")
-alldirExists(rundir,datadir,verbose=TRUE)
+postdir <- "testnew"  # also used to label the output HTML file 
+rundir <- paste0(ddir,postdir)
+datadir <- paste0(ddir,"tasdata")
+alldirExists(rundir,datadir,verbose=verbose)
+source(paste0(datadir,"/TasmanianHS.R"))
+controlfile <- "controlsau.csv"
 # equilibrium zone -------------------------------------------------------------
@@
 
-# TasmanianHS.R should be in rundir, but during development is in data-raw
+# HarvestStrategy.R should be in datadir, if it is to be shared by scenarios
 source(paste0(datadir,"/TasmanianHS.R"))
-controlfile <- "controlsau.csv"
 # run the scenario --------------------- Obviously unhash this to make it work
-#out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,
-#              hcrfun=mcdahcr,sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
-#              getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=32,
-#              cleanslate=TRUE,verbose=TRUE,doproject=TRUE,ndiagprojs=3)
+# out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,hcrfun=mcdahcr,
+#                  sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
+#                  getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=50,
+#                  cleanslate=TRUE,verbose=TRUE,doproject=doproject,ndiagprojs=4,
+#                  openfile=TRUE,savesauout=FALSE)
 # make results webpage ---------------------------------------------------------

### 2021-07-28 data-raw/deprecated.R

@@ -460,6 +460,73 @@ oneyearC <- function(zoneC,zoneDP,catchp,year,iter,sigmar,Ncl,npop,movem) {
   return(zoneD)
 } # end of oneyearC
 
+#' @title oneyearcatold one year's catch dynamics for one abpop
+#'
+#' @description oneyear does one year's dynamics for one input
+#'     population. Where the fishing is based on a given catch per
+#'     population, which would be determined by any harvest control
+#'     rule. It is used to step through populations of a zone.
+#'     Its dynamics are such that each population first undergoes
+#'     growth, then half natural mortality. This allows an estimate of
+#'     exploitable biomass before fishing occurs. The remaining
+#'     dynamics involve the removal of the catch, after estimating the
+#'     harvest rate from catch/exploitb, the application of the last
+#'     half of natural mortality and the addition of recruits. This
+#'     allows the exploitable biomass to be estimated after fishing.
+#'     The cpue uses the average of the pre-fishing and post-fishing
+#'     exploitB to smooth over any changes brought about by fishing.
+#'     The recruitment occurs in oneyearC so that larval dispersal can
+#'     be accounted for. oneyearcat is not used independently of
+#'     oneyearC.
+#'
+#' @param inpopC a single population from a zoneC, an abpop
+#' @param inNt the numbers at size for the year previous to the year
+#'     of dynamics. These are projected into the active year.
+#' @param Nclass the number of size classes used to describe growth.
+#'     used to define vectors
+#' @param incat the literal annual catch from the population. Derived
+#'     from teh harvest control rule for each SAU, then allocated to
+#'     each population with respect to its relative catching
+#'     performance.
+#' @param yr the year in the dynamics being worked on. The first year
+#'     is generated when the zone is defined or when it is initially
+#'     depleted. All dynamics are appllied from year 2 - Nyrs; scalar
+#'
+#' @seealso{
+#'  \link{dohistoricC}, \link{oneyearcat}, \link{oneyearrec}
+#' }
+#'
[…]

### 2021-08-02 R/domse.R
Revised ctrlfiletemplate and datafiletemplate. No EWN

@@ -1,7 +1,9 @@
 #
-# rundir=rundir
-# controlfile="controlsau.csv"
-# datadir=datadir
+# outcond=outcond
+# rundir <- paste0(ddir,postdir)
+# datadir <- paste0(ddir,"tasdata")
+# postdir <- "M125h5"
+# controlfile="controlM125h5.csv"
 # hsargs=hsargs
 # hcrfun=mcdahcr
 # sampleCE=tasCPUE
@@ -10,10 +12,10 @@
 # getdata=tasdata
 # calcpopC=calcexpectpopC
 # varyrs=7
-# startyr=42
+# startyr=48
 # cleanslate=FALSE
 # verbose=TRUE
-# doproject=doproject
+# doproject=TRUE
 # ndiagprojs=4
 
 #' @title do_MSE an encapsulating function to hold the details of a single run
@@ -130,7 +132,7 @@ do_MSE <- function(rundir,controlfile,datadir,hsargs,hcrfun,sampleCE,sampleFIS,
   addtable(round(t(zoneDD$harvestR[(hyrs-9):hyrs,]),4),"final_harvestR.csv",
            rundir,category="zoneDD",
            caption="Last ten years of population vs harvest rate.")
-  popdefs <- getlistvar(zone$zoneC,"popdef")
+  popdefs <- getlistvar(zoneC,"popdef")
   addtable(round(t(popdefs),3),"popdefs.csv",rundir,category="zoneDD",
            caption="Population vs Operating model parameter definitions")
   condout <- plotconditioning(zoneDD,glb,zoneC,condC$histCE,rundir)
@@ -182,7 +184,8 @@ do_MSE <- function(rundir,controlfile,datadir,hsargs,hcrfun,sampleCE,sampleFIS,
   tottime <- round((projtime - starttime),3)
 
   if (savesauout) {
-    sauoutD <- sauout[-c(12,11,10)]
+    sauoutD <- sauout$zonePsau
+    sauoutD <- sauoutD[-c(12,11,10)]
     save(sauoutD,file=paste0(rundir,"/sauoutD.RData"))
   }
   out <- list(tottime=tottime,projtime=projtime,starttime=starttime,glb=glb,
@@ -191,3 +194,170 @@ do_MSE <- function(rundir,controlfile,datadir,hsargs,hcrfun,sampleCE,sampleFIS,
               hcrout=hcrout,production=production,condout=condout)
   return(out)
 } # end of do_MSE
+
+
+
+
+
[…]

### 2021-08-02 R/dynamics.R

@@ -581,7 +581,9 @@ restart <- function(oldzoneD,hyrs,npop,N,zero=TRUE) { # oldzoneD=zoneD; hyrs=hyr
 #'     40 it conducts 3 * 39 years of dynamics (117 years). This is
 #'     not exported. It uses zoneC but always it does this inside
 #'     the environment of another function where zoneC can be found
-#'     Used inside dodepletion and doproduction
+#'     Used inside dodepletion and doproduction. maxiter may need to be
+#'     increased when we introduce a larger movement rate between populations
+#'     for greenlip, or if the number of conditioning years are fewer than 45.
 #'
 #' @param zoneC the constants components of the simulated zone
 #' @param zoneD the dynamics portion of the zone, with matrices and
@@ -592,13 +594,17 @@ restart <- function(oldzoneD,hyrs,npop,N,zero=TRUE) { # oldzoneD=zoneD; hyrs=hyr
 #' @param glob the globals variable from readzonefile
 #' @param maxiter default=3; the number of runs through the equilibrium loop.
 #'
+#' @seealso{
+#'  \link{dodepletion}, \link{doproduction}
+#' }
+#'
 #' @return a list containing a revised dynamics list, zoneD
 #' @export
 #'
 #' @examples
 #' print("wait on built in data sets")
 #' # zoneC=zoneC; zoneD=zoneD; glob=glob; inHarv=rep(initH[aH],numpop); maxiter=2
-runthreeH <- function(zoneC,zoneD,inHarv,glob,maxiter=3) {
+runthreeH <- function(zoneC,zoneD,inHarv,glob,maxiter=2) {
   npop <- glob$numpop
   Nclass <- glob$Nclass
   hyrs <- glob$hyrs

### 2021-08-12 README.Rmd
Revised ctrlfiletemplate and datafiletemplate. No EWN
added getprojyrC and plothsstats. No EWN

@@
 
-* 2021_07-28 aMSE 0.0.0.800 Added functions (changecolumn, getrecdevcolumn, gettasdevssq) to assist in automated operating model conditioning on the fishery.
-Also modified the biology_plots and oneyear and oneyearcat to assist with speeding the processes.
-Lots of additional minor changes but some important ones to do_MSE (read the help, ?do_MSE).
+* 2021-08-12 aMSE 0.0.0.500 Modified do_MSE and added plothsstats, streamlined the conditioning. No EWN.
 
-* 2021-07-20 aMSE 0.0.0.900 Added do_condition and compareCPUE.
-Both used to speed the conditioning of the operating model, though a number of developments are still under development.
-Currently can automatically search for the AvRec value that will optimize the sum-of-squared residual fit between the
-observed CPUE in the historical period, and those predicted by the conditioned model. The undeveloped bit related to the ad hoc recruitment deviates.
+* 2021-08-02 aMSE 0.0.0.700 Revised the 'ctrlfiletemplate' and 'datafiletemplate' functions to reflect the new usage. It should now be possible to run an example scenario related to M15h75.
 
@@
 starttime <- (Sys.time())
+options("show.signif.stars"=FALSE,
+        "stringsAsFactors"=FALSE,
+        "max.print"=50000,
+        "width"=240)
+# declare libraries ------------------------------------------------------------
 library(aMSE)
@@
 library(knitr)
-# OBVIOUSLY you should modify the rundir and datadir to suit your own computer
-if (dir.exists("c:/Users/User/DropBox")) {
-  ddir <- "c:/Users/User/DropBox/A_codeUse/aMSEUse/scenarios/"
-} else {
-  ddir <- "c:/Users/Malcolm/DropBox/A_codeUse/aMSEUse/scenarios/"
-}
-doproject <- TRUE  # change to FALSE if only conditioning is required
+# Obviously you should modify the rundir and datadir to suit your own setup
+prefixdir <- "C:/A_Mal/scenarios/"
 verbose <- TRUE
-postdir <- "testnew"  # also used to label the output HTML file 
-rundir <- paste0(ddir,postdir)
-datadir <- paste0(ddir,"tasdata")
+postfixdir <- "M15h75"
+rundir <- paste0(prefixdir,postfixdir)
+datadir <- rundir
 alldirExists(rundir,datadir,verbose=verbose)
-source(paste0(datadir,"/TasmanianHS.R"))
-controlfile <- "controlsau.csv"
+controlfile <- "controlM15h75.csv"
 # equilibrium zone -------------------------------------------------------------
@@
 # 
-# Of course, usually one would use data files, control.csv and a zone.csv, which
+# Of course, usually one would use data files, control.csv and a saudata.csv, which
 # is listed as the datafile within the control.csv. These must be stored in 
-# rundir. There are example control and data files in the DropBox folder:
-# C:\Users\User\Dropbox\National abalone MSE\aMSE_files. Copy the control2.csv
-# and zonewest.csv into you rundir. Then, assuming yo have the very latest
-# version of aMSE = 5300, the following code should work. Recently both aMSE and
-# makehtml have been altered so that all references to resdir have been changed
-# to rundir, so both packages will need updating. See their respective GitHub
-# readme pages for details.
+# rundir. There are example files in the DropBox folder:
+# .\..\Dropbox\National abalone MSE\aMSE_files\scenarios\examplerun. Read the 
+# 'Using_aMSE' section in the Documentation .docx file in the documentation 
+# directry  Then, assuming you have the very latest version of aMSE = 500, 
+# the code in the documentation file (or in the 'run_aMSE.R' file within the
+# examplerun subdirectory should work.
 
-# HarvestStrategy.R should be in datadir, if it is to be shared by scenarios
-source(paste0(datadir,"/TasmanianHS.R"))
+# HarvestStrategy.R should also be in rundir
+source(paste0(rundir,"/TasmanianHS.R"))
 # run the scenario --------------------- Obviously unhash this to make it work
+#
 # out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,hcrfun=mcdahcr,
 #               sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
-#                  getdata=tasdata,calcpopC=calcexpectpopC,varyrs=7,startyr=50,
-#                  cleanslate=TRUE,verbose=TRUE,doproject=doproject,ndiagprojs=4,
-#                  openfile=TRUE,savesauout=FALSE)
-# make results webpage ---------------------------------------------------------
-# replist <- list(starttime=as.character(out$starttime),
-#                 endtime=as.character(out$projtime))
-# glb <- out$glb
-# projy <- ifelse(doproject,glb$pyrs,0)
-# runnotes <- paste0(out$ctrl$runlabel,":  RunTime = ",out$tottime,
-#                    "  replicates = ",glb$reps,",   years projected = ",projy,
-#                    "  Populations = ",glb$numpop," and SAU = ",glb$nSAU,
-#                    "  Randomseed for conditioning = ",out$ctrl$randseed)
+#               getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
+#               varyrs=7,startyr=48,cleanslate=TRUE,verbose=TRUE,
+#               ndiagprojs=4,savesauout=TRUE)
 # 
-# make_html(replist = replist,  rundir = rundir,  width = 500,  openfile = TRUE,
-#           runnotes = runnotes,   verbose = FALSE,  packagename = "aMSE",
-#           htmlname = postdir)
+# makeoutput(out,rundir,datadir,postfixdir,controlfile,openfile=TRUE,verbose=FALSE)
 

### 2021-08-19 README.Rmd
Added option of using text as saunames in controlfile. No EWN

@@
 
-* 2021-08-12 aMSE 0.0.0.500 Modified do_MSE and added plothsstats, streamlined the conditioning. No EWN.
+* 2021-08-19 aMSE 0.0.0.400 Can now have text as SAU names in the Control file, this will use those names to label all plots (still working on tbles and extra polishing),
 
-* 2021-08-02 aMSE 0.0.0.700 Revised the 'ctrlfiletemplate' and 'datafiletemplate' functions to reflect the new usage. It should now be possible to run an example scenario related to M15h75.
+* 2021-08-13   aMSE 0.0.0.500 removed NumNe (midyear numbers-at-size) and introduced cutcatchN as an argument to do_MSE, which removes all size class data from teh catchN array with size-classes less than cutcatchN, whose default = 56 (2 - 110mm removed).
+This saves a good deal of space.
 
@@
 # Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- "C:/A_Mal/scenarios/"
+prefixdir <- "C:/A_Mal/aa_scenarios/"
 verbose <- TRUE

### 2021-08-23 README.Rmd
Text names for SAU now work in all current plots and tables. No EWN.

@@
 
+* 2021-08-23 aMSE 0.0.0.300 Using text as SAU names in the control file now labels all plots and those labels are not included in the tables.
+glb$saunames now contains the text labels.
+Other changes include generalizing the number of plots to match number of SAU in all cases (I hope), There is also, now, a ConstantCatchHS.R file in the data-raw sub-directory that allows for a constant catch to be applied to each SAU.
+This demonstrates that the issue when using the TasmanianHS.R file is not the aMSE package but how the Tasmanian HS functions work with the available data.
+
 * 2021-08-19 aMSE 0.0.0.400 Can now have text as SAU names in the Control file, this will use those names to label all plots (still working on tbles and extra polishing),
-
-* 2021-08-13   aMSE 0.0.0.500 removed NumNe (midyear numbers-at-size) and introduced cutcatchN as an argument to do_MSE, which removes all size class data from teh catchN array with size-classes less than cutcatchN, whose default = 56 (2 - 110mm removed).
-This saves a good deal of space.
 

### 2021-08-23 R/domse.R

@@ -1,15 +1,15 @@
 #
 # outcond=outcond
+# postfixdir <- "SAfirst"
 # rundir <- paste0(prefixdir,postfixdir)
 # datadir <- rundir
-# postdir <- "M1h75"
-# controlfile="controlM1h75.csv"
-# hsargs=hsargs
-# hcrfun=mcdahcr
-# sampleCE=tasCPUE
-# sampleFIS=tasFIS
-# sampleNaS=tasNaS
-# getdata=tasdata
+# controlfile="controltest_SA.csv"
+# hsargs=c(0,0,0,0,0,0)
+# hcrfun=consthcr
+# sampleCE=constCPUE
+# sampleFIS=constFIS
+# sampleNaS=constNaS
+# getdata=constdata
 # calcpopC=calcexpectpopC
 # varyrs=7
 # startyr=48
@@ -17,6 +17,8 @@
 # verbose=TRUE
 # ndiagprojs=4
 # savesauout=TRUE
+# makehcrout=makeouthcr
+# cutcatchN=56
 
 #' @title do_MSE an encapsulating function to hold the details of a single run
 #'
@@ -93,6 +95,10 @@
 #'     in the catch one can remove all the empty cells below a given size
 #'     class. In the default there are 105 2mm size classes and setting
 #'     cutcatchN = 56 removes size classes 1:55 2 - 110 mm; only from catchN
+#' @param matureL is a vector of 2, default = c(70,200), that define the x-axis
+#'     bounds on the biology maturity-at-Length plots
+#' @param wtatL is a vector of 2, default = c(80,200), that define the x-axis
+#'     bounds on the biology weight-at-Length plots
 #'
 #' @seealso{
 #'  \link{makeequilzone}, \link{dohistoricC}, \link{prepareprojection},
@@ -108,7 +114,7 @@
 do_MSE <- function(rundir,controlfile,datadir,hsargs,hcrfun,sampleCE,sampleFIS,
                    sampleNaS,getdata,calcpopC,makeouthcr,varyrs=7,startyr=42,
                    cleanslate=FALSE,verbose=FALSE,ndiagprojs=3,savesauout=FALSE,
-                   cutcatchN=56) {
+                   cutcatchN=56,matureL=c(70,200),wtatL=c(80,200)) {
   # generate equilibrium zone ----------------------------------------------------
   starttime <- (Sys.time())
   zone <- makeequilzone(rundir,controlfile,datadir,cleanslate=cleanslate,
[…]

### 2021-08-23 R/getfunctions.R

@@ -620,6 +620,7 @@ getzoneprod <- function(product) {
 #' @examples
 #' data(zone)
 #' round(getzoneprops(zone$zoneC,zone$zoneD,zone$glb),4)
+#' # zoneC=zoneC; zoneD=zoneDD;glb=glb;year=hyrs
 getzoneprops <- function(zoneC,zoneD,glb,year=1) { #zoneC=zoneC; zoneD=zoneDD;glb=zone$glb; year=1
   numpop <- glb$numpop
   Nclass <- glb$Nclass
@@ -645,10 +646,10 @@ getzoneprops <- function(zoneC,zoneD,glb,year=1) { #zoneC=zoneC; zoneD=zoneDD;gl
   depletEx <- ExB/ExB0
   legaldepl <- legalmatB/B0
   propprot <- (matB - legalmatB)/matB
-  label <- c("SAU","B0","matureB","legalmatB","propprot","MSY",
+  label <- c("B0","matureB","legalmatB","propprot","MSY",
              "exB0","exploitB","SpBDepl","ExBDepl",
              "legalDepl","MSYDepl","LML","bLML","harvestR","catch")
-  ans <- rbind(sau,B0,matB,legalmatB,propprot,msy,ExB0,ExB,deplet,
+  ans <- rbind(B0,matB,legalmatB,propprot,msy,ExB0,ExB,deplet,
                depletEx,legaldepl,msydepl,lml,blml,harvestR,catch)
   rownames(ans) <- label
   tot <- numeric(length(rownames(ans))); names(tot) <- label
@@ -663,7 +664,7 @@ getzoneprops <- function(zoneC,zoneD,glb,year=1) { #zoneC=zoneC; zoneD=zoneDD;gl
   tot["bLML"] <- sum(wgts * ans["bLML",])
   tot["harvestR"] <- mean(ans["harvestR",])
   tot["catch"] <- sum(ans["catch",])
-  ans <- cbind(ans,tot)
+  ans <- as.data.frame(round(cbind(ans,tot),4))
   colnames(ans) <- c(paste0("p",1:numpop),"zone")
   return(ans)
 }  # end of getzoneprops

### 2021-09-02 README.Rmd
Generalized many of the output labels on plts and tables. No EWN.

@@
 
+* 2021-09-03 aMSE 0.0.0.200 Lots of tidying of help pages, added numbersatsizeSAU. Added a fixed TasmanianHS.R file to the data-raw directory.
+
 * 2021-08-23 aMSE 0.0.0.300 Using text as SAU names in the control file now labels all plots and those labels are not included in the tables.
@@
 This demonstrates that the issue when using the TasmanianHS.R file is not the aMSE package but how the Tasmanian HS functions work with the available data.
-
-* 2021-08-19 aMSE 0.0.0.400 Can now have text as SAU names in the Control file, this will use those names to label all plots (still working on tbles and extra polishing),
 
@@
 # Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- "C:/A_Mal/aa_scenarios/"
-verbose <- TRUE
+prefixdir <- "C:/Users/Malcolm/Dropbox/A_CodeUse/aMSEUse/scenarios/"
 postfixdir <- "M15h75"
@@
 datadir <- rundir
-alldirExists(rundir,datadir,verbose=verbose)
+alldirExists(rundir,datadir,verbose=TRUE)
 controlfile <- "controlM15h75.csv"

### 2021-09-02 R/plotfuns.r

@@ -529,11 +529,11 @@ plotNt <- function(Nt,year,glb,start=3,medcol=0) {
     sdat <- Nt[,year,sau,]
     saumedians[,sau] <- apply(sdat,1,median)
     ymax <- getmax(sdat[start:Nclass,])/1000.0
-    label <- paste0("N '000s  sau",saunames[sau])
+    label <- paste0("N '000s  ",saunames[sau])
     plot(midpts,sdat[,1]/1000.0,type="l",lwd=1,col="grey",panel.first=grid(),
          ylim=c(0,ymax),ylab=label)
     for (i in 1:reps) lines(midpts,sdat[,i]/1000.0,col="grey")
-    lines(midpts,sdat[,1]/1000.0,lwd=2,col=2)
+   # lines(midpts,sdat[,1]/1000.0,lwd=2,col=2)
     lines(midpts,saumedians[,sau]/1000.0,lwd=2,col=medcol)
   }
   return(invisible(saumedians))

### 2021-09-10 README.Rmd
Modified getdata to include NAS data. New saurecdevs plot. NO EWN

@@
 
+* 2021-09-10 aMSE 0.0.0.100 Many small, but significant, changes. Modified getdata which interacts with the ###HS.R file and added the
+zoneDP$NAS data input. Added a summary recdev plot to condition tab.
+
 * 2021-09-03 aMSE 0.0.0.200 Lots of tidying of help pages, added numbersatsizeSAU. Added a fixed TasmanianHS.R file to the data-raw directory.
-
-* 2021-08-23 aMSE 0.0.0.300 Using text as SAU names in the control file now labels all plots and those labels are not included in the tables.
-glb$saunames now contains the text labels.
-Other changes include generalizing the number of plots to match number of SAU in all cases (I hope), There is also, now, a ConstantCatchHS.R file in the data-raw sub-directory that allows for a constant catch to be applied to each SAU.
-This demonstrates that the issue when using the TasmanianHS.R file is not the aMSE package but how the Tasmanian HS functions work with the available data.
 
@@
 # HarvestStrategy.R should also be in rundir
-source(paste0(rundir,"/TasmanianHS.R"))
+source(paste0(rundir,"/TasHS1_Tas.R"))
 # run the scenario --------------------- Obviously unhash this to make it work

### 2021-09-10 data-raw/candidatefunctions.R

@@ -800,3 +800,35 @@ ndiagprojs=4
 savesauout=TRUE
 makehcrout=makeouthcr
 cutcatchN=56
+
+
+dat <- out$condout$sauZone$recruit[,1]
+glb <- out$glb
+loessfit <- loess(dat ~ as.numeric(glb$hyrnames),span=0.625)
+
+plotprep(newdev=FALSE)
+parset()
+ymax <- getmax(dat)
+plot(glb$hyrnames,dat,type="l",ylim=c(0,ymax),panel.first=grid())
+lines(x$x,x$fitted,lwd=2,col=2)
+
+
+
+
+# Plot RecDevs -------------------------------------------------------------
+
+
+
+
+recdevs <- out$condC$recdevs
+glb <- out$glb
+
+saurecdevs(recdevs,glb,rundir)
+
+
+
+
+
+
+
+

### 2021-09-13 README.Rmd
Added option of initial depletion as well as historical catches. Now have TasHS library. No EWN

@@
 # LATEST UPDATE
+
+* 2021-09-13 aMSE 0.0.0.075 Implemented the option of having an initial depletion as well as application of historical catches.
 
@@
 library(aMSE)
+library(TasHS)
 library(rutilsMH)
@@
 # Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- "C:/Users/Malcolm/Dropbox/A_CodeUse/aMSEUse/scenarios/"
+prefixdir <- "C:/Users/user/Dropbox/A_CodeUse/aMSEUse/scenarios/"
 postfixdir <- "M15h75"

### 2021-09-13 R/aMSE_utils.R

@@ -652,7 +652,7 @@ sumpop2sau <- function(invect,sauindex) {
 #'     depleB are the end of year matureB and exploitB divided by their
 #'     respective unfished estimated by SAU obtained using getvar(zoneC,"B0").
 #'
-#' @param inzone one of the zone dynamics objects made up of populations
+#' @param inzone the projected zone dynamics objects made up of populations
 #' @param NAS the numbers-at-size 4D arrays from doprojection; default=NULL so
 #'     it can be ignored during conditioning
 #' @param glb the object containing the global constants

### 2021-09-17 README.Rmd
Initial depletion now operational. No EWN

@@
 
+* 2021-09-17 aMSE 0.0.0.050 Finalized details of the optional initial depletion. Amended the logistic function so it now works when using the knifeedge option!
+
 * 2021-09-13 aMSE 0.0.0.075 Implemented the option of having an initial depletion as well as application of historical catches.
-
-* 2021-09-10 aMSE 0.0.0.100 Many small, but significant, changes. Modified getdata which interacts with the ###HS.R file and added the
-zoneDP$NAS data input. Added a summary recdev plot to condition tab.
-
-* 2021-09-03 aMSE 0.0.0.200 Lots of tidying of help pages, added numbersatsizeSAU. Added a fixed TasmanianHS.R file to the data-raw directory.
 

### 2021-09-17 data-raw/candidatefunctions.R

@@ -89,33 +89,6 @@ for (sau in 1:nsau) { #  sau=1
 }
 
 
-# Read the Obs LF-comp data-----------------------------------------------
-
-options("show.signif.stars"=FALSE,
-        "stringsAsFactors"=FALSE,
-        "max.print"=50000,
-        "width"=240)
-# declare libraries ------------------------------------------------------------
-library(aMSE)
-library(rutilsMH)
-# Obviously you should modify the rundir and datadir to suit your own setup
-if (dir.exists("c:/Users/User/DropBox")) {
-  ddir <- "c:/Users/User/DropBox/A_codeUse/aMSEUse/scenarios/"
-} else {
-  ddir <- "c:/Users/Malcolm/DropBox/A_codeUse/aMSEUse/scenarios/"
-}
-doproject <- FALSE  # change to FALSE if only conditioning is required
-verbose <- TRUE
-rundir <- paste0(ddir,"HS652510")
-datadir <- paste0(ddir,"tasdata")
-
-zone1 <- readctrlfile(rundir,infile="controlsau.csv",datadir=datadir,verbose=verbose)
-
-compdat <- zone1$condC$compdat
-
-
-palfs=compdat$palfs
-palfs
 
 
 # copyto -------------------------------------------------------------
@@ -130,32 +103,6 @@ copyto(rundir,todir=destdir,filename="controlsau.csv")
   #     zoned=zoneDsau;zonep=zonePsau;glb=glb;startyr=30;
   #     picksau=9; histCE=histCE;CIprobs=c(0.05,0.5,0.95); addCI=TRUE
 
-
-
-# sort out LF data-------------------------------------------------
-
-datafile <- paste0(ddir,"aMSEUse/condition/productivity/sau10LF.csv")
-
-lf <- read.csv(file=datafile,header=TRUE)
-str1(lf)
-
-
-yrcount <- tapply(lf[,"counts"],lf[,"year"],sum,na.rm=TRUE)
-lfprops <- cbind(years,yrcount)
-lfprops
-
-range(lf[,"length"])
-mids <- seq(136,210,2)
-nc <- length(mids)
-nc
-
-
-
-
-answer <- makewidedat(lf,mids)
-str(answer)
-
-answer <- makewidedat(lf10,mids)
 
 
 
@@ -380,17 +327,6 @@ for (i in 1:nsau) {
 
[…]

### 2021-09-22 README.Rmd
Simplified modification of multiple control and saudata files. No EWN

@@
 
+* 2021-09-22 aMSE 0.0.0.025 Modifications to do_condition and other functions to facilitate the ease of changing values within an array of control files and saudata files.
+
 * 2021-09-17 aMSE 0.0.0.050 Finalized details of the optional initial depletion. Amended the logistic function so it now works when using the knifeedge option!
-
-* 2021-09-13 aMSE 0.0.0.075 Implemented the option of having an initial depletion as well as application of historical catches.
 
@@
 # Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- "C:/Users/user/Dropbox/A_CodeUse/aMSEUse/scenarios/"
+prefixdir <- "C:/Users/Malcolm/Dropbox/A_CodeUse/aMSEUse/scenarios/"
 postfixdir <- "M15h75"

### 2021-09-22 R/docondition.R

@@ -158,6 +158,7 @@ changeline <- function(indir, filename, linenumber, newline,verbose=FALSE) {
 #'     the run then 'cleanslate' will not affect the data or any .R files.
 #' @param verbose Should progress comments be printed to console, default=FALSE
 #' @param doproduct should production estimates be made. default=FALSE
+#' @param dohistoric should the historical catches be applied. Default=TRUE
 #'
 #' @seealso{
 #'  \link{makeequilzone}, \link{dohistoricC}, \link{prepareprojection},
@@ -170,8 +171,10 @@ changeline <- function(indir, filename, linenumber, newline,verbose=FALSE) {
 #'
 #' @examples
 #' print("wait on suitable data sets in data")
+#' # rundir=rundir; controlfile=controlfile;datadir=datadir;calcpopC=calcexpectpopC
+#' # cleanslate=FALSE; verbose=TRUE; doproduct=TRUE; dohistoric=FALSE
 do_condition <- function(rundir,controlfile,datadir,calcpopC,cleanslate=FALSE,
-                         verbose=FALSE,doproduct=FALSE) {
+                         verbose=FALSE,doproduct=FALSE,dohistoric=TRUE) {
   starttime <- Sys.time()
   zone <- makeequilzone(rundir,controlfile,datadir,doproduct=doproduct,
                         cleanslate=cleanslate,verbose=verbose)
@@ -193,27 +196,35 @@ do_condition <- function(rundir,controlfile,datadir,calcpopC,cleanslate=FALSE,
     if (verbose) cat("Conducting initial depletions  ",initdepl,"\n")
     zoneD <- depleteSAU(zoneC,zoneD,glb,initdepl=initdepl,production)
   }
-  if (verbose) cat("Conditioning on the Fishery data  \n")
-  zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,calcpopC=calcpopC,
-                        sigR=1e-08,sigB=1e-08)
+  if (dohistoric) {
+    if (verbose) cat("Conditioning on the Fishery data  \n")
+    zoneDD <- dohistoricC(zoneD,zoneC,glob=glb,condC,calcpopC=calcpopC,
+                          sigR=1e-08,sigB=1e-08)
+  } else {
+    zoneDD <- zoneD
+  }
   hyrs <- glb$hyrs
   propD <- t(getzoneprops(zoneC,zoneDD,glb,year=hyrs))
[…]

### 2021-09-22 R/plotfuns.r

@@ -33,7 +33,8 @@ compareCPUE <- function(histCE,saucpue,glb,rundir,filen="") {
   ssq <- numeric(nsau)
   doplots=getparplots(nsau)
   plotprep(width=8,height=8,newdev=FALSE,filename=filen,verbose=FALSE)
-  parset(plots=doplots,margin=c(0.3,0.3,0.05,0.05),outmargin=c(0,1,0,0))
+  parset(plots=doplots,margin=c(0.3,0.3,0.05,0.05),outmargin=c(0,1,0,0),
+         byrow=FALSE)
   for (sau in 1:nsau) {
     ssq[sau] <- sum((histCE[,sau] - cpue[,sau])^2,na.rm=TRUE)
     ymax <- getmax(c(cpue[,sau],histCE[,sau]))

### 2021-09-22 data-raw/almostaccepted.R

@@ -36,6 +36,105 @@ combinepopN <- function(inN,glb) {
 } # end of combinepopN
 
 
+
+#' @title getavrec pulls out just the AvRec values for each sau for a scenario
+#'
+#' @description getavrec extracts the AvRec values for each SAU for a given
+#'     scenario. It does this by reading in the saudata file and using grep
+#'     to search for the correct line and returning the line as is. Care is
+#'     required to only take the first record of 'AvRec' so as not to
+#'     consider the variability in sAvRec.
+#'
+#' @param datadir the directory in which one finds the saudata file for the
+#'     given scenario
+#' @param datafile the exact name of the saudata file
+#' @param nsau the number of SAU to be found
+#'
+#' @return a numeric vector of the average recruitment for each SAU
+#' @export
+#'
+#' @examples
+#' \dontrun{
+#'   rundir <- "c:/Users/User/DropBox/A_codeUse/aMSEUse/scenarios/MhLML/M1h5/"
+#'   getavrec(rundir,"saudataM1h5.csv",8)
+#' }
+getavrec <- function(datadir,datafile,nsau) {
+  filen <- filenametopath(datadir,datafile)
+  dat <- readLines(filen)
+  pickA <- grep("AvRec",dat)[1] # ignore sAvRec
+  avrec <- getConst(dat[pickA],nsau)
+  return(avrec)
+} # end of getavrec
+
+
+
+
+#' @title sauavrecssq applies the AvRec and returns the ssq from the cpue
+#'
+#' @description sauavrecssq applies the AvRec vector while adjusting just
+#'     one for the picksau, and then compares the predicted CPUE with the
+#'     observed and returns a sum-of-squared differences. This is used by
+#'     optim while using the Brent option to find an optimum for a single
+#'     parameter.
+#'
+#' @param param is the AvRec for the selected SAU
+#' @param rundir the rundir for the scenario being considered
+#' @param datadir the directory holding the saudata file, usually = rundir
+#' @param controlfile the name of the control file, no path
+#' @param datafile the name of the data file, no path
+#' @param linenum the linenumber containing the AvRec vector
+#' @param calcpopC the HS function that calculates the aspirational catches
+#'     for each SAU
[…]

### 2021-10-17 README.Rmd
Removed reference to cleanslate througout. No EWN.

@@
 # LATEST UPDATE
+
+* 2021-10-17 aMSE 0.0.0.010 As cleanslate, within setuphtml() has been deprecated in the package makehtml, I have removed reference to it throughout aMSE.
+Now, to clean a directory of aMSE files use cleanrundir.
 
@@
 # Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- "C:/Users/Malcolm/Dropbox/A_CodeUse/aMSEUse/scenarios/"
+prefixdir <- "C:/Users/User/Dropbox/A_CodeUse/aMSEUse/scenarios/"
 postfixdir <- "M15h75"
@@
 #               getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
-#               varyrs=7,startyr=48,cleanslate=TRUE,verbose=TRUE,
+#               varyrs=7,startyr=48,verbose=TRUE,
 #               ndiagprojs=4,savesauout=TRUE)

### 2021-11-21 README.Rmd
Added plotsizecomp, popNAStosau, and preparesizecomp. No EWN

@@
 
+* 2021-11-21 aMSE 0.0.0.020 So as to use any observed size-composition data in the conditioning, a number of functions have been added to convert the predicted numbers-at-size in the catch by population into NAS x SAU.
+These can now be plotted against the observed NAS for each SAU and this can be either to the console or to a new tab in the webpage called 'predictedcatchN'.
+This work derived from the progress made with the sizemod package used to fit size-based assessment models to individual SAU.
+Because the MSE sub-divides each SAU into a number of populations each with somewhat different MaxDL, L95, etc, even after fitting the models, when using that data it is then necessary to refit the AvRec values, which, because of the sub-division into populations each having different biology can lead to the MSE AvRec being either smaller or larger than the estimated R0 for a SAU.
+Similar arguments apply to teh recruitment residuals but generally the pattern provides a reasonable fit.
+
 * 2021-10-17 aMSE 0.0.0.010 As cleanslate, within setuphtml() has been deprecated in the package makehtml, I have removed reference to it throughout aMSE.
 Now, to clean a directory of aMSE files use cleanrundir.
-
-* 2021-09-22 aMSE 0.0.0.025 Modifications to do_condition and other functions to facilitate the ease of changing values within an array of control files and saudata files.
-
-* 2021-09-17 aMSE 0.0.0.050 Finalized details of the optional initial depletion. Amended the logistic function so it now works when using the knifeedge option!
 

### 2021-11-21 data-raw/candidatefunctions.R

@@ -1,5 +1,11 @@
 
 
+
+
+
+
+
+
 # prepare size-composition data ------------------------------------------------
 
 zoneCP <- out$zoneCP

### 2021-11-23 README.Rmd
Improved do_condition and simplified optimization of AvRec. No EWN.

@@
 # LATEST UPDATE
+
+* 2021-11-23 aMSE 0.0.0.1 Added functions to simplify the optimization of AvRec relative to the CPUE data after initial parameter estimation for each SAU using the sizemod package.
+Also added a comparison of size-composition data to the do_condition function.
+Next step will be to include the optimization of the recdevs using both CPUE and size-comp data.
 

### 2021-11-23 R/aMSE.R

@@ -3,7 +3,7 @@
 #' @importFrom graphics grid hist lines mtext par plot points title
 #' @importFrom graphics legend abline text layout
 #' @importFrom stats dnorm pnorm rlnorm rnorm quantile sd lm coef median
-#' @importFrom stats loess
+#' @importFrom stats loess optim
 #' @importFrom utils object.size browseURL packageDescription read.csv
 #' @importFrom utils write.table head tail winDialog
 #' @importFrom rutilsMH getmax getmin setpalette plotprep removeEmpty

### 2021-11-23 R/aMSE_utils.R

@@ -347,45 +347,6 @@ makewidedat <- function(inlong,mids,counts=FALSE) { # inlong=lf; mids=mids
   return(answer)
 } # end of makewidedat
 
-#' @title popNAStosau converts NAS data by population to NAS x sau
-#'
-#' @description popNAStosau converts numbers-at-size data x population into
-#'     the same by SAU. This is designed to compare ny observed NAS data from
-#'     a zone with those predicted by the conditioned model, it could also be
-#'     used to improve the conditioning code to avoid fitting only to CPUE
-#'
-#' @param popNAS either the catchN or the Nt from zoneDD, that is the zone
-#'     after the equilibrium zone (zoneD) has had the historical catches
-#'     applied.
-#' @param glb the globals object. Needed for the sauindex and various array
-#'     dimnesion labels.
-#'
-#' @return An Nclass x nyrs x nsau array of numbers-at-size data
-#' @export
-#'
-#' @examples
-#'  #e.g.  popNAStosau(out$zoneDD$catchN,out$glb)
-popNAStosau <- function(popNAS,glb) { # popNAS <- catchN; glb <- out$glb
-  nsau <- glb$nSAU
-  sauindex <- glb$sauindex
-  saunames <- glb$saunames
-  yrnames <- glb$hyrnames
-  N <- glb$Nclass
-  nyrs <- glb$hyrs
-  dims <- dim(popNAS)
-  if (length(dims) == 3) {
-    sauNt <- array(data=0,dim=c(N,nyrs,nsau),
-                   dimnames=list(glb$midpts,yrnames,saunames))
-    for (sau in 1:nsau) { # sau=1; yr=2
-      pick <- which(sauindex == sau)
-      for (yr in 1:nyrs) sauNt[,yr,sau] <- rowSums(popNAS[,yr,pick])
-    }
-    return(sauNt)
-  } else {
-    stop("popNttosau currently designed only for zoneDD  \n")
-  }
-} # end of popNAStosau
-
 
 #' @title poptosau converts projected population dynamics to SAU scale results
 #'

### 2021-11-29 README.Rmd
Removed datadir. Added optimizerecdevs, started implementing hyper-stability of cpue. No EWN.

@@
 # LATEST UPDATE
+
+* 2021-11-29 aMSE 0.0.1 Removed datadir. Started implementing hyper-stability while keeping the interface the same, so that the results from sizemod can be used.
 
@@
 library(knitr)
+dbdir <- getDBdir()
 # Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- "C:/Users/User/Dropbox/A_CodeUse/aMSEUse/scenarios/"
+prefixdir <- paste0(dbdir,"A_CodeUse/aMSEUse/scenarios/")
 postfixdir <- "M15h75"

### 2021-11-29 R/getfunctions.R

@@ -7,7 +7,7 @@
 #'     required to only take the first record of 'AvRec' so as not to
 #'     consider the variability in sAvRec.
 #'
-#' @param datadir the directory in which one finds the saudata file for the
+#' @param rundir the directory in which one finds the saudata file for the
 #'     given scenario
 #' @param datafile the exact name of the saudata file
 #' @param nsau the number of SAU to be found
@@ -20,8 +20,8 @@
 #'   rundir <- "c:/Users/User/DropBox/A_codeUse/aMSEUse/scenarios/MhLML/M1h5/"
 #'   getavrec(rundir,"saudataM1h5.csv",8)
 #' }
-getavrec <- function(datadir,datafile,nsau) {
-  filen <- filenametopath(datadir,datafile)
+getavrec <- function(rundir,datafile,nsau) {
+  filen <- filenametopath(rundir,datafile)
   dat <- readLines(filen)
   pickA <- grep("AvRec",dat)[1] # ignore sAvRec
   avrec <- getConst(dat[pickA],nsau)
@@ -167,7 +167,7 @@ getConst <- function(inline,nb,index=2) { # parses lines containing numbers
 #'     the SAU. See the 'data-file-description.docx', and the example data files
 #'     to see examples of the required format.
 #'
-#' @param datadir the directory containing the LF data files
+#' @param rundir the directory containing the LF data files
 #' @param filename the complete filename for the length-composition data file
 #'
 #' @seealso{
@@ -181,8 +181,8 @@ getConst <- function(inline,nb,index=2) { # parses lines containing numbers
 #'
 #' @examples
 #' print("wait on suitable internal data files")
-getLFdata <- function(datadir,filename) {
-  filen <- filenametopath(datadir,filename)
+getLFdata <- function(rundir,filename) {
+  filen <- filenametopath(rundir,filename)
   lfdat <- read.csv(filen,header=TRUE)
   colnames(lfdat) <- tolower(colnames(lfdat))
   sau <- sort(unique(lfdat[,"sau"]))

### 2021-11-29 R/projection.R

@@ -545,9 +545,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
   exb0 <- getvar(zoneCP,"ExB0")
   hcrout <- makehcrout(glb)
   for (iter in 1:reps) {
-    if (verbose) {
-      if ((iter %% 25) == 0) cat(iter," \n")
-    }
+    if (verbose) if ((iter %% 25) == 0) cat(iter,"   ")
     for (year in startyr:endyr) { # iter=1; year=startyr
       hcrdata <- getdata(sampleCE,sampleFIS,sampleNaS,sauCPUE=zoneDP$cesau,
                          sauacatch=zoneDP$acatch,sauNAS=zoneDP$NAS,
@@ -580,6 +578,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
       zoneDP$TAC[year,iter] <- hcrout$TAC
     } # year loop
   }   # iter loop
+  cat("\n")
   return(list(zoneDP=zoneDP,hcrout=hcrout))
 } # end of doprojections
 

### 2022-05-24 README.Rmd
Bringing everything up to date. Its been a while!

@@
 # declare libraries ------------------------------------------------------------
+suppressPackageStartupMessages({
   library(aMSE)
   library(TasHS)
-library(rutilsMH)
+  library(hutils)
+  library(hplot)
   library(makehtml)
   library(knitr)
+})
 dbdir <- getDBdir()

### 2022-05-24 R/defineZone.R

@@ -1,4 +1,4 @@
-#  rutilsMH::listFunctions("C:/Users/User/Dropbox/A_Code/aMSE/R/defineZone.R")
+#  hutils::listFunctions("C:/Users/User/Dropbox/A_Code/aMSE/R/defineZone.R")
 
 #' @title defineBlock subdivides defined number of populations into blocks
 #'
@@ -596,7 +596,7 @@ makeabpop <- function(popparam,midpts,projLML) {
 #' @examples
 #' print("wait on datafiles")
 makeequilzone <- function(rundir,ctrlfile="control.csv",doproduct=TRUE,verbose=TRUE) {
- #  rundir=rundir;ctrlfile=controlfile;verbose=verbose;doproduct=TRUE; verbose=TRUE
+ #  rundir=rundir;ctrlfile=controlfile;doproduct=FALSE; verbose=TRUE
   zone1 <- readctrlfile(rundir,infile=ctrlfile,verbose=verbose)
   ctrl <- zone1$ctrl
   glb <- zone1$globals     # glb without the movement matrix

### 2022-05-24 R/projection.R

@@ -479,10 +479,11 @@ prepareprojection <- function(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
 #' @title doprojections conducts the replicate model runs for Tasmania
 #'
 #' @description doprojections conducts the replicate model runs for
-#'     Tasmania using the mcdahcr and the input hsargs. The model dynamics and
-#'     the numbers-at-size variables have been separated and returned separately
-#'     because the size components are very large and mean saving such objects
-#'     is both slow and space limiting.
+#'     Tasmania using the mcdahcr and the input hsargs. Note the use of a list
+#'     that takes up both the population and catch numbers-at-size and puts that
+#'     into getdata. By referencing the iteration in the arrays being passed to
+#'     getdata the object sizes are being reduced and hence, hopefully, the
+#'     runtime will be reduced accordingly.
 #'
 #' @param ctrl the ctrl object from readctrlfile
 #' @param zoneDP the object used to contain the dynamics from the replicate
@@ -500,7 +501,8 @@ prepareprojection <- function(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
 #' @param sampleNaS a function that generates the Numbers-at-size samples
 #' @param getdata a function that gathers all the data required by the hcrfun
 #'     and combines it into an hcrdata object ready for the hcrfun. It is
-#'     expected to call sampleCE, sampleFIS, and sampleNAS.
+#'     expected to call sampleCE, sampleFIS, and sampleNAS, even if they only
+#'     return NULL.
 #' @param calcpopC a function that takes the output from hcrfun and generates
 #'     the actual catch per population expected in the current year.
 #' @param makehcrout is a function from HS.R that produces an object that is
@@ -512,7 +514,7 @@ prepareprojection <- function(projC=projC,condC=condC,zoneC=zoneC,glb=glb,
 #'     in the default named collection
 #'
 #' @seealso{
-#'  \link{oneyearsauC}, \link[makehtml]{make_html}. \link{do_MSE}
+#'  \link{oneyearsauC}, \link[makehtml]{make_html}, \link{do_MSE}
 #' }
 #'
[…]

### 2022-06-06 README.Rmd
Added getssqparts to assist conditioning the MSE. No EWN
The option of hyperstability now correctly implemented, as is the expression of exploitable biomass through the year. No EWN.

@@
 # LATEST UPDATE
+
+* 2022-06-06 aMSE 0.0.5.
+Have reviewed the dynamics in detail so that the outputs from the aMSE for each SAU more closely match the outputs from sizemod.
+The original disjunction was partly due to a faulty implementation of the inclusion of the option of hyper-stability, but that has now been amended and the operatnig model is now behaving as it should.
+Breaking the dynamics within each SAU into multiple sub-populations with variation also affects the equivalence between aMSE and sizemod butmostly influences the productivity rather than the recruitment dynamics and the trajectories mainly follow those from sizemod but with the cpue and biomasses being projected incorrectly.
+It appears (still working on a why) that by changing the productivity and sub-dividing that productivity among sub-populations, the overall recruitment levels required to mimic the dynamics in all cases appears to be less in aMSE than in sizemod.
+Otheriwse nothing now needs to change.
+One uses parameters obtained from sizemod, especially the recruitment deviates, and then apply the functions that search for more optimal values for AvRec, the unfished recruitment levles in each SAU.
+The fits obtained should be close to those found in sizemod.
+In 0.0.5 there remain issues relating to the internal data sets, which include one called 'zone', which is not designed to work with the option of hyper-stability (lambda != 1.0).
+Otherwise, No EWN.
+
+* 2022-05-31 aMSE 0.0.4 Added functions that assist in conditioning the operating model by optimizing the fit to CPUE and size-composition data after opimizing the AvRec values for each SAU. No EWN.
 

### 2022-06-06 R/defineZone.R

@@ -77,7 +77,7 @@ definepops <- function(inSAU,inSAUindex,const,glob) {
   hyrs <- glob$hyrs
   columns <- c("DLMax","L50","L95","SigMax","SaMa","SaMb","Wta","Wtb","Me",
                "L50C","deltaC","AvRec","SelP1","SelP2","Nyrs","steeph",
-               "MaxCE","L50mat","SAU","lambda","qest")
+               "MaxCE","L50mat","SAU","lambda","qest","scalece")
   popdefs <- matrix(0,nrow=numpop,ncol=length(columns),
                     dimnames=list(1:numpop,columns))
   popdefs[,"Nyrs"] <- rep(hyrs,numpop) # Num Years - why is this here?
@@ -110,12 +110,11 @@ definepops <- function(inSAU,inSAUindex,const,glob) {
                                    const["sdefsteep",pop])
     popdefs[pop,"AvRec"] <- rlnorm(1,meanlog=const["AvRec",pop],
                                    const["sAvRec",pop])
-    popdefs[pop,"MaxCE"] <- rnorm(1,mean=const["MaxCEpars",pop],
-                                  const["sMaxCEpars",pop])
+    popdefs[pop,"MaxCE"] <- 0 # gets value in makezone
     popdefs[pop,"SAU"] <- const["SAU",pop]
     popdefs[pop,"lambda"] <- const["lambda",pop]
-    meanq <- const["qest",pop]
-    popdefs[pop,"qest"] <- rnorm(1,mean=meanq,sd=meanq/100.0)
+    popdefs[pop,"qest"] <-const["qest",pop]
+    popdefs[pop,"scalece"] <- 0 # gets value in makezone
   }
   test <- popdefs[,"L95"] - popdefs[,"L50"]
   if (any(test < 0)) stop("L95 < L50 - adjust the input data file  \n")
@@ -506,7 +505,7 @@ logistic <- function(inL50,delta,lens,knifeedge=0) {
 #'
 #' @description makeabpop generates the full population structure in an
 #'    unfished state. The structure is pre-determined: Me R0 B0 effB0
-#'    ExB0, effExB0 MSY MSYDepl bLML popq SaM popdef (vector of constants)
[…]

### 2022-06-21 README.Rmd
Amended the template egnerotrs. No EWN.
Now incluodes checkmsedata, comparevar, and plotscene. Can now import outputs from sizemod. No EWN.

@@
 # LATEST UPDATE
+
+* 2022-06-22 aMSE 0.0.7 Now have rewritecontrolfile and rewritesaudatafile and the option to include a table of parameters from sizemod into the inputs of aMSE.
+Have begun development of functions to aid comparing the output from different scenarios, including tables and plots.
+No EWN.
+
+* 2022-06-08 aMSE 0.0.6 Have revised the ctrltemplate and datatemplate generators so that they reflect a lambda = 0.75 and correctly use the hyperstability. No EWN
 

### 2022-07-03 README.Rmd
Fixed plotting of harvest rates in output. Added checkcompdata function.

@@
 # LATEST UPDATE
+
+* 2022-06-29 aMSE 0.0.8 Corrected a large mistake where the estimated SAU harvest rates were being biased high because the SAU's end of year exploitable biomass was being used instead of the mid-year exploitable biomass.
 
@@
 rundir <- paste0(prefixdir,postfixdir)
-datadir <- rundir
-alldirExists(rundir,datadir,verbose=TRUE)
+confirmdir(rundir)
 controlfile <- "controlM15h75.csv"

### 2022-07-03 R/aMSE.R

@@ -24,9 +24,9 @@ NULL
 #'     to age accurately and consistently). The growth dynamics are
 #'     described using the inverse logistic curve (see Haddon et al.
 #'     2008). Earlier versions of this invertebrate MSE are described
-#'     on A development version is #useDynLib aMSE, registration = TRUE
-#'     #importom Rcpp evalCpp
-#'     available on GitHub at github.com/haddonm/aMSE.
+#'     in Haddon et al, (2013) and Haddon and Mundy (2016). A development
+#'     version of this R package is available on GitHub at
+#'     https://github.com/haddonm/aMSE.
 #'
 #' @references Haddon, M., Mundy, C., and D. Tarbath (2008) Using an
 #'     inverse-logistic model to describe growth increments of blacklip

### 2022-07-03 data-raw/cleanR.R

@@ -1,1 +1,93 @@
 
+
+
+
+
+#' @title alldirExists Checks the existence of both a run and data directory
+#'
+#' @description alldirExists answers the questions 'do both a rundir and a
+#'     datadir directory exist?' It uses dir.exists and reports existence if
+#'     already present and provides a stop warning if either does not exist. Of
+#'     course it can also be used to determine whether a single directory exists.
+#'     By default the second directory, indir2 = indir1. This allows for the
+#'     datadir = rundir within aMSE, but also allows for a separate datadir.
+#'     was in aMSE_utils.R
+#'
+#' @param indir1 a character string containing the name of the first directory
+#'     whose existence is to be checked before it is created if it
+#'     does not already exist.
+#' @param indir2 a character string containing the name of a second directory
+#'     whose existence is to be checked, by default this has the same value as
+#'     indir1
+#' @param make if the directory does NOT exist should it be created.
+#'     default = FALSE; if make=FALSE and a directory does not exist
+#'     a warning will be given to the console.
+#' @param verbose default=TRUE, prints directory status to the console,
+#'     If make is set to FALSE and a directory does not exist a
+#'     warning will always be given.
+#'
+#' @return a message to the screen if the directory exists or is
+#'     created; if make is TRUE then it also creates the directory as
+#'     listed in 'indir1'.
+#' @export
+#'
+#' @examples
+#' indirect <- getwd()
+#' alldirExists(indirect)
+alldirExists <- function(indir1,indir2=indir1,make=FALSE,verbose=TRUE) {
+  if (dir.exists(indir1)) {
+    if (verbose) cat("rundir, ",indir1,":  exists  \n")
+  } else {
+    if (make) {
+      dir.create(indir1, recursive = TRUE)
+      if (verbose) cat(indir1,":  created  \n")
+    } else {
+      warning(cat(indir1,":  does not exist \n"))
+    }
+  }
+  if (indir2 != indir1) {
+    if (dir.exists(indir2)) {
+      if (verbose) cat("datadir, ",indir2,":  exists  \n")
+    } else {
+      if (make) {
+        dir.create(indir2, recursive = TRUE)
+        if (verbose) cat("datadir, ",indir2,":  created  \n")
+      } else {
+        warning(cat("datadir, ",indir2,":  does not exist \n"))
+      }
+    }
[…]

### 2022-07-10 data-raw/MakeDatafor_aMSE.r
Ensured temp[late files operational, and all examples work. Production ready. No EWN.

@@ -6,7 +6,8 @@ options("show.signif.stars"=FALSE,"stringsAsFactors"=FALSE,
 #     listFunctions("C:/A_Mal/A_Book/rcode/agestruct/age_utils.r")
 
 library(aMSE)
-library(rutilsMH)
+library(hutils)
+library(hplot)
 library(makehtml)
 
 if (dir.exists("c:/Users/User/DropBox")) {
@@ -34,13 +35,60 @@ save(zone1,file=paste0(datadir,"zone1.RData"))
 save(constants,file=paste0(datadir,"constants.RData"))
 
 
+# make data reday for twosau single population examples-----------------
+options("show.signif.stars"=FALSE,"stringsAsFactors"=FALSE,
+        "max.print"=50000,"width"=240)
+#     listFunctions("C:/A_Mal/A_Book/rcode/agestruct/age_utils.r")
+
+  library(aMSE)
+  library(hutils)
+  library(TasHS)
+  library(makehtml)
+  dropdir <- getDBdir()
+  prefixdir <- paste0(dropdir,"A_codeUse/aMSEUse/scenarios/")
+  postfixdir <- "S21"
+  verbose <- TRUE
+  rundir <- filenametopath(prefixdir,postfixdir)
+  ctrlfile <- paste0("control",postfixdir,".csv")
+
+  doproduct=TRUE
+
+  zone1 <- readctrlfile(rundir,infile=ctrlfile,verbose=verbose)
+  ctrl <- zone1$ctrl
+  glb <- zone1$globals     # glb without the movement matrix
+  bysau <- zone1$ctrl$bysau
+  opar <- NULL
+  parsin <- FALSE
+  saudata <- readsaudatafile(rundir,ctrl$datafile,optpar=opar)
+  constants <- saudata$constants
+  saudat <- saudata$saudat
+  zone1$condC$poprec <- saudata$poprec
+
+  if (verbose) cat("Files read, now making zone \n")
+  out <- setupzone(constants,zone1,doproduct,verbose=verbose) # make operating model
+  zoneC <- out$zoneC
+  zoneD <- out$zoneD
+  glb <- out$glb             # glb now has the movement matrix
+  product <- out$product     # important bits usually saved in rundir
+  zone1$globals <- glb
+
+
+  datadir <- paste0(dropdir,"A_Code/aMSE/data/")
+
+  save(zone1,file=paste0(datadir,"zone1.RData"))
[…]

### 2022-07-11 README.Rmd
Ensured temp[late files operational, and all examples work. Production ready. No EWN.
Switched from hutils to codeutils. Worked on internal consistency. No EWN.

@@
 
+# aMSE
+
+<!-- badges: start -->
+<!-- badges: end -->
+
+This packages up a new Abalone Management Strategy Evaluation framework.
+It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:
+
+* zone - highest geographical level. Is simply the totality of the spatial management units.
+
+* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.
+
+* population - literally a population.
+These are the active components within the simulation framework.
+The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.
+
+This conceptual structure has implications for how the simulations are conditioned during the management strategy testing.
+While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population.
+Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU.
+
+An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant.
+This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.
+
+Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.
+
 # LATEST UPDATE
+
+* 2022-07-10 aMSE 0.0.10 Switched dependency from hutils to codeutils. re-added zone1 for S21 to the data sets. No EWN.
+
+* 2022-07-10 aMSE 0.0.9 Added data sets to allow for examples and documentation to illustrate the conditioning and running of the MSE.
+Added adjustavrec as a function, nade a wrapper rnormz and rlnormz to enable zero variation while still using up a random number so the sequence was maintained.
+Ready for production work.
+No EWN.
 
@@
 Now, to clean a directory of aMSE files use cleanrundir.
-
-# aMSE
-
-<!-- badges: start -->
-<!-- badges: end -->
-
-This packages up a new Abalone Management Strategy Evaluation framework.
-It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:
-
-* zone - highest geographical level. Is simply the totality of the spatial management units.
-
-* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.
-
-* population - literally a population.
-These are the active components within the simulation framework.
-The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.
-
-This conceptual structure has implications for how the simulations are conditioned during the management strategy testing.
-While it is still the case that the general properties of each SAU are used to set the scene for each SAU's component populations, the advent of sufficient GPS logger data now allows a more formal definition of the productivity of each population.
-Currently, the modelling assumes that each population is linearly arranged around the coastline, and this will allow, where available, specific catch/yield data to be used to define the bounds of each population within an SAU.
-
-An important change from previous designs is that larval dispersal is implemented explicitly rather than implicitly being held constant.
-This alters the dynamics so that analytical equilibrium methods no longer work and we need to resort to iterative approaches to equilibrium if the larval dispersal rate is > 0.0.
-
-Of course, adding such a component increases the number of options that may need to be explored, but currently we envisage including a very low, a middle range, and a relatively high level of larval dispersal, to examine its implications.
 

### 2022-07-11 R/docondition.R
Replaced the very large zone data-set with a much smaller one. No EWN.

@@ -18,6 +18,9 @@
 #' @param calcpopC a function that takes the output from hcrfun and generates
 #'     the actual catch per population expected in the current year.
 #' @param verbose Should progress comments be printed to console, default=TRUE
+#' @param lowmult multiplier to determine low range of search
+#' @param highmult multiplier to determine the high range of search
+#' @param iterlim the maximum number of iterations in the search for a solution
 #'
 #' @return nothing though it will write results to the console if verbose=TRUE
 #' @export
@@ -51,7 +54,8 @@
 #'   adjustavrec(rundir,out$glb,out$ctrl,calcpopC=calcexpectpopC,verbose=TRUE)
 #'  # Now repeat the do_condition and makeoutput steps to see the final result
 #' }
-adjustavrec <- function(rundir,glb,ctrl,calcpopC,verbose=TRUE) {
+adjustavrec <- function(rundir,glb,ctrl,calcpopC,verbose=TRUE,
+                        lowmult=0.6,highmult=1.4,iterlim=30) {
   startime <- Sys.time()
   datafile <- ctrl$datafile
   controlfile <- ctrl$controlfile
@@ -61,8 +65,8 @@ adjustavrec <- function(rundir,glb,ctrl,calcpopC,verbose=TRUE) {
   for (sau in 1:nsau) { #  sau=3
     initial <- getavrec(rundir,datafile,nsau=nsau)
     param <- initial[sau]
-    low <- param * 0.6
-    high <- param * 1.4
+    low <- param * lowmult
+    high <- param * highmult
     extra <- initial[-sau]
     if (verbose) cat("Running for sau ",saunames[sau],"\n")
     origssq <- sauavrecssq(param,rundir,controlfile,
@@ -74,7 +78,7 @@ adjustavrec <- function(rundir,glb,ctrl,calcpopC,verbose=TRUE) {
                  rundir=rundir,
                  controlfile=controlfile,datafile=datafile,linenum="AvRec",
                  calcpopC=calcpopC,extra=extra,picksau=sau,nsau=nsau,
-                 control=list(maxit=30))
+                 control=list(maxit=iterlim))
     final[sau] <- round(ans$par,1)
     if (verbose) {
      # cat("ssq = ",ans$value,"  new R0 value = ",ans$par,"\n")

### 2022-08-01 R/defineZone.R
Added a bunch of functions to aid comparing scenarios. No EWN.

@@ -944,7 +944,7 @@ modzoneC <- function(zoneC,zoneD,glob,lowlim=0.0,uplim=0.4,inc=0.005) {
 #' @param glob the globals object from readctrlfile
 #'
 #' @seealso{
-#'   \link{definepops}, \link{readpopdatafile}, \link{readdatafile}
+#'   \link{definepops}, \link{readpopdatafile}, \link{readsaudatafile}
 #' }
 #'
 #' @return a matrix with a row for each population and whose columns

### 2022-08-01 R/projection.R

@@ -196,7 +196,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
                           makehcrout,verbose=FALSE,...) {
   # ctrl=ctrl; zoneDP=zoneDP; zoneCP=zoneCP; glb=glb; hcrfun=mcdahcr; hsargs=hsargs
   # sampleCE=tasCPUE; sampleFIS=tasFIS; sampleNaS=tasNaS;  getdata=tasdata
-  # calcpopC=calcexpectpopC; verbose=TRUE
+  # calcpopC=calcexpectpopC; verbose=TRUE;makehcrout=makeouthcr
   reps <- ctrl$reps
   sigmar <- ctrl$withsigR
   sigmab <- ctrl$withsigB

### 2022-08-05 R/domse.R
added functions to assist with comparing scenarios. How to reference TasHS without requiring TasHS still to be decided.

@@ -93,6 +93,8 @@
 #' @param mincount given size-composition data what minimum sample size will
 #'     be deemed acceptable for inclusion in the plots and conditioning.
 #'     default=100.
+#' @param includeNAS should the NAS projections be included in the output. If
+#'     they are then the final output size is greatly increased. default=FALSE
 #'
 #' @seealso{
 #'  \link{makeequilzone}, \link{dohistoricC}, \link{prepareprojection},
@@ -108,7 +110,8 @@
 do_MSE <- function(rundir,controlfile,hsargs,hcrfun,sampleCE,sampleFIS,
                    sampleNaS,getdata,calcpopC,makeouthcr,varyrs=7,startyr=42,
                    verbose=FALSE,ndiagprojs=3,savesauout=FALSE,cutcatchN=56,
-                   matureL=c(70,200),wtatL=c(80,200),mincount=100) {
+                   matureL=c(70,200),wtatL=c(80,200),mincount=100,
+                   includeNAS=FALSE) {
   # generate equilibrium zone -----------------------------------------------
   starttime <- (Sys.time())
   zone <- makeequilzone(rundir,controlfile,verbose=verbose)
@@ -230,12 +233,17 @@ do_MSE <- function(rundir,controlfile,hsargs,hcrfun,sampleCE,sampleFIS,
   plothsstats(rundir,HSstats,glb)
   addtable(hcrout$refpts,"hcrout_refpts.csv",rundir,category="HSperf",
            caption="HCR reference points")
+  scores <- finalscoreoutputs(rundir=rundir,cpue=sauoutD$cpue,
+                              catches=sauoutD$catch,
+                              glb=glb,yearCE=condC$yearCE,
+                              hsargs=hsargs)
+  if (!includeNAS) NAS=NULL
   out <- list(tottime=tottime,projtime=projtime,starttime=starttime,glb=glb,
               ctrl=ctrl,zoneCP=zoneCP,zoneD=zoneD,zoneDD=zoneDD,zoneDP=zoneDP,
               NAS=NAS,projC=projC,condC=condC,sauout=sauout,outzone=outzone,
               hcrout=hcrout,production=production,condout=condout,
               HSstats=HSstats,saudat=saudat,constants=constants,hsargs=hsargs,
-              sauprod=sauprod)
+              sauprod=sauprod,scores=scores)
   return(out)
 } # end of do_MSE
 

### 2022-08-19 README.Rmd
Added a bunch of functions to aid comparing scenarios. No EWN.
added functions to assist with comparing scenarios. How to reference TasHS without requiring TasHS still to be decided.
More functions to compare scenarios. No EWN
Added do_comparison and expanded comparisons. No EWN.

@@
 # LATEST UPDATE
+
+* 2022-08-19 aMSE 0.0.15 Added do_comparison as a wrapper function for comparing scenarios.
+
+* 2022-08-05 aMSE 0.0.14 Added more functions to assist with comparing scenarios.
+
+* 2022-08-05 aMSE 0.0.12 Added more functions to assist with comparing scenarios. Still sorting out references to other packages.
+
+* 2022-08-01 aMSE 0.0.11 Added a bunch of functions to assist with comparing scenarios. No EWN.
 
@@
 
-```{r example, echo=TRUE}
+```{r example, echo=TRUE, eval=FALSE }
 # a constant TAC example

### 2022-09-26 R/plotfuns.r
Added doquantplot and finalcondyeardepletion functions. No EWN.

@@ -293,6 +293,68 @@ dosauplot <- function(ylabel,postrep,glb,startyr,addCI=FALSE,
   return(invisible(sauCI))
 } # end of dosauplot
 
+#' @title finalcondyeardepletion plots and tabulates final depletion after conditioning
+#'
+#' @description finalcondyeardepletion generates histograms and tabulations of
+#'     either the spawning biomass or exploitable biomass depletion levels in the
+#'     final year of conditioning. This includes the varyrs of recruitment
+#'     variation prior to projections. It plots a histogram for each sau and
+#'     generates a table of quantiles at probs of 0, 0.05, 0.5, 0.95, and 1.0.
+#'
+#' @param rundir the scenario's directory
+#' @param sauzone the output from zonetosau, found in sauplots. This is
+#'     the projeciton results summarized into their respective sau rather than
+#'     the individual populations
+#' @param glb the globals object
+#' @param deplvar which depletion variable to use, either 'sB' or 'eB' are the
+#'     only options. Any other input will use 'sB'
+#' @param console should the plot go to the console or be saved into rundir?
+#'     default=TRUE
+#'
+#' @seealso{
+#'    \link{zonetosau}, \link{sauplots}
+#' }
+#'
+#' @return invisibly returns the matrix of quantiles for each sau
+#' @export
+#'
+#' @examples
+#' print("wait on data sets")
+finalcondyeardepletion <- function(rundir,sauzone,glb,deplvar="sB",console=TRUE) {
+  nsau <- glb$nSAU
+  hyrs <- glb$hyrs
+  saunames <- glb$saunames
+  if (!(deplvar %in% c("sB","eB"))) deplvar <- "sB"
+  depl <- switch(deplvar,"sB" = sauzone$deplsB,
+                 "eB" = sauzone$depleB)
+  deplquant <- matrix(0,nrow=nsau,ncol=5,
+                      dimnames=list(saunames,c("0%","5%","50%","95%","100%")))
+  filen <- ""
+  if (!console) {
+    nfile <- paste0("finalcondyear_",deplvar,"_depletion.png")
+    filen <- filenametopath(rundir,nfile)
+    caption <- "Histograms of the depletion in the final year of conditioning."
+  }
[…]

### 2022-09-26 R/projection.R

@@ -223,8 +223,8 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
                          sauCPUE=zoneDP$cesau[,,iter],
                          sauacatch=zoneDP$acatch[,,iter],
                          sauNAS=list(Nt=zoneDP$Nt[,,,iter],
-                                     catchN=zoneDP$catchN[,,,iter],
-                                     NumNe=zoneDP$NumNe[,,,iter]),
+                         catchN=zoneDP$catchN[,,,iter],
+                         NumNe=zoneDP$NumNe[,,,iter]),
                          year=year)
       hcrout <- hcrfun(hcrdata,hsargs,saunames=glb$saunames)
       popC <- calcpopC(hcrout,exb=zoneDP$exploitB[year-1,,iter],

### 2022-10-18 README.Rmd
Improved help on functions and more functions for comparing scenarios. No EWN.
Added doquantplot and finalcondyeardepletion functions. No EWN.
Altered projections to be all replicates by year rather than all years per replicate.

@@
 # LATEST UPDATE
+
+* 2022-10-18 aMSE 0.0.18 In the projections relicates across years has now been changed to years across replicates to allow for inclusion of FIS performance measures.
+Also, further comparison functions included.
+
+* 2022-09-26 aMSE 0.0.17 Added finalcondyeardepletion to aid in comparisons.
+
+* 2022-08-24 aMSE 0.0.16 Started to add depensation and other refinements.
 

### 2022-10-18 R/projection.R

@@ -193,6 +193,7 @@ calcsau <-  function(invar,saunames,ref0) {# for deplsb depleB
 #'
 #' @examples
 #' print("wait on suitable internal data sets")
+#' # switched year and iter in the loops
 doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
                           sampleCE,sampleFIS,sampleNaS,getdata,calcpopC,
                           makehcrout,verbose=FALSE,...) {
@@ -216,9 +217,12 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
   b0 <- getvar(zoneCP,"B0") #sapply(zoneC,"[[","B0")
   exb0 <- getvar(zoneCP,"ExB0")
   hcrout <- makehcrout(glb,hsargs)
-  for (iter in 1:reps) {
-    if (verbose) if ((iter %% 25) == 0) cat(iter,"   ")
-    for (year in startyr:endyr) { # iter=1; year=startyr
+  for (year in startyr:endyr) { # iter=1; year=startyr
+ #   for (iter in 1:reps) {
+  #  if (verbose) if ((iter %% 25) == 0) cat(iter,"   ")
+    if (verbose) cat(year,"   ")
+      for (iter in 1:reps) {
+    #  for (year in startyr:endyr) { # iter=1; year=startyr
       hcrdata <- getdata(sampleCE,sampleFIS,sampleNaS,
                          sauCPUE=zoneDP$cesau[,,iter],
                          sauacatch=zoneDP$acatch[,,iter],

### 2022-10-31 README.Rmd
Added more functions to aid comparing scenarios. No EWN

@@
 
-* 2022-10-18 aMSE 0.0.18 In the projections relicates across years has now been changed to years across replicates to allow for inclusion of FIS performance measures.
+* 2022-10-31 aMSE 0.0.20 added do_comp_outputs, scenebyvar, scenebyzone, projectiononly, and sauquantbyscene to assist with comparing scenarios.
+
+* 2022-10-26 aMSE 0.0.19 added more functions to facilitate comparing scenarios.
+
+* 2022-10-18 aMSE 0.0.18 In the projections replicates across years has now been changed to years across replicates to allow for inclusion of FIS performance measures.
 Also, further comparison functions included.

### 2022-10-31 R/aMSE_utils.R

@@ -559,6 +559,49 @@ prepareDDNt <- function(inNt,incatchN,glb) { # Nt=zoneDD$Nt; catchN=zoneDD$catch
   return(list(Nt=Nt,catchN=catchN))
 } # end of prepareDDNt
 
+#' @title projectiononly converts hyrs+pyrs_sau_reps to pyrs_sau_reps array
+#'
+#' @description projectiononly is a utility function that takes an output 3D
+#'     array from aMSE with dimensions allyears x nsau x reps and truncates it
+#'     so that only the projection years remain, pyrs x nsau x reps, to simplify
+#'     plotting and tabulating results. This can be used within both aMSE and in
+#'     do_comparison of scenarios.
+#'
+#' @param x a 2D, 3D or 4D array of dimension (hyrs + pyrs) x reps,
+#'     nyrs x nsau x reps, or Nclass x nyrs x reps
+#' @param glb the globals object. Only hyrs and pyrs are used
+#'
+#' @return an array truncated appropriately to include only the projection years
+#' @export
+#'
+#' @examples
+#' print("wait on example data sets")
+#' # x <- catch[[1]];glb <- glbc[[1]]
+projectiononly <- function(x,glb) {
+  beginP <- glb$hyrs + 1
+  endP <- glb$hyrs + glb$pyrs
+  dims <- dim(x)
+  ndim <- length(dims)
+  if (endP %in% dims) {
+    pick <- which(dims == endP)
+    if (pick == 1) {
+      if (ndim == 3) {
+        out <- x[beginP:endP,,]
+      } else {
+        out <- x[beginP:endP,]
+      }
+    }
+    if ((pick == 2) & (ndim == 3)) {
+      out <- x[,beginP:endP,]
+    }
+    if ((pick == 2) & (ndim == 4)) {
+      out <- x[,beginP:endP,,]
+    }
+  } else {
+    out <- x
+  }
+  return(out)
+} # end of projectiononly
 
 #' @title rnormz when sd=0 it returns rep(mean,n) but also uses n random numbers
 #'

### 2022-10-31 R/getfunctions.R

@@ -438,11 +438,11 @@ getprojyraavc <- function(catches,glb) {
 #' @export
 #'
 #' @examples
-#' # sauarr=out$sauout$zonePsau$catch;hyrs=out$glb$hyrs
-#' # pyrs=out$glb$pyrs; startyr=out$glb$hyrs
+#' # sauarr=catch[[1]];hyrs=glbc[[1]]$hyrs
+#' # pyrs=glbc[[1]]$pyrs; startyr=glbc[[1]]$hyrs
 getprojyrs <- function(sauarr,hyrs,pyrs,startyr=hyrs) {
   yrs <- dim(sauarr)[1]
-  if (yrs != hyrs + pyrs)
+  if (yrs != (hyrs + pyrs))
     stop(cat("Input error in getprojyrs, array in has incorrect dimensions \n"))
   return(sauarr[startyr:yrs,,])
 } # end of getprojyrs

### 2022-11-10 README.Rmd
Modified how do_MSE responds to hsargs for new Meta-rules
Added ENVIRON option to control file. Improved compare_scenarios. NoEWN, but vignettes causing trouble

@@
 # LATEST UPDATE
+
+* 2022-11-10 aMSE 0.0.22 now have the capacity to introduce productivity events in specific projection years.
+This can now be done at sau specific levels.
+One defines the survivorship of recruits in each selected year and also the survivorship of all post-settlement animals.
+No changes are needed to one's control file if this option is not being used.
+
+* 2022-11-02 aMSE 0.0.21 added changes to how do_MSE responds to hsargs.
 

### 2022-11-10 R/comparescenarios.R

@@ -65,7 +65,7 @@ boxbysau <- function(rundir,hspm,glbc,scenes,compvar="aavc",filen="",aavcyrs=10,
                   "sum10" = "Sum of Catches over first 10 Years")
   mtext(label,side=1,outer=TRUE,line=-0.3,cex=1.0)
   if (nchar(filen) > 0)
-    addplot(filen=filen,rundir=rundir,category="catches",caption)
+    addplot(filen=filen,rundir=rundir,category="catchBoxPlots",caption)
 } # end of boxbysau
 
 
@@ -608,13 +608,13 @@ cpueboxbysau <- function(rundir,cpue,glbc,scenes,filen="",startyr=0,maxval=0) {
   label <- "Years to Maximum Median CPUE by sau"
   mtext(label,side=1,outer=TRUE,line=-0.3,cex=1.0)
   if (nchar(filen) > 0) {
-    addplot(filen=filen2,rundir=rundir,category="cpue",caption)
+    addplot(filen=filen2,rundir=rundir,category="cpueBoxPlots",caption)
     filen3 <- unlist(strsplit(filen,".",fixed=TRUE))[1]
     filen3 <- paste0(filen3,".csv")
     filen <- filenametopath(rundir,filen3)
     write.csv(cpuebox,file=filen)
     caption <- "Years to maximum median cpue by sau from boxplots."
-    addtable(cpuebox,filen3,rundir=rundir,category="cpue",caption)
+    addtable(cpuebox,filen3,rundir=rundir,category="cpueBoxPlots",caption)
   }
   return(invisible(cpuebox))
 } # end of cpueboxbysau
@@ -677,7 +677,7 @@ cpueHSPM <- function(rundir,cpue,glbc,scenes,filen="",startyr=0) {
     boxresult[[i]] <- qsstats
   }
   if (nchar(filen) > 0)
-    addplot(filen=filen,rundir=rundir,category="cpue",caption)
+    addplot(filen=filen,rundir=rundir,category="cpueBoxPlots",caption)
   return(invisible(boxresult))
 } #end of cpueHSPM
 
@@ -698,9 +698,11 @@ cpueHSPM <- function(rundir,cpue,glbc,scenes,filen="",startyr=0) {
 #' @param files the vector of file names to be used.
 #' @param pickfiles a vector of indices selecting the files to be used.
 #' @param verbose should progress updates be made to the console, default=TRUE
[…]

### 2022-12-13 R/projection.R
Minor edits aimed at improving the comparison of scenarios

@@ -281,7 +281,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
 #'     final scores. Uses getcpueHS from TasHS.
 #'
 #' @param rundir the directory for a given scenario
-#' @param cpueHSPMs is a function defined in the HS package that reconstructs
+#' @param HSPMs is a function defined in the HS package that reconstructs
 #'     the cpue HSPM scores, and other measures ready for plotting
 #' @param cpue the projected cpue dynamics from out$sauout$zonePsau$cpue
 #' @param catches the project catches from out$sauout$zonePsau$catch
@@ -290,6 +290,8 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
 #'     model, found in condC$yearCE.
 #' @param hsargs the list of arguments required by the HS
 #' @param verbose should updates on progress be sent to console? default=TRUE
+#' @param ... required in case whatever function is pointed at HSPMs requires
+#'     more than cpue
 #'
 #' @return a list of two lists. The first being outHS containing all the
 #'     reconstructed HS outputs, and the second being the medians of the
@@ -298,12 +300,12 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
 #'
 #' @examples
 #' print("wait on data sets, example soon")
-#' # rundir=rundir;cpueHSPMs=cpueHSPMs;cpue=sauout$cpue;
+#' # rundir=rundir;HSPMs=HSPMs;cpue=sauout$cpue;
 #' # catches=sauout$catches;glb=glb;yearCE=condC$yearCE
 #' # hsargs=hsargs
-finalscoreoutputs <- function(rundir,cpueHSPMs,cpue,catches,glb,yearCE,hsargs,
-                              verbose=TRUE) {
-  if (!is.null(cpueHSPMs)) {
+finalscoreoutputs <- function(rundir,HSPMs,cpue,catches,glb,yearCE,hsargs,
+                              verbose=TRUE,...) {
+  if (!is.null(HSPMs)) {
     nsau <- glb$nSAU
     reps <- glb$reps
     outHS <- vector(mode="list",length=nsau)
@@ -312,9 +314,8 @@ finalscoreoutputs <- function(rundir,cpueHSPMs,cpue,catches,glb,yearCE,hsargs,
     outmed <- outHS
     if (verbose) cat("Reconstructing HS outputs  \n")
     for (sau in 1:nsau) {
[…]

### 2023-01-17 README.Rmd
added recruits tab to output.
Minor edits aimed at improving the comparison of scenarios
working on do-comparison

@@
 
-# LATEST UPDATE
+# LATEST UPDATES
+
+* 2022-12-14 aMSE 0.0.24 Lots of minor edits and modifications aimed at improving the comparison of scenarios.
+
+* 2022-11-11 aMSE 0.0.23 Added a new tab = recruits, containing plots of the stock recruitment relationships at an sau and at a population within sau level. Added a table to the Tables tab.
 
@@
 * 2022-10-31 aMSE 0.0.20 added do_comp_outputs, scenebyvar, scenebyzone, projectiononly, and sauquantbyscene to assist with comparing scenarios.
-
-* 2022-10-26 aMSE 0.0.19 added more functions to facilitate comparing scenarios.
-
-* 2022-10-18 aMSE 0.0.18 In the projections replicates across years has now been changed to years across replicates to allow for inclusion of FIS performance measures.
-Also, further comparison functions included.
-
-* 2022-09-26 aMSE 0.0.17 Added finalcondyeardepletion to aid in comparisons.
-
-* 2022-08-24 aMSE 0.0.16 Started to add depensation and other refinements.
-
-* 2022-08-19 aMSE 0.0.15 Added do_comparison as a wrapper function for comparing scenarios.
-
-* 2022-08-05 aMSE 0.0.14 Added more functions to assist with comparing scenarios.
-
-* 2022-08-05 aMSE 0.0.12 Added more functions to assist with comparing scenarios. Still sorting out references to other packages.
-
-* 2022-08-01 aMSE 0.0.11 Added a bunch of functions to assist with comparing scenarios. No EWN.
-
-* 2022-07-10 aMSE 0.0.10 Switched dependency from hutils to codeutils. re-added zone1 for S21 to the data sets. No EWN.
-
-* 2022-07-10 aMSE 0.0.9 Added data sets to allow for examples and documentation to illustrate the conditioning and running of the MSE.
-Added adjustavrec as a function, nade a wrapper rnormz and rlnormz to enable zero variation while still using up a random number so the sequence was maintained.
-Ready for production work.
-No EWN.
-
-* 2022-06-29 aMSE 0.0.8 Corrected a large mistake where the estimated SAU harvest rates were being biased high because the SAU's end of year exploitable biomass was being used instead of the mid-year exploitable biomass.
-
-* 2022-06-22 aMSE 0.0.7 Now have rewritecontrolfile and rewritesaudatafile and the option to include a table of parameters from sizemod into the inputs of aMSE.
-Have begun development of functions to aid comparing the output from different scenarios, including tables and plots.
-No EWN.
-
-* 2022-06-08 aMSE 0.0.6 Have revised the ctrltemplate and datatemplate generators so that they reflect a lambda = 0.75 and correctly use the hyperstability. No EWN
-
-* 2022-06-06 aMSE 0.0.5.
-Have reviewed the dynamics in detail so that the outputs from the aMSE for each SAU more closely match the outputs from sizemod.
-The original disjunction was partly due to a faulty implementation of the inclusion of the option of hyper-stability, but that has now been amended and the operatnig model is now behaving as it should.
-Breaking the dynamics within each SAU into multiple sub-populations with variation also affects the equivalence between aMSE and sizemod butmostly influences the productivity rather than the recruitment dynamics and the trajectories mainly follow those from sizemod but with the cpue and biomasses being projected incorrectly.
-It appears (still working on a why) that by changing the productivity and sub-dividing that productivity among sub-populations, the overall recruitment levels required to mimic the dynamics in all cases appears to be less in aMSE than in sizemod.
-Otheriwse nothing now needs to change.
-One uses parameters obtained from sizemod, especially the recruitment deviates, and then apply the functions that search for more optimal values for AvRec, the unfished recruitment levles in each SAU.
-The fits obtained should be close to those found in sizemod.
-In 0.0.5 there remain issues relating to the internal data sets, which include one called 'zone', which is not designed to work with the option of hyper-stability (lambda != 1.0).
-Otherwise, No EWN.
-
-* 2022-05-31 aMSE 0.0.4 Added functions that assist in conditioning the operating model by optimizing the fit to CPUE and size-composition data after opimizing the AvRec values for each SAU. No EWN.
-
-* 2021-11-29 aMSE 0.0.1 Removed datadir. Started implementing hyper-stability while keeping the interface the same, so that the results from sizemod can be used.
-
-* 2021-11-23 aMSE 0.0.0.1 Added functions to simplify the optimization of AvRec relative to the CPUE data after initial parameter estimation for each SAU using the sizemod package.
-Also added a comparison of size-composition data to the do_condition function.
-Next step will be to include the optimization of the recdevs using both CPUE and size-comp data.
-
-* 2021-11-21 aMSE 0.0.0.020 So as to use any observed size-composition data in the conditioning, a number of functions have been added to convert the predicted numbers-at-size in the catch by population into NAS x SAU.
-These can now be plotted against the observed NAS for each SAU and this can be either to the console or to a new tab in the webpage called 'predictedcatchN'.
-This work derived from the progress made with the sizemod package used to fit size-based assessment models to individual SAU.
-Because the MSE sub-divides each SAU into a number of populations each with somewhat different MaxDL, L95, etc, even after fitting the models, when using that data it is then necessary to refit the AvRec values, which, because of the sub-division into populations each having different biology can lead to the MSE AvRec being either smaller or larger than the estimated R0 for a SAU.
-Similar arguments apply to teh recruitment residuals but generally the pattern provides a reasonable fit.
-
-* 2021-10-17 aMSE 0.0.0.010 As cleanslate, within setuphtml() has been deprecated in the package makehtml, I have removed reference to it throughout aMSE.
-Now, to clean a directory of aMSE files use cleanrundir.
 

### 2023-04-04 README.Rmd
Modifed calcpopC to suite SA. No EWN
Added curryear to hcrfun inside doprojections.

@@
 # LATEST UPDATES
+
+* 2023-04-04 aMSE 0.1.0 Added curryear to hcrfun so all HS functions should now include curryear inthier arguments.
+Within doprojecitons() it is set to year inside the loop.
+This was needed to allow the SA HS to use its fleet dynamics model and is not needed in Tasmania, but is required in the argument list.
+
+* 2023-03-31 aMSE 0.0.25 Modified calcpopC througout to allow for SA's harvest strategy and fleet dynamics.
 

### 2023-04-04 R/projection.R

@@ -243,7 +243,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
                          catchN=zoneDP$catchN[,,,iter],
                          NumNe=zoneDP$NumNe[,,,iter]),
                          year=year,decrement=hsargs$decrement)
-      hcrout <- hcrfun(hcrdata,hsargs,saunames=glb$saunames)
+      hcrout <- hcrfun(hcrdata,hsargs,saunames=glb$saunames,curryear=year)
       popC <- calcpopC(hcrout,exb=zoneDP$exploitB[year-1,,iter],
                        sauCPUE= zoneDP$cesau[year-1,,iter],
                        catsau = zoneDP$catsau[year-1,,iter],

### 2023-04-12 R/docondition.R
Started extracting jurisdiction specific functions out of aMSE.

@@ -274,7 +274,7 @@ changeline <- function(indir, filename, linenumber, newline,verbose=FALSE) {
 #' @examples
 #' print("wait on suitable data sets in data")
 #' # rundir=rundir; controlfile=controlfile;calcpopC=calcexpectpopC
-#' # verbose=TRUE; doproduct=FALSE; dohistoric=TRUE; mincount=120
+#' # verbose=TRUE; doproduct=TRUE; dohistoric=TRUE; mincount=120
 #' # matureL=c(70,200);wtatL=c(80,200);mincount=120
 do_condition <- function(rundir,controlfile,calcpopC,
                          verbose=FALSE,doproduct=FALSE,dohistoric=TRUE,

### 2023-04-12 R/projection.R

@@ -277,64 +277,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
   return(list(zoneDP=zoneDP,hcrout=hcrout))
 } # end of doprojections
 
-#' @title finalscoreoutputs generates plots of the PM scores and interactions
-#'
-#' @description finalscoreoutputs is used to reconstruct the harvest strategy
-#'     responses to the dynamics so they can be examined and understood. This
-#'     function generates plots, for each sau, of the projected catches, cpue,
-#'     gard1 scores, grad4 scores, target cpue scores, the cpue target and the
-#'     final scores. Uses getcpueHS from TasHS.
-#'
-#' @param rundir the directory for a given scenario
-#' @param HSPMs is a function defined in the HS package that reconstructs
-#'     the cpue HSPM scores, and other measures ready for plotting
-#' @param cpue the projected cpue dynamics from out$sauout$zonePsau$cpue
-#' @param catches the project catches from out$sauout$zonePsau$catch
-#' @param glb the globals object for the scenario
-#' @param yearCE the year names for the observed CPUE used to condition the
-#'     model, found in condC$yearCE.
-#' @param hsargs the list of arguments required by the HS
-#' @param verbose should updates on progress be sent to console? default=TRUE
-#' @param ... required in case whatever function is pointed at HSPMs requires
-#'     more than cpue
-#'
-#' @return a list of two lists. The first being outHS containing all the
-#'     reconstructed HS outputs, and the second being the medians of the
-#'     various scores and targetcpue.
-#' @export
-#'
-#' @examples
-#' print("wait on data sets, example soon")
-#' # rundir=rundir;HSPMs=HSPMs;cpue=sauout$cpue;
-#' # catches=sauout$catches;glb=glb;yearCE=condC$yearCE
-#' # hsargs=hsargs
-finalscoreoutputs <- function(rundir,HSPMs,cpue,catches,glb,yearCE,hsargs,
-                              verbose=TRUE,...) {
-  if (!is.null(HSPMs)) {
-    nsau <- glb$nSAU
-    reps <- glb$reps
-    outHS <- vector(mode="list",length=nsau)
-    saunames <- glb$saunames
-    names(outHS) <- saunames
-    outmed <- outHS
-    if (verbose) cat("Reconstructing HS outputs  \n")
-    for (sau in 1:nsau) {
[…]

### 2023-04-17 README.Rmd
Started extracting jurisdiction specific functions out of aMSE.
Generalized save-hsargs to allow for data.frames

@@
 # LATEST UPDATES
+
+* 2023-04-17 aMSE 0.1.2 Amended save_hsargs so it can handle data.frame within hsargs.
+
+* 2023-04-12 aMSE 0.1.1 Started generalizing the outputs from aMSE by beginning to separate functions that relate to specific jurisdiction outputs into a separate R package (in Tasmania it is currently called aMSEExtra).
 

### 2023-06-03 R/docondition.R
Some small cahnges to allow SA to run with mcdahcr

@@ -188,12 +188,12 @@ changeline <- function(indir, filename, linenumber, newline,verbose=FALSE) {
   # indir=rundir; filename="saudataM15h5.csv"; linenumber="AvRec"; newline=chgevals
   filen <- filenametopath(indir,filename)
   dat <- readLines(filen)
-  if (class(linenumber) == "numeric") {
+  if (inherits(linenumber,"numeric")) {
     origtext <- dat[linenumber]
     dat[linenumber] <- newline
     writeLines(dat,filen)
   }
-  if (class(linenumber) == "character") { # ie the name of the variable
+  if (inherits(linenumber,"character")) { # ie the name of the variable
     pickL <- grep(linenumber,dat)[1]
     if (length(pickL) == 0)
       stop(cat(paste0(linenumber," does not appear in ",filename),"\n"))

### 2023-06-05 README.Rmd
generalized doprojections and getdata to allow for proper use of startCE.
Minor changes to allow for consthcr scenarios. Vignettes need fixing.

@@
 # LATEST UPDATES
+
+* 2023-06-04 aMSE 0.1.4 Small modifications to generalize across jurisdicitons and allow for the consthcr constant catch harvest strategy; also removed irrelavent messages when plotting the phase plot
+
+* 2023-06-03 aMSE 0.1.3 Modified to allow the mcdahcr and constrefhcr HCR to run successfully with the conditioned SA setup.
 

### 2023-06-12 README.Rmd
Improved conditioning plot, and added option of interim saving of results following projections.

--- 2023-04-17 README.Rmd
@@
 # LATEST UPDATES
+
+* 2023-06-11 aMSE 0.1.5 Modified the condition plot and fixed up a few places where numbers were hard wired into the software.
+Introduced a new option to save output following projections (just in case of experiments with subsequent analyses crash - I'm looking at you SA).
+Now one can set interimout =c(outdir,postfixdir) and that should save a tempout into your outdir directory.
+See the help for do_MSE.
+
+* 2023-06-04 aMSE 0.1.4 Small modifications to generalize across jurisdictions and allow for the consthcr constant catch harvest strategy; also removed irrelavent messages when plotting the phase plot
+
+* 2023-06-03 aMSE 0.1.3 Modified to allow the mcdahcr and constrefhcr HCR to run successfully with the conditioned SA setup.
 

### 2023-07-10 README.Rmd
Vasrious updates. Now us refperiodCE in hsargs instead of startCE and endCE.

@@
 # LATEST UPDATES
+
+* 2023-07-10 aMSE 0.1.6 Modified TasHS and aMSE so that instead of using hsargs$startCE and hsargs$endCE to define a vector of reference years when using the constantrefhcr in Tasmania, one should now use a hsargs$refperiodCE which should be a vector of years describing the constant reference period to be used (assuming one is using the constantrefhcr).
+Within TasHS there is also a checkhsargs() function which provides a brief check of the hsargs being used (strictly that is still under development).
 

### 2023-07-10 R/getfunctions.R

@@ -664,7 +664,9 @@ getunFished <- function(zoneC,zoneD,glb) {  # inzone=zone
 #' @return a numpop vector of the invar constants from zoneC
 #' @export
 #'
-#' @seealso getlistvar
+#' @seealso{
+#'  \link{getlistvar}
+#' }
 #'
 #' @examples
 #' data(zone)

### 2023-07-12 data-raw/vignettefiles/Comparing_Scenarios.Rmd
Now use of refperiodCE is optional. No EWN.

---
title: "Comparison of MSE Scenarios"
author: "Malcolm Haddon, ..."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparison of MSE Scenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

options(knitr.kable.NA = "",
        knitr.table.format = "pandoc")

options("show.signif.stars"=FALSE,
        "stringsAsFactors"=FALSE,
        "max.print"=50000,
        "width"=240)

suppressPackageStartupMessages({
library(aMSE)
library(TasHS)
library(codeutils)
library(hplot)
library(knitr)
library(makehtml)
library(captioner)
})

tab_nums <- captioner(prefix = "Table")
fig_nums <- captioner(prefix = "Figure")

dbdir <- getDBdir()
ddir <- paste0(dbdir,"/A_CodeUse/aMSEDoc/") 
prefixdir <- paste0(dbdir,"/A_CodeUse/aMSEUse/scenarios/")

```

<style type="text/css">
  body, td {
     font-size: 18px;
     font-family: "Times New Roman", Times, serif;
  }
  code.r{
    font-size: 12px;
  }
  pre {
    font-size: 8px
  }
  h1 {
    font-size: 32px
  }
  .inline{
     font-size: 15px;
  }
  .display{
     font-size: 18px;
  }
</style>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: { 
      equationNumbers: { 
            autoNumber: "all",
            formatNumber: function (n) {return '3.'+n}
      } 
  }
});
</script>

# Introduction

This document is part of the more formal documentation for the R package __aMSE__, which aims to provide a common modelling framework to test abalone harvest strategies. Being one of the vignettes included in the package it should be regarded as a living, adaptable document that will be modified as experience increases around different jurisdictions, and the authorship should be expected to alter accordingly as different people contribute. It is assumed in this vignette that the reader has managed to run a number of scenarios that they then wish to compare.

### Saving the Output from aMSE

One concept in the use of __aMSE__ is that of the _rundir_, which is the directory into which all the data files are placed and all the results, plots, and html files will be placed. However, a typical scenario MSE run will include a final line where the output from the _do_MSE()_ function is saved into an _outdir_. The _outdir_ should not be on a cloud service shared with others as the output objects can be large (gigabytes). so typically it will point to a directory on a fast hard-drive or SSD that should be used to identify a given series of scenarios for comparison. The actual scenario name is usually encapsulated in the _postfixdir_ name:

```{r echo=TRUE, eval=FALSE}
# A subset of commands to run a given scenario
postfixdir <- "BC"                   # BC stands for basecase
rundir <- filenametopath(prefixdir,postfixdir)
outdir <- "C:/aMSE_scenarios/BC/"
confirmdir(rundir,ask=FALSE)     # check to see if it already exists
confirmdir(outdir,ask=FALSE)     # and if not create it without asking.

hsargs <- list(mult=0.1, # expansion factor for cpue range when calc the targqnt
               wid = 4, # number of years in the grad4 PM
               targqnt = 0.55, # quantile defining the cpue target
               maxtarg = c(150,150,150,150,150,150,150,150), # max cpue Target
               pmwts = c(0.65,0.25,0.1),  # relative weights of PMs
               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2), # hcr multipliers
               startCE = 2000, # used in constant reference period HS
               endCE = 2019,   # used in constant reference period HS
               metRunder = 1,  # should the metarules be used. o =
               metRover = 1,   # use metarules
               decrement=1, # use fishery data up to the end of the time series
               pmwtSwitch = 0, # number of years after reaching the targCE to
               stablewts = c(0.4, 0.5, 0.1), # replace pmwts with stablewts
               hcrname="mcdahcr",     # the name of the HCR used
               printmat=NULL)

out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=mcdahcr,
              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
              getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
              hcrscoreoutputs=extractandplotscores,HSPMs=getcpueHS,
              fleetdyn=NULL,interimout=c(outdir,postfixdir),
              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
              matureL=c(40,170),wtatL=c(50,200),mincount=120,
              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15))

makeoutput(out,rundir,postfixdir,controlfile,hsfile="TasHS Package",
           openfile=TRUE,verbose=FALSE)

save(out,file=paste0(outdir,postfixdir,".RData"))  # the results are saved
```

## Making comparisons

There are a number of requirements when conducting an Management Strategy Evaluation of alternative harvest strategies on an abalone fishery. It is currently best to use the _RStudio_ integrated development environment, as this facilitates the installation of different R packages, and the editing and running of the R files used to conduct the actual scenario runs.

The R packages required to run a scenario include at least:

* __aMSE__ the main R package containing the operating model and auxillary functions used.
* __hutils__ an R package containing numerous utility functions that are common to many simulation tasks.
* __hplot__ am R pacakge containing functions to facilitate plotting in base R.
* __knitr__ an R package used in vignette preparation and the production of formal tables
* _jurisdictionHS.R_ a source R file containing functions unique to a particular juridiction's harvest strategy.
* other packages may be required depending upon what packages are called within _jurisdictionHS.R_.

Most importantly for this document there are also a number of other data files required for a scenario run and these are all _.csv_ files. In the sections below we will go through their structure and contents in detail to act as a guide to their format and functioning:

* a _control.csv_ file (whose name can be anything the user wishes), that holds control options and many other constants.
* a _saubiology.csv_ file, that holds parameterizations of the biological properties of the _SAU_ and _populations_ that make up the operating model.

There may also be other data files if such data are available, these include:

* a size-composition of commercial catch data file
* a size-composition of fishery-independent survey file

When using the R package __aMSE__ a typical scenario run would entail a number of fundamental steps:

* Initiate each _population_ within each _SAU_ in the operating model with its biological properties.
* Generate the equilibrium unfished zone across all populations and estimate the productivity of each _population_ (and hence each _SAU_ and the total _zone_).
* Condition the operating model using either a hypothetical depletion or apply historical fishery data until the simulated _zone_ is ready to be projected under control of the particular harvest strategy being considered (__`r fig_nums("f1",display="cite")`__).
* Conduct the projections and summarize the results.

<br>

```{r FIG1, echo=FALSE, out.width = '65%', fig.align='center'}
filen <- paste0(ddir,"figures/AlternativeMSE.png")
suppressWarnings(knitr::include_graphics(filen))
```

`r fig_nums("f1", caption=" A spectrum of alternative approaches for initiating and using the MSE framework.")`

<br>

## File Locations for a Scenario Run

Within __aMSE__, it is expected that all data files and all result files will be contained in the same directory. Within the __aMSE__ code, this directory is referred to as the _rundir_ (as in run-directory). It is suggested that each scenario run be given a unique name that is then used as the _rundir_ name. For example, if examining the influence of the legal minimum length (LML) on the outcome of using a particular harvest strategy (perhaps the TasmanianHS) one might have the following scenarios: lml140, lml145, lml150, and so on, all contained in a sub-directory named 'scenarios'. This would make future comparisons simpler to select. But of course, a user may name their _rundir_ whatever they wish.

The intent is that the _control.csv_ and _saubiology.csv_ files (and the _jurisdictionHS.R_ file) be stored in the _rundir_. If this means that the _saubiology.csv_ file would be replicated identically in each scenario sub-directory, then, instead, that file can be stored in a separate _datadir_ as an option. Here we will focus solely on the data files. There are (or will be) functions to generate data file templates for each of the data-file types. These can be run and the resulting templates edited to form the required files. Alternatively, the example data-files, made available with the __aMSE__ package, can be copied and edited appropriately as an alternative option.

Similarly, within __aMSE__, there are functions to read in each type of data-file. If there is a formatting error within the data-file it should throw an error which identifies where it has gone wrong. Further development will be applied to these error checks.

The input data-files all, currently, are required to be _.csv_ files where components with multiple values use a comma as a separator. This makes their preparation using software such as _Microsoft Excel_ relatively simple.

The first of the two files is the _control.csv_, which contains a wide range of settings concerning the structure of the Operating Model as well as any historical conditioning data from a given fishery, if such conditioning is required. The second file is the _saubiology.csv_ file that contains the data that is used to define the biological properties of the _populations_ that are contained in the _SAU_ that make up the simulated _zone_. The rest of this document provides a detailed description of the contents and format of these files.

## The Conditioning Requirements

There is a spectrum to the degree to which the operating model is conditioned on a real fishery. At the least specific one still has to include a spatial structure that reflects a typical abalone stock (in __aMSE__ the top level is the _zone_, which contains one or more _SAU_, which each contain one or more _populations_). This can be customized to suit a particular species, and one could include what might be deemed typical characteristics of growth, maturity, emergence and other details, including variation among the different populations found within each SAU. Even with such details it would remain a generic abalone MSE (__`r fig_nums("f1",display="cite")`__).

It is possible to take such conditioning a step further and customize the details of the biology of each _population_, or at least each _SAU_ making up the spatial structure, to reflect the biology known for a real fishery _zone_. One could also scale predicted productivity and cpue to match a real fishery. This would be less generic but still not closely fitted to a real fishery. The detailed spatial structure internal to the operating model is what makes it impossible to fit the operating model to fishery data as a whole. However, it would still be possible to fit the productivity and history of individual SAU to real fishery data and thereby gain estimates of actual productivity and possibly even recruitment deviates (see _Conditioning the aMSE_). Of course, while this latter approach would appear, at first glance, to be the best strategy for examining a harvest strategy, as it might operate in a real situation, it is also the most demanding in terms of data and difficulty. One can spend large amounts of time iteratively improving an operating model fit to observed catches, length-composition, and CPUE data for each SAU (__`r fig_nums("f1",display="cite")`__). While this is a fine way to spend time it is also a good strategy to remember that there is this spectrum of how closely one reflects nature and that some questions do not require the best possible fit to remain useful.

# The _control.csv_ File

The control file is made up of a series of sections and each will be described in turn, using where possible, the example as generated by the function _ctrlzonetemplate_. Within the _control.csv_ file (which can have any name you wish, it does not have to be _control.csv_) main section headings are written using capital letters for increased clarity. The correct spelling of such section titles is important, as is the spelling of the variable names in the first column. The order of the sections does not need to be strictly maintained but the order of any components within each section must be maintained as described. This is required because in a number of cases they are being parsed in sequence, though many are also being obtained explicitly by name (which is why typing/spelling the names correctly is also important). Where the labels or names are not clear as to the intent of the variable then a description can be given to the right of the actual values.

The intent is to have the input file contents sent to an output text file as they are parsed, so that if the reading of data fails then the point at which it fails is made clear, but this has still to be implemented.

### DESCRIPTION

At the top of the file is an optional description section that can be used to describe the objective of the scenario implied by the control and population files. The author can add as many lines of text to this section as desired as, currently, it is not used in the model run. This may change later so it may act as meta-data for each run. It would be useful to treat this section as such meta-data anyway.

_DESCRIPTION_

_Control file containing details of a particular run. Modify the_

_contents to suit your own situation. In particular. modify the_

_contents of this description to suit the scenario being executed_

_use as many lines as required for your own description_

### START

The start section currently only has three components, a name for the particular run (which no longer gets used, though it remains presents and so could be used) and the full name of the _saubiology.csv_ file (which is expected to be found in the _datadir_, which may or may not be the same as the _rundir_). In each case, note the use of commas to separate the variable's label from the variable's value (cannot be seen if opened in Excel, but as long as the file is saved as a csv file all will be well). The hopefully descriptive text after the variable's value is ignored by the program and can be expanded if desired (as long as it remains as a single line).

_START_

* _runlabel,	 zoneX,	 label for particular run currently ignored_
* _datafile,	 zonewest.csv,	 name of popdefs file = saubiology.csv_
* _bysau,      1, 1=TRUE, 0 = FALSE are we expecting population data by SAU or population_

### zoneCOAST

The zoneCOAST section (an odd name, which may change to become something more descriptive) currently contains five components These determine whether the run will be a batch job (currently not implemented so defaults to 0 = FALSE), how many replicates the projections in the run will have (in a real run this would be something more like 500 or 1000), and a listing of three sources of variation to be included when perfect knowledge (as perfect data) is not to be the case. Other sources of uncertainty will be added here as they are included in the dynamics of projection.

_zoneCOAST_

* _batch,  0, run job interactively or as a batch (0 = FALSE)_ 
* _replicates,  100, number of replicates, usually 500 or 1000_ 
* _withsigR,  0.3, recruitment variability eg 0.3_ 
* _withsigB,  0.05, process error on exploitable biomass_ 
* _withsigCE, 1e-08, process error on cpue calculations_

_withsigR_ is the year-to-year recruitment variability. This does not include any special recruitment events or recruitment deviates that may be included (still to be developed). During the generation of the equilibrium, unfished model recruitment variation is set to a very small number, 1e-08, so that recruitment processes are effectively deterministic off the Beverton-Holt stock recruitment relationship. While there is no evidence of auto-correlated recruitment residuals these will be included as an option, but this is still to be implemented.

$$R_{p,t}=\frac{4h{R_0}{B_{p,t}^{S}}}{(1-h){B_0}+(5h-1){B_{p,t}^{S}}} e^{\varepsilon_{p,t}-\sigma_{R}^2/2}$$

where $\varepsilon_{p,t}$ is defined as:

$$\varepsilon_{p,t}=N \left(0,\sigma_{R}^2 \right)$$

_withsigB_ the fleet dynamics of where divers elect to take their catches within an SAU mean that the actual catches from each SAU differ from the aspirational catches allocated to each SAU according to whatever harvest control rule is used. This will be determined within the _jurisdictionHS.R_ file and may be different to the description given here, which reflects the case in Tasmania.  This is implemented by first allocating catches to the _SAU_ in a deterministic manner and then modifying them by imputing variability to the perceived distribution of biomass among _SAU_ $u$, which is then proagated across compoent _populations_:

$$B_{u,t}^{E,*}={B_{u,t}^{E}}e^{\varepsilon_{u,t}-\sigma_{B}^2/2}$$

where:

$$\varepsilon_{u,t}=N \left(0,\sigma_{B}^2 \right)$$

$\sigma_B$ is the standard deviation of the variation that occurs between the real exploitable biomass and the perceived distribution of exploitable biomass across _SAU's_, and also includes other sources of uncertainty.

This variation is introduced into the potential _SAU_ catches using:

$$C_{u,t}^*=TAC_t \times \frac{B_{u,t}^{E,*}}{\sum{B_{u,t}^{E}}}$$

where $TAC_t$ is the sum of the HCR's aspirational catches. One diagnostic used during conditioning is to examine the variation between the actual catches and those previously determined by the HCR, if one has such data. Otherwise, it should be characterized from the outputs, which will allow its variation to be examined for plausibility.

### ZONE

Again the use of commas to separate variable values is essential. The zone is defined in terms of the number of _SAU_ (spatial assessment units) it is made up of.

_ZONE_

* _nSAU, 8, number of spatial management units eg 2_ 
* _SAUpop, 2, 2, 2, 2, 2, 2, 2, 2, number of populations per SAU in sequence_ 
* _SAUnames, blk6, blk7, blk8, blk9, blk10, blk11, blk12, blk13, labels for each SAU_ 
* _initdepl, 0.4, 0.4, 0.4, 0.35, 0.35, 0.35, 0.35, 0.35, initial depletion levels for each SAU_

The example file from _ctrlzonetemplate_ imputes equal numbers of populations among SAU, this need not be the case. The SAUnames should be only mixtures of letters and numbers but must not contain spaces. The _initdepl_ is the intended initial depletion before either the model run or further conditioning.

### SIZE

The size structure of the operating model was originally built to suit the Tasmanian blacklip abalone fishery. Hence it extends the midpoints of each 2mm size class from 2mm - 210mm, with the 210mm class being a plus group. These values will not necessarily make sense for other jurisdictions or species. _Haliotis roie_, for example, will require a much smaller upper size, and might possibly use 1mm size classes, depending on how data are collected.

_SIZE_

* _minc, 2, centre of minimum size class, classes are 1-3,3-5,5-7..., centered at 2,4,6,..._
* _cw, 2, class width mm_
* _Nclass, 105, number of size classes, leading to a maximum of 210mm_

### RECRUIT

Larval movement between Tasmanian blacklip populations has been demonstrated to be remarkably low. Nevertheless, there will undoubtedly be some small amount of larval drift between populations, which operate on a smaller sale than the SAU within which they are found. With no other real data on this it is assumed that the populations are generally aligned linearly along the coastline and such larval movement is a constant rate.

_RECRUIT_

_larvdisp, 0.01, rate of larval dispersal eg 0.01 = 1.0 percent_

The value of _larvdisp_ is used to construct a simple larval movement matrix that assumes movement only occurs between adjacent populations at a rate of half the _larvdisp_ value. While a value of 0.01 may appear almost trivial this does effect the equilibrium dynamics, but the effects are only minor. However, by including this it is now possible to determine the explicit effect of different levels of larval movement, and this would enable the importance of the relative isolation of the different species to be explored.

### RANDOM

There will often be a need to control the generation of random numbers used during both the definition of the populations and the projections. The use of a random seed is especially important when conditioning the operating model as many random numbers are used during that process and if the conditioning is intending to simulate a specific fishery then the distribution of productivity needs to remain the same during each model run. The truth of this will become more apparent during the description of the _saubiology.csv_ file contents and uses in sections below.

The use of specific random seeds might be wanted, for example, if one wanted to be sure to repeat a particular analysis exactly, of perhaps to see only the effect of changing an argument, perhaps the LML, while making no other changes. So, to ensure the production of the same sequence of pseudo-random numbers one uses _set.seed_. Under _RANDOM_ we have two values, the first, _randomseed_ is used for repeatability of population generation when conditioning the model. By default one would expect this to have a value. The second value, _randomseedP_ is used for repeatability during the projections. Each scenario should have a different _randomseed_ and, optionally, _randomseedP_. One can (should) use _getseed()_ from __rutilsMH__ to generate such seeds. It is recommended that one not try to generate one by oneself. 12345 tends to be used more often than would be expected at random(!) so it, and others akin to it are just not good enough.

If one does not want to set the random seeds and just use any old random sequence that comes along then set _randomseed_ and _randomseedP_ to 0.

_RANDOM, 		Set these to zero for random starting points._

* _randomseed, 3543304, for repeatability of population definitions_
* _randomseedP, 0, for repeatability of Projections_

### initLML

However the model is to be run (__`r fig_nums("f1",display="cite")`__), all scenarios begin by generating an unfished, equilibrium model. This includes a characterization of the equilibrium productivity of this population and this will require a selectivity to be used, hence we need an initial LML. The fishery productivity is altered when the LML is changed because it alters the balance between natural mortality, growth, and fishing mortality. In one of the vignettes, an example is given of running the model to equilibrium under different LML and plotting up the productivity and tabulating the MSY, which shows that smaller LML do not always generate the highest levels of productivity (although that also, of course, depends upon the biological properties of growth).

_initLML, 140,  used to generate unfished zone if no historical catches and LML present_

If the conditioning is to include the application of historical catches the different SAU will still require an initial depletion (which can be set at 1.0) and an initial LML. If conditioning on historical catches then it makes sense to initiate the model at the LML in use at the start of the historical catches. If a generic MSE is being conducted, for example, a generic west coast Tasmania starting projections in 2020 or 2021, then perhaps initiate the model at an LML of 140, although, as in the file generated by _ctrlzonetemplate_, project the dynamics forward with an LML of 145 (see PROJLML later).

### PROJECT

Basic information about the projection period of the simulations. The firstyear and lastyear values are currently not used but remain available. We add together the projection years (here 30) and the inityrs value to obtain the total number of years of the final projections

_PROJECT, 30, number of projection years for each simulation_

* _firstyear, 2020, not currently used_
* _lastyear, 2049, not currently used_

### HARVESTS

This section is no longer used but remains in case a future use arises. It is no longer searched for within the _readctrlfile_ function so nothing placed in this section will influence a run (assuming other key words are not placed here!).

Instead, we now use a _jusrisdictionHS.R_ file (e.g. _tasmanianHS.R_) to contain functions and constants dealing with particular harvest strategies. A separate document will describe the minimum structural requirements of such source files.

### PROJLML

Under the _PROJECT_ heading we read in the number of projection years. Assuming that was > 0, then _readctrlfile_ looks for the heading _PROJLML_. The LML used in the projections are listed here explicitly so that changes through time can be easily implemented. There needs to be at least as many defined as the value of _PROJECT_ (which becomes _projyrs_), here that is 30. The projLML are used to define the selectivity for each of the projyrs during the projections. They are included in the _projC_ object.

_PROJLML, need at least the same number as there are projection years_

* _2020,145,  the year and the Legal Minimum Length (LML, MLL, MLS) e.g. 140_
* _2021,145,_
* _2022,145,_
* _2023,145,_
* _..._
* _..._
* _2074,145,_

### CATCHES

If _CATCHES_ is > 0 then one would need to include at least that number of years of historical catches by SAU along with the year and LML that was used when the catches were taken. An example is given below of the required format. If  _CATCHES_ = 0, then any data here will be ignored. If _CATCHES_ > 0 then _readctrlfile looks for _CondYears_ and then reads in _CATCHES_ rows with the followi9ng format.

_CONDITION,0, if > 1 then how many years in the histLML_

* _CondYears,LML,6,7,8,9,10,11,12,13,  names of the columns_
* _1973,132,0,0,0,0,0,0,0,0,_
* _1974,132,8,1,0,3,24,82,143,16, the year, the LML, the catches by SAU_
* _1975,132,21,36,42,126,130,191,143,43,_
* _..._
* _..._
* _2019,132,30,56,77,252,179,240,153,62,_

### CEYRS

_CEYRS_ is similar to _CATCHES_ in that it defines how many years of cpue data will be read in. The required format is given below for _CEYRS_ = 28. If _CEYRS_ = 0, then any cpue data will be ignored. Note that in the early years there appears to be cpue data missing from SAU6 and SAU13, which are present in the latest years. This is because, in Tasmania, zonation was only introduced in 2000, so cross-zone blocks such as 6 and 13 cannot be included prior to 2000.

_,6,7,8,9,10,11,12,13,_

_CEYRS,28,  if >0 then number reflects number of historical CPUE records by SAU_

* _1992, ,113.4, 94.2, 97.0, 99.4, 98.1, 100.19, ,_
* _1993, ,116.8, 110.38, 109.37, 99.1, 107.88, 102.20, ,_
* _..._
* _..._
* _2018, 106.68, 143.57, 148.43, 133.70, 104.38, 95.01, 106.88, 104.08,_
* _2019, 92.22, 112.70, 107.11, 93.65, 91.89, 86.94, 90.03, 90.78,_

### SIZECOMP

Size-composition data is akin to the biological properties data in being potentially a large amount of data which might confuse the structure and editing of of the _control_ file. Hence we use this to refer to a filename containing the required data. If _SIZECOMP_ = 0 then no file will be read in, otherwise a list of filenames will be expected and these should be stored in the _datadir_. The filenames should begin in the row immediately below _SIZECOMP_.

_SIZECOMP,1_

_laterdata,lateLF-84-20.csv_

<br><br>

# The _saubiology.csv_ File

This file is identified in the _control_ file. This also starts off with some optional descriptive text lines.

_SAU definitions listing Probability density function parameters for each variable_

_These are randomly aligned to each population except for the proportion of recruitment,_

_which is literally allocated down in the popREC section_

As many of these as are required can be included. The real input is signaled by the _SPATIAL_ heading, though this is not currently used explicitly. The _saubiology.csv_ file contents are read in by the function _readsaudatafile_, and each variable is identified and selected by its name (so type carefully), though they are split into groups that relate to their function. First there are lines describing the implied geographical structure being simulated, then comes a matrix of properties for each _SAU_, following on from the _VARS_ heading.

Not all variables input are used to define probability density functions. Thus, each of the _nsau_, _saupop_, _saunames_ are defined explicitly after deciding on the geographical structure to impose. They need to be listed in the order in which they are expected to occur along the coast. Each population is a member of a given _SAU_, and here we are using the Tasmanian block number as the SAU index (and the _saunames_). Each SAU can have a different number of populations.

_SPATIAL_

* _nsau, 8, number of spatial management units eg 2_
* _saupop, 3, 3, 2, 2, 2, 2, 12, 2, number of populations per SAU in sequence_
* _saunames, 6, 7, 8, 9, 10, 11, 12, 13, labels for each SAU_
* _initdepl, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, 0.7, initial depletion levels for each SAU_

The _initdepl_ value is only used if an abstract or truly generic abalone stock is intended. If the intent is to model a plausibly realistic generic stock, or a specific and real abalone stock then this line, while read, is not used.

### PDFS

_PDFs, 30, _

This line is read in to determine the number of variable values to be read (the rows to the matrix). The columns of the matrix are defined by the _nsau_ defined above.

The precision with which this matrix is filled will constitute a large part of any conditioning. Approximate and plausible values will produce a generic MSE whereas, tuning the matrix to the biological properties of a particular zone, as best one can, will generate a much more specific MSE, especially if that is followed up by conditioning the resulting modelled stock on historical fishery data. Thus, if one then included historical catches and other fishery data the result would be as realistic a modelled representation as one could get.

Here we only generate a plausible and generic zone, as can be inferred by the repeated values (which will still lead to different values between populations because the variability in each case - variables prefixed s, are all large enough to influence a random number). If specific values for a given population are wanted then set the respective variation value to 1e-08. All variables are sampled from a Normal distribution except for the average unfished recruitment, which strongly defines the upper productivity of each population, and is sampled from a log-normal distribution (see the function _definepops_).

### Growth

The parameters of the Inverse Logistic model describing growth (Haddon _et al_, 2008; Haddon & Mundy, 2016). Alternative growth functions will be implemented in the future and will be selected in this section. In each case there is the main parameter (MaxDL, L50, L50inc, and SigMAx) followed immediately by the variation to be attributed to a given population's value through sampling from a normal distribution. If values varying around a specific value are wanted for a given _SAU_, then set the respective variation value to 1e-08.

* _MaxDL ,26, 30, 30, 34, 34, 34, 34, 29, maximum growth increment_
* _sMaxDL ,2, 2, 2, 2, 2, 2, 2, 2, variation of MaxDL within each SAU_
* _L50 ,135.276, 135.276, 135.276, ..., ..., ..., ..., ..., Length at 50% MaxDL_
* _sL50 ,2, 2, 2, 2, 2, 2, 2, 2, variation of L50_
* _L50inc ,39, 39, 39, 39, 39, 39, 39, 39, L95 - L50 = delta =L50inc_
* _sL50inc ,1.5, 1.5, 1.25, 1.25, 1.25, 1.25, 1.25, 1.25, variation of L50inc_
* _SigMax  ,3.3784, 3.3784, 3.3784, ..., ..., ..., ..., ..., max var around growth_
* _sSigMax ,0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, var of SigMax_

### LML

It is not impossible that different SAU operate under different LML, this allows for the generation of the equilibrium zone to reflect such variation if present. If, however, historical catches are available and are to be used, they are defined along with the LML used in each year and those LML are used in the _zone_ definition instead.

_LML ,132, 132, 132, 132, 132, 132, 132, 132, Initial legal minimum length_

### Weight-at-Size

The weight-at-size relationships for each population are defined using the standard equation:

$$W_{p,L}=a_pL_{p}^{b_p}$$
which defines the weight of animals in population $p$ at length $L$. For this we need both the $a_p$ and $b_p$ parameters. It was found in Haddon _et al_ (2013; see pages 208-209) that there was a power relationship between the $b_p$ parameter and its corresponding $a_p$ parameter. So that this relationship is not broken when selecting random comnbinations, first the $b_p$ parameter is produced (sampled from a Normal distribution), and that is used to estimate the $a_p$ parameter, with variation. Given the advent of new samples, and different jurisdictions, in each case this relationship should be estimated first. Given the length/size-based dynamics this relationship is highly influential on the outcomes so effort with respect to this relationship is important.

* _Wtb ,3.161963, 3.161963, 3.161963, ..., ..., ..., ..., ..., weight-at-length exponent_
* _sWtb ,0.148, 0.148, 0.148, 0.148, ..., ..., ..., ..., var of Wtb_
* _Wtbtoa ,962.8098, 962.8098, ..., ..., ..., ..., ..., ..., power curve intercept_
* _sWtbtoa ,-14.3526, -14.3526, ..., ..., ..., ..., ..., ..., power curve exponent_

### Natural Mortality

This is a troublesome variable. Previous work has generally assumed a natural mortality rate of about 0.2, but this suggests that most animals would be close to senile by the age of 23 (only 1 percent would be expected to survive that long). However, very many large animals have been found, despite their growth characteristics implying that to reach such sizes would take longer than 23 years. This casts doubt on the notion of M = 0.2, which, if traced, stems from some rather limited tagging data from decades ago. As with the growth characteristics, tagging may negatively affect survivorship, this potentially affecting any early tagging study of natural mortality (biasing it high). Here we have selected 0.15 as the mean, but the uncertainty means that it would be worthwhile comparing the outcomes of any HS when M = c(0.125, 0.15, 0.2).

* _Me ,0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, 0.15, Nat Mort_
* _sMe ,0.003, 0.003, 0.003, 0.003, 0.003, 0.003, 0.003, 0.003, var of Me_

### Recruitment

The average unfished recruitment is a major influence on the productivity of each population. During conditioning it is recommended that this be one of the first variables to adjust to set-up each populations unfished production. At very least it should be possible to determine the relative long-term production of each SAU, and this can be used as a guide to determine the relative productivity of each SAU and this should then be distributed among its constituent populations. If a time-series of GPS data-logger information is available (or potentially of a series of surveys) this can be used to scale the relative productivity of each constituent population within SAU (this is done in the _propREC_ section of the _saubiology.csv_ file, as described below). The _AvRec_ value, in each case, is in linear-space for ease of use, but is commonly expressed as log(AvRec) in the model. This is defined using steepness as shown in the _zoneCOAST_ section above.

* _AvRec,222500, 390000, 128000, 745000, 690000, 1420000, 1250000, 375000, unfished recruitment R0_
* _sAvRec,0.0001, 0.0001, 0.0001, 0.0001, 0.0001, 0.0001, 0.0001, 0.0001, variaiton in R0_
* _defsteep ,0.75, 0.75, 0.75, 0.75, 0.75, 0.75, 0.75, 0.75, Beverton-Holt steepness_
* _sdefsteep ,0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02,_

### Emergence

Emergence has been studied in Tasmania through an examination of the growth of encrusting animals and algae on the animal's shells. The cover by algae (even encrusting algae) being taken as evidence of emergence (exposure to light). This only becomes influential if the emergence curve overlaps with the selectivity curve, which may happen when the LML is low. If no emergence data is present, then just set values so that the emergence curve always runs at smaller sizes than the selectivity curves (see later). These values below are invented (for now).

* _L50C ,126.4222, 126.4222, 126.4222, ..., ..., ..., ..., ..., length at 50% emergent_
* _sL50C ,0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5, 0.5,_
* _deltaC ,10, 10, 10, 10, 10, 10, 10, 10, length at 95% - 50% emergent_
* _sdeltaC ,0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1,_

### CPUE

In order to scale the predicted cpue we use _MaxCEpar_ to define the maximum cpue expected when exploitable biomass is at its greatest density. This is defined for when cpue is linear related to exploitable biomass, which is why there are some extremely high, and implausible values amongst the maxima. The option of including hyper-stability, which would lead to smaller, more plausible maxima, is being explored but still needs to be implemented and that will require an extra couple of parameters to define the relationship between exploitable biomass and cpue.

* _MaxCEpars, 0.475, 0.475, 0.54, 0.6, 0.55, 0.7, 0.43, 0.43, max cpue t-hr_
* _sMaxCEpars, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02,_

### Selectivity

Even though there is an LML, the precision with which abalone are taken with respect to the LML is not always perfect with some legal sized animals being left behind and occasionalyl some just sub-legal being taken (though these may be due to meaasuring error). Here we use a simple addition to the selectivity ogive used for each population.

* _selL50p ,0 ,0, 0, 0, 0, 0, 0, 0, L50 of selectivity, 0 = no bias_
* _selL95p ,1.5, 1.5, 1.5, 1.5, 1.5, 1.5, 1.5, 1.5, L95 of selectivity_

### Size-at-Maturity

It was found in Haddon _et al_ (2013; see pages 210-211) that there were relationships between the two parameters used to define the logistic maturity-at-size relationships for different populations. A fixed parameter value of -16 for the $a_p$ parameter led to a range of plausible values for the $b_p$ parameter

* _SaMa, -16, -16, -16, -16, -16, -16, -16, -16, maturity logistic a par_
* _L50Mat, 114, 114, 114, 114, 114, 114, 114, 114, L50 for maturity b = -1/L50_
* _sL50Mat, 2.5, 2.5, 2.5, 2.5, 2.5, 2.5, 2.5, 2.5,_

## propREC

The _propREC_ section provides details for the distribution of the recruitment levels across the populations within each _SAU_. The three columns of values are the _SAU_, the _population_, and its expected proportion of the _SAU_ recruitment each year. These values need to be set manually during the conditioning and varied until the dynamics approximate the desired dynamics.

* _SAU,pop,propR_
* _6, 1, 0.16, in SAU 6 about 1/6 of recruitment goes to pop 1_
* _6, 2, 0.82, in SAU 6  most recruitment goes to pop2_
* _6, 3, 0.02, in SAU 6 almost no recruitment goes to pop3_ 
* _7, 4, 0.31, in SAU 7 a somewhat more even distribution of recruitment occurs _ 
* _7, 5, 0.46_
* _7, 6, 0.23_
* _8, 7, 0.491_
* _..._
* _..._
* _12 ,26, 0.022_
* _13 ,27, 0.5, in SAU 13 an equal distribution is suggested__
* _13 ,28, 0.5_

<br>

## References

Haddon, M., Mayfield, S., Helidoniotis, F., Chick, R. and C. Mundy (2013) _Identification and Evaluation of Performance Indicators for Abalone Fisheries_. FRDC Final Report 2007/020. CSIRO Oceans and Atmosphere and Fisheries Research Development Corporation. 295 p.

Haddon, M. and C. Mundy (2016) _Testing abalone empirical harvest strategies for setting TACs and associated LMLs, which include the use of novel spatially explicit performance measures._ FRDC Final Report 2011/028. CSIRO Oceans and Atmosphere and Fisheries Research Development Corporation. Hobart 182p.

Haddon, M., Mundy, C., and D. Tarbath (2008) Using an inverse-logistic model to describe growth increments of blacklip abalone (_Haliotis rubra_) in Tasmania. _Fishery Bulletin_ __106__:58-71

### 2023-07-12 data-raw/vignettefiles/the_data_files.Rmd
Now use of refperiodCE is optional. No EWN.

--- 2023-07-12 data-raw/vignettefiles/Comparing_Scenarios.Rmd
@@
 ---
-title: "Comparison of MSE Scenarios"
+title: "Abalone MSE Data files"
 author: "Malcolm Haddon, ..."
@@
 vignette: >
-  %\VignetteIndexEntry{Comparison of MSE Scenarios}
+  %\VignetteIndexEntry{Abalone MSE Data files}
   %\VignetteEngine{knitr::rmarkdown}
@@
 
-suppressPackageStartupMessages({
 library(aMSE)
-library(TasHS)
 library(codeutils)
@@
 library(knitr)
-library(makehtml)
 library(captioner)
-})
 
@@
 
-dbdir <- getDBdir()
-ddir <- paste0(dbdir,"/A_CodeUse/aMSEDoc/") 
-prefixdir <- paste0(dbdir,"/A_CodeUse/aMSEUse/scenarios/")
+if (dir.exists("C:/Users/Malcolm/Dropbox")) {
+  ddir <- "C:/Users/Malcolm/Dropbox/A_CodeUse/aMSEUse/documentation/"
+} else {
+  ddir <- "C:/Users/User/Dropbox/A_CodeUse/aMSEUse/documentation/"
+}
 
@@
 Being one of the vignettes included in the package it should be regarded as a living, adaptable document that will be modified as experience increases around different jurisdictions, and the authorship should be expected to alter accordingly as different people contribute.
-It is assumed in this vignette that the reader has managed to run a number of scenarios that they then wish to compare.
 
-### Saving the Output from aMSE
-
-One concept in the use of __aMSE__ is that of the _rundir_, which is the directory into which all the data files are placed and all the results, plots, and html files will be placed.
-However, a typical scenario MSE run will include a final line where the output from the _do_MSE()_ function is saved into an _outdir_.
-The _outdir_ should not be on a cloud service shared with others as the output objects can be large (gigabytes).
-so typically it will point to a directory on a fast hard-drive or SSD that should be used to identify a given series of scenarios for comparison.
-The actual scenario name is usually encapsulated in the _postfixdir_ name:
-
-```{r echo=TRUE, eval=FALSE}
-# A subset of commands to run a given scenario
-postfixdir <- "BC"                   # BC stands for basecase
-rundir <- filenametopath(prefixdir,postfixdir)
-outdir <- "C:/aMSE_scenarios/BC/"
-confirmdir(rundir,ask=FALSE)     # check to see if it already exists
-confirmdir(outdir,ask=FALSE)     # and if not create it without asking.
-
-hsargs <- list(mult=0.1, # expansion factor for cpue range when calc the targqnt
-               wid = 4, # number of years in the grad4 PM
-               targqnt = 0.55, # quantile defining the cpue target
-               maxtarg = c(150,150,150,150,150,150,150,150), # max cpue Target
-               pmwts = c(0.65,0.25,0.1),  # relative weights of PMs
-               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2), # hcr multipliers
-               startCE = 2000, # used in constant reference period HS
-               endCE = 2019,   # used in constant reference period HS
-               metRunder = 1,  # should the metarules be used. o =
-               metRover = 1,   # use metarules
-               decrement=1, # use fishery data up to the end of the time series
-               pmwtSwitch = 0, # number of years after reaching the targCE to
-               stablewts = c(0.4, 0.5, 0.1), # replace pmwts with stablewts
-               hcrname="mcdahcr",     # the name of the HCR used
-               printmat=NULL)
-
-out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=mcdahcr,
-              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
-              getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
-              hcrscoreoutputs=extractandplotscores,HSPMs=getcpueHS,
-              fleetdyn=NULL,interimout=c(outdir,postfixdir),
-              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
-              matureL=c(40,170),wtatL=c(50,200),mincount=120,
-              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15))
-
-makeoutput(out,rundir,postfixdir,controlfile,hsfile="TasHS Package",
-           openfile=TRUE,verbose=FALSE)
-
-save(out,file=paste0(outdir,postfixdir,".RData"))  # the results are saved
-```
-
-## Making comparisons
+## Stages of an MSE Run
 

### 2023-07-12 R/projection.R

@@ -233,13 +233,8 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
       proprec[i,] <- envimpact[["proprec"]][i,sauindex]
     }
   }
-  if (length(hsargs$refperiodCE) > 0) {
-    beginCE <- hsargs$refperiodCE[1]
-    } else {
-    beginCE <- hsargs$startCE
-  }
   for (year in startyr:endyr) { # iter=1; year=startyr
-    if (verbose) cat(year,"   ")
+    if (verbose) cat(year," ")
       for (iter in 1:reps) {
       hcrdata <- getdata(sampleCE,sampleFIS,sampleNaS,
                          sauCPUE=zoneDP$cesau[,,iter],
@@ -247,7 +242,7 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
                          sauNAS=list(Nt=zoneDP$Nt[,,,iter],
                          catchN=zoneDP$catchN[,,,iter],
                          NumNe=zoneDP$NumNe[,,,iter]),year=year,
-                         startCE=beginCE,decrement=hsargs$decrement)
+                         startCE=hsargs$startCE,decrement=hsargs$decrement)
       hcrout <- hcrfun(hcrdata,hsargs,saunames=glb$saunames,curryear=year)
       calcpopCout <- calcpopC(hcrout,exb=zoneDP$exploitB[year-1,,iter],
                        sauCPUE= zoneDP$cesau[year-1,,iter],

### 2023-07-20 R/projection.R
IMplemented incremental HS score logging and meta-rule flags.

@@ -146,6 +146,9 @@ calcsau <-  function(invar,saunames,ref0) {# for deplsb depleB
   return(res)
 } # end of calcsau
 
+# removed just after calcpopC
+
+
 #' @title doprojections conducts the replicate model runs for Tasmania
 #'
 #' @description doprojections conducts the replicate model runs for
@@ -190,7 +193,7 @@ calcsau <-  function(invar,saunames,ref0) {# for deplsb depleB
 #' }
 #'
 #' @return a list containing the full dynamics across all years zoneDP, and the
-#'     final output from the HCR as hcrout
+#'     final output from the HCR as hcrout and outhcr
 #' @export
 #'
 #' @examples
@@ -198,9 +201,9 @@ calcsau <-  function(invar,saunames,ref0) {# for deplsb depleB
 doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
                           sampleCE,sampleFIS,sampleNaS,getdata,calcpopC,
                           makehcrout,fleetdyn,verbose=FALSE,...) {
-  # ctrl=ctrl; zoneDP=zoneDP; zoneCP=zoneCP; glb=glb; hcrfun=mcdahcr; hsargs=hsargs
+  # ctrl=ctrl; zoneDP=zoneDP; zoneCP=zoneCP; glb=glb; hcrfun=hcrfun; hsargs=hsargs
   # sampleCE=tasCPUE; sampleFIS=tasFIS; sampleNaS=tasNaS;  getdata=tasdata
-  # calcpopC=calcexpectpopC; verbose=TRUE;makehcrout=makeouthcr
+  # calcpopC=calcexpectpopC; verbose=TRUE; fleetdyn=NULL;makehcrout=makeouthcr
   reps <- ctrl$reps
   sigmar <- ctrl$withsigR
   sigmab <- ctrl$withsigB
@@ -211,13 +214,13 @@ doprojections <- function(ctrl,zoneDP,zoneCP,glb,hcrfun,hsargs,
   endyr <- hyrs + pyrs
   sauindex <- glb$sauindex
   saunames <- glb$saunames
+  nsau <- glb$nSAU
   yrnames <- c(glb$hyrnames,glb$pyrnames)
   Nclass <- glb$Nclass
   movem <- glb$move
   r0 <- getvar(zoneCP,"R0") #R0 by population
   b0 <- getvar(zoneCP,"B0") #sapply(zoneC,"[[","B0")
   exb0 <- getvar(zoneCP,"ExB0")
[…]

### 2023-07-26 README.Rmd
IMplemented incremental HS score logging and meta-rule flags.
Revised treament of outputs and HS.
Generalized compare-scenarios. Added tasphaseplot to kobe plots.

@@
 # LATEST UPDATES
+
+* 2023-07-26 aMSE 0.1.9 Added tasphaseplots to the kobe plots for comparison.
+
+* 2023-07-24 aMSE 0.1.8 Revised use of TasHS and of comparescenarios, meta-rule flagging now working correctly.
+
+* 2023-07-20 aMSE 0.1.7 Implemented the use of incremental updates of HS scores in a jurisdiction generic manner. Added a meta-rule flagging system.
 

### 2023-08-01 R/domse.R
Added scoreplot and plotmultflags to arguments for do_MSE

@@ -110,6 +110,14 @@
 #'     it will contain all the HS score values..
 #' @param fleetdyn a function that calculates the distribution of catch across
 #'     the sau and populations. Currently not needed by Tas but needed by SA
+#' @param scoreplot the function that plots the hcr scores and final scores. In
+#'     Tas = plotfinalscores, in SA = SAplotfinalscores. If more inputs than
+#'     outhcr, zoneDP, sau, and filen=filename are needed then define your own
+#'     function using the ellipsis at the end of the argument list.
+#' @param plotmultflags the function that plots the TAC/acatch multipliers, and
+#'     meta rule flags. In Tas = plotmultandflags. If more inputs than
+#'     outhcr, zoneDP, sau, and filen=filename are needed then define your own
+#'     function using the ellipsis at the end of the argument list.
 #' @param interimout should results be saved after projections have finished?
 #'     the default="", which means nothing is saved if no risk of crashing in
 #'     subsequent analyses. But if trying new stuff which may waste lots of time
@@ -159,7 +167,7 @@
 #' print("wait on suitable data sets in data")
 do_MSE <- function(rundir,controlfile,hsargs,hcrfun,sampleCE,sampleFIS,
                    sampleNaS,getdata,calcpopC,makeouthcr,fleetdyn=NULL,
-                   interimout="",
+                   scoreplot,plotmultflags,interimout="",
                    varyrs=7,startyr=42,verbose=FALSE,ndiagprojs=3,
                    cutcatchN=56,matureL=c(70,200),wtatL=c(80,200),mincount=100,
                    includeNAS=FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15)) {
@@ -296,12 +304,12 @@ do_MSE <- function(rundir,controlfile,hsargs,hcrfun,sampleCE,sampleFIS,
     filename <- pathtopath(rundir,filen)
     caption <- paste0("Harvest strategy scores and outputs for ",
                       glb$saunames[sau])
-    scoremed <- plotfinalscores(outhcr,zoneDP,sau,filen=filename)
+    scoremed <- scoreplot(outhcr,zoneDP,sau,filen=filename)
     addplot(filen,rundir=rundir,category="scores",caption)
     filen <- paste0("HS_Mult_MetaFlag_plots_",glb$saunames[sau],".png")
     filename <- pathtopath(rundir,filen)
[…]

### 2023-10-17 README.Rmd
Added scoreplot and plotmultflags to arguments for do_MSE
added poplevelplots for catch and cpue to output

@@
 # LATEST UPDATES
+
+* 2023-10-17 aMSE 0.1.13 added poplevelplot() plots to the output.
+So far this includes catches and cpue.
+Population level plots can give insight into the effectiveness of the fleet dynamics and whether each population is being exploited.
+
+* 2023-08-01 aMSE 0.1.10 Added scoreplots and plotmultflags into the arguments for do_MSE rather than requiring the functions be called plotfinalscores() and plotmultandflags(), so any name can be given to them.
 

### 2023-10-17 R/domse.R

@@ -331,7 +331,8 @@ do_MSE <- function(rundir,controlfile,hsargs,hcrfun,sampleCE,sampleFIS,
   addtable(hcrout$refpts,"hcrout_refpts.csv",rundir,category="HSperf",
            caption="HCR reference points")
   # plot the scores
-  for (sau in 1:glb$nSAU) {
+  nSAU <- glb$nSAU
+  for (sau in 1:nSAU) {
     filen <- paste0("HS_Score_plots_",glb$saunames[sau],".png")
     filename <- pathtopath(rundir,filen)
     caption <- paste0("Harvest strategy scores and outputs for ",
@@ -344,16 +345,26 @@ do_MSE <- function(rundir,controlfile,hsargs,hcrfun,sampleCE,sampleFIS,
     plotmultflags(outhcr,zoneDP,sau,filen=filename)
     addplot(filen,rundir=rundir,category="scores",caption)
   }
+  # generate population plots
+  popmedcatch <- vector(mode="list",length=nSAU)
+  names(popmedcatch) <- glb$saunames
+  popmedcpue <- vector(mode="list",length=nSAU)
+  names(popmedcpue) <- glb$saunames
+  for (sau in 1:nSAU) {
+    saumed <- poplevelplot(rundir=rundir,sau=sau,popvar=zoneDP$catch,glb=glb,
+                 label="Catch",console=FALSE)
+    popmedcatch[[sau]] <- saumed
+    saumed <- poplevelplot(rundir=rundir,sau=sau,popvar=zoneDP$cpue,glb=glb,
+                           label="CPUE",console=FALSE)
+    popmedcpue[[sau]] <- saumed
+  }
   # generate sau phase plots
-  nSAU <- glb$nSAU
   kobedata <- vector(mode="list",length=nSAU)
   names(kobedata) <- glb$saunames
   for (plotsau in 1:glb$nSAU) {
     twophaseplots(dyn=sauout,glb=glb,outhcr=outhcr,sau=plotsau,kobeRP=kobeRP,
                   rundir=rundir,startyr=condC$yearCE[1],console=FALSE,fnt=7)
   }
-
-
   if (!includeNAS) NAS=NULL
   out <- list(tottime=tottime,projtime=projtime,starttime=starttime,glb=glb,
               ctrl=ctrl,zoneCP=zoneCP,zoneD=zoneD,zoneDD=zoneDD,zoneDP=zoneDP,
[…]

### 2023-10-17 R/plotfuns.r

@@ -1240,6 +1240,85 @@ plotzonesau <- function(zonetot,saudat,saunames,label,labelsau,side=3,
   }
 } # end of plotzonesau
 
+#' @title poplevelplot generates population x replicate trajectory plots
+#'
+#' @description poplevelplot is used to produce population level plots of an
+#'     input variable derived from zoneDP. It produces a single plot for a given
+#'     sau which depicts each population's replicate trajectories, all in grey.
+#'     However, medians of the variable through each year are calculated and
+#'     plotted in distinct colours for each population up to a maximum of 12
+#'     populations per sau. If more than 12 then the population index 1:npop is
+#'     used. These plots should be considered as a partial diagnostic for how
+#'     well the implied fleet dynamics is distributing catch among an sau's
+#'     populations. The trajectories are plotted from the last year of
+#'     conditioning to the last year of projections.
+#'
+#' @param rundir the scenario's directory for datafiles and results
+#' @param sau which sau is to be worked on. In Tas = 1:8 in the west
+#' @param popvar which 3D array population variable is to be used. This will be
+#'     generally derived from zoneDP, eg zoneDP$catch or zoneDP$depleB
+#' @param glb the globals object used for sauindex, saunames, hyrs, and reps
+#' @param label the sau name will be added to this label for the y-axis, this
+#'     should be simple, such as 'catch' or have words connected by hyphens or
+#'     underscores as it will also be added to the filename.
+#' @param console should the plot go to the console or be saved into rundir.
+#'     default=TRUE = to the console
+#'
+#' @return invisibly the population medians
+#' @export
+#'
+#' @examples
+#' print("poplevelplot(rundir=rundir,sau=7,popvar=zoneDP$catch,glb=glb,")
+#' print("label='Catch',console=TRUE")
+poplevelplot <- function(rundir,sau,popvar,glb,label="",console=TRUE) {
+  popcol <- c("red","#FF7790","orange","yellow","#80FE90","green","cyan",
+              "#0080FF","blue","#8000FF","magenta","#FF0080")
+  sauindex <- glb$sauindex
+  saunames <- glb$saunames
[…]

### 2023-11-07 README.Rmd
Amended plotmultflags to accept sauout rather than zoneDP.
Enabled upliH and incH to be varied by jurisdiction. No EWN

@@
 
+* 2023-11-07 aMSE 0.1.15 Generalized the estimation of productivity to allow for alternative values of the starting H and upper limit to H to be applied to the dynamics while searching for the productivity curves.
+The production matrix is also used by dodepletion, so uplimH values > 0.4 were required (by SA).
+Only a small cahnge is now required in do_MSE, see the respective help.
+
 * 2023-10-17 aMSE 0.1.13 added poplevelplot() plots to the output.
@@
 Population level plots can give insight into the effectiveness of the fleet dynamics and whether each population is being exploited.
+
+* 2023-10-18 aMSE 0.1.14 A catch-up event where I have added some population level plots, I have fixed up the use of plotmultflags so that the exploitable biomass depletion plots really are at the sau level bu changing the input expected.
 

### 2025-04-09 README.Rmd
Prior to making aMSE public.

--- 2023-07-26 README.Rmd
@@
 
-This packages up a new Abalone Management Strategy Evaluation framework.
-It has a novel structure that is based around the spatial ideas of populations within spatial management units (SAUs), within a zone:
+This is the codebase for an Abalone Management Strategy Evaluation framework, which we call **aMSE**.
+It has a novel structure that is based around the spatial ideas of populations (or areas of persistent production or apps, reflecting how they are defined), within spatial management units (SAUs), within a zone:
 
@@
 
-* population - literally a population.
+* population/app - a population, herein defined as an area of persistent production.
 These are the active components within the simulation framework.
-The dynamics of the simulation are based around the populations, although, with positive larval dispersal (the default) there is some dependency of neighbouring populations on each other.
+The dynamics of the simulation are based around the populations, although, with positive larval dispersal (a low level is the default) there is some dependency of neighbouring populations on each other.
 
@@
 
-# LATEST UPDATES
-
-* 2023-07-26 aMSE 0.1.9 Added tasphaseplots to the kobe plots for comparison.
-
-* 2023-07-24 aMSE 0.1.8 Revised use of TasHS and of comparescenarios, meta-rule flagging now working correctly.
-
-* 2023-07-20 aMSE 0.1.7 Implemented the use of incremental updates of HS scores in a jurisdiction generic manner. Added a meta-rule flagging system.
-
-* 2023-07-10 aMSE 0.1.6 Modified TasHS and aMSE so that instead of using hsargs$startCE and hsargs$endCE to define a vector of reference years when using the constantrefhcr in Tasmania, one should now use a hsargs$refperiodCE which should be a vector of years describing the constant reference period to be used (assuming one is using the constantrefhcr).
-Within TasHS there is also a checkhsargs() function which provides a brief check of the hsargs being used (strictly that is still under development).
-
-* 2023-06-11 aMSE 0.1.5 Modified the condition plot and fixed up a few places where numbers were hard wired into the software.
-Introduced a new option to save output following projections (just in case of experiments with subsequent analyses crash - I'm looking at you SA).
-Now one can set interimout =c(outdir,postfixdir) and that should save a tempout into your outdir directory.
-See the help for do_MSE.
-
-* 2023-06-04 aMSE 0.1.4 Small modifications to generalize across jurisdictions and allow for the consthcr constant catch harvest strategy; also removed irrelavent messages when plotting the phase plot
-
-* 2023-06-03 aMSE 0.1.3 Modified to allow the mcdahcr and constrefhcr HCR to run successfully with the conditioned SA setup.
-
-* 2023-04-17 aMSE 0.1.2 Amended save_hsargs so it can handle data.frame within hsargs.
-
-* 2023-04-12 aMSE 0.1.1 Started generalizing the outputs from aMSE by beginning to separate functions that relate to specific jurisdiction outputs into a separate R package (in Tasmania it is currently called aMSEExtra).
-
-* 2023-04-04 aMSE 0.1.0 Added curryear to hcrfun so all HS functions should now include curryear inthier arguments.
-Within doprojecitons() it is set to year inside the loop.
-This was needed to allow the SA HS to use its fleet dynamics model and is not needed in Tasmania, but is required in the argument list.
-
-* 2023-03-31 aMSE 0.0.25 Modified calcpopC througout to allow for SA's harvest strategy and fleet dynamics.
-
-* 2022-12-14 aMSE 0.0.24 Lots of minor edits and modifications aimed at improving the comparison of scenarios.
-
-* 2022-11-11 aMSE 0.0.23 Added a new tab = recruits, containing plots of the stock recruitment relationships at an sau and at a population within sau level. Added a table to the Tables tab.
-
-* 2022-11-10 aMSE 0.0.22 now have the capacity to introduce productivity events in specific projection years.
-This can now be done at sau specific levels.
-One defines the survivorship of recruits in each selected year and also the survivorship of all post-settlement animals.
-No changes are needed to one's control file if this option is not being used.
-
-* 2022-11-02 aMSE 0.0.21 added changes to how do_MSE responds to hsargs.
-
-* 2022-10-31 aMSE 0.0.20 added do_comp_outputs, scenebyvar, scenebyzone, projectiononly, and sauquantbyscene to assist with comparing scenarios.
-
 ## Installation
 
-Once the MSE framework is made public you can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:
+You can install the development version from [GitHub](https://github.com/haddonm/aMSE) with:
 
@@
 
-devtools::install_github("https://github.com/haddonm/aMSE",build_vignettes = TRUE)
+devtools::install_github("https://github.com/haddonm/aMSE")
+
+# similarly for codeutils, hplot, EGHS, and makehtml
+# knitr you should get from CRAN
 ```
 
-Alternatively, while the development version remains private, you can generate a branch that you can work on by cloning the repository, which, again, can be done very simply within RStudio.
-Open the New Project option in the project dialog at the top right of the RStudio screen and selection Version Control, then use 'https://github.com/haddonm/aMSE' in the top box, identify where you want the new directory put, and press return.
-ALternatively, you could download the zip file from inside the 'code' button and establish an R project from that.
+The working of the **aMSE** software and how to use it has been documented during the developemt and this is available as a GitBook at:
 
-It would be a good idea to read Hadley Wickham's draft chapter on Git and GitHub at https://r-pkgs.org/index.html.
+https://haddonm.github.io/aMSEGuide/
+
+That can ether be read online or a PDF of the whole downloaded (look for the
+small Acrobat symbool at top left).
 
@@
 suppressPackageStartupMessages({
-  library(aMSE)
-  library(TasHS)
-  library(hutils)
-  library(hplot)
-  library(makehtml)
-  library(knitr)
+  library(aMSE)       # from github.com/haddonm MSE codebase
+  library(EGHS)       # from github.com/haddonm example Tas harvest strategy
+  library(codeutils)  # from github.com/haddonm a set of code utilities
+  library(hplot)      # from github.com/haddonm a set of plotting routines
+  library(makehtml)   # from github.com/haddonm used to display results
+  library(knitr)      # from CRAN for generating tables
 })
-dbdir <- getDBdir()
+dbdir <- getDBdir()   # utility to define the dropbox directory
 # Obviously you should modify the rundir and datadir to suit your own setup
 prefixdir <- paste0(dbdir,"A_CodeUse/aMSEUse/scenarios/")
-postfixdir <- "M15h75"
-rundir <- paste0(prefixdir,postfixdir)
+postfixdir <- "M15h75"      # a name for your scenario 
+rundir <- paste0(prefixdir,postfixdir) # files and results directory
 confirmdir(rundir)
-controlfile <- "controlM15h75.csv"
+controlfile <- "controlM15h75.csv"    # name of the control file
 # equilibrium zone -------------------------------------------------------------
-# You now need to ensure that there is, at least, a control.csv, and a 
-# constantsdata.csv file in the data directory plus some other data .csv files
+# You now need to ensure that there is, at least, a control??.csv, and a 
+# saudata??.csv file in the data directory plus some other data .csv files
 # depending on how conditioned you want the model to be. Templates for the
@@
 # is listed as the datafile within the control.csv. These must be stored in 
-# rundir. There are example files in the DropBox folder:
-# .\..\Dropbox\National abalone MSE\aMSE_files\scenarios\examplerun. Read the 
-# 'Using_aMSE' section in the Documentation .docx file in the documentation 
-# directry  Then, assuming you have the very latest version of aMSE = 500, 
-# the code in the documentation file (or in the 'run_aMSE.R' file within the
-# examplerun subdirectory should work.
+# rundir. Read the Using_aMSE' section in the Documentation .docx file in the 
+# documentation. To see examples of code to run the MSE for a real question.
 
-# HarvestStrategy.R should also be in rundir
-source(paste0(rundir,"/TasHS1_Tas.R"))
-# run the scenario --------------------- Obviously unhash this to make it work
-#
-# out <- do_MSE(rundir,controlfile,datadir,hsargs=hsargs,hcrfun=mcdahcr,
+# Define the arguments used in teh harvest strategy
+hsargs <- list(mult=0.1,
+               wid = 4,
+               targqnt = 0.55,
+               maxtarg = c(150,150,150,150,150,150,150,150),
+               pmwts = c(0.65,0.25,0.1),
+               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2),
+               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.15,1.2,1.25),
+               startCE = 1992,
+               endCE = 2020,
+               metRunder = 2,
+               metRover = 2,
+               decrement=1,
+               pmwtSwitch = 4,
+               stablewts = c(0.4, 0.5, 0.1),
+               hcrname="constantrefhcr",
+               printmat=NULL)
+
+hsargs <- checkhsargs(hsargs)
+
+deleteyrs <- 0  # none removed in the western zone
+
+# One would need to unhash this to get it to run (obviously)
+
+#out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=constantrefhcr,
 #              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
 #              getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
-#               varyrs=7,startyr=48,verbose=TRUE,
-#               ndiagprojs=4,savesauout=TRUE)
-# 
-# makeoutput(out,rundir,datadir,postfixdir,controlfile,openfile=TRUE,verbose=FALSE)
+#              fleetdyn=NULL,scoreplot=plotfinalscores,
+#              plotmultflags=plotmultandflags,interimout="",
+#              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
+#              matureL=c(40,170),wtatL=c(50,200),mincount=120,
+#              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15),
+#              nasInterval=1,minsizecomp=c(100,135),uplimH=0.375,incH=0.005,
+#              deleteyrs=deleteyrs,selectyr=0)
+
+#makeoutput(out,rundir,postfixdir,controlfile,hsfile="TasHS Package",
+#           openfile=TRUE,verbose=FALSE)
+
+# save(out,file=paste0(outdir,postfixdir,".RData"))
 
@@
 
-See the vignette Running_aMSE.Rmd for a more detailed example.
-
-See the New.Rd for recent developments
+See the _aMSEGuide_ for detailed examples.

### 2025-04-09 data-raw/some-of-the-history.Rmd
Prior to making aMSE public.

# LATEST UPDATES

* 2023-11-07 aMSE 0.1.15 Generalized the estimation of productivity to allow for alternative values of the starting H and upper limit to H to be applied to the dynamics while searching for the productivity curves. The production matrix is also used by dodepletion, so uplimH values > 0.4 were required (by SA). Only a small cahnge is now required in do_MSE, see the respective help.

* 2023-10-17 aMSE 0.1.13 added poplevelplot() plots to the output. So far this includes catches and cpue. Population level plots can give insight into the effectiveness of the fleet dynamics and whether each population is being exploited.

* 2023-10-18 aMSE 0.1.14 A catch-up event where I have added some population level plots, I have fixed up the use of plotmultflags so that the exploitable biomass depletion plots really are at the sau level bu changing the input expected.

* 2023-08-01 aMSE 0.1.10 Added scoreplots and plotmultflags into the arguments for do_MSE rather than requiring the functions be called plotfinalscores() and plotmultandflags(), so any name can be given to them.

* 2023-07-26 aMSE 0.1.9 Added tasphaseplots to the kobe plots for comparison.

* 2023-07-24 aMSE 0.1.8 Revised use of TasHS and of comparescenarios, meta-rule flagging now working correctly.

* 2023-07-20 aMSE 0.1.7 Implemented the use of incremental updates of HS scores in a jurisdiction generic manner. Added a meta-rule flagging system.

* 2023-07-10 aMSE 0.1.6 Modified TasHS and aMSE so that instead of using hsargs$startCE and hsargs$endCE to define a vector of reference years when using the constantrefhcr in Tasmania, one should now use a hsargs$refperiodCE which should be a vector of years describing the constant reference period to be used (assuming one is using the constantrefhcr). Within TasHS there is also a checkhsargs() function which provides a brief check of the hsargs being used (strictly that is still under development).

* 2023-06-11 aMSE 0.1.5 Modified the condition plot and fixed up a few places where numbers were hard wired into the software. Introduced a new option to save output following projections (just in case of experiments with subsequent analyses crash - I'm looking at you SA). Now one can set interimout =c(outdir,postfixdir) and that should save a tempout into your outdir directory. See the help for do_MSE.

* 2023-06-04 aMSE 0.1.4 Small modifications to generalize across jurisdictions and allow for the consthcr constant catch harvest strategy; also removed irrelavent messages when plotting the phase plot

* 2023-06-03 aMSE 0.1.3 Modified to allow the mcdahcr and constrefhcr HCR to run successfully with the conditioned SA setup.

* 2023-04-17 aMSE 0.1.2 Amended save_hsargs so it can handle data.frame within hsargs.

* 2023-04-12 aMSE 0.1.1 Started generalizing the outputs from aMSE by beginning to separate functions that relate to specific jurisdiction outputs into a separate R package (in Tasmania it is currently called aMSEExtra).

* 2023-04-04 aMSE 0.1.0 Added curryear to hcrfun so all HS functions should now include curryear inthier arguments. Within doprojecitons() it is set to year inside the loop. This was needed to allow the SA HS to use its fleet dynamics model and is not needed in Tasmania, but is required in the argument list.

* 2023-03-31 aMSE 0.0.25 Modified calcpopC througout to allow for SA's harvest strategy and fleet dynamics.

* 2022-12-14 aMSE 0.0.24 Lots of minor edits and modifications aimed at improving the comparison of scenarios.

* 2022-11-11 aMSE 0.0.23 Added a new tab = recruits, containing plots of the stock recruitment relationships at an sau and at a population within sau level. Added a table to the Tables tab.

* 2022-11-10 aMSE 0.0.22 now have the capacity to introduce productivity events in specific projection years. This can now be done at sau specific levels. One defines the survivorship of recruits in each selected year and also the survivorship of all post-settlement animals. No changes are needed to one's control file if this option is not being used.

* 2022-11-02 aMSE 0.0.21 added changes to how do_MSE responds to hsargs.

* 2022-10-31 aMSE 0.0.20 added do_comp_outputs, scenebyvar, scenebyzone, projectiononly, and sauquantbyscene to assist with comparing scenarios.

### 2025-04-09 data-raw/detibble_where_needed.R

@@ -1,0 +1,22 @@
+
+
+library(tibble)
+
+# Creating a tibble
+my_tibble <- tibble(
+  `year` = 1:5,
+  `catch` = letters[1:5],
+  `cpue` = c(TRUE, FALSE, TRUE, FALSE, TRUE)
+)
+
+# Printing the tibble
+print(my_tibble)
+
+# Accessing a column with a non-syntactic name
+my_tibble$year
+#Summary
+#Tibbles are a powerful and user-friendly alternative to traditional 
+#data frames, offering improved printing, subsetting, and error handling, 
+#along with better integration with modern data manipulation packages like 
+#dplyr. They make working with data in R more efficient and less error-prone.
+

### 2025-04-20 README.Rmd
fixed sauribbon, amended readme. No EWN

@@
 This is the codebase for an Abalone Management Strategy Evaluation framework, which we call **aMSE**.
-It has a novel structure that is based around the spatial ideas of populations (or areas of persistent production or apps, reflecting how they are defined), within spatial management units (SAUs), within a zone:
+By 'we' the intent is Malcolm Haddon and Craig Mundy of University of Tasmania, IMAS-FA, Cathy Dichmont of CDC-Consultancy, and Owen Burnell of SARDI, SA.
 
-* zone - highest geographical level. Is simply the totality of the spatial management units.
+The MSE has a novel structure that is based around the spatial ideas of populations (or areas of persistent production or apps, reflecting how they are defined), within spatial assessment units (SAUs), within a zone:
 
-* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks.
+* zone - highest geographical level. Is simply the totality of the spatial assessment units, or more simply a quota zone.
+
+* SAU - spatial assessment units. In Tasmania these would currently be the classical statistical blocks within each quota zone.
 
@@
 These are the active components within the simulation framework.
-The dynamics of the simulation are based around the populations, although, with positive larval dispersal (a low level is the default) there is some dependency of neighbouring populations on each other.
+The dynamics of the simulation are based around the populations, although, with positive larval dispersal (a low level is the default) there is some, although very limited, dependency of neighbouring populations on each other.
 
@@
 
+**aMSE** and the other packages used, all come with a version number.
+This is because they are all under different levels of on-going development.
+For example, **makehtml** hasn't changed since September 2024, and that was only a minor change to the _addtext_ function.
+Other packages changes more often.
+**aMSE** had some additions included on 20 April 2025.
+In fact, **aMSE** may be expected to continue being developed as it is a useful tool in the on-going examination and improvement of existing harvest strategies and the code is added to each time something novel is required.
+However, such additions are designed so as not to disturb the operation and functionality of what is already available.
+This paragraph is simply here to warn that changes may occur but to point out that these should not detract from everyday use.
+
 ## Installation
@@
 
-# similarly for codeutils, hplot, EGHS, and makehtml
+# similarly for codeutils, hplot, EGHS, and makehtml, which are also required
 # knitr you should get from CRAN
@@
 
-The working of the **aMSE** software and how to use it has been documented during the developemt and this is available as a GitBook at:
+The working of the **aMSE** software and how to use it has been documented during the development and this is available as a GitBook at:
 
@@
 
-That can ether be read online or a PDF of the whole downloaded (look for the
-small Acrobat symbool at top left).
+That can either be read online or a PDF of the whole downloaded (look for the small Acrobat symbol at top left and click that to download the latest published version).
+AMSEGuide can be considered as a 'living' document in that it may receive updates and additions as extra functionality is added to the **aMSE** codebase.
+
+The version number of the first publicly available version is 1.0.0, with a date of 20-04-2025. This will alter if changes are made to the software.
 
@@
 
-This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania MCDA using only 50 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000).
-It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological datafile describing each population.
+This is an example which illustrates the generation of an initial equilibrium, which then goes on to apply an early version of the Tasmania Multi-Criterion Decision Analysis (MCDA) using only 50 replicates (that bit takes about 15 seconds on my computer, hence 2.5 minutes for 1000).
+It uses built in data-sets but usually you would read in a control file, which would contain the name of the biological data file describing each population.
 
@@
 suppressPackageStartupMessages({
-  library(aMSE)       # from github.com/haddonm MSE codebase
-  library(EGHS)       # from github.com/haddonm example Tas harvest strategy
-  library(codeutils)  # from github.com/haddonm a set of code utilities
-  library(hplot)      # from github.com/haddonm a set of plotting routines
-  library(makehtml)   # from github.com/haddonm used to display results
+  library(aMSE)       # from github.com/haddonm/aMSE MSE codebase
+  library(EGHS)       # from github.com/haddonm/EGHS example Tas harvest strategy
+  library(codeutils)  # from github.com/haddonm/codeutils a set of code utilities
+  library(hplot)      # from github.com/haddonm/hplot a set of plotting routines
+  library(makehtml)   # from github.com/haddonm/makehtml used to display results
   library(knitr)      # from CRAN for generating tables
@@
 prefixdir <- paste0(dbdir,"A_CodeUse/aMSEUse/scenarios/")
-postfixdir <- "M15h75"      # a name for your scenario 
+postfixdir <- "M15h75"      # name your scenario, here natural mortality=0.15, steepness=0.75 
 rundir <- paste0(prefixdir,postfixdir) # files and results directory
@@
 
-# Define the arguments used in teh harvest strategy
+# Define the arguments used in the harvest strategy
 hsargs <- list(mult=0.1,
@@
 See the _aMSEGuide_ for detailed examples.
+This living document, which is also open to on-going development, provides many more details of the operation and structure of the software components used, but also many more examples of scenarios runs and the comparison of such scenarios, which after all, is the point of the software.
+
+The **EGHS** package is a snapshot of the Tasmanian constant reference period harvest strategy as of December 2024.
+It provides a working HS for examples and for any other user to get started.
+Similar example HS are not currently available for South Australia or Victoria.
+The example HS provides the capacity to explore the effects of changing details within the various HS arguments used in Tasmania (as expressed in the _hsargs_ object).
+As with all these details, see the _aMSEGuide_ for details.
+
+### Acknowledgments
+
+All the coding and documentation for this Management Strategy Evaluation (MSE) framework was part of the FRDC project 2019-118, with funding provided by the Fisheries Research Development Corporation, as well as by the University of Tasmania and the South Australian Research and Development Institute (SARDI).
+The occurrence of the Covid 19 pandemic greatly complicated the development and progress of this project.
+The eventual complexity of coding for the MSE, and the coding of each jurisdiction's harvest strategy took far longer than initial estimates.
+This led to significant additional contributions from Malcolm Haddon and Cathy Dichmont.
+
+Malcolm Haddon takes responsibility for the software in the 'haddonm' GitHub based packages used here (and hence any remaining bugs; although for TasHS any remaining bugs are shared with Craig Mundy!).
+Colleagues in the FRDC MSE project, Cathy Dichmont, Craig Mundy, and Owen Burnell were vital in checking that the MSE code operated as intended, asking the right questions and making the right requests and suggestions.
+By providing the necessary scientifically critical context for the work we managed to succeed with what proved to be a very difficult task.
+Writing the R code to capture the workings of the harvest strategies in each jurisdiction was an especially complicated task.
+For Tasmania this was completed by Malcolm Haddon and Craig Mundy, while that for South Australia was a combination of Johnathon Smart, Cathy Dichmont, and Owen Burnell.
+Cathy Dichmont wrote the R code for the Victorian Harvest Strategy by modifying code written by Jo Potts from MRAG.

### 2025-04-20 R/docomparison.R

@@ -75,7 +75,7 @@
 do_comparison <- function(rundir,postfixdir,outdir,files,pickfiles,verbose=TRUE,
                           intensity=100,zero=FALSE,Q90=TRUE,altscenes=NULL,
                           juris="",ribbonleg="topleft",scencol=NULL,...) {
-  # rundir=rundir;postfixdir=postfixdir;outdir=outdir;files=files;pickfiles=c(1,2,3,4)
+  # rundir=rundir;postfixdir=postfixdir;outdir=outdir;files=files;pickfiles=c(2,3,4,5)
   #  verbose=TRUE; intensity=100; zero=FALSE; altscenes=NULL
   #  juris="";ribbonleg="topleft"; Q90=TRUE; scencol=c(1,2,3,4)
   # get files -------------------
@@ -349,37 +349,42 @@ do_comparison <- function(rundir,postfixdir,outdir,files,pickfiles,verbose=TRUE,
 
   # ribbon plots by sau and dynamic variable
   # Catch ribbon tab -------------------------------------------
-  if (verbose) cat("Now doing the final tabs  \n")
+  if (verbose) cat("Now doing the Catch tab  \n")
   catch <- scenebyvar(dyn,byvar="catch",glb=glbc,projonly = TRUE)
   catqnts <- sauquantbyscene(catch,glbc)
   sauribbon(rundir,scenes=scenes,varqnts=catqnts,
               glbc=glbc,varname="Catch",console=FALSE,
               q90=Q90,intens=intensity,addleg=ribbonleg)
   # cpue ribbon tab------------------------------------------
+  if (verbose) cat("Now doing the cpue tab  \n")
   cpue <- scenebyvar(dyn,byvar="cpue",glb=glbc,projonly = TRUE)
   cpueqnts <- sauquantbyscene(cpue,glbc)
   sauribbon(rundir,scenes=scenes,varqnts=cpueqnts,
               glbc=glbc,varname="cpue",console=FALSE,
               q90=Q90,intens=intensity,addleg=ribbonleg)
   # depletion spawnB ribbon tab--------------------------------------------
+  if (verbose) cat("Now doing the depletion spawnB tab  \n")
   deplsB <- scenebyvar(dyn,byvar="deplsB",glb=glbc,projonly=TRUE)
   deplsBqnts <- sauquantbyscene(deplsB,glbc)
   sauribbon(rundir,scenes=scenes,varqnts=deplsBqnts,
               glbc=glbc,varname="deplsB",console=FALSE,
[…]

### 2025-07-02 README.Rmd
New ReadME

@@
 
-## Example
+## Constant Reference Years Harvest Strategy
 
@@
 ```{r example, echo=TRUE, eval=FALSE }
-# a constant TAC example
+# The following is copied from aMSEGuide Chapter 3
 starttime <- (Sys.time())
@@
 })
-dbdir <- getDBdir()   # utility to define the dropbox directory
-# Obviously you should modify the rundir and datadir to suit your own setup
-prefixdir <- paste0(dbdir,"A_CodeUse/aMSEUse/scenarios/")
-postfixdir <- "M15h75"      # name your scenario, here natural mortality=0.15, steepness=0.75 
-rundir <- paste0(prefixdir,postfixdir) # files and results directory
-confirmdir(rundir)
-controlfile <- "controlM15h75.csv"    # name of the control file
-# equilibrium zone -------------------------------------------------------------
+# The next four code lines define my directory structure. 
+# OBVIOUSLY, modify the rundir definition to suit your own setup!!!
+dbdir <- getDBdir()   # utility to identify the dropbox directory
+prefixdir <- pathtopath(dbdir,"A_codeR/aMSEGuide/runs/")
+postfixdir <- "EG"     # a descriptive name for the rundir 
+rundir <- pathtopath(prefixdir,postfixdir) # define rundir
+startime <- Sys.time() # to document the time taken
+verbose <- TRUE        # send messages to the console about the run progress
+controlfile <- paste0("control",postfixdir,".csv") # match control file name
+outdir <- "C:/aMSE_scenarios/EG/"   # storage on a non-cloud hard-drive
+confirmdir(rundir,ask=FALSE)   # make rundir if it does not exist
+confirmdir(outdir,ask=FALSE)   # to be interactive needs ask = TRUE
 # You now need to ensure that there is, at least, a control??.csv, and a 
 # saudata??.csv file in the data directory plus some other data .csv files
-# depending on how conditioned you want the model to be. Templates for the
-# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
+# Templates for the correct format can be produced using ctrlfiletemplate()
+# and datafiletemplate().
 # 
-# Of course, usually one would use data files, control.csv and a saudata.csv, which
-# is listed as the datafile within the control.csv. These must be stored in 
-# rundir. Read the Using_aMSE' section in the Documentation .docx file in the 
-# documentation. To see examples of code to run the MSE for a real question.
+# Of course, usually one would use data files, control.csv and a saudata.csv, 
+# which is listed as the datafile within the control.csv. These must be stored 
+# in rundir. Read the Using_aMSE' section in aMSEGuide.
+controlfile <- "controlEG.csv" # default example file name, change as needed
+# Now make the controlfile. Read the help file using ?ctrlfiletemplate. This
+# explains what the devrec argument is all about.
 
-# Define the arguments used in the harvest strategy
-hsargs <- list(mult=0.1,
-               wid = 4,
-               targqnt = 0.55,
-               maxtarg = c(150,150,150,150,150,150,150,150),
-               pmwts = c(0.65,0.25,0.1),
+ctrlfiletemplate(indir=rundir,filename=controlfile,devrec=0)
+# Within controlfile, the default data file name is 'saudatapostfixdir.csv'. If
+# you want to call it something else (maybe aloysius.csv or machynlleth.csv?)
+# then edit the seventh line of the controlfile that holds the definition
+# of the name of the SAU data file, and change it in this code.
+
+datafiletemplate(indir=rundir,filename="saudataEG.csv")
+# The program also needs some length composition data. For this we are going 
+# to use one of the inbuilt data-sets called 'lfs' (see its help). The default
+# filen is already written into the controlfile. Change that as necessary.
+
+data(lfs)
+writecompdata(indir=rundir,lfs,filen="lf_WZ90-20.csv")
+
+# Define the arguments used in the constantrefhcr harvest strategy
+hsargs <- list(mult=0.1, #expansion factor for cpue range when calc the targqnt
+               wid = 4, # number of years in the grad4 PM
+               targqnt = 0.55, # quantile defining the cpue target
+               maxtarg = c(150,150,150,150,150,150,150,150), # max cpue Target
+               pmwts = c(0.65,0.25,0.1),  # relative weights of PMs
+               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2), # hcr mults
+               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.2,1.25,1.3),# mr 3
+               startCE = 2000, # used in constant reference period HS
+               endCE = 2019,   # used in constant reference period HS
+               metRunder = 0,  # should the metarules be used. 0 = No
+               metRover = 0,   # use metarules 0 = No
+               decrement=1, # use fishery data up to the end of the time series
+               pmwtSwitch = 0, # n years after reaching the targCE to replace
+               stablewts = c(0.8, 0.15, 0.05), # pmwts with stablewts mr3
+               hcrname="constantrefhcr",     # the name of the HCR used
+               printmat=NULL) # something needed for some HS, not TAS
+
+hsargs <- checkhsargs(hsargs) # lists chosen options before run, only for TAS 
+# if you want to check your own set then include a suitable check function 
+# within your jurisdiction's HS source file or package (see EGHS as example).
+
+deleteyrs <- 0  # no years of data removed from the zone
+
+# ?do_MSE provides more detailed descriptions of each function argument
+# see EGHS package for help with the HS functions
+# do_MSE function can take a few minutes to run so be patient
+out <- do_MSE(rundir,controlfile, # already known
+              hsargs=hsargs,    # defined as global object
+              hcrfun=constantrefhcr,   # the main HS function
+              sampleCE=tasCPUE, # processes cpue data
+              sampleFIS=tasFIS, # processes FIS data (if any)
+              sampleNaS=tasNaS, # processes Numbers-at-Size data
+              getdata=tasdata,  # extracts the data from the zoneDP object
+              calcpopC=calcexpectpopC, #distributes catches to populations
+              makeouthcr=makeouthcr, # generates updateable HS output object
+              fleetdyn=NULL,   # only used in SA and VIC so far
+              scoreplot=plotfinalscores, # a function to plot HS scores
+              plotmultflags=plotmultandflags, # function to plot multipliers
+              interimout="", # save the results after projecting,
+              varyrs=7, # years prior to projections for random recdevs
+              startyr=48, # in plots of projections what year to start
+              verbose=TRUE, # send progress reports to the console
+              ndiagprojs=4, # individual trajectories in DiagProj plots
+              cutcatchN = 56,  # cut size-at-catch array < class = 112mm
+              matureL = c(70, 200), # length range for maturity plots
+              wtatL = c(50, 200),   # length range for weight-at-length plots
+              mincount = 120,   # minimum sample size in size-composition data
+              includeNAS = FALSE, #include the numbers-at-size in saved output?
+              depensate=0, #will depensation occur? see do_MSE help for details
+              kobeRP=c(0.4,0.2,0.15), # ref points for a kobe-like phase plot
+              nasInterval=5,  # year interval to use when plotting pred NaS 
+              minsizecomp=c(100,135), #min size for pred sizes, min for catches
+              uplimH=0.35,incH=0.005, #H range when estimating productivity 
+              deleteyrs=0)    # all length comp years used
+
+makeoutput(out,rundir,postfixdir,controlfile,
+           hsfile="EGHS Package",openfile=TRUE,verbose=FALSE)
+
+# unhash this to save the output to your outdir
+# save(out,file=paste0(outdir,postfixdir,".RData"))
+
+```
+
+After running the whole, even if you do not generate the results webpage, to see the structure of the dynamic object generated by the projections you could try:
+
+```{r, echo=TRUE, eval=FALSE }
+str(out$zoneDP,max.level=1)
+str1(out$hcrout)
+
+```
+
+### Constant Catch Harvest Strategy
+
+Within EGHS there are two harvest control rules.
+The code above uses the constant reference years harvest strategy 'constantrefhcr'.
+There is also the 'consthr' which simply implements a constant unvarying aspirational catch in each SAU each year.
+To run that you should first copy the required data files into a new directory.
+As before, setup the R code to match whatever directory structure you intend to use:
+
+```{r,  echo=TRUE, eval=FALSE}
+library(codeutils)
+library(aMSE)
+dropdir <-getDBdir()   # once again FIX this to suit your own directory setup.
+prefixdir <- pathtopath(dropdir,"/A_codeR/aMSETas/scenarios/")
+filelist=c("controlEG.csv","lf_WZ90-20.csv","saudataEG.csv")
+aMSE::copyto(prefixdir=prefixdir,fromdir="EG",todir="EGconst",
+              filelist=filelist)
+# this copies the files across and renames them appropriately.
+```
+
+The change in the harvest strategy is implemented in the hsargs rather than in the code, although it is necessary to let the do_MSE() funciton know of the change:
+
+```{r,  echo=TRUE, eval=FALSE}
+# amend the run code above with the following
+postfixdir <- "EGconst"
+verbose <- TRUE
+rundir <- pathtopath(prefixdir,postfixdir)
+controlfile <- paste0("control",postfixdir,".csv")
+outdir <- "C:/aMSE_scenarios/EG/"
+confirmdir(rundir,ask=FALSE)
+confirmdir(outdir,ask=FALSE)
+
+# current Tas HS
+hsargs <- list(mult=0.025, # cpue range expansion factor when calc the targqnt
+               wid = 4, # number of years in the grad4 PM
+               targqnt = 0.55, # quantile defining the cpue target
+               maxtarg = c(150,150,150,150,150,150,150,150), # max cpue Target
+               acatch =  c(22,48,26,125,105,210,175,47),#ONLY used for consthcr
+               pmwts = c(0.65,0.25,0.10),  # relative weights of PMs default
                hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2), 
-               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.15,1.2,1.25),
-               startCE = 1992,
-               endCE = 2020,
-               metRunder = 2,
-               metRover = 2,
-               decrement=1,
-               pmwtSwitch = 4,
-               stablewts = c(0.4, 0.5, 0.1),
-               hcrname="constantrefhcr",
+               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.2,1.25,1.3),
+               #hcr = c(1,1,1,1,1,1,1,1,1,1,1), # multipliers for constant catch
+               startCE = 2000, # used in constant reference period HS
+               endCE = 2019,   # used in constant reference period HS
+               #refperiodCE = 2000:2019, # use a vector of years instead
+               metRunder = 2,  # MR2 if CE below targ multTAC = 1
+               metRover = 2,   # MR1 CE>targ for 2+ yrs before multTAC increases
+               decrement=1, # use fishery data up to the end of the time series
+               pmwtSwitch = 4, # number of years after reaching the targCE to
+               stablewts = c(0.65, 0.25, 0.1), # replace pmwts with stablewts
+               hcrname="consthcr", # name of HCR: constantrefhcr / consthcr
                printmat=NULL)
@@
 
-deleteyrs <- 0  # none removed in the western zone
+out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=consthcr,  # constantrefhcr,     #
+              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
+              getdata=constdata,calcpopC=calcexpectpopC,makeouthcr=makeoutconst,
+              fleetdyn=NULL,scoreplot=plotfinalscores,
+              plotmultflags=plotmultandflags,interimout="",
+              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
+              matureL=c(40,170),wtatL=c(50,200),mincount=120,
+              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15),
+              nasInterval=1,minsizecomp=c(100,135),uplimH=0.35,incH=0.005,
+              deleteyrs=0)
 
-# One would need to unhash this to get it to run (obviously)
-
-#out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=constantrefhcr,
-#              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
-#              getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
-#              fleetdyn=NULL,scoreplot=plotfinalscores,
-#              plotmultflags=plotmultandflags,interimout="",
-#              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
-#              matureL=c(40,170),wtatL=c(50,200),mincount=120,
-#              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15),
-#              nasInterval=1,minsizecomp=c(100,135),uplimH=0.375,incH=0.005,
-#              deleteyrs=deleteyrs,selectyr=0)
-
-#makeoutput(out,rundir,postfixdir,controlfile,hsfile="TasHS Package",
-#           openfile=TRUE,verbose=FALSE)
-
-# save(out,file=paste0(outdir,postfixdir,".RData"))
+makeoutput(out,rundir,postfixdir,controlfile,hsfile="EGHS Package",
+           openfile=TRUE,verbose=FALSE)
 
@@
 
-After running the whole, even if you do not generate the results webpage, you could also try:
-
-```{r }
-#str(out$zoneDP,max.level=1)
-#str1(out$hcrout)
-
-```
-
-To see the structure of the dynamic object generated by the projections.
+### Final Words
 

### 2025-07-09 README.Rmd
slightly modeified readme

@@
 
-hsargs <- checkhsargs(hsargs) # lists chosen options before run, only for TAS 
+checkhsargs(hsargs) # lists chosen options before run, only for TAS 
 # if you want to check your own set then include a suitable check function 
@@
 library(aMSE)
-dropdir <-getDBdir()   # once again FIX this to suit your own directory setup.
+
+# ONCE AGAIN YOU WILL NEED TO SETUP YOUR OWN DIRECTORY STRUCTURE  TO DEFINE
+# THE prefixdir
+dropdir <-getDBdir()   
 prefixdir <- pathtopath(dropdir,"/A_codeR/aMSETas/scenarios/")
@@
 # this copies the files across and renames them appropriately.
+# NO CHANGES ARE NEEDED IN THE FILES BECAUSE THE CHANGES OCCUR IN hsargs
 ```
 
-The change in the harvest strategy is implemented in the hsargs rather than in the code, although it is necessary to let the do_MSE() funciton know of the change:
+The change in the harvest strategy is implemented in the hsargs rather than in the code, although it is necessary to let the do_MSE() function know of the change:
 
@@
                hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.2,1.25,1.3),
-               #hcr = c(1,1,1,1,1,1,1,1,1,1,1), # multipliers for constant catch
                startCE = 2000, # used in constant reference period HS
                endCE = 2019,   # used in constant reference period HS
-               #refperiodCE = 2000:2019, # use a vector of years instead
                metRunder = 2,  # MR2 if CE below targ multTAC = 1
@@
 
-hsargs <- checkhsargs(hsargs)
+checkhsargs(hsargs)
 
+# NOTE THE CHANGE FROM tasdata TO constdata AND makeouthcr TO makeoutconst 
+# IN THIE THIRD LINE
 out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=consthcr,  # constantrefhcr,     #

### 2025-07-24 README.Rmd
Modified timevarying options. No EWN

@@
 ```{r,  echo=TRUE, eval=FALSE}
-library(codeutils)
-library(aMSE)
-
+suppressPackageStartupMessages({
+  library(aMSE)       # from github.com/haddonm/aMSE MSE codebase
+  library(EGHS)       # from github.com/haddonm/EGHS example Tas harvest strategy
+  library(codeutils)  # from github.com/haddonm/codeutils a set of code utilities
+  library(hplot)      # from github.com/haddonm/hplot a set of plotting routines
+  library(makehtml)   # from github.com/haddonm/makehtml used to display results
+  library(knitr)      # from CRAN for generating tables
+})
 # ONCE AGAIN YOU WILL NEED TO SETUP YOUR OWN DIRECTORY STRUCTURE  TO DEFINE
@@
 
-checkhsargs(hsargs)
+hsargs <- checkhsargs(hsargs)
 
